### 2013-05-22 trunk/doc/paper- towards the internet of things/paper v1.tex
paper added

\documentclass[conference]{IEEEtran}
\ifCLASSINFOpdf
\else
\fi
\hyphenation{op-tical net-works semi-conduc-tor}
\begin{document}

\title{Towards the Internet of Things: A Real-World Testbed}
\author{\IEEEauthorblockN{Nacer Khalil}
\IEEEauthorblockA{School of Electrical and Engineering\\
Al Akhawayn University In Ifrane\\
Email: n.khalil@aui.ma}
\and
\IEEEauthorblockN{Mohammed Reduane Abid}
\IEEEauthorblockA{School of Electrical and Engineering\\
Al Akhawayn University In Ifrane\\
Email: r.abid@aui.ma}
\and
\IEEEauthorblockN{Driss Benhaddou}
\IEEEauthorblockA{School of Engineering and Technology\\
University of Houston\\
Email: dbenhaddou@uh.edu}}
\maketitle
\begin{abstract}

\end{abstract}
\begin{IEEEkeywords}
WSN, 6lowPAN, TinyOs, IOT, cyber physical systems
\end{IEEEkeywords}
\IEEEpeerreviewmaketitle
\section{Introduction}

\section{Literature review}

\section{Internet of things}

\section{6lowpan}

\section{Architecture}

\section{Implementation}

\section{Measurements}

\section{Conclusion}
The conclusion goes here.

\begin{thebibliography}{1}

\bibitem{IEEEhowto:kopka}
H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

\end{thebibliography}

\end{document}

### 2013-06-02 doc/paper- towards the internet of things/paper v1.tex

\documentclass[conference]{IEEEtran}
\ifCLASSINFOpdf
\else
\fi
\hyphenation{op-tical net-works semi-conduc-tor}
\begin{document}

\title{Towards the Internet of Things: A Real-World Testbed}
\author{\IEEEauthorblockN{Nacer Khalil}
\IEEEauthorblockA{School of Electrical and Engineering\\
Al Akhawayn University In Ifrane\\
Email: n.khalil@aui.ma}
\and
\IEEEauthorblockN{Mohammed Reduane Abid}
\IEEEauthorblockA{School of Electrical and Engineering\\
Al Akhawayn University In Ifrane\\
Email: r.abid@aui.ma}
\and
\IEEEauthorblockN{Driss Benhaddou}
\IEEEauthorblockA{School of Engineering and Technology\\
University of Houston\\
Email: dbenhaddou@uh.edu}}
\maketitle
\begin{abstract}
The internet has seen its importance grow exponentially throughout the previous two decades. It has started as set of computers sharing static information written by human beings and accessed by other human beings. It followed mainly the publisher subscriber model. Later on, with the arrival of internet 2.0, the subscriber had now the ability to affect the content at the publisher's level and this basic idea gave rise to a whole new world of applications such as wikis, blogs, social networks and other applications. The next revolution in the internet is the introduction of machines as players within the internet. This will allow the machines to intercommunicate between them or what is known as machine-to-machine communication. More and more devices will be able to communicate and thus be part of the internet. This will give rise to a whole new dimension of possibilities as the focus of the internet will not only be human interactions but also machine interactions. As an example of autonomous machines communicating through, the components of smart grids will be part of the internet of things. Devices such as AMI, smart homes, wireless sensor networks will all be part of the internet of things. Any device within the smart grid will be uniquely identified through the network and therefore will be able to communicate. In this paper, we provide a back-end architecture and a real-world testbed for the smart home communication.
\end{abstract}

\begin{IEEEkeywords}
WSN, 6lowPAN, TinyOS, IOT, cyber physical systems
\end{IEEEkeywords}

%\IEEEpeerreviewmaketitle
\section{Introduction}
The direction that is taken by the global society is a direction leading to what is known by "always connected" model. People are eager to receive the information as fast as possible and whenever available. They also want to send information as fast as possible and whenever wanted. Also, these requests and responses that are dictated by the human are not all the time directed towards humans , but can be sent also to devices and machines but that is not always possible at this time being. The future  of internet goes into this direction: being able to connect human beings as well as machine in order to provide robust and flexible communication for all. The machines on the other hand need an interface to be able to communicate on the internet. In this paper, we create a small network where a machine-to-machine communication is possible. This can be part of what is known as internet of things. The testbed created is composed of a wireless sensor network(WSN), a middleware station and a mobile client within the field of smart grids where data is collected from the motes within the WSN and sent to the middleware. The mobile client accesses this data and the mobile client can send messages directly to any mote within the WSN network in order to turn on or off any appliance. The issues that were handled are the following:
\begin{itemize}
\item The system is developed using heterogeneous networks, different data link layer technologies
\item The WSN uses IPV6 over 6lowPAN whereas other parts of the system uses IPV4
\end{itemize}
To solve these issues, different programs have been used to deal with these issues.

\section{Literature review}

\section{Internet of things}

\section{6lowpan}
6loWPAN stands for  "IPv6 over Low power Wireless Personal Area Networks".

\section{Architecture}

\section{Implementation}
To have a system that is capable of providing a two-way communication between any host and the any mote within the wireless sensor network, four building blocks had to be provided and which are the following:
\begin{itemize}
\item Mote programming
\item Mote sink packet forwarding
\item Network gateway packet transformation
\item Network gateway sensor data server
\end{itemize}

\subsection{Mote programming}
Each mote within the WSN network is equipped with a current transformer that is attached to data acquisition board through which data is read and transformed to the appropriate format and sent to the network host. In addition to this, the appliance is attached to the mote through the relay pins existing within the data acquisition board in the mote. In other words, the mote can control the electricity going to each mote and can allow it or block it. This means that one can control the appliance by using some of the functionalities provided by the mote. From the mote's perspective there are two parts that are implemented within its TinyOS program. A TCP server that is used to receive on/off requests in order to control the mote and a TCP client used to send sensor data.Once the program installed, an IPV6 address is passed to the installation routine in order to assign a static IPV6 address to the mote in which the program is cross-compiled and installed. The TCP server and TCP client work in parallel as each one's traffic is handled separately.

\subsubsection{TCP server}
The TCP server is an important component in the mote's program. Any one willing to control the appliance must connect to the TCP server and send requests. As a mote may control more than one appliance, we then identify the appliance by a unique id and send a zero to turn off or one to turn on. Requests such as "1 1" means to turn on the appliance with id 1.
\subsubsection{TPC Client}
The TCP Client is an important component in the mote's program. It serves as mean to send sensor data to the gateway reliably. Once the mote is turned on, the client connects to a TCP server that is in the gateway, then the consumption data is sensed periodically (once per second) and sent to the TCP server who deals with the sensor data.

\section{Measurements}

\section{Conclusion}
The conclusion goes here.

\begin{thebibliography}{1}

\bibitem{IEEEhowto:kopka}
H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

\end{thebibliography}

\end{document}

### 2013-06-02 programming/netfiltering/include/sniffer.h

@@ -13,14 +13,13 @@
 #include<stdio.h> //For standard things
 #include<stdlib.h>    //malloc
 #include<string.h>    //strlen
-#include<linux/if.h>
+
 #include<netinet/ip_icmp.h>   //Provides declarations for icmp header
 #include<netinet/udp.h>   //Provides declarations for udp header
 #include<netinet/tcp.h>   //Provides declarations for tcp header
 #include<netinet/ip.h>
 #include<netinet/ip6.h>//Provides declarations for ip header
 #include<netinet/if_ether.h>  //For ETH_P_ALL
-#include<linux/if_packet.h>
 #include<net/ethernet.h>  //For ether_header
 #include<sys/socket.h>
 #include<arpa/inet.h>
@@ -28,6 +27,14 @@
 #include<sys/time.h>
 #include<sys/types.h>
 #include<unistd.h>
+#ifdef __unix
+#include<linux/if_packet.h>
+#include<linux/if.h>
+#else
+#include<net/if_packet.h>
+#include<net/if.h>
+
+#endif
 
 
 // Define some constants.
@@ -40,6 +47,8 @@ int sendIPV6Packet(char* , int ,char* );
 int sendIPV4Packet(char*, int ,char* );
 unsigned short int checksum (unsigned short int *, int );
 char* getIPAddr(char* , int );
+char* 6to4Convert(char* , int );
+char* 4to6Convert(char*, int);
 
 
 

### 2013-06-02 programming/netfiltering/src/sniffer.c

@@ -1,5 +1,5 @@
 #include "sniffer.h"
-
+//HI HI
 struct sockaddr_in source,dest;
 int tcp = 0, udp = 0, icmp = 0, others = 0, igmp = 0,total = 0;
 int i,j;
@@ -700,5 +700,16 @@ char* getIPAddr(char* addr , int version)
 		}
 	}
 	return NULL;
+}
+char* convert6to4(char* addr, int len)
+{
+	return NULL;
+}
+
+char* convert4to6(char*addr, int len)
+{
+	return NULL;
+}
+
 }
 

### 2013-06-03 trunk/doc/paper- towards the internet of things/paper v1.tex

@@
 \begin{abstract}
+The internet has seen its importance grow exponentially throughout the previous two decades.
+It has started as set of computers sharing static information written by human beings and accessed by other human beings.
+It followed mainly the publisher subscriber model.
+Later on, with the arrival of internet 2.0, the subscriber had now the ability to affect the content at the publisher's level and this basic idea gave rise to a whole new world of applications such as wikis, blogs, social networks and other applications.
+The next revolution in the internet is the introduction of machines as players within the internet.
+This will allow the machines to intercommunicate between them or what is known as machine-to-machine communication.
+More and more devices will be able to communicate and thus be part of the internet.
+This will give rise to a whole new dimension of possibilities as the focus of the internet will not only be human interactions but also machine interactions.
+As an example of autonomous machines communicating through, the components of smart grids will be part of the internet of things.
+Devices such as AMI, smart homes, wireless sensor networks will all be part of the internet of things.
+Any device within the smart grid will be uniquely identified through the network and therefore will be able to communicate.
+In this paper, we provide a back-end architecture and a real-world testbed for the smart home communication.
+\end{abstract}
 
-\end{abstract}
 \begin{IEEEkeywords}
-WSN, 6lowPAN, TinyOs, IOT, cyber physical systems
+WSN, 6lowPAN, TinyOS, IOT, cyber physical systems
 \end{IEEEkeywords}
-\IEEEpeerreviewmaketitle
+
+%\IEEEpeerreviewmaketitle
 \section{Introduction}
+The direction that is taken by the global society is a direction leading to what is known by "always connected" model.
+People are eager to receive the information as fast as possible and whenever available.
+They also want to send information as fast as possible and whenever wanted.
+Also, these requests and responses that are dictated by the human are not all the time directed towards humans , but can be sent also to devices and machines but that is not always possible at this time being.
+The future  of internet goes into this direction: being able to connect human beings as well as machine in order to provide robust and flexible communication for all.
+The machines on the other hand need an interface to be able to communicate on the internet.
+In this paper, we create a small network where a machine-to-machine communication is possible.
+This can be part of what is known as internet of things.
+The testbed created is composed of a wireless sensor network(WSN), a middleware station and a mobile client within the field of smart grids where data is collected from the motes within the WSN and sent to the middleware.
+The mobile client accesses this data and the mobile client can send messages directly to any mote within the WSN network in order to turn on or off any appliance.
+The issues that were handled are the following:
+\begin{itemize}
+\item The system is developed using heterogeneous networks, different data link layer technologies
+\item The WSN uses IPV6 over 6lowPAN whereas other parts of the system uses IPV4
+\end{itemize}
+To solve these issues, different programs have been used to deal with these issues.
 
@@
 \section{6lowpan}
+6loWPAN stands for  "IPv6 over Low power Wireless Personal Area Networks".
 
@@
 \section{Implementation}
+To have a system that is capable of providing a two-way communication between any host and the any mote within the wireless sensor network, four building blocks had to be provided and which are the following:
+\begin{itemize}
+\item Mote programming
+\item Mote sink packet forwarding
+\item Network gateway packet transformation
+\item Network gateway sensor data server
+\end{itemize}
 
+\subsection{Mote programming}
+Each mote within the WSN network is equipped with a current transformer that is attached to data acquisition board through which data is read and transformed to the appropriate format and sent to the network host.
+In addition to this, the appliance is attached to the mote through the relay pins existing within the data acquisition board in the mote.
+In other words, the mote can control the electricity going to each mote and can allow it or block it.
+This means that one can control the appliance by using some of the functionalities provided by the mote.
+From the mote's perspective there are two parts that are implemented within its TinyOS program.
+A TCP server that is used to receive on/off requests in order to control the mote and a TCP client used to send sensor data.Once the program installed, an IPV6 address is passed to the installation routine in order to assign a static IPV6 address to the mote in which the program is cross-compiled and installed.
+The TCP server and TCP client work in parallel as each one's traffic is handled separately.
+
+\subsubsection{TCP server}
+The TCP server is an important component in the mote's program.
+Any one willing to control the appliance must connect to the TCP server and send requests.
+As a mote may control more than one appliance, we then identify the appliance by a unique id and send a zero to turn off or one to turn on.
+Requests such as "1 1" means to turn on the appliance with id 1.
+\subsubsection{TPC Client}
+The TCP Client is an important component in the mote's program.
+It serves as mean to send sensor data to the gateway reliably.
+Once the mote is turned on, the client connects to a TCP server that is in the gateway, then the consumption data is sensed periodically (once per second) and sent to the TCP server who deals with the sensor data.
+
+\subsection{Mote sink packet forwarding}
+The mote sink packet forwarding module is a special program installed within a mote that is equipped with a USB port that plays the role of a network interface card.
+The mote is attached to the gateway station and has the BLIP (Berkeley Low power IP stack Protocol) module within in.
+In addition it communicates with the gateway using USB protocol.
+In the gateway station, the network interface module is a sort of a IPV6 over USB tunnel which means that if an IPV6 packet packet is going to the sink, it is encapsulated within a USB frame and once it arrives to the mote, the IPV6 packet is extracted and forwarded to the destination mote holding that IPV6 address.
+The other way around is fairly similar, when a mote wants to send an IPV6 packet to the outside world, the mote creates the packet, sends it to the sink that forwards it by encapsulating it into a USB frame.
+
+\subsection{Network gateway packet transformation}
+This is a very component in the system.
+The issue is the following.
+The WSN network supports only IPV6 while other components such as the middleware and the mobile client do not necessary have an IPV6 address, but we still want all the components to communicate independently of the IP technology to be used.
+To do so, we have created a network packet transformation program.
+This program basically converts IPV4 to IPV6 and vice versa.
+To do so, it assigns virtual IPV4 addresses to IPV6 address holders and and IPV6 address to IPV4 address holders.
+With such a program, each player in the network has both an IPV4 and an IPV6 but is aware of only the one that is assigned to him.
+The other "virtual" address is known only at the level of the program installed at the gateway station between the WSN and the outside world.
+The flow of information works as follow; when a station wants to send requests to a mote, it will send IPV4 packet holding the request to the mote but by addressing it using its virtual IPV4 address.
+This packet will be sniffed by this packet transformation
 \section{Measurements}

### 2013-06-03 programming/ipv4-6/code.c

@@ -1,0 +1,115 @@
+/*
+ * code.c
+ *
+ *  Created on: Jun 3, 2013
+ *      Author: nacerkhalil
+ */
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+
+char* toHex(int );
+int toDec(char , char );
+int convert(char );
+
+int main(int argc, char *argv[])
+{
+	int i;
+	char addr[100];
+	char result[100];
+	char* buffer;
+	int version;
+	int chunk;
+	char* val;
+	version = atoi(argv[1]);
+	strcpy(addr,argv[2]);
+	//scanf("%d",&version);
+	//scanf("%s",addr);
+
+	if(version == 4)
+	{
+		//strcat(result,"fec0:");
+		chunk = atoi(strtok(addr,"."));
+		strcat(result,toHex(chunk));
+
+		chunk = atoi(strtok(NULL,"."));
+		strcat(result,toHex(chunk));
+		strcat(result,":");
+
+		chunk = atoi(strtok(NULL,"."));
+		strcat(result,toHex(chunk));
+
+		chunk = atoi(strtok(NULL,"."));
+		strcat(result,toHex(chunk));
+		strcat(result,"::1");
+
+		printf("the result is %s \n",result);
+
+
+	}
+	else if(version == 6)
+	{
+		val = strtok(addr,":");
+
+
+		//val = strtok(NULL,":");
+		buffer = (char *) calloc(3,sizeof(char));
+		chunk = toDec(val[0],val[1]);
+		sprintf(buffer,"%d",chunk);
+		strcat(result,buffer);
+		strcat(result,".");
+
+		buffer = (char *) calloc(3,sizeof(char));
+		chunk = toDec(val[2],val[3]);
+		sprintf(buffer,"%d",chunk);
+		strcat(result,buffer);
+		strcat(result,".");
+
+		val = strtok(NULL,":");
+		buffer = (char *) calloc(3,sizeof(char));
+		chunk = toDec(val[0],val[1]);
+		sprintf(buffer,"%d",chunk);
+		strcat(result,buffer);
+		strcat(result,".");
+
+		buffer = (char *) calloc(3,sizeof(char));
+		chunk = toDec(val[2],val[3]);
+		sprintf(buffer,"%d",chunk);
+		strcat(result,buffer);
+		//strcat(result,".");
+
+		printf("the result is %s \n",result);
+
+
+
+
+
+
+
+	}
+}
+
[…]

### 2013-06-11 doc/paper- towards the internet of things/paper v1.tex

@@
 \and
-\IEEEauthorblockN{Mohammed Reduane Abid}
+\IEEEauthorblockN{Mohammed Riduane Abid}
 \IEEEauthorblockA{School of Electrical and Engineering\\
@@
 Any device within the smart grid will be uniquely identified through the network and therefore will be able to communicate.
-In this paper, we provide a back-end architecture and a real-world testbed for the smart home communication.
+In this paper, we present a real-world testbed for integrating wireless sensors into the internet.
+The solution uses 6lowPAN, and addresses relevant issues at the level of the gateway.
+we used legacy hardware and proved communication.
 \end{abstract}
@@
 \section{Internet of things}
-
+The internet of things is a new concept based on the idea that there will be much more users that are computers, machine than human beings.
+This means that machines will play an important role in the internet and they will be able to communicate autonomously without the interaction of human beings.
+Wireless sensor network will be part of the internet of things and therefore will be actively communicating in the internet by sending sensor data.
+Nowadays, the internet is mainly based on a human-to-human.
+In the future of internet called the internet of things, we will have in addition to human-to-human communication, human-to-machine communication and machine-to-machine communication.
+An example of human-to-machine communication is when the human will be able to control the wireless sensor network's sampling time.
+An example of machine-to-machine communication which will be discussed in this paper is the middleware being able to turn off the light because no one is in the room.
+The middleware by making use of the sensor data will communicate with motes to turn on or off the light bulb.
 \section{6lowpan}
@@
 
+\subsection{Mote sink packet forwarding}
+The mote sink packet forwarding module is a special program installed within a mote that is equipped with a USB port that plays the role of a network interface card.
+The mote is attached to the gateway station and has the BLIP (Berkeley Low power IP stack Protocol) module within in.
+In addition it communicates with the gateway using USB protocol.
+In the gateway station, the network interface module is a sort of a IPV6 over USB tunnel which means that if an IPV6 packet packet is going to the sink, it is encapsulated within a USB frame and once it arrives to the mote, the IPV6 packet is extracted and forwarded to the destination mote holding that IPV6 address.
+The other way around is fairly similar, when a mote wants to send an IPV6 packet to the outside world, the mote creates the packet, sends it to the sink that forwards it by encapsulating it into a USB frame.
+
+\subsection{Network gateway packet transformation}
+This is a very important component in the system.
+The issue is the following.
+The WSN network supports only IPV6 while other components such as the middleware and the mobile client do not necessary have an IPV6 address, but we still want all the components to communicate independently of the IP technology to be used.
+To do so, we have created a network packet transformation program.
+This program basically converts IPV4 to IPV6 and vice versa.
+To do so, it assigns virtual IPV4 addresses to IPV6 address holders and and IPV6 address to IPV4 address holders.
+With such a program, each player in the network has both an IPV4 and an IPV6 but is aware of only the one that is assigned to him.
+The other "virtual" address is known only at the level of the program installed at the gateway station between the WSN and the outside world.
+The flow of information works as follow; when a station wants to send requests to a mote, it will send IPV4 packet holding the request to the mote but by addressing it using its virtual IPV4 address.
+This packet will be sniffed by this packet transformation program that will extract the TCP datagram from the ip packet, create a new IPV6 packet specifying the source address as the virtual address of the host and the destination as the address of the mote and then append the TCP datagram to the newly created packet and send it to the mote sink packet forwarding component that is seen by the gateway as a network interface card.
+This leads to a complicated issue that is to be handled and which is that the program should keep track of the request responses in order to forward them correctly to the destination.
+To solve such a problem, an algorithm has been created whose sole role is to mechanically compute the IPV6 address of the host based on its IPV4 and vice versa.
+This algorithm is based on function whose primary quality is the fact that it is bijective which means any IPV4 address is mapped to one and only one IPV6 function and vice versa.
+So whenever a request is coming, the source and destination address will be converted using this algorithm and this will make avoid the whole request response tracking part.
 \section{Measurements}

### 2013-06-11 programming/ipv4-6/code.c

@@ -34,14 +34,14 @@ int main(int argc, char *argv[])
 
 		chunk = atoi(strtok(NULL,"."));
 		strcat(result,toHex(chunk));
-		strcat(result,":");
+		strcat(result,"::");
 
 		chunk = atoi(strtok(NULL,"."));
 		strcat(result,toHex(chunk));
 
 		chunk = atoi(strtok(NULL,"."));
 		strcat(result,toHex(chunk));
-		strcat(result,"::1");
+		//strcat(result,"::1");
 
 		printf("the result is %s \n",result);
 

### 2013-06-11 trunk/programming/netfiltering/include/my_sniffer.h
versions merged my_sniffer added...

@@ -1,12 +1,12 @@
 /*
- * sniffer.h
+ * my_sniffer.h
  *
- *  Created on: May 1, 2013
+ *  Created on: Jun 3, 2013
  *      Author: banacer
  */
 
-#ifndef SNIFFER_H_
-#define SNIFFER_H_
+#ifndef MY_SNIFFER_H_
+#define MY_SNIFFER_H_
 #include <netinet/in.h>
 #include <errno.h>
 #include <netdb.h>
@@ -36,10 +36,13 @@
 #define IP6_HDRLEN 40         // IPv6 header length
 #define TCP_HDRLEN 20         // TCP header length, excludes options data
 
-void ProcessPacket(unsigned char* , int);
-int sendIPV6Packet(char* , int ,char* );
-int sendIPV4Packet(char*, int ,char* );
+int sendIPV6Packet(char* , int ,char* ,char* );
+int sendIPV4Packet(char*, int ,char* ,char* );
 unsigned short int checksum (unsigned short int *, int );
-char* getIPAddr(char* , int );
+char* toHex(int );
+int toDec(char , char );
+int convert(char );
+char* translate(int , char* );
+char* cleanIPV6(char* );
 
-#endif /* SNIFFER_H_ */
+#endif /* MY_SNIFFER_H_ */

### 2013-06-11 programming/GatewayServer/src/Code.java
versions merged my_sniffer added...

@@ -1,7 +1,7 @@
 import java.io.BufferedWriter;
 import java.io.OutputStreamWriter;
 import java.net.*;
-
+//hey banacer
 public class Code
 {
         public static void main(String[] args)

### 2013-06-11 programming/netfiltering/src/my_sniffer.c

@@ -1,155 +1,126 @@
-#include "sniffer.h"
-
-struct sockaddr_in source,dest;
-int tcp = 0, udp = 0, icmp = 0, others = 0, igmp = 0,total = 0;
-int i,j;
-char map[10][2][100];
-int map_size;
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include "my_sniffer.h"
 
 int main()
 {
-	FILE* map_file;
-    int saddr_size , data_size;
-    struct sockaddr saddr;
-    unsigned char *buffer = (unsigned char *) malloc(65536);
+	int sockfd,retval,data_size;
+	char *buf;
+	char *addr;
+	char *src_addr;
+	char *dest_addr;
 
+	socklen_t clilen;
+	struct sockaddr_in cliaddr;
+	struct sockaddr_in *in;
+	struct sockaddr_in6 *in6;
+	struct iphdr *ip_hdr;
+	struct ip6_hdr *ip6_hdr;
+	struct ethhdr *header;
 
-    printf("Starting...\n");
-    map_file = fopen("data/ipmapping.dat","r");
-    i = 0;
-    while(!feof(map_file))
+	sockfd = socket( PF_PACKET, SOCK_PACKET , htons(ETH_P_ALL));
+	if (sockfd < 0)
 	{
-    	fscanf(map_file,"%s , %s",map[i][0],map[i][1]);
-    	i++;
+		perror("sock:");
+		exit(1);
 	}
-    map_size = i;
+	buf = (char *)calloc(10000,sizeof(char));
 
-    int sock_raw = socket( PF_PACKET, SOCK_PACKET , htons(ETH_P_ALL)) ;
+	clilen = sizeof(struct sockaddr_in);
+	while(1)
+	{
+		header = (struct ethhdr* ) buf;
 
-    if(sock_raw < 0)
-    {
-        perror("Socket Error");
-        return 1;
-    }
-    while(1)
-    {
-        saddr_size = sizeof(saddr);
-        //Receive a packet
-        data_size = recvfrom(sock_raw , buffer , 65536 , 0 , &saddr , (socklen_t*)&saddr_size);
-        if(data_size < 0 )
-        {
-            printf("Recvfrom error , failed to get packets\n");
-            return 1;
-        }
-        //Now process the packet
-        ProcessPacket(buffer , data_size);
-    }
-    close(sock_raw);
-    free(buffer);
-    printf("Finished");
[…]

### 2013-06-11 programming/netfiltering/src/test.c

@@ -1,0 +1,71 @@
+/*
+ * test.c
+ *
+ *  Created on: Jun 6, 2013
+ *      Author: banacer
+ */
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+
+char* cleanIPV6(char* addr)
+{
+	char* result;
+	char* buffer;
+	int chunk;
+	char *val1, *val2;
+	char* buff;
+	int i;
+	int count = 1;
+	result = (char *) calloc(16,sizeof(char));
+	val1 = strtok(addr,":");
+	val2 = (char*) malloc(sizeof(char));
+	while(val2 != NULL)
+	{
+		val2 = strtok(NULL,":");
+		if(strlen(val1) < 4)
+		{
+			for(i = 0; i < 4 - strlen(val1); i++)
+			{
+				strcat(result,"0");
+			}
+			strcat(result,val1);
+			strcat(result,":");
+		}
+		else
+		{
+			if( val2 == NULL)
+			{
+				strcat(result,":");
+				for(i = 0; i < strlen(val1); i++)
+				{
+					if(val1[i] != '0')
+					{
+						buff = (char *) calloc(2,sizeof(char));
+						buff[0] = val1[i];
+						buff[1] = '\0';
+						strcat(result,buff);
+					}
+
+				}
+			}
+			else
+			{
+				strcat(result,val1);
+				strcat(result,":");
+			}
+		}
+		//printf("val1 = %s , val2 = %s \n",val1,val2);
+		if(val2 != NULL)
+			strcpy(val1,val2);
+		count++;
+	}
+	printf("result is %s\n", result);
+	return result;
+
+}
+
+int main(int argc, char *argv[])
+{
+	cleanIPV6(argv[1]);
+}

### 2013-06-11 doc/paper- towards the internet of things/paper v1.tex

@@
 \documentclass[conference]{IEEEtran}
-\ifCLASSINFOpdf
-\else
-\fi
-\hyphenation{op-tical net-works semi-conduc-tor}
+\usepackage{graphicx}
+
 \begin{document}
@@
 \and
-\IEEEauthorblockN{Mohammed Riduane Abid}
+\IEEEauthorblockN{Mohamed Riduan Abid}
 \IEEEauthorblockA{School of Electrical and Engineering\\
@@
 6loWPAN stands for  "IPv6 over Low power Wireless Personal Area Networks".
+It stands for the name of the working group within IETF.
+It is based on the idea that all the devices independently of their processing power or energy resources should be able to have the TCP-IP protocol stack and be part of the internet of things.
+In order to be able to build the T
+CP-IP protocol stack in low-power devices, there have been multiple aspects of the protocol stack that has to be optimized. As an example, the IP MTU(maximum transmission unit)
 
 \section{Architecture}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.5]{images/network_diagram.jpg}
+\caption{Network diagram}
+\label{fig:network}
+\end{figure}
 

### 2013-06-13 doc/paper- towards the internet of things/paper v1.tex

--- 2013-06-11 doc/paper- towards the internet of things/paper v1.tex
@@
 \item The system is developed using heterogeneous networks, different data link layer technologies
-\item The WSN uses IPV6 over 6lowPAN whereas other parts of the system uses IPV4
+\item The WSN uses IPV6 over 6lowPAN whereas other components of the system use IPV4
 \end{itemize}
@@
 6loWPAN stands for  "IPv6 over Low power Wireless Personal Area Networks".
+It stands for the name of the working group within IETF.
+It is based on the idea that all the devices independently of their processing power or energy resources should be able to have the TCP/IP protocol stack and be part of the internet of things.
+In order to be able to build the TCP/IP protocol stack in low-power devices, there have been multiple aspects of the protocol stack that has to be optimized.
+As an example, the IP MTU(maximum transmission unit) is fixed at 1280 bytes whereas the Zigbee MTU is only 127 bytes.
+This means that not every IPV6 packet can be inserted within a Zigbee frame.
+Another issue is related to the addressing with the 128 bits address.
+in 6lowpan, the addressing via the IPV6 address is performed hierarchically to first identify is the packet is sent to a device within that network but afterwards, addressing is done by counting mainly on the 16 last bits of the IPV6 address.
+These were just two of a large set of issues that 6lowpan tries to solve in order to enable even the low-power devices to join the internet of things.
+TinyOS as being one of the most used operating system comes with its implementation of 6lowpan through a project called BLIP (Berkeley Low-power IP stack) that provides a lightweight implementation of the 6lowpan and that can be installed and used with a variety of motes as long as they support IPV6.
+This project makes use of BLIP to provide the TCP/IP protocol stack to the devices within the wireless sensor network.
+\section{Architecture}
+The project is composed of three building blocks. 
+\begin{itemize}
+\item Wireless sensor network (WSN)
+\item Gateway server
+\item Middleware
+\item Mobile client
+\end{itemize}
+The WSN communicates using zigbee as well as with the gateway server using IPV6 technology whereas whereas the communication between the gateway server, the middleware and the mobile client is done via Wifi using IPV4 technology.
+The architecture enables any component within this system to communicate with any other system independently of the communication medium used (Zigbee or Wifi) or the network technology used (IPV4 or IPV6).
+Figure 1 is a network diagram that depicts the different components of the system as well as the routes that exist for any component to communicate with another.
 
-\section{Architecture}
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.5]{images/network_diagram.jpg}
+\caption{Network diagram}
+\label{fig:network}
+\end{figure}
 
+\subsection{Wireless sensor network}
+The wireless sensor network is based on the mesh topology.
+It is also designed in order to be an ad-hoc network.
+This means that you can place the mote any where as long as there is one link it can attain to communicate with.
+The routing is multi-hop and links are created dynamically between different motes of the WSN.
 \section{Implementation}
@@
 Requests such as "1 1" means to turn on the appliance with id 1.
-\subsubsection{TPC Client}
+\subsubsection{TCP Client}
 The TCP Client is an important component in the mote's program.

### 2013-06-13 trunk/doc/paper- towards the internet of things/paper v1.tex

@@
 \and
-\IEEEauthorblockN{Mohammed Reduane Abid}
+\IEEEauthorblockN{Mohamed Riduan Abid}
 \IEEEauthorblockA{School of Electrical and Engineering\\
@@
 \section{Internet of things}
-
+The internet of things is a new concept based on the idea that there will be much more users that are computers, machine than human beings.
+This means that machines will play an important role in the internet and they will be able to communicate autonomously without the interaction of human beings.
+Wireless sensor network will be part of the internet of things and therefore will be actively communicating in the internet by sending sensor data.
+Nowadays, the internet is mainly based on a human-to-human.
+In the future of internet called the internet of things, we will have in addition to human-to-human communication, human-to-machine communication and machine-to-machine communication.
+An example of human-to-machine communication is when the human will be able to control the wireless sensor network's sampling time.
+An example of machine-to-machine communication which will be discussed in this paper is the middleware being able to turn off the light because no one is in the room.
+The middleware by making use of the sensor data will communicate with motes to turn on or off the light bulb.
 \section{6lowpan}
 6loWPAN stands for  "IPv6 over Low power Wireless Personal Area Networks".
+It stands for the name of the working group within IETF.
+It is based on the idea that all the devices independently of their processing power or energy resources should be able to have the TCP/IP protocol stack and be part of the internet of things.
+In order to be able to build the TCP/IP protocol stack in low-power devices, there have been multiple aspects of the protocol stack that has to be optimized.
+As an example, the IP MTU(maximum transmission unit) is fixed at 1280 bytes whereas the Zigbee MTU is only 127 bytes.
+This means that not every IPV6 packet can be inserted within a Zigbee frame.
+Another issue is related to the addressing with the 128 bits address.
+in 6lowpan, the addressing via the IPV6 address is performed hierarchically to first identify is the packet is sent to a device within that network but afterwards, addressing is done by counting mainly on the 16 last bits of the IPV6 address.
+These were just two of a large set of issues that 6lowpan tries to solve in order to enable even the low-power devices to join the internet of things.
+TinyOS as being one of the most used operating system comes with its implementation of 6lowpan through a project called BLIP (Berkeley Low-power IP stack) that provides a lightweight implementation of the 6lowpan and that can be installed and used with a variety of motes as long as they support IPV6.
+This project makes use of BLIP to provide the TCP/IP protocol stack to the devices within the wireless sensor network.
+\section{Architecture}
+The project is composed of three building blocks. 
+\begin{itemize}
+\item Wireless sensor network (WSN)
+\item Gateway server
+\item Middleware
+\item Mobile client
+\end{itemize}
+The WSN communicates using zigbee as well as with the gateway server using IPV6 technology whereas whereas the communication between the gateway server, the middleware and the mobile client is done via Wifi using IPV4 technology.
+The architecture enables any component within this system to communicate with any other system independently of the communication medium used (Zigbee or Wifi) or the network technology used (IPV4 or IPV6).
+Figure 1 is a network diagram that depicts the different components of the system as well as the routes that exist for any component to communicate with another.
 
-\section{Architecture}
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.5]{images/network_diagram.jpg}
+\caption{Network diagram}
+\label{fig:network}
+\end{figure}
+
+\subsection{Wireless sensor network}
+The wireless sensor network is based on the mesh topology.
+It is also designed in order to be an ad-hoc network.
+This means that you can place the mote any where as long as there is one link it can attain to communicate with.
+The routing is multi-hop and links are created dynamically between different motes of the WSN provided that there frames can reach the destination.
+
+\subsection{Gateway server}
+The gateway server is a crucial component within the system.
+It is responsible for extracting Wifi frames and forwarding them as Zigbee frames and vice versa.
+It is also responsible for receiving IPV4 packet and transforming them into IPV6 and vice versa.
+The gateway server has other functionalities which are the ability to receive sensor data from all the motes in the wireless sensor network and forwarding them to the middleware.
+In case the link between the gateway server and the middleware is lost, the gateway server stores the received data and sends them when the link is up again.
+This feature serves as a way to avoid losing data.
+
+\subsection{Middleware}
+The middleware is a special program that is used to hide the heterogeneity of the system from external users.
+The middleware also provides automation mechanisms to control and reduce the energy consumption, but the main features are the ability to receive data, filter it, transform it and store it in a coherent fashion in order to make the most of this data and use it smartly to reduce consumption.
+Concerning the client access, the middleware provides a set of web services that enable the clients to access all sorts of information such as the real-time consumption, daily consumption and other information inferred from the sensor data coming from the wireless sensor network.
+
+\subsection{Mobile client}
+The mobile client application is an interactive application for android phones that enable users to get access to the real-time consumption of their homes and also to remotely control their appliance by turning on and off any appliance linked to some mote in the wireless sensor network.
+The mobile client when wanting to turn on or off some appliance sends a request directly towards the mote responsible for that action and designate the mote using its "vitual" IPV4 address that it does not own in really since it has an IPV6 address but that virtual IPV4 address is reserved for that mote and the translation is dealt with at the level of the gateway.
+
+\ Now that all components have been explained, the data flow of the information is to be explained.
+As it was stated before, the any component of the system can communicate with any other independently of the data link layer technology or network layer technology.
 
@@
 \subsection{Network gateway packet transformation}
-This is a very component in the system.
+This is a very important component in the system.
 The issue is the following.
@@
 The flow of information works as follow; when a station wants to send requests to a mote, it will send IPV4 packet holding the request to the mote but by addressing it using its virtual IPV4 address.
-This packet will be sniffed by this packet transformation
+This packet will be sniffed by this packet transformation program that will extract the TCP datagram from the ip packet, create a new IPV6 packet specifying the source address as the virtual address of the host and the destination as the address of the mote and then append the TCP datagram to the newly created packet and send it to the mote sink packet forwarding component that is seen by the gateway as a network interface card.
+This leads to a complicated issue that is to be handled and which is that the program should keep track of the request responses in order to forward them correctly to the destination.
+To solve such a problem, an algorithm has been created whose sole role is to mechanically compute the IPV6 address of the host based on its IPV4 and vice versa.
+This algorithm is based on function whose primary quality is the fact that it is bijective which means any IPV4 address is mapped to one and only one IPV6 function and vice versa.
+So whenever a request is coming, the source and destination address will be converted using this algorithm and this will make avoid the whole request response tracking part.
 \section{Measurements}

### 2013-06-14 doc/paper- towards the internet of things/paper v1.tex

@@
 This means that you can place the mote any where as long as there is one link it can attain to communicate with.
-The routing is multi-hop and links are created dynamically between different motes of the WSN.
+The routing is multi-hop and links are created dynamically between different motes of the WSN provided that there frames can reach the destination.
+
+\subsection{Gateway server}
+The gateway server is a crucial component within the system.
+It is responsible for extracting Wifi frames and forwarding them as Zigbee frames and vice versa.
+It is also responsible for receiving IPV4 packet and transforming them into IPV6 and vice versa.
+The gateway server has other functionalities which are the ability to receive sensor data from all the motes in the wireless sensor network and forwarding them to the middleware.
+In case the link between the gateway server and the middleware is lost, the gateway server stores the received data and sends them when the link is up again.
+This feature serves as a way to avoid losing data.
+
+\subsection{Middleware}
+The middleware is a special program that is used to hide the heterogeneity of the system from external users.
+The middleware also provides automation mechanisms to control and reduce the energy consumption, but the main features are the ability to receive data, filter it, transform it and store it in a coherent fashion in order to make the most of this data and use it smartly to reduce consumption.
+Concerning the client access, the middleware provides a set of web services that enable the clients to access all sorts of information such as the real-time consumption, daily consumption and other information inferred from the sensor data coming from the wireless sensor network.
+
+\subsection{Mobile client}
+The mobile client application is an interactive application for android phones that enable users to get access to the real-time consumption of their homes and also to remotely control their appliance by turning on and off any appliance linked to some mote in the wireless sensor network.
+The mobile client when wanting to turn on or off some appliance sends a request directly towards the mote responsible for that action and designate the mote using its "virtual" IPV4 address that it does not own in really since it has an IPV6 address but that virtual IPV4 address is reserved for that mote and the translation is dealt with at the level of the gateway.
+
+\ Now that all components have been explained, the data flow of the information is to be explained.
+As it was stated before, the any component of the system can communicate with any other independently of the data link layer technology or network layer technology.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[height=110mm,width=90mm]{images/control_appliance_data_flow.jpg}
+\caption{Data flow diagram for the mobile client sending on/off commands}
+\label{fig:control_appliance}
+\end{figure}
+
+\
+As one of the main goals of this project is to provide a two-way communication between the client and sensor, we will explain both way's flow of information.
+Figure 2 depicts the flow of information going from the client to the appliance.
+As the mobile client in the system is connected to a Wifi network that uses IPV4 and the mote is a network that uses a IPV6, then there should be a process that does some transformation to incoming and outgoing packets.
+So the client sends the packet to the virtual IPV4 address of the mote.
+The gateway receives it translates the virtual IPV4 address into the real IPV6 address of the mote and sets as source address the virtual IPV6 address of the mobile client.
+The new IPV6 packet is created holding the same payload as the original IPV4 packet.
+This new IPV6 packet is to be forwarded to the wireless sensor network.
+To do so, it is passed to the tunnel that encapsulates the packet into a USB frame and sends it to the mote sink that extracts the IPV6 packet from the USB frame and encapsulates it into a Zigbee frame.
+Once the Zigbee frame arrives to the destination mote, the TCP datagram is extracted from the IPV6 packet and passed to the TCP server of the mote that reads the message and executes it by turning on/off the appliance using I2C protocol.
+
+\
+As a second part of the two-way communication, the mote sends periodically sensor data to the middleware but again the communication is goes through several steps as depicted in figure 3.
+\linebreak 
+The mote reads periodically sends sensor data from the sensor, transforms the data and sends it.
+To send it the mote client connects to a TCP server hosted at the gateway.
+The IPV6 packet is encapsulated in a Zigbee frame that is forwarded to the mote  sink that extracts the IPV6 packet and  encapsulates it into a USB frame and forwarded to the gateway that reads the TCP datagram.
+Once the sensor data is at the gateway, the gateway forwards the sensor data to the middleware and if the link is down, the sensor data is temporarily stored in a database hosted in the gateway and once the link is up, all the stored data is sent to the middleware and cleared from the database.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[height=110mm,width=90mm]{images/send_sensor_data_data_flow.jpg}
+\caption{Data flow diagram for the mobile client sending on/off commands}
+\label{fig:control_appliance}
+\end{figure}
+
 \section{Implementation}

### 2013-06-14 trunk/doc/paper- towards the internet of things/paper v1.tex

@@
 The mobile client application is an interactive application for android phones that enable users to get access to the real-time consumption of their homes and also to remotely control their appliance by turning on and off any appliance linked to some mote in the wireless sensor network.
-The mobile client when wanting to turn on or off some appliance sends a request directly towards the mote responsible for that action and designate the mote using its "vitual" IPV4 address that it does not own in really since it has an IPV6 address but that virtual IPV4 address is reserved for that mote and the translation is dealt with at the level of the gateway.
+The mobile client when wanting to turn on or off some appliance sends a request directly towards the mote responsible for that action and designate the mote using its "virtual" IPV4 address that it does not own in really since it has an IPV6 address but that virtual IPV4 address is reserved for that mote and the translation is dealt with at the level of the gateway.
 
@@
 As it was stated before, the any component of the system can communicate with any other independently of the data link layer technology or network layer technology.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[height=110mm,width=90mm]{images/control_appliance_data_flow.jpg}
+\caption{Data flow diagram for the mobile client sending on/off commands}
+\label{fig:control_appliance}
+\end{figure}
+
+\
+As one of the main goals of this project is to provide a two-way communication between the client and sensor, we will explain both way's flow of information.
+Figure 2 depicts the flow of information going from the client to the appliance.
+As the mobile client in the system is connected to a Wifi network that uses IPV4 and the mote is a network that uses a IPV6, then there should be a process that does some transformation to incoming and outgoing packets.
+So the client sends the packet to the virtual IPV4 address of the mote.
+The gateway receives it translates the virtual IPV4 address into the real IPV6 address of the mote and sets as source address the virtual IPV6 address of the mobile client.
+The new IPV6 packet is created holding the same payload as the original IPV4 packet.
+This new IPV6 packet is to be forwarded to the wireless sensor network.
+To do so, it is passed to the tunnel that encapsulates the packet into a USB frame and sends it to the mote sink that extracts the IPV6 packet from the USB frame and encapsulates it into a Zigbee frame.
+Once the Zigbee frame arrives to the destination mote, the TCP datagram is extracted from the IPV6 packet and passed to the TCP server of the mote that reads the message and executes it by turning on/off the appliance using I2C protocol.
+
+\
+As a second part of the two-way communication, the mote sends periodically sensor data to the middleware but again the communication is goes through several steps as depicted in figure 3.
+\linebreak 
+The mote reads periodically sends sensor data from the sensor, transforms the data and sends it.
+To send it the mote client connects to a TCP server hosted at the gateway.
+The IPV6 packet is encapsulated in a Zigbee frame that is forwarded to the mote  sink that extracts the IPV6 packet and  encapsulates it into a USB frame and forwarded to the gateway that reads the TCP datagram.
+Once the sensor data is at the gateway, the gateway forwards the sensor data to the middleware and if the link is down, the sensor data is temporarily stored in a database hosted in the gateway and once the link is up, all the stored data is sent to the middleware and cleared from the database.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[height=110mm,width=90mm]{images/send_sensor_data_data_flow.jpg}
+\caption{Data flow diagram for the mobile client sending on/off commands}
+\label{fig:control_appliance}
+\end{figure}
 
@@
 The other "virtual" address is known only at the level of the program installed at the gateway station between the WSN and the outside world.
+\\
  The flow of information works as follow; when a station wants to send requests to a mote, it will send IPV4 packet holding the request to the mote but by addressing it using its virtual IPV4 address.
@@
 This leads to a complicated issue that is to be handled and which is that the program should keep track of the request responses in order to forward them correctly to the destination.
+\\
 To solve such a problem, an algorithm has been created whose sole role is to mechanically compute the IPV6 address of the host based on its IPV4 and vice versa.
@@
 
-\begin{thebibliography}{1}
-
-\bibitem{IEEEhowto:kopka}
-H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
-  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
-
-\end{thebibliography}
+\bibliographystyle{plain}
+\bibliography{myrefs}
 

### 2013-06-18 doc/paper- towards the internet of things/paper v1.tex

@@
 \usepackage{graphicx}
+\usepackage[noadjust]{cite}
 
@@
 
-%\IEEEpeerreviewmaketitle
 \section{Introduction}
@@
 \section{Literature review}
-
+Smart grid as being one of the most important applications of the internet of things makes use of different technologies such as Zigbee in the WSN \cite{ref1}.
+\cite{ref2} discusses the need to introduce IPV6 within the wireless sensor networks and discusses the the existing approaches as well as the issues related to introducing IPV6 on top of Zigbee.
+Such issues are fragmentation, frame size, addressing, and security issues.
+This paper introduces a real testbed that includes the whole TCP/IP protocol provided by the BLIP project that takes into consideration most of those issues and implement the two-way communication as needed by smart grids and measures the performance of such a system.
 \section{Internet of things}
@@
 \subsection{Wireless sensor network}
-The wireless sensor network is based on the mesh topology.
+The hardware is composed of motes of type crossbow MPR2600 that are used to build the wireless sensor network as shown in figure 2.From the network aspect, the wireless sensor network is based on the mesh topology.
 It is also designed in order to be an ad-hoc network.
@@
 The routing is multi-hop and links are created dynamically between different motes of the WSN provided that there frames can reach the destination.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.5]{images/micaz.png}
+\caption{Crossbow MRP2600}
+\label{fig:micaz}
+\end{figure}
 
@@
 As one of the main goals of this project is to provide a two-way communication between the client and sensor, we will explain both way's flow of information.
-Figure 2 depicts the flow of information going from the client to the appliance.
+Figure 3 depicts the flow of information going from the client to the appliance.
 As the mobile client in the system is connected to a Wifi network that uses IPV4 and the mote is a network that uses a IPV6, then there should be a process that does some transformation to incoming and outgoing packets.
@@
 \
-As a second part of the two-way communication, the mote sends periodically sensor data to the middleware but again the communication is goes through several steps as depicted in figure 3.
+As a second part of the two-way communication, the mote sends periodically sensor data to the middleware but again the communication is goes through several steps as depicted in figure 4.
 \linebreak 
@@
 
-\begin{thebibliography}{1}
-
-\bibitem{IEEEhowto:kopka}
-H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
-  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
-
-\end{thebibliography}
+%Bibliography
+\mbox{}
+\nocite{*}
+\bibliographystyle{IEEEtran} 
+\bibliography{IEEEabrv,myrefs}
 

### 2013-06-18 doc/paper- towards the internet of things/paper v1.tex

@@
 This paper introduces a real testbed that includes the whole TCP/IP protocol provided by the BLIP project that takes into consideration most of those issues and implement the two-way communication as needed by smart grids and measures the performance of such a system.
+El Kouche also investigates on the widely used architectures and technologies and explains which architecture is the most suitable for a WSN within the internet of things \cite{ref3}.
+\cite{ref4} discusses the requirement of the IOT gateway as well as a proposed architecture for the system to be deployed in the gateway.
 \section{Internet of things}

### 2013-06-24 trunk/doc/paper- towards the internet of things/paper v1.tex
THESIS DIR ADDED WITH AN INITIAL OUTLINE OF THE REPORT

@@
 The internet has seen its importance grow exponentially throughout the previous two decades.
-It has started as set of computers sharing static information written by human beings and accessed by other human beings.
+It has started as a set of computers sharing static information written by human beings and accessed by other human beings.
 It followed mainly the publisher subscriber model.
@@
 The future  of internet goes into this direction: being able to connect human beings as well as machine in order to provide robust and flexible communication for all.
-The machines on the other hand need an interface to be able to communicate on the internet.
+The machines on the other hand need an interface to be able to communicate on the internet.There will be millions of devices in the internet both producing and consuming information \cite{ref8}.
 In this paper, we create a small network where a machine-to-machine communication is possible.
-This can be part of what is known as internet of things.
+This can be part of what is known as internet of things \cite{ref6}.
 The testbed created is composed of a wireless sensor network(WSN), a middleware station and a mobile client within the field of smart grids where data is collected from the motes within the WSN and sent to the middleware.
@@
 \section{Literature review}
+Smart grid as being one of the most important applications of the internet of things makes use of different technologies such as Zigbee in the WSN \cite{ref1}.
+\cite{ref2} discusses the need to introduce IPV6 within the wireless sensor networks and discusses the the existing approaches as well as the issues related to introducing IPV6 on top of Zigbee.
+Such issues are fragmentation, frame size, addressing, and security issues.
+This paper introduces a real testbed that includes the whole TCP/IP protocol provided by the BLIP project that takes into consideration most of those issues and implement the two-way communication as needed by smart grids and measures the performance of such a system.
+El Kouche also investigates on the widely used architectures and technologies and explains which architecture is the most suitable for a WSN within the internet of things \cite{ref3}.
+\cite{ref4} discusses the requirement of the IOT gateway as well as a proposed architecture for the system to be deployed in the gateway.
+A rather similar architecture to what is done in this paper is proposed by Yerra, Bharathi, Rajalakshmi and Desai \cite{ref7}.
+They also use GSM technology to transmit information whereas in our case, internet is used for this purpose.
+\
 
@@
 
-\bibliographystyle{plain}
-\bibliography{myrefs}
+%Bibliography
+\mbox{}
+\nocite{*}
+\bibliographystyle{IEEEtran} 
+\bibliography{IEEEabrv,myrefs}
 

### 2013-06-24 doc/thesis/report.tex
THESIS DIR ADDED WITH AN INITIAL OUTLINE OF THE REPORT

\documentclass[12pt,a4paper,final]{report}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\newcommand{\HRule}{\rule{\linewidth}{0.5mm}}

\author{Nacer Khalil}
\title{An internet of things testbed for the smart grid application}

\begin{document}

\begin{titlepage}
\begin{center}
\includegraphics[width=0.6\textwidth]{img/aui_logo.jpg}
%\includegraphics[width=0.45\textwidth]{img/tum_logo.jpg}
\linebreak
\linebreak
\linebreak
\linebreak
\linebreak
\linebreak
% Title
\HRule \\[0.4cm]
{ \huge \bfseries An Internet Of Things Testbed For The Smart Grid Application}\\[0.4cm]

\HRule \\[1.5cm]

% Author and supervisor
\begin{minipage}{0.4\textwidth}
\begin{flushleft} \large
\emph{Author:}\\
Nacer \textsc{Khalil}
\end{flushleft}
\end{minipage}
\begin{minipage}{0.4\textwidth}
\begin{flushright} \large
\emph{Supervisor:} \\
Dr.~Riduan \textsc{Abid} \\
Dr.~Driss \textsc{Benhaddou} \\
Dr.~Michael \textsc{Gerndt} 
\end{flushright}
\end{minipage}

\vfill

% Bottom of the page
{\large \today}

\end{center}
\end{titlepage}

\chapter{Abstract}

\chapter{Acknowledgements}

%THIS IS THE TABLE OF CONTENT PART
\tableofcontents
\newpage
\chapter{Introduction}

\section{Problem statement}
\section{Purpose of study}
\section{Importance of study}
\section{scope of study}
\section{Rationale of study}
\section{Definition of terms}

\chapter{Literature review}

\chapter{Methodology}

\chapter{Findings}

\chapter{Conclusion}

\chapter{Recommendations}

\mbox{}
\nocite{*}
\bibliographystyle{IEEEtran} 
\bibliography{IEEEabrv,myrefs}
\newpage
\listoffigures
\chapter{Appendix}

\end{document}

### 2013-06-26 trunk/doc/thesis/report.tex

--- 2013-06-24 doc/thesis/report.tex
@@
-\documentclass[12pt,a4paper,final]{report}
+\documentclass[oneside,12pt,a4paper,final]{book}
 \usepackage[utf8]{inputenc}
@@
 \usepackage{graphicx}
+\usepackage{setspace}
 \newcommand{\HRule}{\rule{\linewidth}{0.5mm}}
+%\usepackage{hyperref}
+\usepackage[acronym]{glossaries}
 
@@
 
+\makeglossaries
+
 \begin{document}
+\doublespacing
+\input{acronym}
+\frontmatter
 
-\begin{titlepage}
-\begin{center}
-\includegraphics[width=0.6\textwidth]{img/aui_logo.jpg}
-%\includegraphics[width=0.45\textwidth]{img/tum_logo.jpg}
-\linebreak
-\linebreak
-\linebreak
-\linebreak
-\linebreak
-\linebreak
-% Title
-\HRule \\[0.4cm]
-{ \huge \bfseries An Internet Of Things Testbed For The Smart Grid Application}\\[0.4cm]
+%COVER PAGE
+\input{cover_page}
 
-\HRule \\[1.5cm]
-
-% Author and supervisor
-\begin{minipage}{0.4\textwidth}
-\begin{flushleft} \large
-\emph{Author:}\\
-Nacer \textsc{Khalil}
-\end{flushleft}
-\end{minipage}
-\begin{minipage}{0.4\textwidth}
-\begin{flushright} \large
-\emph{Supervisor:} \\
-Dr.~Riduan \textsc{Abid} \\
-Dr.~Driss \textsc{Benhaddou} \\
-Dr.~Michael \textsc{Gerndt} 
-\end{flushright}
-\end{minipage}
-
-\vfill
-
-% Bottom of the page
-{\large \today}
-
-\end{center}
-\end{titlepage}
+\chapter{Acknowledgements}
+I am deeply thankful to my supervisor Dr.
+Mohamed Riduan Abid for his very close supervision, guidance and for the advice he kept giving me during the lifetime of the project.
+Your advices whether they are computer science related, or even every day life related wise words things I will never forget.
+\\
+I am also thankful for having Dr.
+Driss Benhaddou as a supervisor.
+His advice, insight, guidance and very enlightening ideas.
+I wouldn't have been so interested in smart grids without the discussions we had on the topic.
+\\
+I would also like to thank Dr. Michael Gerndt for his guidance in this thesis and also for his help and advising throughout my masters degree in Technische Universät München.
+\\
+I  also thank Dr. Hamid Harroud who suggested several directions to take throughout the course of this project.
+\\
+My special thanks to my close friends who helped relax in times that seemed very stressful.
+\\
+I also greatly thank all excellent professors of whom I had the chance to be a student.
 
 \chapter{Abstract}
+\paragraph{}
+The information technology is advancing exponentially and is having bigger impact on our lives as time advances.
+The next revolution in computing is called the \gls{iot}.
+It is basically the internet we know nowadays except that physical objects (fridge, television, air conditioning) will be the major players in this complex system.
+In this context, the project builds on this idea a test bed for smart home energy optimization.
+The project allows users to monitor their energy consumption via their smart phone and also the users can control their appliances by turning on and off appliances remotely.
+\paragraph{}
+The system is divided into three main components.
+Each of them is using different technologies such as \gls{tinyos}, \gls{nesc}, \gls{j2se}, \gls{http} and other technologies that will be discussed in the thesis report.
+\paragraph{}
+At this stage, the system allows monitoring, but turning on and off appliances is not yet there mainly because of the absence of the material but the communication for that purpose is there.
 
-\chapter{Acknowledgements}
+\chapter{Résumé}
+La technologie de l'information progresse de façon exponentielle et impact sur nos vies prend de l'importance plus le temps avance.
+La prochaine révolution de l'informatique est appelé le \glsreset {iot}.
+Il s'agit essentiellement de Internet tel que nous la connaissons aujourd'hui, sauf que les objets physiques (réfrigérateur, television, climatiseur) seront les principaux acteurs de ce complexe système.
+Dans ce contexte, le projet s'appuie sur cette idée un prototype qui a pour but l'optimisation intelligente de l'énergie dans les domiciles.
+Le projet permet aux utilisateurs de suivre de près leur consommation d'énergie via leur smartphone et permet de contrôler leurs appareils en allumand et éteignant les appareils à distance.
+\paragraph {}
+Le système est divisé en trois composantes principales.
+Chacun d'entre eux  utilise différentes technologies telles que \gls{tinyos}, \gls{nesc}, \gls{j2se}, \gls{http} et d'autres technologies qui seront présentées dans le rapport de thèse.
+\paragraph{}
+A ce stade, le système permet de surveiller, mais allumer et éteindre les appareils n'est pas encore mise en place principalement en raison de l'absence de matériel, mais l'infrastructure de communication pour ce but là est en place.
+%\chapter{ملخص}
+
+\printglossaries
 
@@
 \tableofcontents
-\newpage
+
+\mainmatter
 \chapter{Introduction}
 
+\section{Importance of study}
+\section{Rationale of study}
 \section{Problem statement}
 \section{Purpose of study}
-\section{Importance of study}
 \section{scope of study}
-\section{Rationale of study}
-\section{Definition of terms}
+\section{Outcome of the study}
+\section{Outline of the Thesis}
 
@@
 
-\chapter{Methodology}
+\chapter{Internet Of Things}
+
+\chapter{System Architecture}
+
+\chapter{Implementation}
 
 \chapter{Findings}
+\section{Experiment}
+%\subsection{Methodology}
+\section{Data gathering}
+\section{Interpretation of result}
 
@@
 
-\chapter{Recommendations}
+\chapter{Future Work}
 
-\mbox{}
-\nocite{*}
+\appendix
+\chapter{TinyOs program}
+\chapter{6to4/4to6 packet transformation}
+
+%\backmatter
+%\mbox{}
+%\nocite{*}
 \bibliographystyle{IEEEtran} 
 \bibliography{IEEEabrv,myrefs}
+
 \newpage
+
 \listoffigures
-\chapter{Appendix}
+\listoftables
 

### 2013-06-26 trunk/doc/thesis/report.tex

@@
 \usepackage{setspace}
-\newcommand{\HRule}{\rule{\linewidth}{0.5mm}}
+\newcommand{\HRule}{\rule{\linewidth}{0.6mm}}
 %\usepackage{hyperref}
@@
 \author{Nacer Khalil}
-\title{An internet of things testbed for the smart grid application}
+\title{Internet of Things: a Real-World Deployment Towards Smart Grids Integration}
 
@@
 The next revolution in computing is called the \gls{iot}.
-It is basically the internet we know nowadays except that physical objects (fridge, television, air conditioning) will be the major players in this complex system.
+It is basically the internet we know nowadays except that physical objects (fridge, television, air conditioning...) will be the major players in this complex system.
 In this context, the project builds on this idea a test bed for smart home energy optimization.
@@
 \chapter{Résumé}
-La technologie de l'information progresse de façon exponentielle et impact sur nos vies prend de l'importance plus le temps avance.
-La prochaine révolution de l'informatique est appelé le \glsreset {iot}.
-Il s'agit essentiellement de Internet tel que nous la connaissons aujourd'hui, sauf que les objets physiques (réfrigérateur, television, climatiseur) seront les principaux acteurs de ce complexe système.
-Dans ce contexte, le projet s'appuie sur cette idée un prototype qui a pour but l'optimisation intelligente de l'énergie dans les domiciles.
-Le projet permet aux utilisateurs de suivre de près leur consommation d'énergie via leur smartphone et permet de contrôler leurs appareils en allumand et éteignant les appareils à distance.
+\paragraph{}
+La technologie de l'information progresse de façon exponentielle et son impact sur nos vies prend de l'importance plus le temps avance.
+La prochaine révolution de l'informatique est appelé \glsreset{iot} \gls{iot}.
+Il s'agit essentiellement de Internet tel que nous la connaissons aujourd'hui, sauf que les objets physiques (réfrigérateur, television, climatiseur...) seront les principaux acteurs de ce complexe système.
+Dans ce contexte, le projet s'appuie sur cette idée et présente un prototype qui a pour but l'optimisation intelligente de l'énergie dans les domiciles.
+Le projet permet aux utilisateurs de suivre de près leur consommation d'énergie via leur smartphone et permet de contrôler leurs appareils en allumant et éteignant les appareils à distance.
 \paragraph {}
@@
 
+\singlespacing
+
 \printglossaries
-
-%THIS IS THE TABLE OF CONTENT PART
 \tableofcontents
@@
 \mainmatter
+\doublespacing
 \chapter{Introduction}
@@
 \section{Importance of study}
+\paragraph{}
+The global electrical grid one the bases that gives us this quality of life that most people around the world enjoy.
+The global electrical grid is sometimes called the largest machine ever made by man.
+\cite{ref1}.
+It a widespread machine, distributed but interconnected system that provides power to households, companies and other organizations and therefore the world's economy is counting on it.
+The electrical grid knows numerous problems such as blackouts, current instabilities that costs billions of dollars yearly to both the utilities, utilities and insurances.
+The grid suffers from the problem of peak hours where energy consumption where in many case the one single utility cannot provide such energy.
+\paragraph{}
+The \gls{sg} comes to provide solutions to the problems that suffers from the traditional electrical grid.
+\gls{sg} is an intelligent, auto-balancing, and self-monitoring electrical grid \cite{ref2} and is and whose energy sources are distributed between utilities and consumers.
+The nature of the \gls{sg} as having distributed energy sources reduces numerous losses that arises in the traditional grid such as the energy loss in the energy transmission lines and also in the overload of the electrical lines.
+\paragraph{}
+Another revolution is about to begin and it is taking place in the information technology and is called \glsreset{iot}\gls{iot}.
+\gls{iot} is the era of autonomous machine able to produce and consume information and being able to act autonomously without the guidance of humans to performs their tasks.
+Cisco predicts that the number of devices to be connected in 2020 will break the 50 billions devices \cite{ref3} whereas the world population will be around 7.6 billion.
+That make 6.58 devices connected for each person in the world whereas the number of devices connected in 2010 12.5 billion.
+This clearly shows that the shift to the \gls{iot} has already started.
+\paragraph{}
+\glsreset{sg}\gls{sg} can and will benefit from the \gls{iot}.
+The integration of the \gls{sg} and and \gls{iot} makes this study and thesis of high importance in order to solve as much as possible the problems of the traditional grid but at the same bring a new dimension to the power grid by benefiting from the numerous possibilities offered by the \gls{iot}.
+
 \section{Rationale of study}
+\paragraph{}
+This project has been done in order to investigate ways the \gls{iot} can benefit the \gls{sg}.
+The smart grid as a set of philosophical ideas counts on a set of novel functionalities.
+Once of these ideas is the two-way communication that will make possible the utility to communicate with consumers and vice versa.
+This concept is possible with the Internet Of Things as devices in \gls{iot} communicate between them by producing and consuming information.
+Numerous ideas are based on this two-way communication such as the \gls{dr} which offers a set of services for both the power provider and the consumer.
+\paragraph{}
+The \glsreset{iot}\gls{iot} counts on different technologies such as \gls{wsn}, \gls{rfid}, \gls{ipv6} and other technologies which make the internet of things a very heterogeneous and complex system \cite{ref4}.
+One of the challenges in \gls{iot} is to deals with such heterogeneity and take the most out of it in order to make the communication as smooth and robust as possible.
+This heterogeneity should be made transparent to any two parties trying to communicate in \gls{iot}.
+
 \section{Problem statement}
+\paragraph{}
+The \glsreset{sg} \gls{sg} counts on the two-way communication to build to a whole communication and protocol stack namely the \glsreset{dr}\gls{dr} and other services built on top of the \gls{dr}.
+From the consumer's side or more precisely the household's side, there are smart home energy management in the market nowadays but none of these products offers integration with \gls{sg}.
+To offer such an integration with \gls{sg}, there is a need to first build this two-way communication and make it possible for any appliance to report its real-time energy consumption and also to be able to be controlled remotely with this two-way communication.
+Such infrastructure is missing for the time being and therefore the integration of the smart home within the smart grid is not possible.
+Also, in order for \gls{sg} to be integrated into the \gls{iot}, all devices should be able to communicate in the two directions.
+\paragraph{}
+
 \section{Purpose of study}
@@
 
-%\backmatter
-%\mbox{}
-%\nocite{*}
+\backmatter
+\singlespacing
+
 \bibliographystyle{IEEEtran} 
 \bibliography{IEEEabrv,myrefs}
-
-\newpage
 

### 2013-06-27 doc/thesis/report.tex

@@
 \usepackage{graphicx}
-\newcommand{\HRule}{\rule{\linewidth}{0.5mm}}
-
+\usepackage{setspace}
+\newcommand{\HRule}{\rule{\linewidth}{0.6mm}}
+%\usepackage{hyperref}
+\usepackage[acronym]{glossaries} 
+\makeglossaries
 \author{Nacer Khalil}
-\title{An internet of things testbed for the smart grid application}
+\title{Internet of Things: a Real-World Deployment Towards Smart Grids Integration}
+\newenvironment{dedication}{\vspace{6ex}\begin{quotation}\begin{center}\begin{em}}{\par\end{em}\end{center}\end{quotation}}
 
 \begin{document}
+\doublespacing
+\input{acronym}
+\frontmatter
 
-\begin{titlepage}
-\begin{center}
-\includegraphics[width=0.6\textwidth]{img/aui_logo.jpg}
-%\includegraphics[width=0.45\textwidth]{img/tum_logo.jpg}
-\linebreak
-\linebreak
-\linebreak
-\linebreak
-\linebreak
-\linebreak
-% Title
-\HRule \\[0.4cm]
-{ \huge \bfseries An Internet Of Things Testbed For The Smart Grid Application}\\[0.4cm]
-
-\HRule \\[1.5cm]
-
-% Author and supervisor
-\begin{minipage}{0.4\textwidth}
-\begin{flushleft} \large
-\emph{Author:}\\
-Nacer \textsc{Khalil}
-\end{flushleft}
-\end{minipage}
-\begin{minipage}{0.4\textwidth}
-\begin{flushright} \large
-\emph{Supervisor:} \\
-Dr.~Riduan \textsc{Abid} \\
-Dr.~Driss \textsc{Benhaddou} \\
-Dr.~Michael \textsc{Gerndt} 
-\end{flushright}
-\end{minipage}
-
-\vfill
-
-% Bottom of the page
-{\large \today}
-
-\end{center}
-\end{titlepage}
+%COVER PAGE
+\input{cover_page}
+\chapter{} % dedications chapter
+\begin{dedication}
+To my dear parents for their patience, love and support \\
+To my brothers Hossein, Omar and Saad \\
+To my cousin Sara and her family \\
+To my family \\
+To my friends
+\end{dedication}
+\chapter{Acknowledgements}
+\paragraph{}
+I am deeply thankful to my supervisor Dr.
+Mohamed Riduan Abid for his very close supervision, guidance and for the advice he kept giving me during the lifetime of the project.
+Your advices whether they are computer science related, or even every day life related wise words things I will never forget.
+\paragraph*{}
+I am also thankful for having Dr.
+Driss Benhaddou as a supervisor.
+His advice, insight, guidance and very enlightening ideas.
+I wouldn't have been so interested in smart grids without the discussions we had on the topic.
+\paragraph{}
+I would also like to thank Dr. Michael Gerndt for his guidance in this thesis and also for his help and advising throughout my masters degree in Technische Universät München.
+\paragraph{}
+I  also thank Dr. Hamid Harroud who suggested several directions to take throughout the course of this project.
+\paragraph{}
+My special thanks to my close friends who helped relax in times that seemed very stressful. I wouldn't have made it without their help.
+\paragraph{}
+I also greatly thank all excellent professors of whom I had the chance to be a student. I learned things in class  I would have never learnt elsewhere.
 
 \chapter{Abstract}
+\paragraph{}
+The information technology is advancing exponentially and is having bigger impact on our lives as time advances.
+The next revolution in computing is called the \gls{iot}.
+It is basically the internet we know nowadays except that physical objects (fridge, television, air conditioning...) will be the major players in this complex system.
+In this context, the project builds on this idea a test bed for smart home energy optimization.
+The project allows users to monitor their energy consumption via their smart phone and also the users can control their appliances by turning on and off appliances remotely.
+\paragraph{}
+The system is divided into three main components.
+Each of them is using different technologies such as \gls{tinyos}, \gls{nesc}, \gls{j2se}, \gls{http} and other technologies that will be discussed in the thesis report.
+\paragraph{}
+At this stage, the system allows monitoring, but turning on and off appliances is not yet there mainly because of the absence of the material but the communication for that purpose is there.
 
-\chapter{Acknowledgements}
+\chapter{Résumé}
+\paragraph{}
+La technologie de l'information progresse de façon exponentielle et son impact sur nos vies prend de l'importance plus le temps avance.
+La prochaine révolution de l'informatique est appelé \glsreset{iot} \gls{iot}.
+Il s'agit essentiellement de Internet tel que nous la connaissons aujourd'hui, sauf que les objets physiques  seront les principaux acteurs de ce système.
+Ainsi, le projet s'appuie sur cette idée et présente un prototype qui a pour but l'optimisation intelligente de l'énergie dans les domiciles.
+Le projet permet aux utilisateurs de suivre de près leur consommation d'énergie via leur smartphone et permet de contrôler leurs appareils en allumant et éteignant les appareils à distance.
+\paragraph {}
+Le système est divisé en trois composantes principales.
+Chacun d'entre eux  utilise différentes technologies telles que \gls{tinyos}, \gls{nesc}, \gls{j2se}, \gls{http} et d'autres technologies qui seront présentées dans le rapport de thèse.
+\paragraph{}
+A ce stade, le système permet de surveiller, mais controler les appareils n'est pas encore mise en place principalement en raison de l'absence de matériel, mais l'infrastructure de communication pour ce but là est en place.
+%\chapter{ملخص}
 
-%THIS IS THE TABLE OF CONTENT PART
+\singlespacing
+
 \tableofcontents
-\newpage
+\listoffigures
+\listoftables
+
+\mainmatter
+\doublespacing
 \chapter{Introduction}
 
+\section{Importance of study}
+\paragraph{}
+The global electrical grid one the bases that gives us this quality of life that most people around the world enjoy.
+The global electrical grid is sometimes called the largest machine ever made by man.
+\cite{ref1}.
+It a widespread machine, distributed but interconnected system that provides power to households, companies and other organizations and therefore the world's economy is counting on it.
+The electrical grid knows numerous problems such as blackouts, current instabilities that costs billions of dollars yearly to both the utilities, utilities and insurances.
+The grid suffers from the problem of peak hours where energy consumption where in many case the one single utility cannot provide such energy.
+\paragraph{}
+The \gls{sg} comes to provide solutions to the problems that suffers from the traditional electrical grid.
+\gls{sg} is an intelligent, auto-balancing, and self-monitoring electrical grid \cite{ref2} and is and whose energy sources are distributed between utilities and consumers.
+The nature of the \gls{sg} as having distributed energy sources reduces numerous losses that arises in the traditional grid such as the energy loss in the energy transmission lines and also in the overload of the electrical lines.
+\paragraph{}
+Another revolution is about to begin and it is taking place in the information technology and is called \glsreset{iot}\gls{iot}.
+\gls{iot} is the era of autonomous machine able to produce and consume information and being able to act autonomously without the guidance of humans to performs their tasks.
+Cisco predicts that the number of devices to be connected in 2020 will break the 50 billions devices \cite{ref3} whereas the world population will be around 7.6 billion.
+That make 6.58 devices connected for each person in the world whereas the number of devices connected in 2010 12.5 billion.
+This clearly shows that the shift to the \gls{iot} has already started.
+\paragraph{}
+\glsreset{sg}\gls{sg} can and will benefit from the \gls{iot}.
+The integration of the \gls{sg} and and \gls{iot} makes this study and thesis of high importance in order to solve as much as possible the problems of the traditional grid but at the same bring a new dimension to the power grid by benefiting from the numerous possibilities offered by the \gls{iot}.
+
+\section{Rationale of study}
+\paragraph{}
+This project has been done in order to investigate ways the \gls{iot} can benefit the \gls{sg}.
+The smart grid as a set of philosophical ideas counts on a set of novel functionalities.
+Once of these ideas is the two-way communication that will make possible the utility to communicate with consumers and vice versa.
+This concept is possible with the Internet Of Things as devices in \gls{iot} communicate between them by producing and consuming information.
+Numerous ideas are based on this two-way communication such as the \gls{dr} which offers a set of services for both the power provider and the consumer.
+\paragraph{}
+The \glsreset{iot}\gls{iot} counts on different technologies such as \gls{wsn}, \gls{rfid}, \gls{ipv6} and other technologies which make the Internet of things a very heterogeneous and complex system \cite{ref4}.
+One of the challenges in \gls{iot} is to deals with such heterogeneity and take the most out of it in order to make the communication as smooth and robust as possible.
+This heterogeneity should be made transparent to any two parties trying to communicate in \gls{iot}.
+
 \section{Problem statement}
+\paragraph{}
+The \glsreset{sg} \gls{sg} counts on the two-way communication to build to a whole communication and protocol stack namely the \glsreset{dr}\gls{dr} and other services built on top of the \gls{dr}.
+From the consumer's side or more precisely the household's side, there are smart home energy management software in the market nowadays but none of these products offers integration with \gls{sg}.
+To have such an integration with \gls{sg}, there is a need to first build this two-way communication and make it possible for any appliance to report its real-time energy consumption and also to be able to be controlled remotely with this two-way communication.
+Such infrastructure is missing for the time being and therefore the integration of the smart home within the smart grid is not possible.
+Also, in order for \gls{sg} to be integrated into the \gls{iot}, all devices should be able to communicate in the two directions.
+\paragraph{}
+As \gls{sg} counts on numerous technologies to achieve its main goals.
+This usage of different technologies imposes another problem which is the heterogeneity of the system.
+As numerous components count on different technologies, there is a need to interface them in order to make this heterogeneity transparent to the all components of the system.
+This means that there is a need to build a software within the whole system that would constitute an important part of it and will take care of making the communication and processes technology-transparent and give the possibility to all components of the system to work hand-in-hand in order their common goals independently of the data link layer technology (Ethernet, Wifi or Zigbee) they are counting on, the network layer technology (\gls{ipv4} or \gls{ipv6}).
+
 \section{Purpose of study}
-\section{Importance of study}
+ \paragraph{}
+The goal of the study is to solve the problems introduced in the problem statement section.
+This means creating the two-way communication, and also solving the heterogeneity that arises from integrating smart homes with the smart grid and making each components in the smart home as part of the Internet of things by making it communicate with any device in the Internet and having an \gls{ipv6} address that is built by making use of \gls{6lowpan} whose goal is to provide the TCP/IP to the smallest devices and with the least computational and energy rich.
+\paragraph{}
+The first purpose of this study is to create a testbed having the two-way communication and also being able to make use of it in order to build richer applications based on this possibility.
+To do it, there was a need to first introduce \gls{6lowpan} within every device in the \gls{wsn}.
+This would make every sensor node also known as mote identified and ready to communicate and be part of the \gls{iot}.
+This communication will go in the two directions and based on that the two-way communication will have to be built.
+\paragraph{}
+The second purpose of this study is to solve the heterogeneity problem that arises with the integration of smart homes within the smart grid and also when making the whole system as part of the \glsreset{iot}\gls{iot}.
 \section{scope of study}
-\section{Rationale of study}
-\section{Definition of terms}
+\section{Outcome of the study}
+\section{Outline of the Thesis}
 
@@
 
-\chapter{Recommendations}
+\chapter{Future Work}
 
-\mbox{}
-\nocite{*}
+\appendix
+\chapter{TinyOs program}
+\chapter{6to4/4to6 packet transformation}
+
+\backmatter
+
+\singlespacing
+\printglossaries
 \bibliographystyle{IEEEtran} 
 \bibliography{IEEEabrv,myrefs}
-\newpage
-\listoffigures
-\chapter{Appendix}
-
 \end{document}

### 2013-06-27 trunk/doc/thesis/report.tex
chapter 1 done

@@
 \usepackage[acronym]{glossaries} 
-
+\makeglossaries
 \author{Nacer Khalil}
 \title{Internet of Things: a Real-World Deployment Towards Smart Grids Integration}
-
-\makeglossaries
+\newenvironment{dedication}{\vspace{6ex}\begin{quotation}\begin{center}\begin{em}}{\par\end{em}\end{center}\end{quotation}}
 
@@
 \input{cover_page}
-
+\chapter{} % dedications chapter
+\begin{dedication}
+To my dear parents for their patience, love and support \\
+To my brothers Hossein, Omar and Saad \\
+To my cousin Sara and her family \\
+To my family \\
+To my friends
+\end{dedication}
 \chapter{Acknowledgements}
+\paragraph{}
 I am deeply thankful to my supervisor Dr.
@@
 Your advices whether they are computer science related, or even every day life related wise words things I will never forget.
-\\
+\paragraph*{}
 I am also thankful for having Dr.
@@
 I wouldn't have been so interested in smart grids without the discussions we had on the topic.
-\\
+\paragraph{}
 I would also like to thank Dr. Michael Gerndt for his guidance in this thesis and also for his help and advising throughout my masters degree in Technische Universät München.
-\\
+\paragraph{}
 I  also thank Dr. Hamid Harroud who suggested several directions to take throughout the course of this project.
-\\
-My special thanks to my close friends who helped relax in times that seemed very stressful.
-\\
-I also greatly thank all excellent professors of whom I had the chance to be a student.
+\paragraph{}
+My special thanks to my close friends who helped relax in times that seemed very stressful. I wouldn't have made it without their help.
+\paragraph{}
+I also greatly thank all excellent professors of whom I had the chance to be a student. I learned things in class  I would have never learnt elsewhere.
 
@@
 \paragraph{}
-At this stage, the system allows monitoring, but turning on and off appliances is not yet there mainly because of the absence of the material but the communication for that purpose is there.
+At this stage, the system allows monitoring, but turning on and off appliances is not yet there mainly because of the absence of the equipment but the communication for that purpose is there.
 
@@
 La prochaine révolution de l'informatique est appelé \glsreset{iot} \gls{iot}.
-Il s'agit essentiellement de Internet tel que nous la connaissons aujourd'hui, sauf que les objets physiques (réfrigérateur, television, climatiseur...) seront les principaux acteurs de ce complexe système.
-Dans ce contexte, le projet s'appuie sur cette idée et présente un prototype qui a pour but l'optimisation intelligente de l'énergie dans les domiciles.
+Il s'agit essentiellement de Internet tel que nous la connaissons aujourd'hui, sauf que les objets physiques  seront les principaux acteurs de ce système.
+Ainsi, le projet s'appuie sur cette idée et présente un prototype qui a pour but l'optimisation intelligente de l'énergie dans les domiciles.
 Le projet permet aux utilisateurs de suivre de près leur consommation d'énergie via leur smartphone et permet de contrôler leurs appareils en allumant et éteignant les appareils à distance.
@@
 \paragraph{}
-A ce stade, le système permet de surveiller, mais allumer et éteindre les appareils n'est pas encore mise en place principalement en raison de l'absence de matériel, mais l'infrastructure de communication pour ce but là est en place.
+A ce stade, le système permet de surveiller, mais controler les appareils n'est pas encore mise en place principalement en raison de l'absence de matériel, mais l'infrastructure de communication pour ce but là est en place.
 %\chapter{ملخص}
@@
 
-\printglossaries
 \tableofcontents
+\listoffigures
+\listoftables
 
@@
 \paragraph{}
-The \glsreset{iot}\gls{iot} counts on different technologies such as \gls{wsn}, \gls{rfid}, \gls{ipv6} and other technologies which make the internet of things a very heterogeneous and complex system \cite{ref4}.
+The \glsreset{iot}\gls{iot} counts on different technologies such as \gls{wsn}, \gls{rfid}, \gls{ipv6} and other technologies which make the Internet of things a very heterogeneous and complex system \cite{ref4}.
 One of the challenges in \gls{iot} is to deals with such heterogeneity and take the most out of it in order to make the communication as smooth and robust as possible.
@@
 The \glsreset{sg} \gls{sg} counts on the two-way communication to build to a whole communication and protocol stack namely the \glsreset{dr}\gls{dr} and other services built on top of the \gls{dr}.
-From the consumer's side or more precisely the household's side, there are smart home energy management in the market nowadays but none of these products offers integration with \gls{sg}.
-To offer such an integration with \gls{sg}, there is a need to first build this two-way communication and make it possible for any appliance to report its real-time energy consumption and also to be able to be controlled remotely with this two-way communication.
+From the consumer's side or more precisely the household's side, there are smart home energy management software in the market nowadays but none of these products offers integration with \gls{sg}.
+To have such an integration with \gls{sg}, there is a need to first build this two-way communication and make it possible for any appliance to report its real-time energy consumption and also to be able to be controlled remotely with this two-way communication.
 Such infrastructure is missing for the time being and therefore the integration of the smart home within the smart grid is not possible.
@@
 \paragraph{}
+As \gls{sg} counts on numerous technologies to achieve its main goals.
+This usage of different technologies imposes another problem which is the heterogeneity of the system.
+As numerous components count on different technologies, there is a need to interface them in order to make this heterogeneity transparent to the all components of the system.
+This means that there is a need to build a software within the whole system that would constitute an important part of it and will take care of making the communication and processes technology-transparent and give the possibility to all components of the system to work hand-in-hand in order their common goals independently of the data link layer technology (Ethernet, Wifi or Zigbee) they are counting on, the network layer technology (\gls{ipv4} or \gls{ipv6}).
 
 \section{Purpose of study}
+ \paragraph{}
+The goal of the study is to solve the problems introduced in the problem statement section.
+This means creating the two-way communication, and also solving the heterogeneity that arises from integrating smart homes with the smart grid and making each components in the smart home as part of the Internet of things by making it communicate with any device in the Internet and having an \gls{ipv6} address that is built by making use of \gls{6lowpan} whose goal is to provide the TCP/IP to the smallest devices and with the least computational and energy rich.
+\paragraph{}
+The first purpose of this study is to create a testbed having the two-way communication and also being able to make use of it in order to build richer applications based on this possibility.
+To do it, there was a need to first introduce \gls{6lowpan} within every device in the \gls{wsn}.
+This would make every sensor node also known as mote identified and ready to communicate and be part of the \gls{iot}.
+This communication will go in the two directions and based on that the two-way communication will have to be built.
+\paragraph{}
+The second purpose of this study is to solve the heterogeneity problem that arises with the integration of smart homes within the smart grid and also when making the whole system as part of the \glsreset{iot}\gls{iot}.
+Solving such a problem means having an additional component that will track all communication packets and change them to fit the needs of the destination component.
+This will make of this additional component a middleman and therefore make the heterogeneity problem disappear.
+
 \section{scope of study}
+\paragraph{}
+The thesis aims to solves the two problems presented in section 1.3.
+To do so, a variety of tools and technologies were used for this purpose.The two problems were not solved one independently of the other but even though the problems seem separate one cannot solve without having to solve the other.
+In other words, after the creation of the two-way communication in the \glsreset{wsn}\gls{wsn} which supported \gls{ipv6} over \gls{6lowpan}, other network only supported \gls{ipv4} and therefore could not communicate, and this was due to the heterogeneity problem.
+This is why the two-way communication was created and then the heterogeneity problem had to be solved in order to see everything working.
+The system is divided into four components:
+\begin{itemize}
+\item The fully \gls{ipv6} supported \gls{wsn}
+\item The \gls{tinyos} program for sensor transmission and appliance request handler
+\item The heterogeneity handler and middleman gateway server
+\item The mobile client
+\end{itemize}
+\paragraph{}
+Each of these four component exist to solve part of the problem.
+Each of these components will be discussed extensively in the Chapter 4 from the architectural aspect of the component and in Chapter 5 from the implementation level of the system.
+\paragraph{}
+It is to be noted that some of the concepts will be discussed from the architectural point in Chapter 4 but not from an implementation point in Chapter 5.
+The sole reason behind is the fact that some of the equipment required for this thesis did not arrive on time and therefore the parts in the implementation requiring such equipment was omitted to meet the deadlines.
 \section{Outcome of the study}
+\paragraph{}
+The outcome of the thesis is to design and implement a test-bed that is able to show the two-way communication in action.
+It will also have to show concepts of the \glsreset{iot}\gls{iot} which are that every node in the system can communicate autonomously and with any other node in the Internet.
+Another \gls{iot} concept is the ability of any node in the \gls{iot} to consume and produce information that is communicated through the network.
+The autonomous production of information is shown by the fact that every sensor node in the \gls{wsn} reads sensory data from its sensors that are attached to appliance in order to sense their consumption and this information is produced and transmitted periodically.
+The consumption of data by is shown by external nodes in the \gls{iot} sending request to the sensor node in the \gls{wsn} in order to control and appliance's status i.e.
+turn on the television.
+This will prove the consumption of information by the nodes in the \gls{iot}.
+\paragraph{}
+The test-bed should also show that communication between any two nodes in the \gls{iot} is successful, robust and reliable and that the heterogeneity of the system is made transparent to the nodes in the system by the creation of the software tool residing in the gateway which is between the \gls{iot} and the outside world.
+This software tool is seen as a heterogeneity handler and also a communication middleman that translates the communication to the appropriate technology needed by the recipient.
+\paragraph{}
+The testbed will able to show the two-way communication in action and will also show that the heterogeneity does not form a barrier by its successful handling.
+
 \section{Outline of the Thesis}
-
+\paragraph{}
+The thesis is is divided into eight chapters.
+Each of the chapter approaches thesis from a different aspect.
+The first chapter is an introductory chapter that gives a flavor of the project that is to be studied in this thesis and also introduces the problem and the solution proposed in this topic.
+This chapter also puts the reader into perspective by introducing some of the main concepts related to the \glsreset{sg}\gls{sg} and the \glsreset{iot}\gls{iot}.
+\paragraph{}
+The second chapter presents the background of the work by revising the literature related to the topics of \gls{iot} and \gls{sg}.
+It also discussed related work and how it was tackled in different research papers, how does it relate to the work done in this thesis and also how does it differ.
+\paragraph{}
+The third chapter is dedicated to the \glsreset{iot}\gls{iot}.
+It presents the topic, talks about the importance of \gls{iot}, its impact on the world, the actual Internet, the future of Internet and also how will \gls{sg} benefit from it.
+\paragraph{}
+The fourth chapter will present the architecture of the system and the test-bed. It will discuss the system from different aspects, and how were the two problems solved from the architectural aspect.
+\paragraph{}
+The fifth chapter will mostly discuss the implementation system.
+It will start by presenting the hardware used in this project, then move to the technology enablers.
+Each component of the system will be discussed extensively and move to the integration of the components to form the test-bed.
+\paragraph{}
+The sixth chapter will start first by presenting the findings then move to explain the experiment that was done to measure the performance of the system.
+The data gathering section will show how and what data was gathered and then move to analyzing and interpreting the data.
+\paragraph{}
+Chapter seven and eight are the last two chapters of the thesis that will conclude the thesis, provide directions about future work and recommendations to the reader what parts of the system could be ameliorated and how it can be achieved.
 \chapter{Literature review}
@@
 \chapter{Internet Of Things}
+
+\section{The actual Internet}
+\section{Introduction to \gls{iot}}
+\section{Importance of \gls{iot}}
+\section{The impact of \gls{iot}}
+\section{The benefits of \gls{iot}}
+\section{The future of Internet}
 
@@
 \backmatter
+
 \singlespacing
-
+\printglossaries
 \bibliographystyle{IEEEtran} 
 \bibliography{IEEEabrv,myrefs}
-
-\listoffigures
-\listoftables
-
 \end{document}

### 2013-06-27 doc/thesis/report.tex

@@
 \usepackage{setspace}
+\usepackage{arabtex}
+\usepackage{utf8}
 \newcommand{\HRule}{\rule{\linewidth}{0.6mm}}
@@
 \title{Internet of Things: a Real-World Deployment Towards Smart Grids Integration}
+
 \newenvironment{dedication}{\vspace{6ex}\begin{quotation}\begin{center}\begin{em}}{\par\end{em}\end{center}\end{quotation}}
@@
 A ce stade, le système permet de surveiller, mais controler les appareils n'est pas encore mise en place principalement en raison de l'absence de matériel, mais l'infrastructure de communication pour ce but là est en place.
+\setcode{utf8}
+
 %\chapter{ملخص}
+
+\chapter{Arabic}
+\paragraph{}
+\begin{arabtex}
+تكنولوجيا المعلومات تتقدم باطراد وجود أكبر تأثير على حياتنا كما تقدم الوقت.
+الثورة القادمة في مجال الحوسبة يسمى إنترنت الأشياء (قام المحفل).
+انها اساسا شبكة الانترنت ونحن نعلم في الوقت الحاضر إلا أن الأشياء المادية (ثلاجة، تلفزيون، تكييف هواء ...) وسوف تكون القوى الفاعلة الرئيسية في هذا النظام المعقد.
+وفي هذا السياق، فإن المشروع يبني على هذه الفكرة على سرير اختبار لترشيد استهلاك الطاقة الذكية الرئيسية.
+ال يسمح المشروع للمستخدمين لرصد استهلاك الطاقة لديها عبر الذكية الخاصة بهم الهاتف وأيضا يمكن للمستخدمين التحكم في الأجهزة عن طريق تشغيل وإيقاف الأجهزة عن بعد.
+\end{arabtex}
+\paragraph{}
+\begin{arabtex}
+وينقسم النظام إلى ثلاثة عناصر رئيسية. كل واحد منهم يستخدم
+التكنولوجيات المختلفة، مث لTinyOS، NESC، ومعيار جافا
+طبعة J2SE، بروتوكول نقل النص التشعبي HTTP وغيرها من التكنولوجيات
+التي سيتم مناقشتها في التقرير أطروحة.
+\end{arabtex}
+\paragraph{}
+\begin{arabtex}
+
+في هذه المرحلة، ونظام مراقبة يسمح، ولكن تشغيل وإيقاف تطبيق صحيفة-
+ ليس هناك حتى الآن وذلك أساسا بسبب غياب المعدات ولكن البلاغ لهذا الغرض هناك.
+\end{arabtex}
 
@@
 The second purpose of this study is to solve the heterogeneity problem that arises with the integration of smart homes within the smart grid and also when making the whole system as part of the \glsreset{iot}\gls{iot}.
+Solving such a problem means having an additional component that will track all communication packets and change them to fit the needs of the destination component.
+This will make of this additional component a middleman and therefore make the heterogeneity problem disappear.
+
 \section{scope of study}
+\paragraph{}
+The thesis aims to solves the two problems presented in section 1.3.
+To do so, a variety of tools and technologies were used for this purpose.The two problems were not solved one independently of the other but even though the problems seem separate one cannot solve without having to solve the other.
+In other words, after the creation of the two-way communication in the \glsreset{wsn}\gls{wsn} which supported \gls{ipv6} over \gls{6lowpan}, other network only supported \gls{ipv4} and therefore could not communicate, and this was due to the heterogeneity problem.
+This is why the two-way communication was created and then the heterogeneity problem had to be solved in order to see everything working.
+The system is divided into four components:
+\begin{itemize}
+\item The fully \gls{ipv6} supported \gls{wsn}
+\item The \gls{tinyos} program for sensor transmission and appliance request handler
+\item The heterogeneity handler and middleman gateway server
+\item The mobile client
+\end{itemize}
+\paragraph{}
+Each of these four component exist to solve part of the problem.
+Each of these components will be discussed extensively in the Chapter 4 from the architectural aspect of the component and in Chapter 5 from the implementation level of the system.
+\paragraph{}
+It is to be noted that some of the concepts will be discussed from the architectural point in Chapter 4 but not from an implementation point in Chapter 5.
+The sole reason behind is the fact that some of the equipment required for this thesis did not arrive on time and therefore the parts in the implementation requiring such equipment was omitted to meet the deadlines.
 \section{Outcome of the study}
+\paragraph{}
+The outcome of the thesis is to design and implement a test-bed that is able to show the two-way communication in action.
+It will also have to show concepts of the \glsreset{iot}\gls{iot} which are that every node in the system can communicate autonomously and with any other node in the Internet.
+Another \gls{iot} concept is the ability of any node in the \gls{iot} to consume and produce information that is communicated through the network.
+\paragraph{}
+The autonomous production of information is shown by the fact that every sensor node in the \gls{wsn} reads sensory data from its sensors that are attached to appliance in order to sense their consumption and this information is produced and transmitted periodically.
+The consumption of data by is shown by external nodes in the \gls{iot} sending request to the sensor node in the \gls{wsn} in order to control and appliance's status i.e.
+turn on the television.
+This will prove the consumption of information by the nodes in the \gls{iot}.
+\paragraph{}
+The test-bed should also show that communication between any two nodes in the \gls{iot} is successful, robust and reliable and that the heterogeneity of the system is made transparent to the nodes in the system by the creation of the software tool residing in the gateway which is between the \gls{iot} and the outside world.
+This software tool is seen as a heterogeneity handler and also a communication middleman that translates the communication to the appropriate technology needed by the recipient.
+\paragraph{}
+The testbed will able to show the two-way communication in action and will also show that the heterogeneity does not form a barrier by its successful handling.
+
 \section{Outline of the Thesis}
-
+\paragraph{}
+The thesis is is divided into eight chapters. Each of the chapter approaches thesis from a different aspect. 
+\paragraph{Chapter 1}
+It is an introductory chapter that gives a flavor of the project that is to be studied in this thesis and also introduces the problem and the solution proposed in this topic.
+This chapter also puts the reader into perspective by introducing some of the main concepts related to the \glsreset{sg}\gls{sg} and the \glsreset{iot}\gls{iot}.
+\paragraph{Chapter 2}
+It presents the background of the work by revising the literature related to the topics of \gls{iot} and \gls{sg}.
+It also discussed related work and how it was tackled in different research papers, how does it relate to the work done in this thesis and also how does it differ.
+\paragraph{Chapter 3}
+It is dedicated to the \glsreset{iot}\gls{iot}.
+It presents the topic, talks about the importance of \gls{iot}, its impact on the world, the actual Internet, the future of Internet and also how will \gls{sg} benefit from it.
+\paragraph{Chapter 4}
+It will present the architecture of the system and the test-bed. It will discuss the system from different aspects, and how were the two problems solved from the architectural aspect.
+\paragraph{Chapter 5}
+This chapter will mostly discuss the implementation system.
+It will start by presenting the hardware used in this project, then move to the technology enablers.
+Each component of the system will be discussed extensively and move to the integration of the components to form the test-bed.
+\paragraph{Chapter 6}
+It will start first by presenting the findings then move to explain the experiment that was done to measure the performance of the system.
+The data gathering section will show how and what data was gathered and then move to analyzing and interpreting the data.
+\paragraph{Chapter 7, 8}
+These are the last two chapters of the thesis that will conclude the thesis, provide directions about future work and recommendations to the reader what parts of the system could be ameliorated and how it can be achieved.
 \chapter{Literature review}
+\section{The need for smart grid}
+\paragraph{}
+The actual electrical grid is based on a technology that is at least one century old \cite{ref5}.
+As a result, there is a need to change as to meet to 21st century requirement such as having more green energy in order to reduce the pollution caused mainly by the use of fossil energy.
+As a result, research has introduced a new technology called \glsreset{sg}\gls{sg}.
+Although it is still a research project and numerous issues related to \gls{sg} are left unanswered, nations across the globe started to deploy \gls{sg} project.
+It is also noted that in order for the grid to become more reliable, secure and efficient, a bi-directional information and communication infrastructure should be embedded within the grid \cite{ref5}.
+\paragraph{}
+The smart grid brings around new concepts that will achieve efficiency, reliable and security to the power grid. The first one is the introduction of the \gls{dg} by exploiting \gls{res} \cite{ref6}
 
 \chapter{Internet Of Things}
+
+\section{The actual Internet}
+\section{Introduction to \gls{iot}}
+\section{Importance of \gls{iot}}
+\section{The impact of \gls{iot}}
+\section{The benefits of \gls{iot}}
+\section{The future of Internet}
 
@@
 \bibliography{IEEEabrv,myrefs}
+
 \end{document}

### 2013-06-28 doc/thesis/report.tex

@@
 To my family \\
-To my friends
+To my friends \\
+
 \end{dedication}
@@
 \chapter{Literature review}
-\section{The need for smart grid}
+\section{The need for Smart Grid}
 \paragraph{}
@@
 \paragraph{}
-The smart grid brings around new concepts that will achieve efficiency, reliable and security to the power grid. The first one is the introduction of the \gls{dg} by exploiting \gls{res} \cite{ref6}
+The smart grid brings around new concepts that will achieve efficiency, reliable and security to the power grid.
+The first one is the introduction of the \gls{dg} by exploiting \gls{res} \cite{ref6} and this reduce the losses of energy in the transmission.
+This distribution by itself enables the power grid this to move from the one way energy distribution to a two-way energy distribution where the utility will not be the sole energy producer but households can now produce and even sell the extra power.
+This is known nowadays as the feed-in-tariff which is a policy that allows the consumers to become producer and sell the excess power to the grid \cite{ref7}.
+This will allow the power grid to provide power to consumer in a "free market" fashion which means that it will obey to the laws of supply and demand \cite{ref7}.\glsreset{iot} \glsreset{sg}
+\section{\gls{iot} and \gls{sg}}
+\paragraph{}
+The \glsreset{iot}\gls{iot} has known high acceptance by the players in the Internet.
+Nowadays, it is estimated that 12.5 billion devices are connected to the Internet and that this trend will continue to read 50 billion connected devices by the end of the decade \cite{ref3}.
+\gls{iot} brings several benefits to the actual Internet and these benefits will pull \gls{sg} to be part of it.
+Consequently, research has taken this direction which is to integrate the \gls{sg} within the \gls{iot} as most aspects of \gls{sg} will be enhanced by \gls{iot} \cite{ref8}.
+\paragraph{}
+\gls{iot} counts on a set of technologies which are \glsreset{rfid}\gls{rfid}, \gls{m2m} \cite{ref9}, public communication networks such as 2G/3G mobile networks, power lines, WI-FI, Zigbee and other technologies \cite{ref8}.
+The smart grid's architecture based on \gls{iot} will be divided into three layers:
+\begin{itemize}
+\item Perception layer: Devices will autonomously perform their tasks such as sensing and executing orders.
+\item Communication layer: The communication infrastructure is defined here and issues related to how to communicate information is handled.
+\item Application layer: Applications can be built to make use of the underlying data.
+\end{itemize}
+\paragraph{}
+Some of the research is discussing the need to replace Zigbee by Wi-Fi as the Wi-Fi offers higher bandwidth, non-line transmission ability, large-scale data collection and is highly cost-effective \cite{ref10} but still Zigbee has the sole advantage of very low energy consumption which is something required in \gls{sg} and \gls{iot}, because in systems that are installed far away from any maintenance team such as smart grid monitoring and self-healing devices, one cannot manage to move for the sole purpose of providing energy often.
+\paragraph{}
+From the architectural point of view, integrating \gls{sg} within \gls{iot} means having to address heterogeneity issues.
+An \gls{iot} gateway system that is based on Zigbee and GPRS protocols \cite{ref11}.
+This enables then to deal partly with the heterogeneity problem and therefore enable the \gls{wsn} to communicate with the mobile telecommunication network.
+Another solution to the heterogeneity problem is proposed where the solution is to present a new, light-weight web service transport protocol called Lean Transport Protocol (LTP) that will allow transparent exchange of web service messages between all kinds of devices \cite{ref12}.
+This protocol is platform-independent, low energy consumer.
+Some researchers claim that the heterogeneity is mainly coming from the fact that there are different \gls{wsn} devices that do not necessarily use the same standards and protocols.
+They propose to move all the \glspl{wsn} to all-IP networks as this would remove most of the heterogeneity but there are already deployed legacy \glspl{wsn} \cite{ref13}.
+They sketch an architecture capable of converting all the \glspl{wsn}, new and legacy, to support \gls{ipv6}.\glsreset{sg}
+\section{The \gls{sg} two-way communication}
+\paragraph{}
 

### 2013-07-01 doc/thesis/report.tex

@@
 \end{dedication}
-\chapter{Acknowledgements}
+\chapter{Acknowledgments}
 \paragraph{}
@@
 \chapter{System Architecture}
+\paragraph{}
+The thesis is based on the study and creation of a test-bed system whose goal is to integrate smart home with the \glsreset{sg}\gls{sg} and all within the \glsreset{iot}\gls{iot}.
+To do so, we first define the system's objectives and goals which do align  with the thesis's problem statement and purpose of study.
+Then each part of the system is presented, and a section of this chapter is dedicated to each part explaining the architecture of each component and how it aligns and integrates with the whole system.
+\section{System objectives}
+\paragraph{Objective 1}
+The goal of the system is to present a working test-bed that is able to interact with any appliance in the smart home.
+The system makes this possible by making all appliances part of the \glsreset{wsn}\gls{wsn} where each sensor node is addressable and may be used to control the appliance.
+\paragraph{Objective 2}
+Another system objective is to be able to monitor in real-time the consumption of each appliance and send this to a central repository where this data is to be stored.
+This will enable the sensor node to keep track of the appliance it is controlling and monitoring and collect consumption data that is sent and stored in a \gls{rdbms}.
+\paragraph{Objective 3}
+The third objective is to make this system hide its heterogeneity from the outside world and also from different component belonging to it.
+This means that the system should be able to make any two components communicate independently of the technology they are using and also making each component think the other is using the same technology as his.
+
+\section{Division of components}
+\paragraph{}
+The system is divided into nine components where every component is in charge of a part of the system's objectives.
+When these nine components work hand-in-hand, they achieve all the system's and project's objectives and goals.
+The components that are defined for this system are the following:
+
+\begin{itemize}
+\item Full \gls{ipv6} \glsreset{wsn}\gls{wsn}
+\item Mote sensor data collection TCP client
+\item Mote appliance control TCP server
+\item Gateway packet transformation process
+\item \gls{ipv6} over USB tunnel
+\item mote \gls{ipv6} backstation
+\item Sensor data storage process
+\item HTTP sensor data sender
+\item Mobile client
+\end{itemize}
+
+\section{General architecture of the system}
+\paragraph{}
+The general architecture is depicted in figure \ref{fig:gen_architecture}.
+This figure shows the different components of the system and their interactions.
+Each set of components is appearing within a box and that box is labeled.
+This label refers to the hardware deployment of the component within the whole system.
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.40]{img/general_architecture.jpg}
+\caption{General architecture depicting the main components of the system}
+\label{fig:gen_architecture}
+\end{figure}
+\paragraph{}
+Now that the general architecture has been defined, the question that should be asked is whether all these components reside in the same layer in the \gls{osi} model and if not, in which layer does each of these components reside in the \gls{osi} model.
+As an answer to such a question, figure \ref{fig:layered_view} is here to show on what layer does every component act on.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.40]{img/layered_view.jpg}
+\caption{Layered view showing where every component appears in the \gls{osi} model}
+\label{fig:layered_view}
+\end{figure}
+\paragraph{}
+From figure \ref{fig:layered_view}, one can see that there are three components in performing in the data link layer and network layer.
+The first one is the packet transformation process whose goal is to hide the communication heterogeneity by sniffing incoming packets, transforming them, and forwarding them to the destination to meet the destination's communication technology and standards used.
+The second component performing in the data link layer and network is the \gls{ipv6} over USB tunnel.
+This component is in charge of getting raw \gls{ipv6} packets and encapsulate them through USB to reach the destination.
+The third component is deployed in the mote sink which is the entry and exit point of the \gls{wsn}.
+This special mote receives \gls{ipv6} packets encapsulated in USB frames that are forwarded to the destination mote using Zigbee.
+
+\paragraph{}
+There are four components that perform in the session layer, presentation layer and application layer and are existing in most of the devices needed in the system.
+The first one is the client mobile App.
+This component is mostly a TCP client and uses many of the services provided by the layers it is performing on.
+The second component is the sensor data storage which is a TCP server receiving data from the \gls{wsn} and sending storing it in a \gls{rdbms}.
+The third and fourth components are the Mote TCP client and Mote TCP server that send sensor data and through the TCP client and receive orders to control the appliance using the TCP server.
+\paragraph{}
+The only component acting in the application layer exclusively is the HTTP sensor data sender which provides a gateway to other applications willing to use such data and later on make use of the system.
+This component receives data and sends to the destination by using HTTP protocol.
+The middleware application was using HTTP to get sensor data from this system.
+\paragraph{}
+Now that the general structure of the software system has been established, the hardware structure is to be presented and explained further.
+figure \ref{fig:network_diagram} is a network diagram that shows the of the hardware components that exist within this system.
+The \glsreset{wsn}\gls{wsn} appears on the right of the document.
+Attached to it a mote sink that exist both in the \gls{wsn} and outside it.
+This mote sink is attached to the gateway server through USB.
+The gateway has also a Wi-Fi interface card that enables it to communicate with the other parts of the system such as the middleware and the mobile client.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.8]{img/network_diagram.jpg}
+\caption{Network diagram of the system}
+\label{fig:network_diagram}
+\end{figure}
+\paragraph{}
+Now that the network diagram has been established in \ref{fig:network_diagram} where the hardware component appears, the next thing is to show where each software component will be deployed.
+In other words, what are the software components that exist on which hardware machine.
+This is the goal of table \ref{table:soft_hard_deployment}.
+It clearly shows what software components exist in what hardware.
+From the table \ref{table:soft_hard_deployment}, it is clear that there are hardware machines that have more than one component deployed and this is also the case for small devices such as sensor nodes within the \gls{wsn}.
+
+\begin{table}
+    \begin{tabular}{lllll}
+    		  \hline
+    Software component - Hardware component & Mote & Mote sink & Gateway & Mobile \\ \hline
+    Mote sensor data collection TCP client  & X    & X         & ~       & ~      \\ 
+    Mote appliance control TCP server       & X    & X         & ~       & ~      \\
+    Gateway packet transformation process   & ~    & ~         & X       & ~      \\
+    IPv6 over USB tunnel                    & ~    & ~         & X       & ~      \\
+    Mote IPv6 backstation                   & ~    & X         & ~       & ~      \\
+    Sensor data storage process             & ~    & ~         & X       & ~      \\
+    HTTP sensor data sender                 & ~    & ~         & X       & ~      \\
+    Mobile client                           & ~    & ~         & ~       & X      \\
+    \end{tabular}
+    \caption{Software-hardware deployment table}
+    \label{table:soft_hard_deployment}
+\end{table}
+\paragraph{}
+In the following sections, each software component will be described in more details showing why such a component is crucial and what are the architectural decisions made to be what it is.
+
+\section{Mote sensor data collection TCP client}
+The mote sensor data collection TCP client is nothing mote but a \gls{tcp} client deployed in a mote that is attached to an appliance that reads sensor data periodically from the consumption sensor more known as the \gls{ct}  as showns in figure \ref{fig:current_transformer} that returns after a few transformations of the data the exact amount of power consumer in \gls{wh}.
+The original amount returned a voltage value in \gls{v} that is then transformed into the electric intensity in \gls{a}.
+To transform, the voltage into the actual current intensity we make use of the data provided by the \gls{ct} manufacturer.
+By making use of an experiment whose goals is to measure the slope and the constant of the voltage to intensity function.
+This function resulted in the following values:
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.8]{img/ct.jpg}
+\caption{Current transformer}
+\label{fig:current_transformer}
+\end{figure}
+
+\begin{table}
+    \begin{tabular}{ll}
+    \hline
+    slope    & 66.7 \gls{a} \\ \hline
+    Constant & 0 \gls{a}    \\ \hline
+    \end{tabular}
+    \caption{Current Transformer values}
+        \label{table:CT_vals}
+\end{table}
+The transformation function is then: \\
+\boldmath{$I = 66.7U_0$} \\
+I: Intensity in \glsreset{a}\gls{a} \\
+$U_0$: The voltage read from the \gls{ct}
+
+\paragraph{}
+Afterwards, as the voltage in Moroccan households is 220 \gls{v} we get the power using the following function:
+\\
+\boldmath{$ P = U_1 I $} \\
+P: Power in \gls{w} \\
+I: Intensity in \glsreset{a}\gls{a} \\
+$U_1$: Electrical potential difference or Voltage in household in \glsreset{v}\gls{v}
+\paragraph{}
+P refers to the power, not the actual consumption of the appliance.
+To get the consumption we use an additional parameter which is the sampling rate.
+The sampling rate is defined as the number of sensor reading actions per hour.
+If one reads sensor data every second, then the sampling rate is 3600 samples per second.
+\paragraph{}
+To find the actual consumption, we need to consecutive sensor readings that we call $P_0$ and $P_1$ and use the following function:
+\\
+$ C = \cfrac{P_0 + \cfrac{P_1 - P_0}{2}}{S} $ \\
+C: Current Consumption in \glsreset{wh}\gls{wh} \\
+P: Power sample number 1 in \glsreset{w}\gls{w} \\
+P': Power sample number 2 in \glsreset{w}\gls{w} \\
+$U_1$: Electrical potential difference or Voltage in households in \glsreset{v}\gls{v} \\
+S: Sampling rate \\
+The assumption taken is that if the power varies between two consecutive samples, the change was made linear throughout the interval of time between the two.
+\paragraph{}
+Once the consumption data in hand after this series of transformation, the next thing left to do is to send this data.
+To send it, the mote make use of the \gls{tcp}  client built on top of \gls{blip} package that implements the TCP/IP protocol stack.
+The system is made in a way that all the sensor data is sent to the gateway services that will decide to store it in a data store or send it to another system outside of the scope of this project.
 
@@
 \appendix
+\chapter{How to install the system}
 \chapter{TinyOs program}

### 2013-07-01 doc/thesis/report.tex

@@
 \section{Mote sensor data collection TCP client}
-The mote sensor data collection TCP client is nothing mote but a \gls{tcp} client deployed in a mote that is attached to an appliance that reads sensor data periodically from the consumption sensor more known as the \gls{ct}  as showns in figure \ref{fig:current_transformer} that returns after a few transformations of the data the exact amount of power consumer in \gls{wh}.
+The mote sensor data collection TCP client is nothing more but a \gls{tcp} client deployed in a mote that is attached to an appliance that reads sensor data periodically from the consumption sensor more known as the \gls{ct}  as showns in figure \ref{fig:current_transformer} that returns after a few transformations of the data the exact amount of power consumer in \gls{wh}.
 The original amount returned a voltage value in \gls{v} that is then transformed into the electric intensity in \gls{a}.
@@
 The system is made in a way that all the sensor data is sent to the gateway services that will decide to store it in a data store or send it to another system outside of the scope of this project.
+\section{ Mote appliance control TCP server}
+\paragraph{}
+The mote appliance control \glsreset{tcp}\gls{tcp} server is a component that is deployed in the sensor nodes of the \gls{wsn} to get commands from the outside world or at least outside of the \gls{wsn} in order control the appliance attached it.
+To control an appliance, one has access to two commands which are the ability to turn on and off a device.
+\paragraph{}
+Figure \ref{fig:appliance_control} depicts the architecture as well as the data flow of the command to turn on/ off an appliance.
+The TCP server hosted in the mote receives on/off commands from and external TCP client.
+The command is then extracted from the protocol of communication, and based on that, the PIN is known.
+the PIN information as well as the functionality is called as it is implemented in the MDA320CA that applies it on the pin and opens or closes the circuit depending on the command specified by the external TCP client.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.35]{img/appliance_control.jpg}
+\caption{TCP Server controling appliance}
+\label{fig:appliance_control}
+\end{figure}
+
+\paragraph{}
+The mote by itself does not have the necessary capabilities needed to control a mote.
+In order for a mote to have such capabilities, it needs to be equipped with an additional board, a Data Acquisition board also called a extension board that enable the mote to have more capabilities and therefore the ability to control the power flowing into the mote.
+The extension board used in the course of this project is the Crossbow MDA320CA as depicted in figure \ref{fig:mda320ca}.
+This board has pins which are electrical inputs and outputs that make it possible to have external analog sensors, digital sensors and also be able to power the appliance.
+But the powering of the appliances is not as straightforward as it seems especially that sensor nodes are very low-power devices that cannot provide enough energy to appliances such as a television or a fridge.
+\paragraph{}
+To make a mote control the appliance, the mote will control the power going in a very small circuit and this circuit is responsible for letting the power flow through the appliance from another source of energy which is naturally far more important than the one provided in the mote.
+In summary, the mote has a the extension board of type MDA320CA that has a small electric circuit attached to one of its digital pins that then controls the flow of electricity going to the appliance.
+By turning off a pin, the electricity stops flowing through the circuit and becomes open which stops the electricity to go to the appliance.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.8]{img/mda320.jpg}
+\caption{Crossbow MDA320CA Data acquisition board}
+\label{fig:mda320ca}
+\end{figure}
+
+\section{Mote IPv6 backstation}
+\paragraph{} 
+This component that exist in one special mote in the \glsreset{wsn}\gls{wsn}.
+This mote is able to communicate with the rest of the \gls{wsn} using Zigbee and also with the external world using \gls{usb}.
+This make this mote a sink and a router because it has two interface cards.
+The software component deployed in this sink is a packet forwarder.
+This means that once it receives an \gls{ipv6} packet that is not in the \gls{wsn}, it is forwarded through \gls{usb}.
+\paragraph{}
+The mote's sole goal is forwarding packets in and outside of the \gls{wsn} while other sensor node do not need to worry about how the packet will arrive to its destination.
+This mote is part of the components aiming to solve the heterogeneity problem stated in the problem statement.
+By making the sensor nodes talk to any device as if it is part of the \gls{wsn}, they do not really have to worry about how it will arrive or even how did the packet get to them in the first place.
+
+\section{IPv6 over USB tunnel}
+\paragraph{}
+The gateway server machine is equipped with two interface cards: the Wi-Fi network interface card and the Zigbee network interface card.
+The Zigbee network interface card is the one on focus in this section.
+To have a device communicating using Zigbee, one sensor node within the \gls{wsn} was used for this purpose and the software component implemented in this sensor node is the one described in section 4.6.
+That said, the packet is received through \gls{usb} but not read and recognized as a packet by the TCP/IP protocol stack.
+To do so, a tunnel had to be created for the purpose of extracting packets from the \gls{usb} port.
+Another goal is to forward \gls{ipv6} packets through that same \gls{usb} and this can only be done by the use of an \gls{ipv6} over \gls{usb} tunnel.
+
+\section{Gateway packet transformation process}
+The gateway packet transformation process is one of the main components that hides the heterogeneity from the rest of the system and the outside world. The
 
@@
 \chapter{How to install the system}
-\chapter{TinyOs program}
-\chapter{6to4/4to6 packet transformation}
 

### 2013-07-01 trunk/doc/thesis/report.tex

@@
 To my family \\
-To my friends
+To my friends \\
+
 \end{dedication}
-\chapter{Acknowledgements}
+\chapter{Acknowledgments}
 \paragraph{}
@@
 Another \gls{iot} concept is the ability of any node in the \gls{iot} to consume and produce information that is communicated through the network.
+\paragraph{}
 The autonomous production of information is shown by the fact that every sensor node in the \gls{wsn} reads sensory data from its sensors that are attached to appliance in order to sense their consumption and this information is produced and transmitted periodically.
@@
 \paragraph{}
-The thesis is is divided into eight chapters.
-Each of the chapter approaches thesis from a different aspect.
-The first chapter is an introductory chapter that gives a flavor of the project that is to be studied in this thesis and also introduces the problem and the solution proposed in this topic.
+The thesis is is divided into eight chapters. Each of the chapter approaches thesis from a different aspect. 
+\paragraph{Chapter 1}
+It is an introductory chapter that gives a flavor of the project that is to be studied in this thesis and also introduces the problem and the solution proposed in this topic.
 This chapter also puts the reader into perspective by introducing some of the main concepts related to the \glsreset{sg}\gls{sg} and the \glsreset{iot}\gls{iot}.
-\paragraph{}
-The second chapter presents the background of the work by revising the literature related to the topics of \gls{iot} and \gls{sg}.
+\paragraph{Chapter 2}
+It presents the background of the work by revising the literature related to the topics of \gls{iot} and \gls{sg}.
 It also discussed related work and how it was tackled in different research papers, how does it relate to the work done in this thesis and also how does it differ.
-\paragraph{}
-The third chapter is dedicated to the \glsreset{iot}\gls{iot}.
+\paragraph{Chapter 3}
+It is dedicated to the \glsreset{iot}\gls{iot}.
 It presents the topic, talks about the importance of \gls{iot}, its impact on the world, the actual Internet, the future of Internet and also how will \gls{sg} benefit from it.
-\paragraph{}
-The fourth chapter will present the architecture of the system and the test-bed. It will discuss the system from different aspects, and how were the two problems solved from the architectural aspect.
-\paragraph{}
-The fifth chapter will mostly discuss the implementation system.
+\paragraph{Chapter 4}
+It will present the architecture of the system and the test-bed. It will discuss the system from different aspects, and how were the two problems solved from the architectural aspect.
+\paragraph{Chapter 5}
+This chapter will mostly discuss the implementation system.
 It will start by presenting the hardware used in this project, then move to the technology enablers.
 Each component of the system will be discussed extensively and move to the integration of the components to form the test-bed.
+\paragraph{Chapter 6}
+It will start first by presenting the findings then move to explain the experiment that was done to measure the performance of the system.
+The data gathering section will show how and what data was gathered and then move to analyzing and interpreting the data.
+\paragraph{Chapter 7, 8}
+These are the last two chapters of the thesis that will conclude the thesis, provide directions about future work and recommendations to the reader what parts of the system could be ameliorated and how it can be achieved.
+\chapter{Literature review}
+\section{The need for Smart Grid}
 \paragraph{}
-The sixth chapter will start first by presenting the findings then move to explain the experiment that was done to measure the performance of the system.
-The data gathering section will show how and what data was gathered and then move to analyzing and interpreting the data.
+The actual electrical grid is based on a technology that is at least one century old \cite{ref5}.
+As a result, there is a need to change as to meet to 21st century requirement such as having more green energy in order to reduce the pollution caused mainly by the use of fossil energy.
+As a result, research has introduced a new technology called \glsreset{sg}\gls{sg}.
+Although it is still a research project and numerous issues related to \gls{sg} are left unanswered, nations across the globe started to deploy \gls{sg} project.
+It is also noted that in order for the grid to become more reliable, secure and efficient, a bi-directional information and communication infrastructure should be embedded within the grid \cite{ref5}.
 \paragraph{}
-Chapter seven and eight are the last two chapters of the thesis that will conclude the thesis, provide directions about future work and recommendations to the reader what parts of the system could be ameliorated and how it can be achieved.
-\chapter{Literature review}
+The smart grid brings around new concepts that will achieve efficiency, reliable and security to the power grid.
+The first one is the introduction of the \gls{dg} by exploiting \gls{res} \cite{ref6} and this reduce the losses of energy in the transmission.
+This distribution by itself enables the power grid this to move from the one way energy distribution to a two-way energy distribution where the utility will not be the sole energy producer but households can now produce and even sell the extra power.
+This is known nowadays as the feed-in-tariff which is a policy that allows the consumers to become producer and sell the excess power to the grid \cite{ref7}.
+This will allow the power grid to provide power to consumer in a "free market" fashion which means that it will obey to the laws of supply and demand \cite{ref7}.\glsreset{iot} \glsreset{sg}
+\section{\gls{iot} and \gls{sg}}
+\paragraph{}
+The \glsreset{iot}\gls{iot} has known high acceptance by the players in the Internet.
+Nowadays, it is estimated that 12.5 billion devices are connected to the Internet and that this trend will continue to read 50 billion connected devices by the end of the decade \cite{ref3}.
+\gls{iot} brings several benefits to the actual Internet and these benefits will pull \gls{sg} to be part of it.
+Consequently, research has taken this direction which is to integrate the \gls{sg} within the \gls{iot} as most aspects of \gls{sg} will be enhanced by \gls{iot} \cite{ref8}.
+\paragraph{}
+\gls{iot} counts on a set of technologies which are \glsreset{rfid}\gls{rfid}, \gls{m2m} \cite{ref9}, public communication networks such as 2G/3G mobile networks, power lines, WI-FI, Zigbee and other technologies \cite{ref8}.
+The smart grid's architecture based on \gls{iot} will be divided into three layers:
+\begin{itemize}
+\item Perception layer: Devices will autonomously perform their tasks such as sensing and executing orders.
+\item Communication layer: The communication infrastructure is defined here and issues related to how to communicate information is handled.
+\item Application layer: Applications can be built to make use of the underlying data.
+\end{itemize}
+\paragraph{}
+Some of the research is discussing the need to replace Zigbee by Wi-Fi as the Wi-Fi offers higher bandwidth, non-line transmission ability, large-scale data collection and is highly cost-effective \cite{ref10} but still Zigbee has the sole advantage of very low energy consumption which is something required in \gls{sg} and \gls{iot}, because in systems that are installed far away from any maintenance team such as smart grid monitoring and self-healing devices, one cannot manage to move for the sole purpose of providing energy often.
+\paragraph{}
+From the architectural point of view, integrating \gls{sg} within \gls{iot} means having to address heterogeneity issues.
+An \gls{iot} gateway system that is based on Zigbee and GPRS protocols \cite{ref11}.
+This enables then to deal partly with the heterogeneity problem and therefore enable the \gls{wsn} to communicate with the mobile telecommunication network.
+Another solution to the heterogeneity problem is proposed where the solution is to present a new, light-weight web service transport protocol called Lean Transport Protocol (LTP) that will allow transparent exchange of web service messages between all kinds of devices \cite{ref12}.
+This protocol is platform-independent, low energy consumer.
+Some researchers claim that the heterogeneity is mainly coming from the fact that there are different \gls{wsn} devices that do not necessarily use the same standards and protocols.
+They propose to move all the \glspl{wsn} to all-IP networks as this would remove most of the heterogeneity but there are already deployed legacy \glspl{wsn} \cite{ref13}.
+They sketch an architecture capable of converting all the \glspl{wsn}, new and legacy, to support \gls{ipv6}.\glsreset{sg}
+\section{The \gls{sg} two-way communication}
+\paragraph{}
 
@@
 \chapter{System Architecture}
+\paragraph{}
+The thesis is based on the study and creation of a test-bed system whose goal is to integrate smart home with the \glsreset{sg}\gls{sg} and all within the \glsreset{iot}\gls{iot}.
+To do so, we first define the system's objectives and goals which do align  with the thesis's problem statement and purpose of study.
+Then each part of the system is presented, and a section of this chapter is dedicated to each part explaining the architecture of each component and how it aligns and integrates with the whole system.
+\section{System objectives}
+\paragraph{Objective 1}
+The goal of the system is to present a working test-bed that is able to interact with any appliance in the smart home.
+The system makes this possible by making all appliances part of the \glsreset{wsn}\gls{wsn} where each sensor node is addressable and may be used to control the appliance.
+\paragraph{Objective 2}
+Another system objective is to be able to monitor in real-time the consumption of each appliance and send this to a central repository where this data is to be stored.
+This will enable the sensor node to keep track of the appliance it is controlling and monitoring and collect consumption data that is sent and stored in a \gls{rdbms}.
+\paragraph{Objective 3}
+The third objective is to make this system hide its heterogeneity from the outside world and also from different component belonging to it.
+This means that the system should be able to make any two components communicate independently of the technology they are using and also making each component think the other is using the same technology as his.
+
+\section{Division of components}
+\paragraph{}
+The system is divided into nine components where every component is in charge of a part of the system's objectives.
+When these nine components work hand-in-hand, they achieve all the system's and project's objectives and goals.
+The components that are defined for this system are the following:
+
+\begin{itemize}
+\item Full \gls{ipv6} \glsreset{wsn}\gls{wsn}
+\item Mote sensor data collection TCP client
+\item Mote appliance control TCP server
+\item Gateway packet transformation process
+\item \gls{ipv6} over USB tunnel
+\item mote \gls{ipv6} backstation
+\item Sensor data storage process
+\item HTTP sensor data sender
+\item Mobile client
+\end{itemize}
+
+\section{General architecture of the system}
+\paragraph{}
+The general architecture is depicted in figure \ref{fig:gen_architecture}.
+This figure shows the different components of the system and their interactions.
+Each set of components is appearing within a box and that box is labeled.
+This label refers to the hardware deployment of the component within the whole system.
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.40]{img/general_architecture.jpg}
+\caption{General architecture depicting the main components of the system}
+\label{fig:gen_architecture}
+\end{figure}
+\paragraph{}
+Now that the general architecture has been defined, the question that should be asked is whether all these components reside in the same layer in the \gls{osi} model and if not, in which layer does each of these components reside in the \gls{osi} model.
+As an answer to such a question, figure \ref{fig:layered_view} is here to show on what layer does every component act on.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.40]{img/layered_view.jpg}
+\caption{Layered view showing where every component appears in the \gls{osi} model}
+\label{fig:layered_view}
+\end{figure}
+\paragraph{}
+From figure \ref{fig:layered_view}, one can see that there are three components in performing in the data link layer and network layer.
+The first one is the packet transformation process whose goal is to hide the communication heterogeneity by sniffing incoming packets, transforming them, and forwarding them to the destination to meet the destination's communication technology and standards used.
+The second component performing in the data link layer and network is the \gls{ipv6} over USB tunnel.
+This component is in charge of getting raw \gls{ipv6} packets and encapsulate them through USB to reach the destination.
+The third component is deployed in the mote sink which is the entry and exit point of the \gls{wsn}.
+This special mote receives \gls{ipv6} packets encapsulated in USB frames that are forwarded to the destination mote using Zigbee.
+
+\paragraph{}
+There are four components that perform in the session layer, presentation layer and application layer and are existing in most of the devices needed in the system.
+The first one is the client mobile App.
+This component is mostly a TCP client and uses many of the services provided by the layers it is performing on.
+The second component is the sensor data storage which is a TCP server receiving data from the \gls{wsn} and sending storing it in a \gls{rdbms}.
+The third and fourth components are the Mote TCP client and Mote TCP server that send sensor data and through the TCP client and receive orders to control the appliance using the TCP server.
+\paragraph{}
+The only component acting in the application layer exclusively is the HTTP sensor data sender which provides a gateway to other applications willing to use such data and later on make use of the system.
+This component receives data and sends to the destination by using HTTP protocol.
+The middleware application was using HTTP to get sensor data from this system.
+\paragraph{}
+Now that the general structure of the software system has been established, the hardware structure is to be presented and explained further.
+figure \ref{fig:network_diagram} is a network diagram that shows the of the hardware components that exist within this system.
+The \glsreset{wsn}\gls{wsn} appears on the right of the document.
+Attached to it a mote sink that exist both in the \gls{wsn} and outside it.
+This mote sink is attached to the gateway server through USB.
+The gateway has also a Wi-Fi interface card that enables it to communicate with the other parts of the system such as the middleware and the mobile client.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.8]{img/network_diagram.jpg}
+\caption{Network diagram of the system}
+\label{fig:network_diagram}
+\end{figure}
+\paragraph{}
+Now that the network diagram has been established in \ref{fig:network_diagram} where the hardware component appears, the next thing is to show where each software component will be deployed.
+In other words, what are the software components that exist on which hardware machine.
+This is the goal of table \ref{table:soft_hard_deployment}.
+It clearly shows what software components exist in what hardware.
+From the table \ref{table:soft_hard_deployment}, it is clear that there are hardware machines that have more than one component deployed and this is also the case for small devices such as sensor nodes within the \gls{wsn}.
+
+\begin{table}
+    \begin{tabular}{lllll}
+    		  \hline
+    Software component - Hardware component & Mote & Mote sink & Gateway & Mobile \\ \hline
+    Mote sensor data collection TCP client  & X    & X         & ~       & ~      \\ 
+    Mote appliance control TCP server       & X    & X         & ~       & ~      \\
+    Gateway packet transformation process   & ~    & ~         & X       & ~      \\
+    IPv6 over USB tunnel                    & ~    & ~         & X       & ~      \\
+    Mote IPv6 backstation                   & ~    & X         & ~       & ~      \\
+    Sensor data storage process             & ~    & ~         & X       & ~      \\
+    HTTP sensor data sender                 & ~    & ~         & X       & ~      \\
+    Mobile client                           & ~    & ~         & ~       & X      \\
+    \end{tabular}
+    \caption{Software-hardware deployment table}
+    \label{table:soft_hard_deployment}
+\end{table}
+\paragraph{}
+In the following sections, each software component will be described in more details showing why such a component is crucial and what are the architectural decisions made to be what it is.
+
+\section{Mote sensor data collection TCP client}
+The mote sensor data collection TCP client is nothing more but a \gls{tcp} client deployed in a mote that is attached to an appliance that reads sensor data periodically from the consumption sensor more known as the \gls{ct}  as showns in figure \ref{fig:current_transformer} that returns after a few transformations of the data the exact amount of power consumer in \gls{wh}.
+The original amount returned a voltage value in \gls{v} that is then transformed into the electric intensity in \gls{a}.
+To transform, the voltage into the actual current intensity we make use of the data provided by the \gls{ct} manufacturer.
+By making use of an experiment whose goals is to measure the slope and the constant of the voltage to intensity function.
+This function resulted in the following values:
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.8]{img/ct.jpg}
+\caption{Current transformer}
+\label{fig:current_transformer}
+\end{figure}
+
+\begin{table}
+    \begin{tabular}{ll}
+    \hline
+    slope    & 66.7 \gls{a} \\ \hline
+    Constant & 0 \gls{a}    \\ \hline
+    \end{tabular}
+    \caption{Current Transformer values}
+        \label{table:CT_vals}
+\end{table}
+The transformation function is then: \\
+\boldmath{$I = 66.7U_0$} \\
+I: Intensity in \glsreset{a}\gls{a} \\
+$U_0$: The voltage read from the \gls{ct}
+
+\paragraph{}
+Afterwards, as the voltage in Moroccan households is 220 \gls{v} we get the power using the following function:
+\\
+\boldmath{$ P = U_1 I $} \\
+P: Power in \gls{w} \\
+I: Intensity in \glsreset{a}\gls{a} \\
+$U_1$: Electrical potential difference or Voltage in household in \glsreset{v}\gls{v}
+\paragraph{}
+P refers to the power, not the actual consumption of the appliance.
+To get the consumption we use an additional parameter which is the sampling rate.
+The sampling rate is defined as the number of sensor reading actions per hour.
+If one reads sensor data every second, then the sampling rate is 3600 samples per second.
+\paragraph{}
+To find the actual consumption, we need to consecutive sensor readings that we call $P_0$ and $P_1$ and use the following function:
+\\
+$ C = \cfrac{P_0 + \cfrac{P_1 - P_0}{2}}{S} $ \\
+C: Current Consumption in \glsreset{wh}\gls{wh} \\
+P: Power sample number 1 in \glsreset{w}\gls{w} \\
+P': Power sample number 2 in \glsreset{w}\gls{w} \\
+$U_1$: Electrical potential difference or Voltage in households in \glsreset{v}\gls{v} \\
+S: Sampling rate \\
+The assumption taken is that if the power varies between two consecutive samples, the change was made linear throughout the interval of time between the two.
+\paragraph{}
+Once the consumption data in hand after this series of transformation, the next thing left to do is to send this data.
+To send it, the mote make use of the \gls{tcp}  client built on top of \gls{blip} package that implements the TCP/IP protocol stack.
+The system is made in a way that all the sensor data is sent to the gateway services that will decide to store it in a data store or send it to another system outside of the scope of this project.
+\section{ Mote appliance control TCP server}
+\paragraph{}
+The mote appliance control \glsreset{tcp}\gls{tcp} server is a component that is deployed in the sensor nodes of the \gls{wsn} to get commands from the outside world or at least outside of the \gls{wsn} in order control the appliance attached it.
+To control an appliance, one has access to two commands which are the ability to turn on and off a device.
+\paragraph{}
+Figure \ref{fig:appliance_control} depicts the architecture as well as the data flow of the command to turn on/ off an appliance.
+The TCP server hosted in the mote receives on/off commands from and external TCP client.
+The command is then extracted from the protocol of communication, and based on that, the PIN is known.
+the PIN information as well as the functionality is called as it is implemented in the MDA320CA that applies it on the pin and opens or closes the circuit depending on the command specified by the external TCP client.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.35]{img/appliance_control.jpg}
+\caption{TCP Server controling appliance}
+\label{fig:appliance_control}
+\end{figure}
+
+\paragraph{}
+The mote by itself does not have the necessary capabilities needed to control a mote.
+In order for a mote to have such capabilities, it needs to be equipped with an additional board, a Data Acquisition board also called a extension board that enable the mote to have more capabilities and therefore the ability to control the power flowing into the mote.
+The extension board used in the course of this project is the Crossbow MDA320CA as depicted in figure \ref{fig:mda320ca}.
+This board has pins which are electrical inputs and outputs that make it possible to have external analog sensors, digital sensors and also be able to power the appliance.
+But the powering of the appliances is not as straightforward as it seems especially that sensor nodes are very low-power devices that cannot provide enough energy to appliances such as a television or a fridge.
+\paragraph{}
+To make a mote control the appliance, the mote will control the power going in a very small circuit and this circuit is responsible for letting the power flow through the appliance from another source of energy which is naturally far more important than the one provided in the mote.
+In summary, the mote has a the extension board of type MDA320CA that has a small electric circuit attached to one of its digital pins that then controls the flow of electricity going to the appliance.
+By turning off a pin, the electricity stops flowing through the circuit and becomes open which stops the electricity to go to the appliance.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.8]{img/mda320.jpg}
+\caption{Crossbow MDA320CA Data acquisition board}
+\label{fig:mda320ca}
+\end{figure}
+
+\section{Mote IPv6 backstation}
+\paragraph{} 
+This component that exist in one special mote in the \glsreset{wsn}\gls{wsn}.
+This mote is able to communicate with the rest of the \gls{wsn} using Zigbee and also with the external world using \gls{usb}.
+This make this mote a sink and a router because it has two interface cards.
+The software component deployed in this sink is a packet forwarder.
+This means that once it receives an \gls{ipv6} packet that is not in the \gls{wsn}, it is forwarded through \gls{usb}.
+\paragraph{}
+The mote's sole goal is forwarding packets in and outside of the \gls{wsn} while other sensor node do not need to worry about how the packet will arrive to its destination.
+This mote is part of the components aiming to solve the heterogeneity problem stated in the problem statement.
+By making the sensor nodes talk to any device as if it is part of the \gls{wsn}, they do not really have to worry about how it will arrive or even how did the packet get to them in the first place.
+
+\section{IPv6 over USB tunnel}
+\paragraph{}
+The gateway server machine is equipped with two interface cards: the Wi-Fi network interface card and the Zigbee network interface card.
+The Zigbee network interface card is the one on focus in this section.
+To have a device communicating using Zigbee, one sensor node within the \gls{wsn} was used for this purpose and the software component implemented in this sensor node is the one described in section 4.6.
+That said, the packet is received through \gls{usb} but not read and recognized as a packet by the TCP/IP protocol stack.
+To do so, a tunnel had to be created for the purpose of extracting packets from the \gls{usb} port.
+Another goal is to forward \gls{ipv6} packets through that same \gls{usb} and this can only be done by the use of an \gls{ipv6} over \gls{usb} tunnel.
+
+\section{Gateway packet transformation process}
+The gateway packet transformation process is one of the main components that hide the heterogeneity from the rest of the system and the outside world.
+This component mainly deals with packet transformation which means that it extracts the data from the packet and create a completely new one that holds the data and forward it.
+This component is composed of three subcomponents:
+\begin{itemize}
+\item Packet sniffer
+\item Packet transformer
+\item Address translator
+\end{itemize}
+\paragraph{}
+Figure \ref{fig:gateway_transformation} depicts the architecture of this component by showing the the different subcomponents that reside in this component and their interaction.
+It also shows the data flow within this component and what are the inputs and outputs of the system.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.6]{img/packet_transformation.jpg}
+\caption{Gateway packet transformation subsystem data flow}
+\label{fig:gateway_transformation}
+\end{figure}
+
+\subsection{Packet sniffer}
 
@@
 \appendix
-\chapter{TinyOs program}
-\chapter{6to4/4to6 packet transformation}
+\chapter{How to install the system}
 

### 2013-07-01 trunk/doc/thesis/report.tex

@@
 \end{figure}
+\paragraph{}
+The subsections 4.8.1 - 4.8.3 will go more into details in explaining the three subcomponents of the system.
 
 \subsection{Packet sniffer}
+\paragraph{}
+The packet sniffer subcomponent is the one responsible for sniffing and forwarding the packets to the component.
+It does not forward all the packets that are not going to the gateway machine.
+In other words, all the packets that are not meant to be received by the gateway server are sniffed and forwarded to the packet transformer that will continue the work.
+\subsection{Packet transformer}
+\paragraph{}
+Once the packet is received, the first thing this subcomponent does is check from where did it come, was it from the \gls{wsn} or the outside world.
+If it came from the outside world, the packet will only be examined if it is meant to go the \gls{wsn}.
+But an important and crucial problem presents itself:
+\paragraph{Problem}
+\underline{The outside world mostly uses \gls{ipv4} whereas the \gls{wsn} uses exclusively \gls{ipv6}} \\
+Two networks using different network layer technology cannot communicate even though \gls{ipv6} is nothing more but a new version of the very known \gls{ipv4} and this is because \gls{ipv6} has no backward compatibility with \gls{ipv4}.
+\paragraph{Solution}
+To solve such a problem, we decide to provide all the nodes in the this system with both an \gls{ipv4} and \gls{ipv6} address.
+Some of the nodes have either \gls{ipv4} or \gls{ipv6} addresses, but for the technology missing we assign the missing one virtually.
+For example, a sensor node in the \gls{wsn} with a real \gls{ipv6} address equal to FEC0::1 will be assigned a virtual \gls{ipv4} address equal to 192.168.0.1 but he will not be aware of it.
+Only the Gateway packet transformation component will be aware of it and this is the mechanism used to transform the packets and hide the heterogeneity from the rest of the components of the system.
+\paragraph{}
+Now that a packet is received, it checks whether this component came from the \gls{wsn} or the outside world, depending on that, it extracts the data from the packet which can be a \gls{tcp} or \gls{udp} datagram.
+The source and destination addresses are extracted from the packet and translated to the other network technology.
+Afterwards, a new packet referring to the destination technology used is created by having as a source and destination addresses the translated addresses and as payload the the \gls{tcp} \gls{udp} datagram coming from the original packet.
+Finally, the packet is forwarded to the network.
+\subsection{Address translator}
+The address translator is a very specialized subcomponent whose sole goal is to translate addresses between \gls{ipv4} and \gls{ipv6}.
+That is said, this component receives the address to translate and returns the translated one.
+Its concern is that the translation algorithm yields a bijective function which means that for any \gls{ipv4} address there is one and only one \gls{ipv6} address resulting from that source and vice versa.
+This statement compresses a set of statements which are the following:
 
+\begin{itemize}
+\item Any address belonging to one technology can be translated to the other technology. This is known as surjectivity.
+\item Any address cannot be translated into more than one address from the other technology. This is known as injectivity.
+\end{itemize}
+\paragraph{}
+The gateway packet transformation process is an important component responsible for dealing with the heterogeneity problem.
+It hides most of the heterogeneity in the system and make communication go independently of the data link layer or network layer technologies used.
+
+\section{Sensor data storage process}
+\paragraph{}
+The sensor data storage process is a component small component within the system whose sole purpose is to store such data.
+It seen by the \gls{wsn} as a TCP server, where sensor nodes get connected to it as soon as they are turned ON.
+The sensor node starts sending packet periodically (depending on the sampling rate specified) and whenever the sensor data is received, it is stored into a \gls{rdbms} where the schema is created as shown in figure \ref{fig:erd}.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.5]{img/smart_grid_db.jpg}
+\caption{Entity relationship diagram}
+\label{fig:erd}
+\end{figure}
+
+\section{HTTP sensor data sender} 
+The HTTP sensor data sender is TCP server to which sensor nodes get connected to and start sending their data periodically.
+Once the data is received they are able to send data to the outside world periodically depending on the needs of the clients outside the system.
+This system counts on the deployment of an HTTP server at the level of outside the system client and send an HTTP request containing the sensor data received from the \gls{wsn}.
+\paragraph{}
+The sensor nodes are able to send this data by themselves directly to the outside as this is part of the two-way communication but for convenience, it was decided that the \gls{wsn} sends its data to the gateway and it is up to the gateway to send such data or store it in the \gls{rdbms}.
+\section{Mobile client}
+The mobile client is a component of the system but can be seen as an outside world client as it is here to receive sensor data and also send ON/OFF commands to the system.
 \chapter{Implementation}

### 2013-07-02 trunk/doc/thesis/report.tex
chapter 5 almost done

@@
 \listoftables
+\printglossaries
 
@@
 
-\section{Importance of study}
+\section{Importance of Study}
 \paragraph{}
@@
 
-\section{Rationale of study}
+\section{Rationale of Study}
 \paragraph{}
@@
 
-\section{Problem statement}
+\section{Problem Statement}
 \paragraph{}
@@
 
-\section{Purpose of study}
+\section{Purpose of Study}
  \paragraph{}
@@
 
-\section{scope of study}
+\section{Scope of Study}
 \paragraph{}
@@
 The sole reason behind is the fact that some of the equipment required for this thesis did not arrive on time and therefore the parts in the implementation requiring such equipment was omitted to meet the deadlines.
-\section{Outcome of the study}
+\section{Outcome of the Study}
 \paragraph{}
@@
 These are the last two chapters of the thesis that will conclude the thesis, provide directions about future work and recommendations to the reader what parts of the system could be ameliorated and how it can be achieved.
-\chapter{Literature review}
+\chapter{Literature Review}
 \section{The need for Smart Grid}
@@
 They sketch an architecture capable of converting all the \glspl{wsn}, new and legacy, to support \gls{ipv6}.\glsreset{sg}
-\section{The \gls{sg} two-way communication}
+\section{The \gls{sg} Two-Way Communication}
 \paragraph{}
@@
 \section{Importance of \gls{iot}}
-\section{The impact of \gls{iot}}
-\section{The benefits of \gls{iot}}
-\section{The future of Internet}
+\section{The Impact of \gls{iot}}
+\section{The Benefits of \gls{iot}}
+\section{The Future of Internet}
 
@@
 Then each part of the system is presented, and a section of this chapter is dedicated to each part explaining the architecture of each component and how it aligns and integrates with the whole system.
-\section{System objectives}
+\section{System Objectives}
 \paragraph{Objective 1}
@@
 
-\section{Division of components}
+\section{System Components}
 \paragraph{}
@@
 
-\section{General architecture of the system}
+\section{General Architecture of the System}
 \paragraph{}
@@
 
-\section{Mote sensor data collection TCP client}
+\section{Mote Sensor Data Collection TCP Client}
 The mote sensor data collection TCP client is nothing more but a \gls{tcp} client deployed in a mote that is attached to an appliance that reads sensor data periodically from the consumption sensor more known as the \gls{ct}  as showns in figure \ref{fig:current_transformer} that returns after a few transformations of the data the exact amount of power consumer in \gls{wh}.
@@
 
-\section{Mote IPv6 backstation}
+\section{Mote IPv6 Backstation}
 \paragraph{} 
@@
 
-\section{IPv6 over USB tunnel}
+\section{IPv6 over USB Tunnel}
 \paragraph{}
@@
 
-\section{Gateway packet transformation process}
+\section{Gateway Packet Transformation Process}
 The gateway packet transformation process is one of the main components that hide the heterogeneity from the rest of the system and the outside world.
@@
 
-\subsection{Packet sniffer}
+\subsection{Packet Sniffer}
 \paragraph{}
@@
 In other words, all the packets that are not meant to be received by the gateway server are sniffed and forwarded to the packet transformer that will continue the work.
-\subsection{Packet transformer}
+\subsection{Packet Transformer}
 \paragraph{}
@@
 Finally, the packet is forwarded to the network.
-\subsection{Address translator}
+\subsection{Address Translator}
 The address translator is a very specialized subcomponent whose sole goal is to translate addresses between \gls{ipv4} and \gls{ipv6}.
@@
 
-\section{Sensor data storage process}
+\section{Sensor Data Storage Process}
 \paragraph{}
@@
 
-\section{HTTP sensor data sender} 
+\section{HTTP Sensor Data Sender} 
 The HTTP sensor data sender is TCP server to which sensor nodes get connected to and start sending their data periodically.
@@
 The sensor nodes are able to send this data by themselves directly to the outside as this is part of the two-way communication but for convenience, it was decided that the \gls{wsn} sends its data to the gateway and it is up to the gateway to send such data or store it in the \gls{rdbms}.
-\section{Mobile client}
+\section{Mobile Client}
 The mobile client is a component of the system but can be seen as an outside world client as it is here to receive sensor data and also send ON/OFF commands to the system.
-\chapter{Implementation}
+
+\chapter{Deployment}
+\paragraph{}
+In this chapter, the test-bed will be discussed from the implementation point of view.
+That is said, each component within the system will be given as much details as possible in order for the reader to open the code and understand what is going.
+Chapter 4 discussed the components from the architectural aspect but no implementation and technology decisions were discussed and the goal of this chapter is to explain these details.
+\section{Technology Enablers}
+The system's implementation was built using a set of software and hardware technologies and standards that were used throughout the course of this project to make the test-bed achieve its objectives.
+%\glsreset{wsn}
+\subsection{Technologies used in \gls{wsn}}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.6]{img/crossbow.jpg}
+\caption{Crossbow Professional Kit}
+\label{fig:crossbow}
+\end{figure}
+
+The \gls{wsn} was built using Crossbow professional kit as shown in figure \ref{fig:crossbow} which counts 7 Micaz sensor nodes of type MPR2600 as shows in figure \ref{fig:micaz}.
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.8]{img/micaz.png}
+\caption{Crossbow MPR2600 Micaz Mote}
+\label{fig:micaz}
+\end{figure}
+
+In addition, each mote is supplied with a sensor board of type Crossbow MTS400CA which contains temperature, humidity, barometric pressure, ambient light and infrared light sensors that are used to send different types of sensor data to the gateway.
+This board is shown in figure \ref{fig:mts400}.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.5]{img/mts400ca.jpg}
+\caption{Crossbow MTS400CA Sensor Board}
+\label{fig:mts400}
+\end{figure}
+\paragraph{}
+The WSN is run using \gls{tinyos} which is one of the most used opensource \gls{os} in \glspl{wsn}.
+Programs are implemented using \gls{nesc} which is the sole programming language used in the \gls{tinyos} environment.
+This programming language is event-oriented and its syntax is very influenced by C.
+In addition, C libraries can be used in \gls{nesc} programs.
+
+\subsection{Technologies used in the Gateway}
+\paragraph{}
+The gateway's hardware machine is Dell Optiplex 9010 running a Linux environment more specifically the Ubuntu 13.04 distribution.
+The gateway has a set of technologies used to provide numerous services the system counts on.
+The programming languages used in the components hosted in this machine are C and \gls{j2se}.
+Netfilter is also used to control incoming and outgoing packets.
+Concerning the storage of sensor data MySQl is the \gls{rdbms} used.
+
+\subsection{Technologies used in the Mobile Client}
+\paragraph{}
+The mobile used is an Android smart phone with Android 2.3. The application developed is using \gls{j2se}, PHP for database connectivity.
+
+\section{Mote TCP Client And TCP Server}
+\paragraph{}
+Section 4.4 and 4.5 were dedicated to explain the TCP Client and TCP server separately.
+In this section, one section will explain both of them because from the implementation point of view,  both of them are implemented in the \gls{tinyos} program.
+\paragraph{}
+A typical \gls{tinyos} program has two parts: the components definition part and the module part.
+In the components definition part, the main components that will be used.
+These components refer to already implemented modules whose functionalities will be used in the in-development module.
+In the module part, we provide define the interfaces we will be using and the interfaces we will be providing if any \cite{ref14}.
+Back to definition part, after defining components we will be using, we then link the interfaces needed in the module to the components providing that interface.
+This is known as wiring.
+\paragraph{}
+The components needed in this program are the following:
+\begin{itemize}
+\item MainC: provides the essential functionalities such as the initialization and closing of the program.
+\item LedsC: provides access to the leds existing in the mote
+\item Taos2550C: is a light sensor component that enable to read data
+\item TimerMillic: A Timer that will be used for sampling
+\item TCPSocket as client: Is the TCP client
+\item TCPSocket as Server: Is the TCP server
+\end{itemize}
+In the modules part we define the components that will be used which are the following:
+\begin{itemize}
+\item Boot: This is the interface that provides main functions of booting and closing
+\item TCP as server: The main interface for TCP
+\item TCP as client: The main interface for TCP
+\item Leds: The interface to access the leds
+\item timer as sensingTimer: The interface to program and use the timer
+\item Read as Sensor: The interface to talk with the sensor
+\end{itemize}
+Some of the interface do not have all the functions provided and the not implemented function are implemented in the module. The following part will talk about each function that is to be implemented
+\paragraph{After boot of the mote}
+The first thing done after the mote is booted is to initialize the TCP client and server.
+The Client is given a random port by using the rand function.
+The TCP server is binded to port 7.
+The client is set to connect to the Data store in the gateway by providing the \gls{ipv6} address of the gateway and port number of the server.
+Afterwards, the timer is set to be triggered every 10 seconds which means that the sampling rate will be 360 sensor readings per hour.
+\paragraph{TCP Server data received}
+Once the data is received from a client, it assumes that there are two data items separated by a space.
+The first item refers to the led number(1,2,3) and the second refers to the command 0 meaning turn Off and 1 means turn On.
+The string is tokenized and the data items are extracted and the command is executed accordingly.
+\paragraph{Timer fired}
+Once the timer is fired, this means it is time to read sensor data from the sensor.
+The command to read from the sensor is called which goes and reads data and once the data is read the event "read done" is triggered.
+\paragraph{Read done}
+Once the data is read and ready, the next thing is to transform it and then send it.
+Concerning the transformation, a series of equations are applied as seen in section 4.4 which build the data value from the original value gotten and it is ready to be sent.
+A message is created containing two values: the mote NODE\_ID which is the identifier of the mote predefined in the installation of the program in the mote and the sensor data value.
+These two values are separated by a space, encapsulated within a \gls{tcp} datagram and sent.
+\paragraph{}
+During the installation of the software the \glsreset{blip}\gls{blip} package should specified as it is a dependency of the software system.
+\gls{blip} provides the TCP/IP protocol stack that implements \gls{6lowpan} standard by \gls{ietf}.
+In  the installation the NODE\_ID is specified.
+If NODE\_ID is 1, this will make the mote have an \gls{ipv6} of FEC0::1, as the NODE\_ID is appended to the predefined prefix of \gls{wsn} during the installation.
+\section{Mote IPv6 Backstation}
+\paragraph{}
+This component is provided as part of the \gls{blip} software package.
+It is \gls{tinyos} program whose goal is to forward packet back and forth between the \gls{wsn} and the outside.
+This mote is connected to the \gls{wsn} using Zigbee and to the outside world using \gls{usb}.
+This special mote has its \gls{ipv6} being FEC0::64.
+It receives \gls{ipv6} packets encapsulated in \gls{usb} frames, extracts them and encapsulates them in a Zigbee frame and vice versa.
+\paragraph{}
+During the installation of this \gls{tinyos} program into the mote sink, one should make sure that it is being installed in the mote sink as \gls{usb} is a necessity.
+The \gls{ipv6} address does not need to be provided as it is already predefined in the program.
+
+\section{IPv6 over USB Tunnel}
+\paragraph{}
+This is a program that is installed in the gateway server machine.
+As the mote sink is connected to the gateway using \gls{usb}, the gateway will recognize it as a network interface card using \gls{ipv6} as a network technology.
+That is said, this component is part of the \gls{blip} software package and creates the network interface card module and starts extracting incoming \gls{ipv6} packets from the \gls{usb} frames and encapsulates outgoing \gls{ipv6} into \gls{usb} frames.
+\paragraph{}
+The installation and initialization of this component will be given more details in appendix A.
+\section{Gateway Packet Transformation Process}
+\paragraph{}
+This component is in charge of hiding the heterogeneity of the system from the rest of the the components of the system and also the outside world.
+It does that by intercepting all the packet moving and transforms them according to the standards used by the recipient.
+\paragraph{}
 
@@
 %\subsection{Methodology}
-\section{Data gathering}
-\section{Interpretation of result}
+\section{Data Gathering}
+\section{Interpretation of Results}
 
@@
 \singlespacing
-\printglossaries
 \bibliographystyle{IEEEtran} 

### 2013-07-02 doc/thesis/report.tex
implementation and appendices done

@@
 %\usepackage{hyperref}
+\usepackage{listings}
+\lstset{language=C}
 \usepackage[acronym]{glossaries} 
@@
 
-\section{Gateway packet transformation process}
-The gateway packet transformation process is one of the main components that hides the heterogeneity from the rest of the system and the outside world. The
+\section{Gateway Packet Transformation Process}
+The gateway packet transformation process is one of the main components that hide the heterogeneity from the rest of the system and the outside world.
+This component mainly deals with packet transformation which means that it extracts the data from the packet and create a completely new one that holds the data and forward it.
+This component is composed of three subcomponents:
+\begin{itemize}
+\item Packet sniffer
+\item Packet transformer
+\item Address translator
+\end{itemize}
+\paragraph{}
+Figure \ref{fig:gateway_transformation} depicts the architecture of this component by showing the the different subcomponents that reside in this component and their interaction.
+It also shows the data flow within this component and what are the inputs and outputs of the system.
 
-\chapter{Implementation}
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.6]{img/packet_transformation.jpg}
+\caption{Gateway packet transformation subsystem data flow}
+\label{fig:gateway_transformation}
+\end{figure}
+\paragraph{}
+The subsections 4.8.1 - 4.8.3 will go more into details in explaining the three subcomponents of the system.
+
+\subsection{Packet Sniffer}
+\paragraph{}
+The packet sniffer subcomponent is the one responsible for sniffing and forwarding the packets to the component.
+It does not forward all the packets that are not going to the gateway machine.
+In other words, all the packets that are not meant to be received by the gateway server are sniffed and forwarded to the packet transformer that will continue the work.
+\subsection{Packet Transformer}
+\paragraph{}
+Once the packet is received, the first thing this subcomponent does is check from where did it come, was it from the \gls{wsn} or the outside world.
+If it came from the outside world, the packet will only be examined if it is meant to go the \gls{wsn}.
+But an important and crucial problem presents itself:
+\paragraph{Problem}
+\underline{The outside world mostly uses \gls{ipv4} whereas the \gls{wsn} uses exclusively \gls{ipv6}} \\
+Two networks using different network layer technology cannot communicate even though \gls{ipv6} is nothing more but a new version of the very known \gls{ipv4} and this is because \gls{ipv6} has no backward compatibility with \gls{ipv4}.
+\paragraph{Solution}
+To solve such a problem, we decide to provide all the nodes in the this system with both an \gls{ipv4} and \gls{ipv6} address.
+Some of the nodes have either \gls{ipv4} or \gls{ipv6} addresses, but for the technology missing we assign the missing one virtually.
+For example, a sensor node in the \gls{wsn} with a real \gls{ipv6} address equal to FEC0::1 will be assigned a virtual \gls{ipv4} address equal to 192.168.0.1 but he will not be aware of it.
+Only the Gateway packet transformation component will be aware of it and this is the mechanism used to transform the packets and hide the heterogeneity from the rest of the components of the system.
+\paragraph{}
+Now that a packet is received, it checks whether this component came from the \gls{wsn} or the outside world, depending on that, it extracts the data from the packet which can be a \gls{tcp} or \gls{udp} datagram.
+The source and destination addresses are extracted from the packet and translated to the other network technology.
+Afterwards, a new packet referring to the destination technology used is created by having as a source and destination addresses the translated addresses and as payload the the \gls{tcp} \gls{udp} datagram coming from the original packet.
+Finally, the packet is forwarded to the network.
+\subsection{Address Translator}
+The address translator is a very specialized subcomponent whose sole goal is to translate addresses between \gls{ipv4} and \gls{ipv6}.
+That is said, this component receives the address to translate and returns the translated one.
+Its concern is that the translation algorithm yields a bijective function which means that for any \gls{ipv4} address there is one and only one \gls{ipv6} address resulting from that source and vice versa.
+This statement compresses a set of statements which are the following:
+
+\begin{itemize}
+\item Any address belonging to one technology can be translated to the other technology. This is known as surjectivity.
+\item Any address cannot be translated into more than one address from the other technology. This is known as injectivity.
+\end{itemize}
+\paragraph{}
+The gateway packet transformation process is an important component responsible for dealing with the heterogeneity problem.
+It hides most of the heterogeneity in the system and make communication go independently of the data link layer or network layer technologies used.
+
+\section{Sensor Data Storage Process}
+\paragraph{}
+The sensor data storage process is a component small component within the system whose sole purpose is to store such data.
+It seen by the \gls{wsn} as a TCP server, where sensor nodes get connected to it as soon as they are turned ON.
+The sensor node starts sending packet periodically (depending on the sampling rate specified) and whenever the sensor data is received, it is stored into a \gls{rdbms} where the schema is created as shown in figure \ref{fig:erd}.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.5]{img/smart_grid_db.jpg}
+\caption{Entity relationship diagram}
+\label{fig:erd}
+\end{figure}
+
+\section{HTTP Sensor Data Sender} 
+The HTTP sensor data sender is TCP server to which sensor nodes get connected to and start sending their data periodically.
+Once the data is received they are able to send data to the outside world periodically depending on the needs of the clients outside the system.
+This system counts on the deployment of an HTTP server at the level of outside the system client and send an HTTP request containing the sensor data received from the \gls{wsn}.
+\paragraph{}
+The sensor nodes are able to send this data by themselves directly to the outside as this is part of the two-way communication but for convenience, it was decided that the \gls{wsn} sends its data to the gateway and it is up to the gateway to send such data or store it in the \gls{rdbms}.
+\section{Mobile Client}
+The mobile client is a component of the system but can be seen as an outside world client as it is here to receive sensor data and also send ON/OFF commands to the system.
+
+\chapter{Deployment}
+\paragraph{}
+In this chapter, the test-bed will be discussed from the implementation point of view.
+That is said, each component within the system will be given as much details as possible in order for the reader to open the code and understand what is going.
+Chapter 4 discussed the components from the architectural aspect but no implementation and technology decisions were discussed and the goal of this chapter is to explain these details.
+\section{Technology Enablers}
+The system's implementation was built using a set of software and hardware technologies and standards that were used throughout the course of this project to make the test-bed achieve its objectives.
+%\glsreset{wsn}
+\subsection{Technologies used in \gls{wsn}}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.6]{img/crossbow.jpg}
+\caption{Crossbow Professional Kit}
+\label{fig:crossbow}
+\end{figure}
+
+The \gls{wsn} was built using Crossbow professional kit as shown in figure \ref{fig:crossbow} which counts 7 Micaz sensor nodes of type MPR2600 as shows in figure \ref{fig:micaz}.
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.8]{img/micaz.png}
+\caption{Crossbow MPR2600 Micaz Mote}
+\label{fig:micaz}
+\end{figure}
+
+In addition, each mote is supplied with a sensor board of type Crossbow MTS400CA which contains temperature, humidity, barometric pressure, ambient light and infrared light sensors that are used to send different types of sensor data to the gateway.
+This board is shown in figure \ref{fig:mts400}.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.5]{img/mts400ca.jpg}
+\caption{Crossbow MTS400CA Sensor Board}
+\label{fig:mts400}
+\end{figure}
+\paragraph{}
+The WSN is run using \gls{tinyos} which is one of the most used opensource \gls{os} in \glspl{wsn}.
+Programs are implemented using \gls{nesc} which is the sole programming language used in the \gls{tinyos} environment.
+This programming language is event-oriented and its syntax is very influenced by C.
+In addition, C libraries can be used in \gls{nesc} programs.
+
+\subsection{Technologies used in the Gateway}
+\paragraph{}
+The gateway's hardware machine is Dell Optiplex 9010 running a Linux environment more specifically the Ubuntu 13.04 distribution.
+The gateway has a set of technologies used to provide numerous services the system counts on.
+The programming languages used in the components hosted in this machine are C and \gls{j2se}.
+Netfilter is also used to control incoming and outgoing packets.
+Concerning the storage of sensor data MySQl is the \gls{rdbms} used.
+
+\subsection{Technologies used in the Mobile Client}
+\paragraph{}
+The mobile used is an Android smart phone with Android 2.3. The application developed is using \gls{j2se}, PHP for database connectivity.
+
+\section{Mote TCP Client And TCP Server}
+\paragraph{}
+Section 4.4 and 4.5 were dedicated to explain the TCP Client and TCP server separately.
+In this section, one section will explain both of them because from the implementation point of view,  both of them are implemented in the \gls{tinyos} program.
+\paragraph{}
+A typical \gls{tinyos} program has two parts: the components definition part and the module part.
+In the components definition part, the main components that will be used.
+These components refer to already implemented modules whose functionalities will be used in the in-development module.
+In the module part, we provide define the interfaces we will be using and the interfaces we will be providing if any \cite{ref14}.
+Back to definition part, after defining components we will be using, we then link the interfaces needed in the module to the components providing that interface.
+This is known as wiring.
+\paragraph{}
+The components needed in this program are the following:
+\begin{itemize}
+\item MainC: provides the essential functionalities such as the initialization and closing of the program.
+\item LedsC: provides access to the leds existing in the mote
+\item Taos2550C: is a light sensor component that enable to read data
+\item TimerMillic: A Timer that will be used for sampling
+\item TCPSocket as client: Is the TCP client
+\item TCPSocket as Server: Is the TCP server
+\end{itemize}
+In the modules part we define the components that will be used which are the following:
+\begin{itemize}
+\item Boot: This is the interface that provides main functions of booting and closing
+\item TCP as server: The main interface for TCP
+\item TCP as client: The main interface for TCP
+\item Leds: The interface to access the leds
+\item timer as sensingTimer: The interface to program and use the timer
+\item Read as Sensor: The interface to talk with the sensor
+\end{itemize}
+Some of the interface do not have all the functions provided and the not implemented function are implemented in the module. The following part will talk about each function that is to be implemented
+\paragraph{After boot of the mote}
+The first thing done after the mote is booted is to initialize the TCP client and server.
+The Client is given a random port by using the rand function.
+The TCP server is binded to port 7.
+The client is set to connect to the Data store in the gateway by providing the \gls{ipv6} address of the gateway and port number of the server.
+Afterwards, the timer is set to be triggered every 10 seconds which means that the sampling rate will be 360 sensor readings per hour.
+\paragraph{TCP Server data received}
+Once the data is received from a client, it assumes that there are two data items separated by a space.
+The first item refers to the led number(1,2,3) and the second refers to the command 0 meaning turn Off and 1 means turn On.
+The string is tokenized and the data items are extracted and the command is executed accordingly.
+\paragraph{Timer fired}
+Once the timer is fired, this means it is time to read sensor data from the sensor.
+The command to read from the sensor is called which goes and reads data and once the data is read the event "read done" is triggered.
+\paragraph{Read done}
+Once the data is read and ready, the next thing is to transform it and then send it.
+Concerning the transformation, a series of equations are applied as seen in section 4.4 which build the data value from the original value gotten and it is ready to be sent.
+A message is created containing two values: the mote NODE\_ID which is the identifier of the mote predefined in the installation of the program in the mote and the sensor data value.
+These two values are separated by a space, encapsulated within a \gls{tcp} datagram and sent.
+\paragraph{}
+During the installation of the software the \glsreset{blip}\gls{blip} package should specified as it is a dependency of the software system.
+\gls{blip} provides the TCP/IP protocol stack that implements \gls{6lowpan} standard by \gls{ietf}.
+In  the installation the NODE\_ID is specified.
+If NODE\_ID is 1, this will make the mote have an \gls{ipv6} of FEC0::1, as the NODE\_ID is appended to the predefined prefix of \gls{wsn} during the installation.
+\section{Mote IPv6 Backstation}
+\paragraph{}
+This component is provided as part of the \gls{blip} software package.
+It is \gls{tinyos} program whose goal is to forward packet back and forth between the \gls{wsn} and the outside.
+This mote is connected to the \gls{wsn} using Zigbee and to the outside world using \gls{usb}.
+This special mote has its \gls{ipv6} being FEC0::64.
+It receives \gls{ipv6} packets encapsulated in \gls{usb} frames, extracts them and encapsulates them in a Zigbee frame and vice versa.
+\paragraph{}
+During the installation of this \gls{tinyos} program into the mote sink, one should make sure that it is being installed in the mote sink as \gls{usb} is a necessity.
+The \gls{ipv6} address does not need to be provided as it is already predefined in the program.
+
+\section{IPv6 over USB Tunnel}
+\paragraph{}
+This is a program that is installed in the gateway server machine.
+As the mote sink is connected to the gateway using \gls{usb}, the gateway will recognize it as a network interface card using \gls{ipv6} as a network technology.
+That is said, this component is part of the \gls{blip} software package and creates the network interface card module and starts extracting incoming \gls{ipv6} packets from the \gls{usb} frames and encapsulates outgoing \gls{ipv6} into \gls{usb} frames.
+\paragraph{}
+The installation and initialization of this component will be given more details in appendix A.
+\section{Gateway Packet Transformation Process}
+\paragraph{}
+This component is in charge of hiding the heterogeneity of the system from the rest of the the components of the system and also the outside world.
+It does that by intercepting all the packet moving and transforms them according to the standards used by the recipient.
+\paragraph{}
+The program works with Netfilter in parallel. This means that Netfilter should drop all incoming and this program will take care of dealing with incoming packets separately. 
+\paragraph{}
+To make use of this process, there is an initial configuration that should be done to all the components of the system in order for the system to work.
+The gateway has two network interface cards: the Wi-Fi interface and the Zigbee interface provided through the tunnel.
+The first one is using \gls{ipv4} and the second is using \gls{ipv6}.
+\paragraph{}
+The program has three subcomponents as seen in section 4.8. The three components are the following:
+\begin{itemize}
+\item Packet sniffer
+\item Packet transformer
+\item Address translator
+\end{itemize}
+The following subsections will discuss each of these subcomponent separately from the implementation point of view.
+
+\subsection{Packet sniffer}
+\paragraph{}
+The packet sniffer is the one in charge of sniffing packet from both network interface card, as Raw Frames.
+The code snippet shows the functions called to sniff the packets.
+We make use of the raw socket library that enables us to sniff data link layer frames before being dropped by Netfilter.
+The main reason of dropping the packets by Netfilter is that the packets will not addressed to the computer and therefore the computer make send ICMP messages to the sender saying that the route is not the best and therefore the sender will never send packets again which is something we need to avoid to have.
+
+\begin{lstlisting}
+sockfd = socket( PF_PACKET, SOCK_PACKET , htons(ETH_P_ALL));
+if (sockfd < 0)
+	exit(1);
+
+while(1)
+{
+	data_size = recvfrom(sockfd,buf,10000,0
+		,(struct sockaddr *)&cliaddr,&clilen);
+}
+
+\end{lstlisting}
+\paragraph{}
+The first line of code is the initialization of the raw socket, we define the type, scope and name space of the packets to be sniffed.
+This will return a file descriptor that will allow us to call the socket.
+Afterwards, we start reading packets in an infinite loop.
+
+\paragraph{}
+After sniffing a packet, we must check if it is an \gls{ipv4} or \gls{ipv6} packet.
+The problem is that when reading from the Wi-Fi interface card, we get the whole frame where the network packet is encapsulated, but the Zigbee Tunnel only gives the raw \gls{ipv6} packet.
+Once a packet is read by the frame, from which interface card came is a problem and therefore a partial solution is found.
+We cast the packet to Wi-Fi frame and whenever we check for the network protocol, we get two solutions, 0 referring to \gls{ipv6} and 8 referring to \gls{ipv4}.
+If we get zero, the whole data should be casted to the \gls{ipv6} data structure whereas if it is 8, we should move the pointer forward to drop the Wi-Fi header and cast the following part to the \gls{ipv4} data structure.
+\paragraph{}
+In each of those conditions we only select the TCP datagrams and this is why we check if the next header in the IP structure is equal to 6.
+After that we move the pointer to drop the IP header and only keep the TCP datagram that will have to be sent in a the newly transformed packet.
+Depending on the network technology of the sniffed packet, we should build a new packet with the technology that is recognized by the recipient.
+
+\begin{lstlisting}
+if(header->h_proto == 0) 
+{
+	if(ip6_hdr->ip6_nxt == 6)
+	{
+			buf += IP6_HDRLEN;
+			... 			
+			sendIPV4Packet(..)
+	}
+}
+else if(header->h_proto == 8)
+{
+	if(ip_hdr->protocol == 6)
+	{
+		buf += (sizeof(struct ethhdr)+ IP4_HDRLEN);		
+		...
+		sendIPV6Packet(...)
+	}
+}
+\end{lstlisting}
+
+\subsection{Packet transformer}
+\paragraph{}
+The packet tranformer subcomponent is in charge of building two types of packets: \gls{ipv4} and \gls{ipv6} packets.
+As stated in section 4.8, all devices in this system have an \gls{ipv4} and \gls{ipv6} address where one of them is not recognized only at the gateway.
+As the \gls{wsn} excelusively supports \gls{ipv6} and they cannot be addressed without using their \gls{ipv6} address and any other device not using \gls{ipv6} may communicate with them by using their virtual \gls{ipv4} address at once the packets reach the gateway, it transform so that the data carried will be delivered and the packet will have instead of \gls{ipv4} source and destination addresses, \gls{ipv6} addresses that will be translated using the Address translator subcomponent.
+\paragraph{}
+This subcomponent is based on two important functions:
+\begin{itemize}
+\item sendIPv6Packet(..)
+\item sendIPv4Packet(..)
+\end{itemize}
+\paragraph{}
+The idea is the same for except that each technology has its own details such as filling different fields and computing the checksum.
+Another difference is when creating the \gls{ipv4} one should also create the Wi-Fi frame otherwise it won't be able to be sent through the corresponding network interface card whereas when building \gls{ipv6} packet one should just send the packet and the tunnel will take care of encapsulating it into a \gls{usb} frame.
+The function receives the data to be sent which is in all cases a \gls{tcp} datagram along with the size of the datagram, the source and destination addresses that are already translated.
+Afterwards, the transformator creates the packet appends the datagram, fills the source and destination address and create a socket attached to the corresponding interface card and send the packet.
+\begin{lstlisting}
+int sendIPV6Packet(char* buffer,int size,char *src ,char *dest)
+{
+	//CREATE PACKET WITH ALL ITS DETAILS
+	...
+	memcpy (frame + IP6_HDRLEN, buffer, size);
+	sd = socket (PF_PACKET, SOCK_RAW, htons (ETH_P_ALL))) < 0);
+	sendto (sd, rame, length...);
+}
+\end{lstlisting}
+
+\subsection{Address translator}
+\paragraph{}
+The address translator is the subcomponent that assigns \gls{ipv4} and \gls{ipv6} to all devices in this system.
+For instance, sensor nodes in the \gls{wsn} have \gls{ipv6} addresses but this subcomponent assigns to them \gls{ipv4} addresses.
+\paragraph{}
+To do so, the translation algorithm is purely bijective.
+That is said, any \gls{ipv4} is translated to one and only one \gls{ipv6} and vice versa.
+For a mobile client with \gls{ipv4} address of 10.50.1.50 to talk with a mote whose \gls{ipv6} address is FEC0::1, the mobile client should be aware of its translated to \gls{ipv4} address which is 192.168.0.1.
+Once the packet is received and knowing it is not addressed to it.
+The source and destination \gls{ipv4} addresses are translated to \gls{ipv6} making the source address equal to 0A32::0132 and the destination FEC0::1.
+The \gls{ipv6} packet is created and reaches the destination which is the sensor node.
+Once the response is going back the scenario is the same.
+\paragraph{}
+The translation algorithm has been tested extensively to make sure it is bijective and that any address can be translated to one and no more and than one address.
+When configuring the system, one should make sure that there is no two devices that have the same address real or virtual.
+Any \gls{ipv6} and \gls{ipv4} should be unique throughout the whole system.
+The \gls{wsn} should be granted an interval of \gls{ipv4} address so that no one will be able to have them besides the \gls{wsn} that uses them virtually.
 
 \chapter{Findings}
+
 \section{Experiment}
 %\subsection{Methodology}
-\section{Data gathering}
-\section{Interpretation of result}
+\section{Data Gathering}
+\section{Interpretation of Results}
 
@@
 \chapter{Future Work}
+\paragraph{}
+This thesis project's was to establish the infrastructure for integrating smart home within \glsreset{sg}\gls{sg} and make it part of the \glsreset{iot}\gls{iot}.
+This project enabled the deployment of the two-way communication as an infrastructure that can be used to build the \glsreset{dr}\gls{dr}.
+\gls{dr} is one of the most important topics in smart grids.
+It enables the utility and households to communicate with a set of services that any of them may request from the other.
+The current state of the art does not incorporate this communication as part of the smart home systems \cite{ref15}.
+But as there is a need for this communication, \gls{dr} will have to be implementation as part of the
+
+\backmatter
 
@@
 \chapter{How to install the system}
+\section{Hardware needed}
+In order for the user to install the system, one should make sure he/she is having a computer that will be used for the sole purpose of hosting different services.
+The crossbow professional kit is a necessity and a router must be available to connect the different components through Wi-Fi.
+A smartphone or a separate computer with Android emulator may de used.
 
-\backmatter
+\section{Software installation}
+
+\subsection{Operating system}
+The machine should run Linux Ubuntu distribution (11.10 or more recent). You way use Windows as long as you Install a Virtual machine using Cygwin or other tools
+\subsection{TinyOs Installation}
+\begin{enumerate}
+\item Execute: sudo gedit /etc/apt/source.list
+\item Add to the opened file: deb http://hinrg.cs.jhu.edu/tinyos hardy main
+\item sudo apt-get update 
+\item sudo apt-get install tinyos-2.1.1
+\item sudo gedit ~/.bashrc
+\item add the following line: \\
+export TOSDIR=\$TOSROOT/tos \\
+export CLASSPATH=\$TOSROOT/support/sdk/java/tinyos.jar:.\$CLASSPATH \\
+export MAKERULES=\$TOSROOT/support/make/Makerules \\
+export PATH=/opt/msp430/bin:\$PATH \\
+source /opt/tinyos-2.1.1/tinyos.sh \\
+
+\item sudo chmod 777 \$TOSROOT
+\end{enumerate}
+
+\subsection{Configuring \gls{blip}}
+\glsreset{blip}\gls{blip} should be compiled before being used. To do so, follow these instructions:
+\begin{enumerate}
+\item cd \$TOSROOT/support/sdk/c/sf \\
+./bootstrap \\
+./configure \\
+make 
+\item cd \$TOSROOT/support/sdk/c/blip \\
+./bootstrap.sh \\
+./configure \\
+make 
+\end{enumerate}
+
+\subsection{Install Software in Motes}
+The first thing to do is to install the backstation within the Mote sink (the mote with the \gls{usb} IO) To install it we do the following:
+\begin{enumerate}
+\item Plug the mote and type the command dmesg to see what \gls{usb} ID was assigned to it. It will generally be USB0.
+\item cd \$TOSROOT/apps/IPBaseStation
+\item make micaz blip install  mib520, /dev/ttyUSB0
+\end{enumerate}
+The backstation is now installed. We will now install the Mote TCP Client and Server in each sensor node. To do it, we should first download the code from the SVN repository.
+\begin{enumerate}
+\item Make sure you have SVN by typing svn --version
+\item If you do not have SVN type sudo apt-get install subversion
+\item download the file:
+\item cd
+\item svn checkout http://grid-aui-tinyos.googlecode.com/svn/trunk/ grid-aui-tinyos
+\item cd grid-aui-tinyos//programming/TCPEcho
+\end{enumerate}
+Now take the MIB520 programming card, plug it and type dmesg to see where it plugged. lets call the location *USB. Now plug each mote in the MIB520 card and for each plugged mote do the following:
+
+\begin{enumerate}
+\item  make micaz blip install.*NUM mib520,/dev/tty*USB
+\item make clean
+\end{enumerate}
+Note that *NUM refers to the NODE\_ID. For the first mote choose 1, for the second choose 2...
+
+\subsection{Install Packet Transformation process}
+To install this component, you must do the following:
+\begin{enumerate}
+\item Make sure you have CMake, type sudo apt-get install cmake
+\item cd
+\item grid-aui-tinyos//programming/netfiltering
+\item cmake CMakeLists.txt
+\item make
+\end{enumerate}
+
+\subsection{Install Gateway Sensor Data TCP Server}
+To install the server you must do the following:
+\begin{enumerate}
+\item cd
+\item grid-aui-tinyos//programming/GatewayServer/src
+\item make
+\end{enumerate}
+
+\chapter{Run the System}
+There is a set of instructions to follow in order as some of the components must be alive only when others are already alive. To run the system, do the following:
+\begin{enumerate}
+\item To turn on the Tunnel do the following
+\subitem cd \$TOSROOT/support/sdk/c/blip/ 
+\subitem sudo driver/ip-driver /dev/ttyUSB0 micaz. \\
+ You must check that the mote sink is attached to USB0 by using dmesg, if not change it accordingly. After that, you should have an output that looks similar to the figure \ref{fig:tunnel_output}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.9]{img/tunnel_output.jpg}
+\caption{IPv6 over USB tunnel execution output}
+\label{fig:tunnel_output}
+\end{figure}
+
+\item To Run the Gateway Sensor Data TCP server do the following:
+\subitem cd
+\subitem cd grid-aui-tinyos/programming/GatewayServer/src
+\subitem java Server
+
+\item To Enable IP forwarding do the following:
+\subitem sudo bash
+\subitem echo 1 \textgreater /proc/sys/net/ipv4/ip\_forward
+\subitem exit
+
+\item  To drop all incoming packets: sudo iptables -P INPUT DROP
+\item To disable ICPMP REDIRECT TO HOST:
+\subitem sudo bash
+\subitem /sbin/sysctl -w net.ipv4.conf.all.accept\_redirects=0
+\subitem /sbin/sysctl -w net.ipv4.conf.all.send\_redirects=0
+\subitem /sbin/sysctl -w net.ipv6.conf.all.accept\_redirects=0
+\subitem /sbin/sysctl -w net.ipv6.conf.all.send\_redirects=0
+\subitem exit
+
+\item Run the Packet transformation process
+\subitem cd
+\subitem cd grid-aui-tinyos/programming/netfiltering
+\subitem sudo my\_sniffer
+
+\item Turn on the wireless sensor nodes one by one
+\end{enumerate}
+
+\paragraph{}
+You should be able to get sensor data periodically. The mobile client access for turning On and Off will be explained in the Appendix C.
 
 \singlespacing
-\printglossaries
 \bibliographystyle{IEEEtran} 

### 2013-07-03 programming/netfiltering/src/my_sniffer.c
Only chapter 3 missing

@@ -1,6 +1,7 @@
 #include <stdio.h>
 #include <stdlib.h>
 #include <string.h>
+#include <time.h>
 #include "my_sniffer.h"
 
 int main()
@@ -10,7 +11,8 @@ int main()
 	char *addr;
 	char *src_addr;
 	char *dest_addr;
-
+	struct timespec start, stop;
+	double accum;
 	socklen_t clilen;
 	struct sockaddr_in cliaddr;
 	struct sockaddr_in *in;
@@ -30,8 +32,8 @@ int main()
 	clilen = sizeof(struct sockaddr_in);
 	while(1)
 	{
-		header = (struct ethhdr* ) buf;
 
+		header = (struct ethhdr* ) buf;
 		data_size = recvfrom(sockfd,buf,10000,0,(struct sockaddr *)&cliaddr,&clilen);
 
 		if(header->h_proto == 0) // MOST PROBABLY IPV6 WE SHOULD MAKE SURE IT IS
@@ -109,7 +111,6 @@ int main()
 
 			}
 		}
-
 	}
 }
 int sendIPV6Packet(char* buffer,int size,char *src ,char *dest)

### 2013-07-03 trunk/doc/thesis/report.tex
Only chapter 3 missing

@@
 \documentclass[oneside,12pt,a4paper,final]{book}
+%\usepackage[margin=3cm]{geometry}
 \usepackage[utf8]{inputenc}
@@
 \paragraph{}
+The program works with Netfilter in parallel. This means that Netfilter should drop all incoming and this program will take care of dealing with incoming packets separately. 
+\paragraph{}
+To make use of this process, there is an initial configuration that should be done to all the components of the system in order for the system to work.
+The gateway has two network interface cards: the Wi-Fi interface and the Zigbee interface provided through the tunnel.
+The first one is using \gls{ipv4} and the second is using \gls{ipv6}.
+\paragraph{}
+The program has three subcomponents as seen in section 4.8. The three components are the following:
+\begin{itemize}
+\item Packet sniffer
+\item Packet transformer
+\item Address translator
+\end{itemize}
+The following subsections will discuss each of these subcomponent separately from the implementation point of view.
+
+\subsection{Packet sniffer}
+\paragraph{}
+The packet sniffer is the one in charge of sniffing packet from both network interface card, as Raw Frames.
+The code snippet shows the functions called to sniff the packets.
+We make use of the raw socket library that enables us to sniff data link layer frames before being dropped by Netfilter.
+The main reason of dropping the packets by Netfilter is that the packets will not addressed to the computer and therefore the computer make send ICMP messages to the sender saying that the route is not the best and therefore the sender will never send packets again which is something we need to avoid to have.
+
+\begin{lstlisting}
+sockfd = socket( PF_PACKET, SOCK_PACKET , htons(ETH_P_ALL));
+if (sockfd < 0)
+	exit(1);
+
+while(1)
+{
+	data_size = recvfrom(sockfd,buf,10000,0
+		,(struct sockaddr *)&cliaddr,&clilen);
+}
+
+\end{lstlisting}
+\paragraph{}
+The first line of code is the initialization of the raw socket, we define the type, scope and name space of the packets to be sniffed.
+This will return a file descriptor that will allow us to call the socket.
+Afterwards, we start reading packets in an infinite loop.
+
+\paragraph{}
+After sniffing a packet, we must check if it is an \gls{ipv4} or \gls{ipv6} packet.
+The problem is that when reading from the Wi-Fi interface card, we get the whole frame where the network packet is encapsulated, but the Zigbee Tunnel only gives the raw \gls{ipv6} packet.
+Once a packet is read by the frame, from which interface card came is a problem and therefore a partial solution is found.
+We cast the packet to Wi-Fi frame and whenever we check for the network protocol, we get two solutions, 0 referring to \gls{ipv6} and 8 referring to \gls{ipv4}.
+If we get zero, the whole data should be casted to the \gls{ipv6} data structure whereas if it is 8, we should move the pointer forward to drop the Wi-Fi header and cast the following part to the \gls{ipv4} data structure.
+\paragraph{}
+In each of those conditions we only select the TCP datagrams and this is why we check if the next header in the IP structure is equal to 6.
+After that we move the pointer to drop the IP header and only keep the TCP datagram that will have to be sent in a the newly transformed packet.
+Depending on the network technology of the sniffed packet, we should build a new packet with the technology that is recognized by the recipient.
+
+\begin{lstlisting}
+if(header->h_proto == 0) 
+{
+	if(ip6_hdr->ip6_nxt == 6)
+	{
+			buf += IP6_HDRLEN;
+			... 			
+			sendIPV4Packet(..)
+	}
+}
+else if(header->h_proto == 8)
+{
+	if(ip_hdr->protocol == 6)
+	{
+		buf += (sizeof(struct ethhdr)+ IP4_HDRLEN);		
+		...
+		sendIPV6Packet(...)
+	}
+}
+\end{lstlisting}
+
+\subsection{Packet transformer}
+\paragraph{}
+The packet tranformer subcomponent is in charge of building two types of packets: \gls{ipv4} and \gls{ipv6} packets.
+As stated in section 4.8, all devices in this system have an \gls{ipv4} and \gls{ipv6} address where one of them is not recognized only at the gateway.
+As the \gls{wsn} excelusively supports \gls{ipv6} and they cannot be addressed without using their \gls{ipv6} address and any other device not using \gls{ipv6} may communicate with them by using their virtual \gls{ipv4} address at once the packets reach the gateway, it transform so that the data carried will be delivered and the packet will have instead of \gls{ipv4} source and destination addresses, \gls{ipv6} addresses that will be translated using the Address translator subcomponent.
+\paragraph{}
+This subcomponent is based on two important functions:
+\begin{itemize}
+\item sendIPv6Packet(..)
+\item sendIPv4Packet(..)
+\end{itemize}
+\paragraph{}
+The idea is the same for except that each technology has its own details such as filling different fields and computing the checksum.
+Another difference is when creating the \gls{ipv4} one should also create the Wi-Fi frame otherwise it won't be able to be sent through the corresponding network interface card whereas when building \gls{ipv6} packet one should just send the packet and the tunnel will take care of encapsulating it into a \gls{usb} frame.
+The function receives the data to be sent which is in all cases a \gls{tcp} datagram along with the size of the datagram, the source and destination addresses that are already translated.
+Afterwards, the transformator creates the packet appends the datagram, fills the source and destination address and create a socket attached to the corresponding interface card and send the packet.
+\begin{lstlisting}
+int sendIPV6Packet(char* buffer,int size,char *src ,char *dest)
+{
+	//CREATE PACKET WITH ALL ITS DETAILS
+	...
+	memcpy (frame + IP6_HDRLEN, buffer, size);
+	sd = socket (PF_PACKET, SOCK_RAW, htons (ETH_P_ALL))) < 0);
+	sendto (sd, rame, length...);
+}
+\end{lstlisting}
+
+\subsection{Address translator}
+\paragraph{}
+The address translator is the subcomponent that assigns \gls{ipv4} and \gls{ipv6} to all devices in this system.
+For instance, sensor nodes in the \gls{wsn} have \gls{ipv6} addresses but this subcomponent assigns to them \gls{ipv4} addresses.
+\paragraph{}
+To do so, the translation algorithm is purely bijective.
+That is said, any \gls{ipv4} is translated to one and only one \gls{ipv6} and vice versa.
+For a mobile client with \gls{ipv4} address of 10.50.1.50 to talk with a mote whose \gls{ipv6} address is FEC0::1, the mobile client should be aware of its translated to \gls{ipv4} address which is 192.168.0.1.
+Once the packet is received and knowing it is not addressed to it.
+The source and destination \gls{ipv4} addresses are translated to \gls{ipv6} making the source address equal to 0A32::0132 and the destination FEC0::1.
+The \gls{ipv6} packet is created and reaches the destination which is the sensor node.
+Once the response is going back the scenario is the same.
+\paragraph{}
+The translation algorithm has been tested extensively to make sure it is bijective and that any address can be translated to one and no more and than one address.
+When configuring the system, one should make sure that there is no two devices that have the same address real or virtual.
+Any \gls{ipv6} and \gls{ipv4} should be unique throughout the whole system.
+The \gls{wsn} should be granted an interval of \gls{ipv4} address so that no one will be able to have them besides the \gls{wsn} that uses them virtually.
 
 \chapter{Findings}
+\paragraph{}
+The system was built to provide monitor and control home appliances and also provide an infrastructure for future platforms to be integrated in this system.
+Such infrastructures can be the integration of the \glsreset{dr}\gls{dr}.
+To build overlying system's, one should make sure that the system has a certain level of performance that one can count on to build overlying services.
+The purpose of this chapter is to measure the performance of the system.
+It is done by first doing and experiment which will help us gather data and then based on this data, an interpretation of this data will be done to see whether the system is shows an acceptable level of performance.
 \section{Experiment}
-%\subsection{Methodology}
+\paragraph{}
+To measure the performance of the system, an experiment was conducted where the goal is to quantify the performance of the system.
+\gls{6lowpan} was built in order for the mote receiving the packet to forward it is not directed to it before dealing with the one directed to it.
+This means that there is more priority given to routed packet that the processed ones.
+As a result, this clearly pushes us to think that the delay of the system will not be much affected as the traffic in the network grows.
+\paragraph{Experiment 1:}
+We will first have one mote in the \gls{wsn} that will be interacting with the outside world.
+We will start sending ping command periodically (once every 1 second) and record the delay and the jitter.
+Once every two minutes, we will add another mote to the system and continue pinging the same mote.
+We keep adding motes every 2 minutes until we fill the \gls{wsn} with 6 motes which is the maximum number of motes available in the system.
+This will enable us to collect data about the mote delay and jitter as we add more traffic to the system.
+\paragraph{Experiment 2: }
+The second experience is meant to measure the time it takes to transform the packet.
+In other words, measure the average delay and jitter that is added when having the Gateway Transformation Process as part of the system.
+The timer starts before the sniffing of the packet and ends right after the packet is transformed and sent.
+We believe that there will be an additional delay to the communication but we are hoping that it will not affect much the whole delay of the communication between different components of the system. 
 \section{Data Gathering}
+\paragraph{}
+After both experiments were conducted, we collected a significant amount of data that was used to measure the performance and an interpretation was done in section 6.3 to have an idea about the performance of the system.
+\paragraph{Experiment 1}
+In the first experiment, the same mote was pinged once every second and we added new motes to the system every two minutes.
+The experiment as a whole took 12 minutes and 720 pings we done.
+Table \ref{table:exp1} shows the result of the experiment.
+
+\begin{table}[htbp]
+    \begin{tabular}{lll}
+    \hline
+    number of motes & Delay (ms) & Jitter (ms) \\\hline
+    1               & 87.5         & 1.2         \\ 
+    2               & 88.1         & 2.0         \\
+    3               & 88.3        & 2.7         \\
+    4               & 88.8         & 3.1         \\
+    5               & 90.0         & 3.3         \\
+    6               & 89.5         & 3.5         \\
+    \end{tabular}
+    \caption{Experiment 1 delay and jitter}
+    \label{table:exp1}
+\end{table}
+For each row in the table, the average delay is computed using the pings that were performed in the 2 minutes interval which means 120 pings per row.
+The jitter is also computed in a similar as the average variation from the mean delay in that 2 minutes interval.
+The result is displayed in the table \ref{table:exp1}.
+
+\paragraph{Experiment 2}
+The second experiment's goal is to measure the additional delay coming from the Gateway Packet transformation process whose goal to handle and hide the heterogeneity of the system.
+The result gathered reflects the average delay and jitter computed over elapsed time from the sniffing of the packet to the transformation and sending to the recipient.
+Table \ref{table:exp2} shows the result that was collected from this experiment as it was done on 200 packets that were sniffed and transformed by the system.
+
+\begin{table}[htbp]
+    \begin{tabular}{ll}
+    \hline
+    Delay (ms) & Jitter (ms) \\ \hline
+    0.1        & 0.0         \\ 
+    \end{tabular}
+    \caption{Experiment 2 delay and jitter}
+        \label{table:exp2}
+\end{table}
+The transformation process's elapsed time is counted in nanoseconds and from \ref{table:exp2}, we can see that the average delay is on average 0.1 milliseconds whereas the jitter is 0 milliseconds meaning that the process's time varies between a few nanoseconds to at most 150 nano seconds.
+Depending on the machine's load.
 \section{Interpretation of Results}
-
+\paragraph{}
+The data collected in section 6.2 shows that the performance of the system is very good. Even though the delay in the \gls{wsn} is high, 
+ nothing can be done about that as even though the traffic in the \gls{wsn} is low, the delay is still high and this has to do with the computational time taken by the TCP/IP protocol protocol stack provided by the
 \chapter{Conclusion}
-
+\paragraph{}
+This project enabled us to establish an infrastructure as part of the \glsreset{iot}\gls{iot} and move towards having the smart home system's integrated within the smart.
+At this stage and counting solely on this work, one cannot clearly see the integration of smart home systems within the \gls{iot}, but this is an infrastructure that provides two main services which are the establishment of the two-way communication between the smart home devices and the outside world which may be the utility for instance.
+This two-way communication enables the \glsreset{wsn}\gls{wsn} to monitor home appliances and send sensor data to a repository which monitors the home's consumption and this device can be used for simulating a real world \gls{ami}.
+The other part of the communication is the ability of clients to control appliances status by turning them On and Off.
+\paragraph{}
+The second contribution of this work the smart management of the heterogeneity.
+The communication from the mote to the mobile client is built on a set of technologies and the journey of a simple \gls{tcp} datagram move through many different data link layer standards and different network technologies.
+Starting from the mote sending sensor data, The TCP data is encapsulated into an \gls{ipv6} packet that is encapsulated into a Zigbee frame.
+Afterward, the packet is extracted from the Zigbee frame and encapsulated into a \gls{usb} frame.
+Once it arrives to the gateway, the \gls{tcp} datagram is extracted from the \gls{ipv6} which is extracted from the \gls{usb} frame and encapsulated in an \gls{ipv4} packet that encapsulated into a Wi-Fi frame and sent until it reaches the mobile.
+That is said, it clear that the system is heterogeneous and making this heterogeneity transparent was a contribution of this work.
+\paragraph{}
+The \glsreset{wsn}\gls{wsn} was built making use of the very recent \glsreset{6lowpan}\gls{6lowpan} whose goal is to provide the TCP/IP protocol stack to the smallest devices.
+The \gls{wsn} is autonomous in the sense that it can work independently and indefinitely by monitoring the home appliances.
+\paragraph{}
+All in all, this project as a big picture is a testbed providing the main components of a smart home management with an intelligent \gls{wsn}, a robust and powerful middleware application that is able to use the sensory data and use it to control the home's appliances and a mobile application able to provide the users remote monitoring of their smart homes and control.
+This was possible by making the system as part of the \glsreset{iot}\gls{iot}.
 \chapter{Future Work}
-
-\appendix
-\chapter{How to install the system}
+\paragraph{}
+This thesis project's was to establish the infrastructure for integrating smart home within \glsreset{sg}\gls{sg} and make it part of the \glsreset{iot}\gls{iot}.
+This project enabled the deployment of the two-way communication as an infrastructure that can be used to build the \glsreset{dr}\gls{dr}.
+\gls{dr} is one of the most important topics in smart grids.
+It enables the utility and households to communicate with a set of services that any of them may request from the other.
+The current state of the art does not incorporate this communication as part of the smart home systems \cite{ref15}.
+But as there is a need for this communication, \gls{dr} will have to be implemented as part of the smart home package.
+\paragraph{}
+Demand reponse integration is a future work for this study.
+\gls{dr} has a set of service that have to be implemented in the smart home packages.
+In \gls{dr}, the consumer has the possibility to ask the utility about the instantaneous price quote for electricity.
+The consumer when in need of high priority in energy may ask the utility to provide such to provide a minimum of energy that will be guaranteed.
+Of course this means at a cost, but this agreement is part of the \gls{dr} and should be implemented at the level of smart home systems.
+Consumers can also help utilities at difficult times.
+For instance, when the utility cannot provide enough energy at peak times, the utility may ask some consumers to significantly cut their consumption and as a result be billed a reduced cost for the electricity that is consumed.
+In general, the smart home aims to reduce energy consumption and as the utility at times may need to reduce its energy production especially in peak hours, the \gls{dr} will help the utility ask the smart home to reduce the home consumption.
+\paragraph{}
+As this project aims to integrate the smart home devices within the \glsreset{iot}\gls{iot}, there is a need to integrate \glsreset{rfid}\gls{rfid} within this system as a future work.
+The \gls{rfid} is a building block within the \gls{iot} and will equip the devices with a low-power radio identification that will allow more flexibility to the system, as the external systems may be able to recognize uniquely each device in the \gls{iot} without counting primarily on the \gls{ipv6} which may change from time to time and therefore cannot keep track of the each device's information and monitor it reliably.
+\paragraph{}
+This system provides near real-time sensory data that is not much exploited intelligently.
+The only use of this data is to enable users to monitor their home consumption.
+As far as the system is concerned, there is a need for a more intelligent exploitation of this very rich data that can be used by intelligent algorithm to automatically and autonomously reduce the household's energy consumption.
+Such algorithms can be based on machine learning and data mining techniques making the direction of this a cross disciplinary project aiming to reduce the consumption of household without reducing the utility of users.
+In other words, build algorithms that will reduce consumption of households while the user is not noticing change.
+\paragraph{}
+Before building new services on top of this system, one should test different aspects of the system.
+In this study, the only thing that was tested, measured and quantified was the performance of the system which gave good results about the system's overall performance.
+As a future work, one should test different aspects of the system such as the reliability and availability of the system.
+More experiments should be conducted to see how well the system acts in high traffic networks and how good could it support new overlying systems such as \gls{dr}.
+\paragraph{}
+The system provides different services, but there is not authentication or authorization involved.
+This means that any users knowing the mote connected to the appliance he/she is interested in, he/she can control it and there is no existing party or component that will control such a communication.
+As a future work, the control of the system's access should be monitored in order for the system to be more secure and robust to external attacks.
 
 \backmatter
-
 \singlespacing
+\addcontentsline{toc}{section}{References}
 \bibliographystyle{IEEEtran} 
@@
 
+\doublespacing
+\appendix
+\chapter{How to install the system}
+\section{Hardware needed}
+In order for the user to install the system, one should make sure he/she is having a computer that will be used for the sole purpose of hosting different services.
+The crossbow professional kit is a necessity and a router must be available to connect the different components through Wi-Fi.
+A smartphone or a separate computer with Android emulator may de used.
+
+\section{Software installation}
+
+\subsection{Operating system}
+The machine should run Linux Ubuntu distribution (11.10 or more recent). You way use Windows as long as you Install a Virtual machine using Cygwin or other tools
+\subsection{TinyOs Installation}
+\begin{enumerate}
+\item Execute: sudo gedit /etc/apt/source.list
+\item Add to the opened file: deb http://hinrg.cs.jhu.edu/tinyos hardy main
+\item sudo apt-get update 
+\item sudo apt-get install tinyos-2.1.1
+\item sudo gedit ~/.bashrc
+\item add the following line: \\
+export TOSDIR=\$TOSROOT/tos \\
+export CLASSPATH=\$TOSROOT/support/sdk/java/tinyos.jar:.\$CLASSPATH \\
+export MAKERULES=\$TOSROOT/support/make/Makerules \\
+export PATH=/opt/msp430/bin:\$PATH \\
+source /opt/tinyos-2.1.1/tinyos.sh \\
+
+\item sudo chmod 777 \$TOSROOT
+\end{enumerate}
+
+\subsection{Configuring \gls{blip}}
+\glsreset{blip}\gls{blip} should be compiled before being used. To do so, follow these instructions:
+\begin{enumerate}
+\item cd \$TOSROOT/support/sdk/c/sf \\
+./bootstrap \\
+./configure \\
+make 
+\item cd \$TOSROOT/support/sdk/c/blip \\
+./bootstrap.sh \\
+./configure \\
+make 
+\end{enumerate}
+
+\subsection{Install Software in Motes}
+The first thing to do is to install the backstation within the Mote sink (the mote with the \gls{usb} IO) To install it we do the following:
+\begin{enumerate}
+\item Plug the mote and type the command dmesg to see what \gls{usb} ID was assigned to it. It will generally be USB0.
+\item cd \$TOSROOT/apps/IPBaseStation
+\item make micaz blip install  mib520, /dev/ttyUSB0
+\end{enumerate}
+The backstation is now installed. We will now install the Mote TCP Client and Server in each sensor node. To do it, we should first download the code from the SVN repository.
+\begin{enumerate}
+\item Make sure you have SVN by typing svn --version
+\item If you do not have SVN type sudo apt-get install subversion
+\item download the file:
+\item cd
+\item svn checkout http://grid-aui-tinyos.googlecode.com/svn/trunk/ grid-aui-tinyos
+\item cd grid-aui-tinyos//programming/TCPEcho
+\end{enumerate}
+Now take the MIB520 programming card, plug it and type dmesg to see where it plugged. lets call the location *USB. Now plug each mote in the MIB520 card and for each plugged mote do the following:
+
+\begin{enumerate}
+\item  make micaz blip install.*NUM mib520,/dev/tty*USB
+\item make clean
+\end{enumerate}
+Note that *NUM refers to the NODE\_ID. For the first mote choose 1, for the second choose 2...
+
+\subsection{Install Packet Transformation process}
+To install this component, you must do the following:
+\begin{enumerate}
+\item Make sure you have CMake, type sudo apt-get install cmake
+\item cd
+\item grid-aui-tinyos//programming/netfiltering
+\item cmake CMakeLists.txt
+\item make
+\end{enumerate}
+
+\subsection{Install Gateway Sensor Data TCP Server}
+To install the server you must do the following:
+\begin{enumerate}
+\item cd
+\item grid-aui-tinyos//programming/GatewayServer/src
+\item make
+\end{enumerate}
+
+\chapter{Run the System}
+There is a set of instructions to follow in order as some of the components must be alive only when others are already alive. To run the system, do the following:
+\begin{enumerate}
+\item To turn on the Tunnel do the following
+\subitem cd \$TOSROOT/support/sdk/c/blip/ 
+\subitem sudo driver/ip-driver /dev/ttyUSB0 micaz. \\
+ You must check that the mote sink is attached to USB0 by using dmesg, if not change it accordingly. After that, you should have an output that looks similar to the figure \ref{fig:tunnel_output}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.9]{img/tunnel_output.jpg}
+\caption{IPv6 over USB tunnel execution output}
+\label{fig:tunnel_output}
+\end{figure}
+
+\item To Run the Gateway Sensor Data TCP server do the following:
+\subitem cd
+\subitem cd grid-aui-tinyos/programming/GatewayServer/src
+\subitem java Server
+
+\item To Enable IP forwarding do the following:
+\subitem sudo bash
+\subitem echo 1 \textgreater /proc/sys/net/ipv4/ip\_forward
+\subitem exit
+
+\item  To drop all incoming packets: sudo iptables -P INPUT DROP
+\item To disable ICPMP REDIRECT TO HOST:
+\subitem sudo bash
+\subitem /sbin/sysctl -w net.ipv4.conf.all.accept\_redirects=0
+\subitem /sbin/sysctl -w net.ipv4.conf.all.send\_redirects=0
+\subitem /sbin/sysctl -w net.ipv6.conf.all.accept\_redirects=0
+\subitem /sbin/sysctl -w net.ipv6.conf.all.send\_redirects=0
+\subitem exit
+
+\item Run the Packet transformation process
+\subitem cd
+\subitem cd grid-aui-tinyos/programming/netfiltering
+\subitem sudo my\_sniffer
+
+\item Turn on the wireless sensor nodes one by one
+\end{enumerate}
+
+\paragraph{}
+You should be able to get sensor data periodically. The mobile client access for turning On and Off will be explained in the Appendix C.
+
 \end{document}

### 2013-07-03 doc/thesis/report.tex

@@
 \chapter{Findings}
-
+\paragraph{}
+The system was built to provide monitor and control home appliances and also provide an infrastructure for future platforms to be integrated in this system.
+Such infrastructures can be the integration of the \glsreset{dr}\gls{dr}.
+To build overlying system's, one should make sure that the system has a certain level of performance that one can count on to build overlying services.
+The purpose of this chapter is to measure the performance of the system.
+It is done by first doing and experiment which will help us gather data and then based on this data, an interpretation of this data will be done to see whether the system is shows an acceptable level of performance.
 \section{Experiment}
-%\subsection{Methodology}
+\paragraph{}
+To measure the performance of the system, an experiment was conducted where the goal is to quantify the performance of the system.
+\gls{6lowpan} was built in order for the mote receiving the packet to forward it is not directed to it before dealing with the one directed to it.
+This means that there is more priority given to routed packet that the processed ones.
+As a result, this clearly pushes us to think that the delay of the system will not be much affected as the traffic in the network grows.
+\paragraph{Experiment 1:}
+We will first have one mote in the \gls{wsn} that will be interacting with the outside world.
+We will start sending ping command periodically (once every 1 second) and record the delay and the jitter.
+Once every two minutes, we will add another mote to the system and continue pinging the same mote.
+We keep adding motes every 2 minutes until we fill the \gls{wsn} with 6 motes which is the maximum number of motes available in the system.
+This will enable us to collect data about the mote delay and jitter as we add more traffic to the system.
+\paragraph{Experiment 2: }
+The second experience is meant to measure the time it takes to transform the packet.
+In other words, measure the average delay and jitter that is added when having the Gateway Transformation Process as part of the system.
+The timer starts before the sniffing of the packet and ends right after the packet is transformed and sent.
+We believe that there will be an additional delay to the communication but we are hoping that it will not affect much the whole delay of the communication between different components of the system. 
 \section{Data Gathering}
+\paragraph{}
+After both experiments were conducted, we collected a significant amount of data that was used to measure the performance and an interpretation was done in section 6.3 to have an idea about the performance of the system.
+\subsection{Experiment 1}
+\paragraph{}
+In the first experiment, the same mote was pinged once every second and we added new motes to the system every two minutes.
+The experiment as a whole took 12 minutes and 720 pings we done.
+Table \ref{table:exp1} shows the result of the experiment.
+
+\begin{table}[htbp]
+    \begin{tabular}{lll}
+    \hline
+    number of motes & Delay (ms) & Jitter (ms) \\\hline
+    1               & 87.5         & 1.2         \\ 
+    2               & 88.1         & 2.0         \\
+    3               & 88.3        & 2.7         \\
+    4               & 88.8         & 3.1         \\
+    5               & 90.0         & 3.3         \\
+    6               & 89.5         & 3.5         \\
+    \end{tabular}
+    \caption{Experiment 1 delay and jitter}
+    \label{table:exp1}
+\end{table}
+\paragraph{}
+For each row in the table, the average delay is computed using the pings that were performed in the 2 minutes interval which means 120 pings per row.
+The jitter is also computed in a similar as the average variation from the mean delay in that 2 minutes interval.
+The result is displayed in the table \ref{table:exp1}.
+
+\subsection{Experiment 2}
+\paragraph{}
+The second experiment's goal is to measure the additional delay coming from the Gateway Packet transformation process whose goal to handle and hide the heterogeneity of the system.
+The result gathered reflects the average delay and jitter computed over elapsed time from the sniffing of the packet to the transformation and sending to the recipient.
+Table \ref{table:exp2} shows the result that was collected from this experiment as it was done on 200 packets that were sniffed and transformed by the system.
+
+\begin{table}[htbp]
+    \begin{tabular}{ll}
+    \hline
+    Delay (ms) & Jitter (ms) \\ \hline
+    0.1        & 0.0         \\ 
+    \end{tabular}
+    \caption{Experiment 2 delay and jitter}
+        \label{table:exp2}
+\end{table}
+\paragraph{}
+The transformation process's elapsed time is counted in nanoseconds and from \ref{table:exp2}, we can see that the average delay is on average 0.1 milliseconds whereas the jitter is 0 milliseconds meaning that the process's time varies between a few nanoseconds to at most 150 nanoseconds.
+Depending on the machine's load.
 \section{Interpretation of Results}
-
+\subsection{Experiment 1}
+\paragraph{}
+The data collected in section 6.2 shows that the performance of the system is very good. Even though the delay in the \gls{wsn} is high, 
+ nothing can be done about that as even though the traffic in the \gls{wsn} is low, the delay is still high and this has to do with the computational time taken by the TCP/IP protocol protocol stack provided implemented by \glsreset{blip}\gls{blip} using \glsreset{6lowpan}\gls{6lowpan}.
+To reduce the delay, one should probably rethink the design of the \gls{blip} or find a way to omit it, but that is not a good idea since we want the \gls{wsn} to be part of \gls{iot}.
+ \paragraph{}
+ The good news in experiment 1 is that the privatization of packets brings benefit to the network.
+Even though the traffic grows exponentially, the delay does not grow much and the even if the jitter follows to some extent the traffic's trend but still the average delay shows good result.
+ \subsection{Experiment 2}
+ \paragraph{}
+ The Packet transformation process is a component that shows excellent results.
+Before conducting experiment 2, we were persuaded that there will be additional delay originating from this process and it could be responsible for the high delay seen in experiment 2 but the results shows otherwise and that the component is not to be put responsible for such important delay seen in experiment 1.
 \chapter{Conclusion}
-
+\paragraph{}
+This project enabled us to establish an infrastructure as part of the \glsreset{iot}\gls{iot} and move towards having the smart home system's integrated within the smart.
+At this stage and counting solely on this work, one cannot clearly see the integration of smart home systems within the \gls{iot}, but this is an infrastructure that provides two main services which are the establishment of the two-way communication between the smart home devices and the outside world which may be the utility for instance.
+This two-way communication enables the \glsreset{wsn}\gls{wsn} to monitor home appliances and send sensor data to a repository which monitors the home's consumption and this device can be used for simulating a real world \gls{ami}.
+The other part of the communication is the ability of clients to control appliances status by turning them On and Off.
+\paragraph{}
+The second contribution of this work the smart management of the heterogeneity.
+The communication from the mote to the mobile client is built on a set of technologies and the journey of a simple \gls{tcp} datagram move through many different data link layer standards and different network technologies.
+Starting from the mote sending sensor data, The TCP data is encapsulated into an \gls{ipv6} packet that is encapsulated into a Zigbee frame.
+Afterward, the packet is extracted from the Zigbee frame and encapsulated into a \gls{usb} frame.
+Once it arrives to the gateway, the \gls{tcp} datagram is extracted from the \gls{ipv6} which is extracted from the \gls{usb} frame and encapsulated in an \gls{ipv4} packet that encapsulated into a Wi-Fi frame and sent until it reaches the mobile.
+That is said, it clear that the system is heterogeneous and making this heterogeneity transparent was a contribution of this work.
+\paragraph{}
+The \glsreset{wsn}\gls{wsn} was built making use of the very recent \glsreset{6lowpan}\gls{6lowpan} whose goal is to provide the TCP/IP protocol stack to the smallest devices.
+The \gls{wsn} is autonomous in the sense that it can work independently and indefinitely by monitoring the home appliances.
+\paragraph{}
+All in all, this project as a big picture is a testbed providing the main components of a smart home management with an intelligent \gls{wsn}, a robust and powerful middleware application that is able to use the sensory data and use it to control the home's appliances and a mobile application able to provide the users remote monitoring of their smart homes and control.
+This was possible by making the system as part of the \glsreset{iot}\gls{iot}.
 \chapter{Future Work}
@@
 The current state of the art does not incorporate this communication as part of the smart home systems \cite{ref15}.
-But as there is a need for this communication, \gls{dr} will have to be implementation as part of the
+But as there is a need for this communication, \gls{dr} will have to be implemented as part of the smart home package.
+\paragraph{}
+Demand reponse integration is a future work for this study.
+\gls{dr} has a set of service that have to be implemented in the smart home packages.
+In \gls{dr}, the consumer has the possibility to ask the utility about the instantaneous price quote for electricity.
+The consumer when in need of high priority in energy may ask the utility to provide such to provide a minimum of energy that will be guaranteed.
+Of course this means at a cost, but this agreement is part of the \gls{dr} and should be implemented at the level of smart home systems.
+Consumers can also help utilities at difficult times.
+For instance, when the utility cannot provide enough energy at peak times, the utility may ask some consumers to significantly cut their consumption and as a result be billed a reduced cost for the electricity that is consumed.
+In general, the smart home aims to reduce energy consumption and as the utility at times may need to reduce its energy production especially in peak hours, the \gls{dr} will help the utility ask the smart home to reduce the home consumption.
+\paragraph{}
+As this project aims to integrate the smart home devices within the \glsreset{iot}\gls{iot}, there is a need to integrate \glsreset{rfid}\gls{rfid} within this system as a future work.
+The \gls{rfid} is a building block within the \gls{iot} and will equip the devices with a low-power radio identification that will allow more flexibility to the system, as the external systems may be able to recognize uniquely each device in the \gls{iot} without counting primarily on the \gls{ipv6} which may change from time to time and therefore cannot keep track of the each device's information and monitor it reliably.
+\paragraph{}
+This system provides near real-time sensory data that is not much exploited intelligently.
+The only use of this data is to enable users to monitor their home consumption.
+As far as the system is concerned, there is a need for a more intelligent exploitation of this very rich data that can be used by intelligent algorithm to automatically and autonomously reduce the household's energy consumption.
+Such algorithms can be based on machine learning and data mining techniques making the direction of this a cross disciplinary project aiming to reduce the consumption of household without reducing the utility of users.
+In other words, build algorithms that will reduce consumption of households while the user is not noticing change.
+\paragraph{}
+Before building new services on top of this system, one should test different aspects of the system.
+In this study, the only thing that was tested, measured and quantified was the performance of the system which gave good results about the system's overall performance.
+As a future work, one should test different aspects of the system such as the reliability and availability of the system.
+More experiments should be conducted to see how well the system acts in high traffic networks and how good could it support new overlying systems such as \gls{dr}.
+\paragraph{}
+The system provides different services, but there is not authentication or authorization involved.
+This means that any users knowing the mote connected to the appliance he/she is interested in, he/she can control it and there is no existing party or component that will control such a communication.
+As a future work, the control of the system's access should be monitored in order for the system to be more secure and robust to external attacks.
 
 \backmatter
-
+\singlespacing
+
+\addcontentsline{toc}{chapter}{Bibliography}
+\renewcommand{\bibname}{References}
+\bibliographystyle{plain} 
+\bibliography{myrefs}
+
+\doublespacing
 \appendix
@@
 
-\singlespacing
-\bibliographystyle{IEEEtran} 
-\bibliography{IEEEabrv,myrefs}
-
 \end{document}

### 2013-07-03 doc/thesis/report.tex
report DONE :D

@@
 \paragraph{}
-The information technology is advancing exponentially and is having bigger impact on our lives as time advances.
-The next revolution in computing is called the \gls{iot}.
-It is basically the internet we know nowadays except that physical objects (fridge, television, air conditioning...) will be the major players in this complex system.
-In this context, the project builds on this idea a test bed for smart home energy optimization.
-The project allows users to monitor their energy consumption via their smart phone and also the users can control their appliances by turning on and off appliances remotely.
-\paragraph{}
-The system is divided into three main components.
-Each of them is using different technologies such as \gls{tinyos}, \gls{nesc}, \gls{j2se}, \gls{http} and other technologies that will be discussed in the thesis report.
-\paragraph{}
-At this stage, the system allows monitoring, but turning on and off appliances is not yet there mainly because of the absence of the equipment but the communication for that purpose is there.
+The Information Technology is advancing exponentially and largely bigger impacting our daily lives.
+The current evolution in computing is called the \gls{iot}.
+It is smoothly evolving from an Internet of people to and Internet of Things.
+By 2020, it is expected to have 50 billion "Things" connected to the Internet.
+\paragraph{}
+However such a migration induces a strong level of complexity when handling interoperability between heterogeneous things, i.e. Wireless Sensors, RFID, and mobile computing.
+\paragraph{}
+In this thesis, we focus on the  integration of wireless sensors into \gls{iot}.
+This prologues for a future integration of \gls{sg}\glsreset{sg} into \gls{iot}.
+In this work, we deployed a smart building testbed for energy optimization.
+\paragraph{}
+The system is using different technologies such as \gls{tinyos}, \gls{nesc}, \gls{j2se} and  \gls{http}.
 
@@
 \paragraph{}
-La technologie de l'information progresse de façon exponentielle et son impact sur nos vies prend de l'importance plus le temps avance.
-La prochaine révolution de l'informatique est appelé \glsreset{iot} \gls{iot}.
-Il s'agit essentiellement de Internet tel que nous la connaissons aujourd'hui, sauf que les objets physiques  seront les principaux acteurs de ce système.
-Ainsi, le projet s'appuie sur cette idée et présente un prototype qui a pour but l'optimisation intelligente de l'énergie dans les domiciles.
-Le projet permet aux utilisateurs de suivre de près leur consommation d'énergie via leur smartphone et permet de contrôler leurs appareils en allumant et éteignant les appareils à distance.
-\paragraph {}
-Le système est divisé en trois composantes principales.
-Chacun d'entre eux  utilise différentes technologies telles que \gls{tinyos}, \gls{nesc}, \gls{j2se}, \gls{http} et d'autres technologies qui seront présentées dans le rapport de thèse.
-\paragraph{}
-A ce stade, le système permet de surveiller, mais controler les appareils n'est pas encore mise en place principalement en raison de l'absence de matériel, mais l'infrastructure de communication pour ce but là est en place.
+La technologie de l'information progresse de façon exponentielle et prend largement plus impact sur nos vies quotidiennes.
+L'évolution actuelle de l'informatique est appelé \glsreset{iot}\gls{iot}.
+il est prévu d'avoir en 2020 50 milliards "choses" connecté à l'Internet.
+\paragraph{}
+Cependant, une telle migration induit un fort niveau de complexité lorsque l'interopérabilité entre en jeu entre differentes choses hétérogènes, à savoir capteurs sans fil, RFID et l'informatique mobile.
+\paragraph{}
+Dans cette thèse, nous nous concentrons sur l'intégration de capteurs sans fil dans \gls{iot}.
+Cela conduit vers une future intégration de Smart Grid (SG) dans \gls{iot}.
+Au cours de ce projet, nous avons déployé un banc d'essai intelligente du bâtiment pour l'optimisation de l'énergie.
+\paragraph{}
+Ce système d'information utilise différentes technologies  telles que TinyOS (TinyOS), \glsreset{nesc}\gls{nesc}, Java Standard Edition (J2SE), HTTP.
+
 \setcode{utf8}
@@
 They sketch an architecture capable of converting all the \glspl{wsn}, new and legacy, to support \gls{ipv6}.\glsreset{sg}
-\section{The \gls{sg} Two-Way Communication}
-\paragraph{}
 
 \chapter{Internet Of Things}
-
 \section{The actual Internet}
+\paragraph{}
+The \glsreset{iot}\gls{iot} refers to the evolution of the actual Internet.The term \gls{iot} was used for the first time by Kevin Ashton in 1999 \cite{ref17}.
+The human beings will represent a minority in this network.
+\gls{iot} will become the biggest machine ever created by human beings.
+In the actual internet nowadays
 \section{Introduction to \gls{iot}}
+\paragraph{}
+The \glsreset{iot}\gls{iot} is a network mostly ruled my machines, from the most powerful supercomputer and mainframes to the smallest devices such as sensor nodes.
+All will have one common point, being all connected to the same network and able to communicate, produce and consume information.
+The \gls{iot} is changing the world of computing.
+New paradigms, architectures, frameworks and prototypes are being built everyday to support it \cite{ref15}.
+New technologies such as \glsreset{rfid}\gls{rfid}, \glsreset{m2m}\gls{m2m}, \glsreset{ipv6}\gls{ipv6}, \glsreset{6lowpan}\gls{6lowpan} and \gls{ims} have come to place to support the \gls{iot}.
+The whole world is now being connected including humans and objects \cite{ref18}.
+\paragraph{}
+The \glsreset{iot}\gls{iot} brings numerous advantages but numerous issues will have to solved before this network explodes to trillions of devices connected to it.
+The first issue is the interconnection of all the devices into the \gls{iot}.
+What hardware will support such an ultra high traffic network with an ultra high performance and availability requirements.
+Will ubiquitous computing or pervasive computing have architectures that deal with such infrastructure which will clearly form a bottleneck in this whole complex system.
+\paragraph{}
+There are three steps in the flow of event in the \gls{iot} \cite{ref17}.
+\begin{enumerate}
+\item The object must by identified. \gls{rfid} is used for this purpose in addition to the feature of the objects should be recognized and converted to electronic format
+\item  The information must be transported to the information consumer.
+\item The information should be processed intelligently.
+\end{enumerate}
+
 \section{Importance of \gls{iot}}
-\section{The Impact of \gls{iot}}
+\paragraph{}
+The actual Internet is evolving into the \glsreset{iot}\gls{iot} because it allows things that are not possible in the Internet we all know.
+Autonomy of devices, production and consumption by things other than humans is a need.
+New high level communication language such as \gls{owl} are being created and studied extensively to enable objects to communicate freely without the need to interact with humans to get directives and configurations.
+\paragraph{}
+The world economy is having more and more requirements from the Internet.
+Requirements such as the identification of every device in the world and its interconnection to the Internet.
+The ability to get real-time information from devices throughout the globe.
+\paragraph{}
+
 \section{The Benefits of \gls{iot}}
+\paragraph{}
+\gls{iot} brings numerous solutions to the new requirements.
+In the Internet of Things, objects will work and communicate freely.
+Any device will be able to send its sensory information and consume data according to its needs.
+From this simple idea arise a whole new set of applications that will benefit the world.
+Smart grids exploit this idea and make it part of its requirement \cite{ref8}.
+More and more applications are now being developed.
+Ideas such as smart farming, smart transportation, smart homes, smart health, smart postal and smart refrigerator are only a small set of applications that count on the evolution of Internet towards the Internet of Things \cite{ref19}.
+\paragraph{}
+The Internet of Things will be able to help the world community solve and minimize the impact of numerous problems it faces.
+With the deployment of applications as part of \gls{iot}, we can predict natural disasters by sending real-time sensory data about geological, volcanic, seismic, meteorological and spacial variables and help communities avoid loosing their lives and being affected by these natural disasters.
+We can monitor water scarcity on a real-time basis.
+We can monitor in real-time our health by incorporating microscopic sensors that will be part of the Internet of Things.
+
 \section{The Future of Internet}
+\paragraph{}
+The Internet's counts billions of connected devices nowadays.
+This number is evolving exponentially and is not going to stop in the coming decades.
+It is estimated that there will be trillions of objects connected by the end of the century \cite{ref3}.
+In the next decade, we will watch a major change in this network.
+The network architectures we know nowadays will changes radically.
+The hardware architectures used now to evolve to meet the expectations of \gls{iot}.
+Applications based on \gls{iot} will explode and new possibilities will be open to world community.
+\paragraph{}
+Privacy of information is going to be a serious issue in \gls{iot}.
+As everything is monitored and controlled, privacy becomes an issue.
+The question is how will the Internet of Things evolved without having to be affected by privacy and security issues \cite{ref20}.
 
@@
 \backmatter
-\singlespacing
 
@@
 \renewcommand{\bibname}{References}
-\bibliographystyle{plain} 
-\bibliography{myrefs}
-
-\doublespacing
+\bibliographystyle{IEEEtran} 
+\bibliography{IEEEabrv,myrefs}
+
 \appendix

### 2013-07-04 trunk/doc/thesis/report.tex
chapter 1 reviewed

@@
 \paragraph{}
-The global electrical grid one the bases that gives us this quality of life that most people around the world enjoy.
+The global electrical grid is one the bases that gives us this quality of life that most people around the world enjoy.
 The global electrical grid is sometimes called the largest machine ever made by man.
 \cite{ref1}.
-It a widespread machine, distributed but interconnected system that provides power to households, companies and other organizations and therefore the world's economy is counting on it.
-The electrical grid knows numerous problems such as blackouts, current instabilities that costs billions of dollars yearly to both the utilities, utilities and insurances.
-The grid suffers from the problem of peak hours where energy consumption where in many case the one single utility cannot provide such energy.
-\paragraph{}
-The \gls{sg} comes to provide solutions to the problems that suffers from the traditional electrical grid.
-\gls{sg} is an intelligent, auto-balancing, and self-monitoring electrical grid \cite{ref2} and is and whose energy sources are distributed between utilities and consumers.
-The nature of the \gls{sg} as having distributed energy sources reduces numerous losses that arises in the traditional grid such as the energy loss in the energy transmission lines and also in the overload of the electrical lines.
-\paragraph{}
-Another revolution is about to begin and it is taking place in the information technology and is called \glsreset{iot}\gls{iot}.
-\gls{iot} is the era of autonomous machine able to produce and consume information and being able to act autonomously without the guidance of humans to performs their tasks.
-Cisco predicts that the number of devices to be connected in 2020 will break the 50 billions devices \cite{ref3} whereas the world population will be around 7.6 billion.
-That make 6.58 devices connected for each person in the world whereas the number of devices connected in 2010 12.5 billion.
+It a widespread machine, distributed but an interconnected system that provides power to households, companies and other organisations and therefore the world's economy is counting on it.
+The electrical grid knows numerous problems such as blackouts, current instabilities that cost billions of dollars yearly to both the utilities and insurances.
+The grid suffers from the problem of peak hours where energy consumption is high and in most cases, one single utility cannot provide.
+\paragraph{}
+The \gls{sg} comes as a solution to the problems that suffers from the traditional electrical grid.
+\gls{sg} is an intelligent, auto-balancing, and self-monitoring electrical grid \cite{ref2} and whose energy sources are distributed between utilities and consumers.
+The nature of the \gls{sg} as having distributed energy sources reduces numerous losses that arise from long energy transmission lines and also from the overload of the electrical lines.
+\paragraph{}
+Another revolution began and it is taking place in the information technology and is called \glsreset{iot}\gls{iot}.
+\gls{iot} is the era of autonomous machine able to produce and consume information and being able to act autonomously, without the guidance of humans, to performs their tasks.
+Cisco predicts that the number of devices to be connected to the Internet in 2020 will break the 50 billions devices \cite{ref3} whereas the world population will be around 7.6 billion.
+That makes 6.58 devices connected for each person in the world whereas in 2010, the number of devices connected was 12.5 billion.
 This clearly shows that the shift to the \gls{iot} has already started.
@@
 \glsreset{sg}\gls{sg} can and will benefit from the \gls{iot}.
-The integration of the \gls{sg} and and \gls{iot} makes this study and thesis of high importance in order to solve as much as possible the problems of the traditional grid but at the same bring a new dimension to the power grid by benefiting from the numerous possibilities offered by the \gls{iot}.
+The integration of the \gls{sg} into the \gls{iot} makes this study and thesis of high importance in order to solve as much as possible the problems of the traditional grid but at the same bring a new dimension to the power grid by benefiting from the numerous possibilities offered by the \gls{iot}.
 
@@
 This project has been done in order to investigate ways the \gls{iot} can benefit the \gls{sg}.
-The smart grid as a set of philosophical ideas counts on a set of novel functionalities.
-Once of these ideas is the two-way communication that will make possible the utility to communicate with consumers and vice versa.
+The Smart Grid as a set of philosophical ideas counts on a set of novel functionalities.
+One of these ideas is the two-way communication that will make possible the utility to communicate with consumers and vice versa.
 This concept is possible with the Internet Of Things as devices in \gls{iot} communicate between them by producing and consuming information.
@@
 \paragraph{}
-The \glsreset{iot}\gls{iot} counts on different technologies such as \gls{wsn}, \gls{rfid}, \gls{ipv6} and other technologies which make the Internet of things a very heterogeneous and complex system \cite{ref4}.
-One of the challenges in \gls{iot} is to deals with such heterogeneity and take the most out of it in order to make the communication as smooth and robust as possible.
+The \glsreset{iot}\gls{iot} counts on different technologies such as \gls{wsn}, \gls{rfid}, \gls{ipv6} and \gls{6lowpan}\glsreset{6lowpan} which make the Internet of Things a very heterogeneous and complex system \cite{ref4}.
+One of the challenges in \gls{iot} is to deal with such heterogeneity and take the most out of it in order to make the communication as smooth and transparent as possible.
 This heterogeneity should be made transparent to any two parties trying to communicate in \gls{iot}.
@@
 \paragraph{}
-The \glsreset{sg} \gls{sg} counts on the two-way communication to build to a whole communication and protocol stack namely the \glsreset{dr}\gls{dr} and other services built on top of the \gls{dr}.
-From the consumer's side or more precisely the household's side, there are smart home energy management software in the market nowadays but none of these products offers integration with \gls{sg}.
-To have such an integration with \gls{sg}, there is a need to first build this two-way communication and make it possible for any appliance to report its real-time energy consumption and also to be able to be controlled remotely with this two-way communication.
-Such infrastructure is missing for the time being and therefore the integration of the smart home within the smart grid is not possible.
+The \glsreset{sg} \gls{sg} counts on the two-way communication to build to a whole communication framework and a protocol stack namely the \glsreset{dr}\gls{dr} and other services built on top of this last.
+From the consumer's side or more precisely the household's side, there are smart home energy management systems in the market nowadays but none of them offers an integration with \gls{sg}.
+To have such an integration with \gls{sg}, there is a need to first build this two-way communication, to make it possible for any appliance to report its real-time energy consumption, and to be able to be controlled remotely with this infrastructure.
+Such pillar is missing for the time being, and therefore the integration of the smart home within the smart grid is not possible.
 Also, in order for \gls{sg} to be integrated into the \gls{iot}, all devices should be able to communicate in the two directions.
 \paragraph{}
-As \gls{sg} counts on numerous technologies to achieve its main goals.
-This usage of different technologies imposes another problem which is the heterogeneity of the system.
+\gls{sg} counts on various technologies to achieve its main goals.
+This usage of different technologies brings another problem which is the heterogeneous nature of the system.
 As numerous components count on different technologies, there is a need to interface them in order to make this heterogeneity transparent to the all components of the system.
-This means that there is a need to build a software within the whole system that would constitute an important part of it and will take care of making the communication and processes technology-transparent and give the possibility to all components of the system to work hand-in-hand in order their common goals independently of the data link layer technology (Ethernet, Wifi or Zigbee) they are counting on, the network layer technology (\gls{ipv4} or \gls{ipv6}).
+As a matter of fact, a software should be built within the system, and would constitute an important part of it, whose main goal is to take care of making the communication technology-transparent and give the possibility to all components of the system to work hand-in-hand in order to achieve their common goals independently of the data link layer (Ethernet, Wifi or Zigbee) or the network layer technology (\gls{ipv4} or \gls{ipv6}) they are using.
 
@@
 The goal of the study is to solve the problems introduced in the problem statement section.
-This means creating the two-way communication, and also solving the heterogeneity that arises from integrating smart homes with the smart grid and making each components in the smart home as part of the Internet of things by making it communicate with any device in the Internet and having an \gls{ipv6} address that is built by making use of \gls{6lowpan} whose goal is to provide the TCP/IP to the smallest devices and with the least computational and energy rich.
+This means creating the two-way communication, and solving the heterogeneity that arises from integrating smart homes with the smart grid and making each components in the smart home as part of the Internet of Things by making it communicate with any device in the Internet by having an \gls{ipv6} address that is built by making use of \gls{6lowpan} whose goal is to provide the TCP/IP to the smallest devices and with the least computational resources.
 \paragraph{}
@@
 This would make every sensor node also known as mote identified and ready to communicate and be part of the \gls{iot}.
-This communication will go in the two directions and based on that the two-way communication will have to be built.
+This communication will go in the two directions and based on that, the two-way communication will be built.
 \paragraph{}
@@
 The thesis aims to solves the two problems presented in section 1.3.
-To do so, a variety of tools and technologies were used for this purpose.The two problems were not solved one independently of the other but even though the problems seem separate one cannot solve without having to solve the other.
-In other words, after the creation of the two-way communication in the \glsreset{wsn}\gls{wsn} which supported \gls{ipv6} over \gls{6lowpan}, other network only supported \gls{ipv4} and therefore could not communicate, and this was due to the heterogeneity problem.
-This is why the two-way communication was created and then the heterogeneity problem had to be solved in order to see everything working.
-The system is divided into four components:
+To do so, a variety of tools and technologies were used for this purpose.The two problems were not solved one independently of the other but even though the problems seem separate, one cannot solve without having to solve the other.
+In other words, after the creation of the two-way communication in the \glsreset{wsn}\gls{wsn} which supported \gls{ipv6} over \gls{6lowpan}, other networks only supported \gls{ipv4} and therefore could not communicate, and this was due to the heterogeneity problem.
+This is why the two-way communication was created but only after solving the heterogeneity problem, everything worked.
+The system is divided into four subsystems:
 \begin{itemize}
@@
 \paragraph{}
-Each of these four component exist to solve part of the problem.
-Each of these components will be discussed extensively in the Chapter 4 from the architectural aspect of the component and in Chapter 5 from the implementation level of the system.
-\paragraph{}
-It is to be noted that some of the concepts will be discussed from the architectural point in Chapter 4 but not from an implementation point in Chapter 5.
-The sole reason behind is the fact that some of the equipment required for this thesis did not arrive on time and therefore the parts in the implementation requiring such equipment was omitted to meet the deadlines.
+Each of these four component is there to solve part of the problem.
+Each of these subsystems will be discussed extensively in the Chapter 4 from the architectural aspect of the component, and in Chapter 5 from the deployment point of view.
+\paragraph{}
+It is to be noted that some of the concepts will be discussed from the architectural point in Chapter 4 but not from an implementation aspect in Chapter 5.
+The sole reason behind is the fact that some of the equipment required for this thesis did not arrive on time and therefore the parts in the implementation requiring such equipment were omitted to meet the deadlines.
 \section{Outcome of the Study}
@@
 The autonomous production of information is shown by the fact that every sensor node in the \gls{wsn} reads sensory data from its sensors that are attached to appliance in order to sense their consumption and this information is produced and transmitted periodically.
-The consumption of data by is shown by external nodes in the \gls{iot} sending request to the sensor node in the \gls{wsn} in order to control and appliance's status i.e.
-turn on the television.
+The consumption of data by is shown by external nodes in the \gls{iot} sending requests to the sensor node in the \gls{wsn} in order to control the appliance's status i.e.
+turn On the television.
 This will prove the consumption of information by the nodes in the \gls{iot}.
@@
 After both experiments were conducted, we collected a significant amount of data that was used to measure the performance and an interpretation was done in section 6.3 to have an idea about the performance of the system.
-\paragraph{Experiment 1}
+\subsection{Experiment 1}
+\paragraph{}
 In the first experiment, the same mote was pinged once every second and we added new motes to the system every two minutes.
@@
 \end{table}
+\paragraph{}
 For each row in the table, the average delay is computed using the pings that were performed in the 2 minutes interval which means 120 pings per row.
@@
 
-\paragraph{Experiment 2}
+\subsection{Experiment 2}
+\paragraph{}
 The second experiment's goal is to measure the additional delay coming from the Gateway Packet transformation process whose goal to handle and hide the heterogeneity of the system.
@@
 \end{table}
-The transformation process's elapsed time is counted in nanoseconds and from \ref{table:exp2}, we can see that the average delay is on average 0.1 milliseconds whereas the jitter is 0 milliseconds meaning that the process's time varies between a few nanoseconds to at most 150 nano seconds.
+\paragraph{}
+The transformation process's elapsed time is counted in nanoseconds and from \ref{table:exp2}, we can see that the average delay is on average 0.1 milliseconds whereas the jitter is 0 milliseconds meaning that the process's time varies between a few nanoseconds to at most 150 nanoseconds.
 Depending on the machine's load.
 \section{Interpretation of Results}
+\subsection{Experiment 1}
 \paragraph{}
 The data collected in section 6.2 shows that the performance of the system is very good. Even though the delay in the \gls{wsn} is high, 
- nothing can be done about that as even though the traffic in the \gls{wsn} is low, the delay is still high and this has to do with the computational time taken by the TCP/IP protocol protocol stack provided by the
+ nothing can be done about that as even though the traffic in the \gls{wsn} is low, the delay is still high and this has to do with the computational time taken by the TCP/IP protocol protocol stack provided implemented by \glsreset{blip}\gls{blip} using \glsreset{6lowpan}\gls{6lowpan}.
+To reduce the delay, one should probably rethink the design of the \gls{blip} or find a way to omit it, but that is not a good idea since we want the \gls{wsn} to be part of \gls{iot}.
+ \paragraph{}
+ The good news in experiment 1 is that the privatization of packets brings benefit to the network.
+Even though the traffic grows exponentially, the delay does not grow much and the even if the jitter follows to some extent the traffic's trend but still the average delay shows good result.
+ \subsection{Experiment 2}
+ \paragraph{}
+ The Packet transformation process is a component that shows excellent results.
+Before conducting experiment 2, we were persuaded that there will be additional delay originating from this process and it could be responsible for the high delay seen in experiment 2 but the results shows otherwise and that the component is not to be put responsible for such important delay seen in experiment 1.
 \chapter{Conclusion}
@@
 \backmatter
-\singlespacing
-\addcontentsline{toc}{section}{References}
+
+\addcontentsline{toc}{chapter}{Bibliography}
+\renewcommand{\bibname}{References}
 \bibliographystyle{IEEEtran} 
@@
 
-\doublespacing
 \appendix

### 2013-07-05 trunk/doc/thesis/report.tex
major correction done. Not yet done

@@
 \newcommand{\HRule}{\rule{\linewidth}{0.6mm}}
+\usepackage[titletoc]{appendix}
 %\usepackage{hyperref}
@@
 \newenvironment{dedication}{\vspace{6ex}\begin{quotation}\begin{center}\begin{em}}{\par\end{em}\end{center}\end{quotation}}
+
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd{\thebibliography}{%
+  \chapter*{\bibname}\@mkboth{\MakeUppercase\bibname}{\MakeUppercase\bibname}}{%
+  \chapter{Index}}{}{}
+\makeatother
 
@@
 \paragraph{}
-The Information Technology is advancing exponentially and largely bigger impacting our daily lives.
-The current evolution in computing is called the \gls{iot}.
+Information Technology is advancing exponentially and largely impacting our daily lives.
+The current evolution in computing is the \gls{iot}.
 It is smoothly evolving from an Internet of people to and Internet of Things.
@@
 \paragraph{}
-However such a migration induces a strong level of complexity when handling interoperability between heterogeneous things, i.e. Wireless Sensors, RFID, and mobile computing.
+However such a migration induces a strong level of complexity when handling interoperability between heterogeneous Things, i.g. Wireless Sensors, RFID, and mobile devices.
 \paragraph{}
 In this thesis, we focus on the  integration of wireless sensors into \gls{iot}.
-This prologues for a future integration of \gls{sg}\glsreset{sg} into \gls{iot}.
-In this work, we deployed a smart building testbed for energy optimization.
+This is q prologue for the future integration of \gls{sg} into \gls{iot}.
+In this work, we deployed a smart building testbed for energy optimization, and for integration into \gls{sg}\glsreset{sg}.
+\paragraph{}
+The testbed built provides IPv6 gatewaying, remote control of wireless sensor nodes which is built on top of \gls{6lowpan}\glsreset{6lowpan}, and mobile access to the system.
 \paragraph{}
@@
 \paragraph{}
-The global electrical grid is one the bases that gives us this quality of life that most people around the world enjoy.
-The global electrical grid is sometimes called the largest machine ever made by man.
-\cite{ref1}.
-It a widespread machine, distributed but an interconnected system that provides power to households, companies and other organisations and therefore the world's economy is counting on it.
-The electrical grid knows numerous problems such as blackouts, current instabilities that cost billions of dollars yearly to both the utilities and insurances.
+The global electrical grid is the bases that gives us this quality of life that most people around the world enjoy.
+The global electrical grid is sometimes called the largest machine ever made by man \cite{ref1}.
+It is a widespread and distributed machine that provides power to households, companies and organizations.
+Therefore the whole world's economy counts on it.
+The electrical grid knows numerous problems such as blackouts, and current instabilities that cost billions of dollars yearly to both the utilities and insurances.
 The grid suffers from the problem of peak hours where energy consumption is high and in most cases, one single utility cannot provide.
 \paragraph{}
-The \gls{sg} comes as a solution to the problems that suffers from the traditional electrical grid.
-\gls{sg} is an intelligent, auto-balancing, and self-monitoring electrical grid \cite{ref2} and whose energy sources are distributed between utilities and consumers.
-The nature of the \gls{sg} as having distributed energy sources reduces numerous losses that arise from long energy transmission lines and also from the overload of the electrical lines.
-\paragraph{}
-Another revolution began and it is taking place in the information technology and is called \glsreset{iot}\gls{iot}.
-\gls{iot} is the era of autonomous machine able to produce and consume information and being able to act autonomously, without the guidance of humans, to performs their tasks.
-Cisco predicts that the number of devices to be connected to the Internet in 2020 will break the 50 billions devices \cite{ref3} whereas the world population will be around 7.6 billion.
-That makes 6.58 devices connected for each person in the world whereas in 2010, the number of devices connected was 12.5 billion.
-This clearly shows that the shift to the \gls{iot} has already started.
-\paragraph{}
-\glsreset{sg}\gls{sg} can and will benefit from the \gls{iot}.
-The integration of the \gls{sg} into the \gls{iot} makes this study and thesis of high importance in order to solve as much as possible the problems of the traditional grid but at the same bring a new dimension to the power grid by benefiting from the numerous possibilities offered by the \gls{iot}.
+The \glsreset{sg}\gls{sg} comes as a solution to the problems of the traditional electrical grid.
+\gls{sg} is an intelligent, auto-balancing, self-monitoring, and self-healing electrical grid \cite{ref2} and whose energy sources are distributed between utilities and consumers.
+The nature of the \gls{sg} as having distributed energy sources reduces numerous losses that arise from the long distance transmission lines, and also from the stress and the overload of the electrical lines.
+\paragraph{}
+Another revolution began, and it is taking place in the Information Technology.
+It is called the \glsreset{iot}\gls{iot}.
+\gls{iot} is the era of autonomous machine being able to produce and consume information, and being able to act autonomously, without the guidance of humans, to performs their tasks.
+Cisco predicts that the number of devices to be connected to the Internet in 2020 will break the 50 billions devices from 500 million in 2003.
+That would make 6.58 devices connected for each person in the world whereas that same ration was 0.07 17 years earlier \cite{ref3}.
+This clearly shows that the shift to the \gls{iot} will be exponential.
+\paragraph{}
+\glsreset{sg}\gls{sg} should benefit from the \gls{iot}.
+The integration of the \gls{sg} into the \gls{iot} constitutes the main topic of this study and goes along with increasing interest recently raised in both academia and industry, on the inherent correspondence between \gls{sg} and \gls{iot}.
+This  correspondence stems mainly from the need to disseminate information between the \gls{sg} components in order to make optimal decisions about energy usage.
+This information dissemination could only be optimal when deploying ubiquitous sensoring and metering, and using the available network infrastructure, i.e.
+the Internet.
 
@@
 This project has been done in order to investigate ways the \gls{iot} can benefit the \gls{sg}.
-The Smart Grid as a set of philosophical ideas counts on a set of novel functionalities.
-One of these ideas is the two-way communication that will make possible the utility to communicate with consumers and vice versa.
-This concept is possible with the Internet Of Things as devices in \gls{iot} communicate between them by producing and consuming information.
-Numerous ideas are based on this two-way communication such as the \gls{dr} which offers a set of services for both the power provider and the consumer.
-\paragraph{}
-The \glsreset{iot}\gls{iot} counts on different technologies such as \gls{wsn}, \gls{rfid}, \gls{ipv6} and \gls{6lowpan}\glsreset{6lowpan} which make the Internet of Things a very heterogeneous and complex system \cite{ref4}.
-One of the challenges in \gls{iot} is to deal with such heterogeneity and take the most out of it in order to make the communication as smooth and transparent as possible.
-This heterogeneity should be made transparent to any two parties trying to communicate in \gls{iot}.
+The Smart Grid needs a set of novel functionalities.
+One of these ideas is the two-way communication that will make it possible for the utility to communicate with consumers and vice versa.
+This is possible in the Internet Of Things as devices in \gls{iot} communicate by producing and consuming information.
+Numerous ideas assume the existence of the two-way communication such as the \gls{dr} which offers a set of services for both the power provider and the consumer.
+\paragraph{}
+The \glsreset{iot}\gls{iot} uses different technologies such as \gls{wsn}, \gls{rfid}, \gls{ipv6} and \gls{6lowpan}\glsreset{6lowpan} which makes the Internet of Things a very heterogeneous and complex system \cite{ref4}.
+One of the challenges in \gls{iot} is to deal with this heterogeneity in order to make the communication smooth and transparent to the parties involved.
+This heterogeneity should become invisible to any two devices willing to communicate in the \gls{iot}.
 
@@
 \paragraph{}
-The \glsreset{sg} \gls{sg} counts on the two-way communication to build to a whole communication framework and a protocol stack namely the \glsreset{dr}\gls{dr} and other services built on top of this last.
-From the consumer's side or more precisely the household's side, there are smart home energy management systems in the market nowadays but none of them offers an integration with \gls{sg}.
-To have such an integration with \gls{sg}, there is a need to first build this two-way communication, to make it possible for any appliance to report its real-time energy consumption, and to be able to be controlled remotely with this infrastructure.
+The \glsreset{sg} \gls{sg} relies on the two-way communication to build a whole communication framework and a protocol stack namely the \glsreset{dr}\gls{dr} and other services built on top of this last.
+From the consumer's/household's side or more precisely the household's side,there are smart home energy management systems in the market nowadays, but none of them offers an integration with \gls{sg}.
+To have such an integration with \gls{sg}, there is a demand for the two-way communication, which is to make it possible for any appliance to report its real-time energy consumption, and to be able to be controlled remotely.
 Such pillar is missing for the time being, and therefore the integration of the smart home within the smart grid is not possible.
@@
 \gls{sg} counts on various technologies to achieve its main goals.
-This usage of different technologies brings another problem which is the heterogeneous nature of the system.
+This use of different technologies comes with a problem which is the heterogeneous nature of the system.
 As numerous components count on different technologies, there is a need to interface them in order to make this heterogeneity transparent to the all components of the system.
-As a matter of fact, a software should be built within the system, and would constitute an important part of it, whose main goal is to take care of making the communication technology-transparent and give the possibility to all components of the system to work hand-in-hand in order to achieve their common goals independently of the data link layer (Ethernet, Wifi or Zigbee) or the network layer technology (\gls{ipv4} or \gls{ipv6}) they are using.
+As a matter of fact, a software should be built within the system, and would constitute an important part of it, whose main goal is to take care of making the communication technology-transparent and give the possibility to all components of the system to work hand-in-hand in order to achieve their goals independently of the data link layer (Ethernet, Wifi or Zigbee) or the network layer technology (\gls{ipv4} or \gls{ipv6}) they are using.
 
@@
 \paragraph{}
-The first purpose of this study is to create a testbed having the two-way communication and also being able to make use of it in order to build richer applications based on this possibility.
-To do it, there was a need to first introduce \gls{6lowpan} within every device in the \gls{wsn}.
-This would make every sensor node also known as mote identified and ready to communicate and be part of the \gls{iot}.
-This communication will go in the two directions and based on that, the two-way communication will be built.
-\paragraph{}
-The second purpose of this study is to solve the heterogeneity problem that arises with the integration of smart homes within the smart grid and also when making the whole system as part of the \glsreset{iot}\gls{iot}.
-Solving such a problem means having an additional component that will track all communication packets and change them to fit the needs of the destination component.
+The first purpose of this study is to create a testbed having the two-way communication and also being able to make use of it in order to build richer applications based on it.
+To do so, \gls{6lowpan} had to be integrated within every device in the \gls{wsn}.
+This would make every sensor node (mote) identifiable in the network and ready to communicate and be part of the \gls{iot}.
+This communication will go in both directions and based on \gls{6lowpan}, the two-way communication will be built.
+\paragraph{}
+The second purpose of this study is to solve the heterogeneity problem that arises with the integration of smart homes into the grid and also when making the whole system part of the \glsreset{iot}\gls{iot}.
+Solving such a problem means having an extra component that will track all  packets and change them to fit the standards of the destination component.
 This will make of this additional component a middleman and therefore make the heterogeneity problem disappear.
@@
 \paragraph{}
-The thesis aims to solves the two problems presented in section 1.3.
-To do so, a variety of tools and technologies were used for this purpose.The two problems were not solved one independently of the other but even though the problems seem separate, one cannot solve without having to solve the other.
-In other words, after the creation of the two-way communication in the \glsreset{wsn}\gls{wsn} which supported \gls{ipv6} over \gls{6lowpan}, other networks only supported \gls{ipv4} and therefore could not communicate, and this was due to the heterogeneity problem.
-This is why the two-way communication was created but only after solving the heterogeneity problem, everything worked.
+This study aims to solves the two problems presented in section 1.3.
+To do so, a variety of tools and technologies were used for this purpose.The two problems could not be solved one independently of the other even though the problems seem separate, one cannot solve  the a problem without having to solve the other.
+In other words, after the creation of the two-way communication in the \glsreset{wsn}\gls{wsn} which adopts \gls{ipv6} over \gls{6lowpan}, other networks only support \gls{ipv4} and therefore the communication is not possible, and this was due to the heterogeneous nature of the system.
+This is why the two-way communication was created, but only after solving the heterogeneity problem, everything worked.
 The system is divided into four subsystems:
@@
 \item The fully \gls{ipv6} supported \gls{wsn}
-\item The \gls{tinyos} program for sensor transmission and appliance request handler
 \item The heterogeneity handler and middleman gateway server
+\item The middle-ware
 \item The mobile client
@@
 \paragraph{}
-Each of these four component is there to solve part of the problem.
-Each of these subsystems will be discussed extensively in the Chapter 4 from the architectural aspect of the component, and in Chapter 5 from the deployment point of view.
-\paragraph{}
-It is to be noted that some of the concepts will be discussed from the architectural point in Chapter 4 but not from an implementation aspect in Chapter 5.
-The sole reason behind is the fact that some of the equipment required for this thesis did not arrive on time and therefore the parts in the implementation requiring such equipment were omitted to meet the deadlines.
+Each of these four components is there to solve part of the problem.
+Each of these subsystems will be discussed extensively in the Chapter 4 from its architectural aspect of the component, and in Chapter 5 from its deployment point of view.
+\paragraph{}
+It is to be noted that some of the components will be discussed from the architectural point in Chapter 4 but not from an implementation aspect in Chapter 5.
+The sole reason behind is the fact that some of the equipment required for this thesis were not delivered on time and therefore the parts in the implementation requiring such equipment were omitted to meet the deadlines.
 \section{Outcome of the Study}
@@
 The outcome of the thesis is to design and implement a test-bed that is able to show the two-way communication in action.
-It will also have to show concepts of the \glsreset{iot}\gls{iot} which are that every node in the system can communicate autonomously and with any other node in the Internet.
-Another \gls{iot} concept is the ability of any node in the \gls{iot} to consume and produce information that is communicated through the network.
-\paragraph{}
-The autonomous production of information is shown by the fact that every sensor node in the \gls{wsn} reads sensory data from its sensors that are attached to appliance in order to sense their consumption and this information is produced and transmitted periodically.
-The consumption of data by is shown by external nodes in the \gls{iot} sending requests to the sensor node in the \gls{wsn} in order to control the appliance's status i.e.
-turn On the television.
-This will prove the consumption of information by the nodes in the \gls{iot}.
-\paragraph{}
-The test-bed should also show that communication between any two nodes in the \gls{iot} is successful, robust and reliable and that the heterogeneity of the system is made transparent to the nodes in the system by the creation of the software tool residing in the gateway which is between the \gls{iot} and the outside world.
-This software tool is seen as a heterogeneity handler and also a communication middleman that translates the communication to the appropriate technology needed by the recipient.
-\paragraph{}
-The testbed will able to show the two-way communication in action and will also show that the heterogeneity does not form a barrier by its successful handling.
+It will also have to show its integration into the \glsreset{iot}\gls{iot} by showing that every node in the system can communicate autonomously and with any other node in the Internet.
+Another \gls{iot} concept is the ability of any node in the \gls{wsn} to consume and produce information that is communicated through the network.
+\paragraph{}
+The autonomous production of information is shown by the fact that every sensor node in the \gls{wsn} reads sensory data from its sensors, that are attached to appliance, in order to sense their consumption and this information is produced and communicated periodically.
+The consumption of data is proved by showing the execution of the commands communicated through the network by external nodes in the \gls{iot}.
+These commands request the sensor node to turn On or Off an appliance.
+\paragraph{}
+The test-bed should also show that communication between any two nodes in the \gls{iot} is successful and reliable, and that the heterogeneity of the system is transparent to the nodes in the system and this is done by creating a software tool deployed in the gateway which is between the \gls{wsn} and the outside world.
+This software tool is perceived as a heterogeneity handler and also a communication middleman that translates the communication to the appropriate technology needed by the recipient.
+\paragraph{}
+The testbed will be able to show the two-way communication in action and will also show that the heterogeneity is not a barrier.
 
@@
 \paragraph{}
-The thesis is is divided into eight chapters. Each of the chapter approaches thesis from a different aspect. 
+The thesis is divided into eight chapters. Each of the chapter approaches thesis from a different aspect. 
 \paragraph{Chapter 1}
-It is an introductory chapter that gives a flavor of the project that is to be studied in this thesis and also introduces the problem and the solution proposed in this topic.
-This chapter also puts the reader into perspective by introducing some of the main concepts related to the \glsreset{sg}\gls{sg} and the \glsreset{iot}\gls{iot}.
+It is an introductory chapter that gives a flavor of the project that is  studied in this thesis and also introduces the reader to the problem and the solution proposed in this context.
+This chapter also puts into perspective by introducing some of the main concepts related to the \glsreset{sg}\gls{sg} and the \glsreset{iot}\gls{iot}.
 \paragraph{Chapter 2}
@@
 \paragraph{Chapter 3}
-It is dedicated to the \glsreset{iot}\gls{iot}.
-It presents the topic, talks about the importance of \gls{iot}, its impact on the world, the actual Internet, the future of Internet and also how will \gls{sg} benefit from it.
+It is dedicated to the \glsreset{iot}\gls{iot}. It presents the topic, talks about the importance of \gls{iot}, its impact on the world, the actual Internet and the future of Internet.
 \paragraph{Chapter 4}
-It will present the architecture of the system and the test-bed. It will discuss the system from different aspects, and how were the two problems solved from the architectural aspect.
+It will present the architecture of the test-bed. It will discuss the system from different aspects, and how were the problems stated in section 1.3 solved from the architectural aspect.
 \paragraph{Chapter 5}
-This chapter will mostly discuss the implementation system.
+This chapter will mostly discuss the system deployment.
 It will start by presenting the hardware used in this project, then move to the technology enablers.
-Each component of the system will be discussed extensively and move to the integration of the components to form the test-bed.
+Afterward, Each component of the system will be discussed extensively and move to the integration of the components to form the test-bed.
 \paragraph{Chapter 6}
-It will start first by presenting the findings then move to explain the experiment that was done to measure the performance of the system.
-The data gathering section will show how and what data was gathered and then move to analyzing and interpreting the data.
+It will start first by presenting the findings then move to explain the experiments that were conducted. Afterwards, the data gathered will be shown, analyzed and interpreted.
 \paragraph{Chapter 7, 8}
-These are the last two chapters of the thesis that will conclude the thesis, provide directions about future work and recommendations to the reader what parts of the system could be ameliorated and how it can be achieved.
+These are concluding chapters of this study. They provide directions about future work, recommendations to the reader about what parts of the system could be ameliorated and how it can be achieved.
 \chapter{Literature Review}
@@
 \paragraph{}
-The actual electrical grid is based on a technology that is at least one century old \cite{ref5}.
-As a result, there is a need to change as to meet to 21st century requirement such as having more green energy in order to reduce the pollution caused mainly by the use of fossil energy.
-As a result, research has introduced a new technology called \glsreset{sg}\gls{sg}.
-Although it is still a research project and numerous issues related to \gls{sg} are left unanswered, nations across the globe started to deploy \gls{sg} project.
-It is also noted that in order for the grid to become more reliable, secure and efficient, a bi-directional information and communication infrastructure should be embedded within the grid \cite{ref5}.
-\paragraph{}
-The smart grid brings around new concepts that will achieve efficiency, reliable and security to the power grid.
+The actual electrical grid is based on a technology that is at least one century old and its problems are becoming more and more costly \cite{ref5}.
+Therefore, there is a need to change it as to meet to 21st century requirements such as having more green energy in order to reduce pollution caused mainly by the use of fossil energy.
+As a matter of fact, research introduced us to new technology: the \glsreset{sg}\gls{sg}.
+Although it is still a research project and has not reached maturity yet, nations across the globe started to deploy \gls{sg} projects.
+The \gls{sg} is relying on Information Technology to solve many of the problem existing in the traditional grid.
+Consequently, in order for the grid to become more reliable, secure and efficient, a bi-directional information and communication infrastructure should be embedded within the grid \cite{ref5}.
+\paragraph{}
+The smart grid brings new concepts that will make it achieve greater efficiency, reliability and security compared to the traditional grid.
 The first one is the introduction of the \gls{dg} by exploiting \gls{res} \cite{ref6} and this reduce the losses of energy in the transmission.
-This distribution by itself enables the power grid this to move from the one way energy distribution to a two-way energy distribution where the utility will not be the sole energy producer but households can now produce and even sell the extra power.
-This is known nowadays as the feed-in-tariff which is a policy that allows the consumers to become producer and sell the excess power to the grid \cite{ref7}.
-This will allow the power grid to provide power to consumer in a "free market" fashion which means that it will obey to the laws of supply and demand \cite{ref7}.\glsreset{iot} \glsreset{sg}
+This distribution by itself enables the power grid to move from the one way energy distribution to a two-way energy distribution where the utility will not be the sole energy producer, but households can now produce and even sell the extra electricity.
+This is known nowadays as the feed-in-tariff which is a policy that allows the consumers to become producers and sell the excess power to the grid \cite{ref7}.
+This will allow the power grid to provide electricity to consumer in a "free market" fashion which means that it will obey to the laws of supply and demand \cite{ref7}.\glsreset{iot} \glsreset{sg}
 \section{\gls{iot} and \gls{sg}}
 \paragraph{}
-The \glsreset{iot}\gls{iot} has known high acceptance by the players in the Internet.
-Nowadays, it is estimated that 12.5 billion devices are connected to the Internet and that this trend will continue to read 50 billion connected devices by the end of the decade \cite{ref3}.
-\gls{iot} brings several benefits to the actual Internet and these benefits will pull \gls{sg} to be part of it.
-Consequently, research has taken this direction which is to integrate the \gls{sg} within the \gls{iot} as most aspects of \gls{sg} will be enhanced by \gls{iot} \cite{ref8}.
-\paragraph{}
-\gls{iot} counts on a set of technologies which are \glsreset{rfid}\gls{rfid}, \gls{m2m} \cite{ref9}, public communication networks such as 2G/3G mobile networks, power lines, WI-FI, Zigbee and other technologies \cite{ref8}.
+The \glsreset{iot}\gls{iot} has been highly accepted by the players in the Internet as its adoption is starting smoothly.
+Nowadays, it is estimated that 12.5 billion devices are connected to the Internet and that this trend will continue to reach 50 billion devices by the end of the decade \cite{ref3}.
+\gls{iot} brings several benefits to the actual Internet and will pull the \gls{sg} to make this part of it.
+Consequently, research has taken the direction to integrate the \gls{sg} within the \gls{iot} as most of its aspects favor  the \gls{sg} \cite{ref8}.
+\paragraph{}
+\gls{iot} counts on a set of technologies which are \glsreset{rfid}\gls{rfid}, \gls{m2m} \cite{ref9}, public communication networks such as 2G/3G mobile networks, power line communication, WI-FI and Zigbee\cite{ref8}.
 The smart grid's architecture based on \gls{iot} will be divided into three layers:
@@
 \paragraph{}
-Some of the research is discussing the need to replace Zigbee by Wi-Fi as the Wi-Fi offers higher bandwidth, non-line transmission ability, large-scale data collection and is highly cost-effective \cite{ref10} but still Zigbee has the sole advantage of very low energy consumption which is something required in \gls{sg} and \gls{iot}, because in systems that are installed far away from any maintenance team such as smart grid monitoring and self-healing devices, one cannot manage to move for the sole purpose of providing energy often.
+Research discusses the benefits to replace Zigbee by Wi-Fi as this last offers higher bandwidth, non-line transmission ability, large-scale data collection and is highly cost-effective \cite{ref10}.
+Still, Zigbee has the exclusive advantage of comsuming the least energy of the communication technologies that exist nowadays which is something required in \gls{sg} and \gls{iot}, especially for systems that are installed far away from any maintenance team such as smart grid monitoring and self-healing devices where one cannot manage to move for the sole purpose of charging these devices.
 \paragraph{}
 From the architectural point of view, integrating \gls{sg} within \gls{iot} means having to address heterogeneity issues.
-An \gls{iot} gateway system that is based on Zigbee and GPRS protocols \cite{ref11}.
-This enables then to deal partly with the heterogeneity problem and therefore enable the \gls{wsn} to communicate with the mobile telecommunication network.
-Another solution to the heterogeneity problem is proposed where the solution is to present a new, light-weight web service transport protocol called Lean Transport Protocol (LTP) that will allow transparent exchange of web service messages between all kinds of devices \cite{ref12}.
-This protocol is platform-independent, low energy consumer.
-Some researchers claim that the heterogeneity is mainly coming from the fact that there are different \gls{wsn} devices that do not necessarily use the same standards and protocols.
-They propose to move all the \glspl{wsn} to all-IP networks as this would remove most of the heterogeneity but there are already deployed legacy \glspl{wsn} \cite{ref13}.
-They sketch an architecture capable of converting all the \glspl{wsn}, new and legacy, to support \gls{ipv6}.\glsreset{sg}
+An \gls{iot} gateway system based on Zigbee and GPRS protocols helps to deal partly with the heterogeneity problem and therefore enable the \gls{wsn} to communicate with the mobile telecommunication network \cite{ref11}.
+Another solution to the heterogeneity problem is proposed with a new, light-weight web service transport protocol called Lean Transport Protocol (LTP) that will allow transparent exchange of web service messages between all kinds of devices.
+This protocol is platform-independent, low energy communication \cite{ref12}.
+Some researchers claim that the major source of heterogeneity the fact that there are different types of \gls{wsn} devices (Micaz, Mica2, Telosb...) that do not use the same standards and protocols.
+They propose to move the communication in \glspl{wsn} to an "all-IP" as this would remove most of the heterogeneity but as there are already deployed legacy \glspl{wsn}, the task is complicated to conduct.
+Fortunately, the researchers sketch an architecture capable of converting all the \glspl{wsn}, new and legacy, to support \gls{ipv6} \cite{ref13}.\glsreset{sg}
 
 \chapter{Internet Of Things}
-\section{The actual Internet}
+\section{The Actual Internet}
 \paragraph{}
 The \glsreset{iot}\gls{iot} refers to the evolution of the actual Internet.The term \gls{iot} was used for the first time by Kevin Ashton in 1999 \cite{ref17}.
-The human beings will represent a minority in this network.
-\gls{iot} will become the biggest machine ever created by human beings.
-In the actual internet nowadays
+The human beings will represent in the future a very small minority in the Internet and as it will be rules by autonomous machines.
+\gls{iot} will become the biggest machine ever created by human beings as the smart grid will be part of it.
 \section{Introduction to \gls{iot}}
@@
 The \glsreset{iot}\gls{iot} is a network mostly ruled my machines, from the most powerful supercomputer and mainframes to the smallest devices such as sensor nodes.
-All will have one common point, being all connected to the same network and able to communicate, produce and consume information.
+All will have one common point, being all connected to the Internet and able to communicate, produce and consume information.
 The \gls{iot} is changing the world of computing.
-New paradigms, architectures, frameworks and prototypes are being built everyday to support it \cite{ref15}.
-New technologies such as \glsreset{rfid}\gls{rfid}, \glsreset{m2m}\gls{m2m}, \glsreset{ipv6}\gls{ipv6}, \glsreset{6lowpan}\gls{6lowpan} and \gls{ims} have come to place to support the \gls{iot}.
+New paradigms, architectures, frameworks and prototypes are built and refined everyday to support it \cite{ref15}.
+New technologies such as \glsreset{rfid}\gls{rfid}, \glsreset{m2m}\gls{m2m}, \glsreset{ipv6}\gls{ipv6}, \glsreset{6lowpan}\gls{6lowpan} and \gls{ims} have come to place to establish the infrastructure for \gls{iot}.
 The whole world is now being connected including humans and objects \cite{ref18}.
 \paragraph{}
-The \glsreset{iot}\gls{iot} brings numerous advantages but numerous issues will have to solved before this network explodes to trillions of devices connected to it.
+The \glsreset{iot}\gls{iot} brings numerous advantages but many issues will have to solved before this network explodes to connecting trillions.
 The first issue is the interconnection of all the devices into the \gls{iot}.
 What hardware will support such an ultra high traffic network with an ultra high performance and availability requirements.
-Will ubiquitous computing or pervasive computing have architectures that deal with such infrastructure which will clearly form a bottleneck in this whole complex system.
-\paragraph{}
-There are three steps in the flow of event in the \gls{iot} \cite{ref17}.
+Will ubiquitous computing or pervasive computing have architectures that deal with such infrastructure?
+\paragraph{}
+To better understand \gls{iot},The following are the three steps in its flow of information:
 \begin{enumerate}
-\item The object must by identified. \gls{rfid} is used for this purpose in addition to the feature of the objects should be recognized and converted to electronic format
-\item  The information must be transported to the information consumer.
-\item The information should be processed intelligently.
+\item The object is identified by making use of \gls{rfid}. In addition, the objects should be recognized by sensing the environment and this information should be converted to electronic format
+\item  The information is transported to the information consumer.
+\item The information is processed intelligently \cite{ref17}.
 \end{enumerate}
@@
 \paragraph{}
-The actual Internet is evolving into the \glsreset{iot}\gls{iot} because it allows things that are not possible in the Internet we all know.
-Autonomy of devices, production and consumption by things other than humans is a need.
-New high level communication language such as \gls{owl} are being created and studied extensively to enable objects to communicate freely without the need to interact with humans to get directives and configurations.
-\paragraph{}
-The world economy is having more and more requirements from the Internet.
-Requirements such as the identification of every device in the world and its interconnection to the Internet.
-The ability to get real-time information from devices throughout the globe.
+The actual Internet is evolving into the \glsreset{iot}\gls{iot} because it opens possibilities that are not in the actual Internet.
+The ability of Things to produce and consume information is a requirement.
+New high-level communication languages such as \gls{owl} were invented and studied extensively to enable objects to communicate freely without the need to interact with humans.
+\paragraph{}
+The world's requirements from the Internet are growing exponentially.
+Such new requirement are the accurate identification of Things, and the ability to get real-time information from devices throughout the globe.
 \paragraph{}
@@
 \paragraph{}
-\gls{iot} brings numerous solutions to the new requirements.
+\gls{iot} is armed with solutions to the new requirements.
 In the Internet of Things, objects will work and communicate freely.
-Any device will be able to send its sensory information and consume data according to its needs.
-From this simple idea arise a whole new set of applications that will benefit the world.
-Smart grids exploit this idea and make it part of its requirement \cite{ref8}.
-More and more applications are now being developed.
-Ideas such as smart farming, smart transportation, smart homes, smart health, smart postal and smart refrigerator are only a small set of applications that count on the evolution of Internet towards the Internet of Things \cite{ref19}.
-\paragraph{}
-The Internet of Things will be able to help the world community solve and minimize the impact of numerous problems it faces.
-With the deployment of applications as part of \gls{iot}, we can predict natural disasters by sending real-time sensory data about geological, volcanic, seismic, meteorological and spacial variables and help communities avoid loosing their lives and being affected by these natural disasters.
-We can monitor water scarcity on a real-time basis.
-We can monitor in real-time our health by incorporating microscopic sensors that will be part of the Internet of Things.
+Any device will be able to send its sensory information, and consume data.
+From this simple idea arises a whole new set of applications that will benefit the world.
+Smart grids exploit this idea and make it part of its requirement in order to solve the problem in the electrical grid\cite{ref8}.
+More and more applications based on \gls{iot} are now being developed.
+Ideas such as smart farming, smart transportation, smart homes, smart health, smart postal and smart refrigerators are only a small subset of a myriad of applications that will come to life with \gls{iot} \cite{ref19}.
+\paragraph{}
+The Internet of Things will be able to help the world's community solve and minimize the impact of hazards it faces.
+With the deployment of applications which will be part of \gls{iot}, we can predict natural disasters by sending real-time sensory data about geological, volcanic, seismic, meteorological and spacial variables and help communities avoid losing their lives.
+We can also monitor water scarcity on a real-time basis.Finally, each one's health can be monitored in real-time by incorporating microscopic sensors that are able to communicate this Information through Internet.
 
@@
 \paragraph{}
-The Internet's counts billions of connected devices nowadays.
-This number is evolving exponentially and is not going to stop in the coming decades.
+The Internet counts billions of connected devices nowadays.
+This number is growing exponentially and this is not going to stop in the coming decades.
 It is estimated that there will be trillions of objects connected by the end of the century \cite{ref3}.
-In the next decade, we will watch a major change in this network.
+In the next decade, we will watch a major change in the Internet.
 The network architectures we know nowadays will changes radically.
-The hardware architectures used now to evolve to meet the expectations of \gls{iot}.
-Applications based on \gls{iot} will explode and new possibilities will be open to world community.
+The hardware architectures used now will evolve to meet the expectations of \gls{iot}.
+Applications based on \gls{iot} will explode and new possibilities will be open to the world community.
 \paragraph{}
 Privacy of information is going to be a serious issue in \gls{iot}.
-As everything is monitored and controlled, privacy becomes an issue.
-The question is how will the Internet of Things evolved without having to be affected by privacy and security issues \cite{ref20}.
+As every device will be monitored and controlled, privacy becomes an issue.
+The question is how will the Internet of Things keep its evolution without having to be affected by privacy and security arguments \cite{ref20}.
 
@@
 \paragraph{}
-The thesis is based on the study and creation of a test-bed system whose goal is to integrate smart home with the \glsreset{sg}\gls{sg} and all within the \glsreset{iot}\gls{iot}.
-To do so, we first define the system's objectives and goals which do align  with the thesis's problem statement and purpose of study.
-Then each part of the system is presented, and a section of this chapter is dedicated to each part explaining the architecture of each component and how it aligns and integrates with the whole system.
+The thesis is based on the study and creation of a test-bed whose goal is to interconnect smart homes with the \glsreset{sg}\gls{sg} and integrate both into the \glsreset{iot}\gls{iot}.
+To do so, we first define the system's objectives and goals.
+Then each part of the system is presented, and a section of this chapter is dedicated to each part explaining the architecture of each component and how it aligns and integrates into the whole system.
 \section{System Objectives}
@@
 \paragraph{Objective 2}
-Another system objective is to be able to monitor in real-time the consumption of each appliance and send this to a central repository where this data is to be stored.
-This will enable the sensor node to keep track of the appliance it is controlling and monitoring and collect consumption data that is sent and stored in a \gls{rdbms}.
+The second objective is to monitor in real-time the consumption of each appliance and communicate this information to a central repository where it is stored.
+The sensor node will be able to control, monitor and sense the consumption data that is communicated and stored in a \gls{rdbms}.
 \paragraph{Objective 3}
-The third objective is to make this system hide its heterogeneity from the outside world and also from different component belonging to it.
-This means that the system should be able to make any two components communicate independently of the technology they are using and also making each component think the other is using the same technology as his.
+The third objective is to make this system hide its heterogeneity from the outside world and also from the different components that it accounts.
+This means that the system should be able to make any two components communicate independently of the technology they are using and also make each component believe the other is using the same technology as his.
 
@@
 \paragraph{}
-The system is divided into nine components where every component is in charge of a part of the system's objectives.
+The system is divided into nine components, as depicted in figure \ref{fig:gen_architecture}, where every component is in charge of a part of the system's objectives.
 When these nine components work hand-in-hand, they achieve all the system's and project's objectives and goals.
@@
 \item Mote appliance control TCP server
-\item Gateway packet transformation process
+\item \gls{ipv4}/\gls{ipv6} Gatewaying Process
 \item \gls{ipv6} over USB tunnel
@@
 The general architecture is depicted in figure \ref{fig:gen_architecture}.
-This figure shows the different components of the system and their interactions.
-Each set of components is appearing within a box and that box is labeled.
-This label refers to the hardware deployment of the component within the whole system.
-\begin{figure}[htbp]
-\centering
-\includegraphics[scale=0.40]{img/general_architecture.jpg}
+This figure shows the existing components and their interactions.
+Each set of components appear within a box as this last refers to what hardware are the components deployed in as it is labeled in the box.
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=17cm,height=17cm,keepaspectratio]{img/general_architecture.jpg}
 \caption{General architecture depicting the main components of the system}
@@
 \paragraph{}
-Now that the general architecture has been defined, the question that should be asked is whether all these components reside in the same layer in the \gls{osi} model and if not, in which layer does each of these components reside in the \gls{osi} model.
+Now that the general architecture has been defined, the question that should be asked is whether all these components operate in the same layer in the \gls{osi} model and if not, in which layer does each of these components reside.
 As an answer to such a question, figure \ref{fig:layered_view} is here to show on what layer does every component act on.
@@
 \end{figure}
+
 \paragraph{}
 From figure \ref{fig:layered_view}, one can see that there are three components in performing in the data link layer and network layer.
-The first one is the packet transformation process whose goal is to hide the communication heterogeneity by sniffing incoming packets, transforming them, and forwarding them to the destination to meet the destination's communication technology and standards used.
-The second component performing in the data link layer and network is the \gls{ipv6} over USB tunnel.
-This component is in charge of getting raw \gls{ipv6} packets and encapsulate them through USB to reach the destination.
-The third component is deployed in the mote sink which is the entry and exit point of the \gls{wsn}.
+The first one is the \gls{ipv4}/\gls{ipv6} gatewaying process whose goal is to hide the communication heterogeneity by sniffing incoming packets, transforming them in order to meet the recipient's communication technology and standards.
+The second component performing in the data link layer and network is the \gls{ipv6} over USB tunnel; This component is in charge of receiving raw \gls{ipv6} packets and encapsulate them in  USB frames to reach the destination.
+The third component is deployed in the sink which is the entry and exit point of the \gls{wsn}.
 This special mote receives \gls{ipv6} packets encapsulated in USB frames that are forwarded to the destination mote using Zigbee.
@@
 \paragraph{}
-There are four components that perform in the session layer, presentation layer and application layer and are existing in most of the devices needed in the system.
+There are four components that operate in the session layer, presentation layer and application layer and are existing in most of the devices needed in the system.
 The first one is the client mobile App.
-This component is mostly a TCP client and uses many of the services provided by the layers it is performing on.
-The second component is the sensor data storage which is a TCP server receiving data from the \gls{wsn} and sending storing it in a \gls{rdbms}.
-The third and fourth components are the Mote TCP client and Mote TCP server that send sensor data and through the TCP client and receive orders to control the appliance using the TCP server.
-\paragraph{}
-The only component acting in the application layer exclusively is the HTTP sensor data sender which provides a gateway to other applications willing to use such data and later on make use of the system.
-This component receives data and sends to the destination by using HTTP protocol.
-The middleware application was using HTTP to get sensor data from this system.
-\paragraph{}
-Now that the general structure of the software system has been established, the hardware structure is to be presented and explained further.
-figure \ref{fig:network_diagram} is a network diagram that shows the of the hardware components that exist within this system.
-The \glsreset{wsn}\gls{wsn} appears on the right of the document.
-Attached to it a mote sink that exist both in the \gls{wsn} and outside it.
-This mote sink is attached to the gateway server through USB.
-The gateway has also a Wi-Fi interface card that enables it to communicate with the other parts of the system such as the middleware and the mobile client.
+This component is implements a TCP client and uses most of the services provided by the layers it is operating in.
+The second component is the sensor data storage which constitutes of a TCP server receiving data from the \gls{wsn}, transforming and storing it in a \gls{rdbms}.
+The third and fourth components are the Mote TCP client and server that send sensor data and through the TCP client and receive commands using the TCP server.
+\paragraph{}
+The only component acting in the application layer exclusively is the HTTP sensor data sender which provides a gateway to other applications that need this data.
+This component receives data and communicates it to the destination using HTTP.
+The middle-ware application used HTTP to get sensor data from this system.
+\paragraph{}
+Now that the general structure of the software system has been established, the hardware structure will be presented and explained further.
+figure \ref{fig:network_diagram} is a network diagram that shows the hardware components in this system.
+The \glsreset{wsn}\gls{wsn} appears on the right of the figure.
+Attached to it the sink that which constitutes and entry and exit point for the \gls{wsn}.
+This sink is attached to the gateway server through USB a the gateway sees it as a network interface card.
+In addition, the gateway is equipped with a Wi-Fi interface card that enables it to communicate with the other parts of the system such as the middle-ware and the mobile client, and connect to the Internet.
 
@@
 \end{figure}
-\paragraph{}
-Now that the network diagram has been established in \ref{fig:network_diagram} where the hardware component appears, the next thing is to show where each software component will be deployed.
+
+\paragraph{}
+Now that the network diagram has been established in \ref{fig:network_diagram} where the hardware component are depicted, we need to show where each software component is deployed.
 In other words, what are the software components that exist on which hardware machine.
-This is the goal of table \ref{table:soft_hard_deployment}.
-It clearly shows what software components exist in what hardware.
+This is the goal of the matrix apearing in table \ref{table:soft_hard_deployment}.
+It shows where very software component is deployed.
 From the table \ref{table:soft_hard_deployment}, it is clear that there are hardware machines that have more than one component deployed and this is also the case for small devices such as sensor nodes within the \gls{wsn}.
@@
     Mote appliance control TCP server       & X    & X         & ~       & ~      \\
-    Gateway packet transformation process   & ~    & ~         & X       & ~      \\
+    \gls{ipv4}/\gls{ipv6} Gatewaying Process   & ~    & ~         & X       & ~      \\
     IPv6 over USB tunnel                    & ~    & ~         & X       & ~      \\
@@
 \end{table}
-\paragraph{}
-In the following sections, each software component will be described in more details showing why such a component is crucial and what are the architectural decisions made to be what it is.
-
-\section{Mote Sensor Data Collection TCP Client}
-The mote sensor data collection TCP client is nothing more but a \gls{tcp} client deployed in a mote that is attached to an appliance that reads sensor data periodically from the consumption sensor more known as the \gls{ct}  as showns in figure \ref{fig:current_transformer} that returns after a few transformations of the data the exact amount of power consumer in \gls{wh}.
-The original amount returned a voltage value in \gls{v} that is then transformed into the electric intensity in \gls{a}.
-To transform, the voltage into the actual current intensity we make use of the data provided by the \gls{ct} manufacturer.
-By making use of an experiment whose goals is to measure the slope and the constant of the voltage to intensity function.
-This function resulted in the following values:
+
+\paragraph{}
+In the following sections, each software component will be described in more details showing why such a component is crucial and what are the architectural decisions made to become what it is.
+
+\section{Mote Sensor Data Collection-TCP Client}
+\paragraph{}
+The mote sensor data collection TCP client is a simple \gls{tcp} client deployed in a mote that is equipped with a \gls{ct} attached to an appliance that reads sensory data periodically as shown in figure \ref{fig:current_transformer}.
+This data needs to be transformed using a set of transformation functions to get the consumption in \gls{wh}.
+The original value returned is a voltage value in \gls{v} that is then transformed into the current intensity in \gls{a} and with the use of second transformation function, we get the consumption.
 
@@
 \end{table}
-The transformation function is then: \\
+To transform the initial voltage into the current intensity, we use constant variables provided by the \gls{ct} manufacturer.
+But these value cannot be taken for granted.
+To measure them, we conduct an experiment to using the Voltmeter and Amperemeter to measure the slope, $s$, and the function constant, $i_0$, of the intensity transformation function.
+This function resulted in the following values:
+\begin{itemize}
+\item $s = 66.70$
+\item $u_0 = 0$
+\end{itemize}
+Therefore, the first transformation function is: \\
 \boldmath{$I = 66.7U_0$} \\
@@
 \paragraph{}
-Afterwards, as the voltage in Moroccan households is 220 \gls{v} we get the power using the following function:
+Afterwards, we assume the voltage in Moroccan households is 220\gls{v} we get the power using the following function:
 \\
-\boldmath{$ P = U_1 I $} \\
+\boldmath{$ P = 220I $} \\
 P: Power in \gls{w} \\
-I: Intensity in \glsreset{a}\gls{a} \\
-$U_1$: Electrical potential difference or Voltage in household in \glsreset{v}\gls{v}
-\paragraph{}
-P refers to the power, not the actual consumption of the appliance.
-To get the consumption we use an additional parameter which is the sampling rate.
-The sampling rate is defined as the number of sensor reading actions per hour.
+I: Intensity in \glsreset{a}\gls{a}
+\paragraph{}
+To get the consumption we use an the sampling rate.
+The sampling rate is as the number of sensory readings per hour.
 If one reads sensor data every second, then the sampling rate is 3600 samples per second.
 \paragraph{}
-To find the actual consumption, we need to consecutive sensor readings that we call $P_0$ and $P_1$ and use the following function:
+To find the consumption, we need to have consecutive sensor readings that we call $P_0$ and $P_1$ and use the following function:
 \\
@@
 S: Sampling rate \\
-The assumption taken is that if the power varies between two consecutive samples, the change was made linear throughout the interval of time between the two.
-\paragraph{}
-Once the consumption data in hand after this series of transformation, the next thing left to do is to send this data.
-To send it, the mote make use of the \gls{tcp}  client built on top of \gls{blip} package that implements the TCP/IP protocol stack.
-The system is made in a way that all the sensor data is sent to the gateway services that will decide to store it in a data store or send it to another system outside of the scope of this project.
-\section{ Mote appliance control TCP server}
-\paragraph{}
-The mote appliance control \glsreset{tcp}\gls{tcp} server is a component that is deployed in the sensor nodes of the \gls{wsn} to get commands from the outside world or at least outside of the \gls{wsn} in order control the appliance attached it.
-To control an appliance, one has access to two commands which are the ability to turn on and off a device.
-\paragraph{}
-Figure \ref{fig:appliance_control} depicts the architecture as well as the data flow of the command to turn on/ off an appliance.
-The TCP server hosted in the mote receives on/off commands from and external TCP client.
-The command is then extracted from the protocol of communication, and based on that, the PIN is known.
-the PIN information as well as the functionality is called as it is implemented in the MDA320CA that applies it on the pin and opens or closes the circuit depending on the command specified by the external TCP client.
+The assumption made is that if the power varies between two consecutive samples, the change was linear throughout the interval of time between the two sensory readings.
+\paragraph{}
+Once the consumption data in hand after this series of transformation, the next thing left to do is to communicate this data.
+To send it, the mote uses the \gls{tcp} client built on top of the \gls{blip} package that implements the TCP/IP protocol stack.
+Most important, the system is built in a way that forces all motes to communicate the sensor data gateway services by default, and it will be the gateway's responsibility to manage it (store or communicate it to an external system).
+
+\section{ Mote Appliance Control-TCP server}
+\paragraph{}
+The mote appliance control \glsreset{tcp}\gls{tcp} server is deployed in all the sensor nodes of the \gls{wsn} including the sink .
+The goal of this TCP server is be ready to receive commands from external node in the Internet in order to turn On or Off the appliance it oversees.
+To control an appliance.
+\paragraph{}
+Figure \ref{fig:appliance_control} depicts the architecture as well as the data flow of the command to turn On/Off an appliance.
+The TCP server hosted in the mote receives On/Off commands from an external TCP client.
+The command is then extracted from the IP packet, and based on the command sent, the PIN is identified.
+Afterward, the command is delegated the MDA320CA that executes it on the chosen PIN by opening (Off) or closing (On) the circuit who is feeding the appliance with electricity.
 
@@
 \paragraph{}
-The mote by itself does not have the necessary capabilities needed to control a mote.
-In order for a mote to have such capabilities, it needs to be equipped with an additional board, a Data Acquisition board also called a extension board that enable the mote to have more capabilities and therefore the ability to control the power flowing into the mote.
-The extension board used in the course of this project is the Crossbow MDA320CA as depicted in figure \ref{fig:mda320ca}.
-This board has pins which are electrical inputs and outputs that make it possible to have external analog sensors, digital sensors and also be able to power the appliance.
-But the powering of the appliances is not as straightforward as it seems especially that sensor nodes are very low-power devices that cannot provide enough energy to appliances such as a television or a fridge.
-\paragraph{}
-To make a mote control the appliance, the mote will control the power going in a very small circuit and this circuit is responsible for letting the power flow through the appliance from another source of energy which is naturally far more important than the one provided in the mote.
-In summary, the mote has a the extension board of type MDA320CA that has a small electric circuit attached to one of its digital pins that then controls the flow of electricity going to the appliance.
-By turning off a pin, the electricity stops flowing through the circuit and becomes open which stops the electricity to go to the appliance.
+The mote on itself does not have the necessary hardware capabilities to control an appliance.
+It needs to be equipped with a Data Acquisition board also known as Extension Board that enable the mote to have more capabilities and therefore the ability to control the power flowing into the appliance.
+The extension board used in this project is the Crossbow MDA320CA as depicted in figure \ref{fig:mda320ca}.
+This board has PINs which are electrical inputs and outputs that permit possible to have external analog sensors, digital sensors and also the possibility feeding electricity to the appliance.
+However, the power supply of the appliances from the mote is not straightforward.
+\paragraph{}
+To make a mote control the appliance, the mote will govern the supply of electricity by controlling a small circuit that is responsible for pipelining the power flowing to the appliance.
+In short, the mote is equipped with an extension board to which an electric circuit is attached and whose responsible for the flow of power going to the appliance.
+Thus, the turn Off command will open the circuit and the appliance will stop getting fed electricity.
 
@@
 \paragraph{} 
-This component that exist in one special mote in the \glsreset{wsn}\gls{wsn}.
-This mote is able to communicate with the rest of the \gls{wsn} using Zigbee and also with the external world using \gls{usb}.
-This make this mote a sink and a router because it has two interface cards.
+This component is deployed only the sink of the \glsreset{wsn}\gls{wsn}.
+This mote is able to communicate with the sensor nodes using Zigbee and with the gateway using \gls{usb}.
 The software component deployed in this sink is a packet forwarder.
-This means that once it receives an \gls{ipv6} packet that is not in the \gls{wsn}, it is forwarded through \gls{usb}.
-\paragraph{}
-The mote's sole goal is forwarding packets in and outside of the \gls{wsn} while other sensor node do not need to worry about how the packet will arrive to its destination.
-This mote is part of the components aiming to solve the heterogeneity problem stated in the problem statement.
-By making the sensor nodes talk to any device as if it is part of the \gls{wsn}, they do not really have to worry about how it will arrive or even how did the packet get to them in the first place.
+In fact, once it receives an \gls{ipv6} packet that is going outside the \gls{wsn}, it is forwarded through \gls{usb}.
+\paragraph{}
+The component's goal is to forward packets in and out of the \gls{wsn} while other sensor node do not need to worry about how the packet will be communicated.
+This mote as part of the components aiming to solve the heterogeneity problem, provides communication transparency.
 
@@
 \paragraph{}
-The gateway server machine is equipped with two interface cards: the Wi-Fi network interface card and the Zigbee network interface card.
-The Zigbee network interface card is the one on focus in this section.
-To have a device communicating using Zigbee, one sensor node within the \gls{wsn} was used for this purpose and the software component implemented in this sensor node is the one described in section 4.6.
-That said, the packet is received through \gls{usb} but not read and recognized as a packet by the TCP/IP protocol stack.
-To do so, a tunnel had to be created for the purpose of extracting packets from the \gls{usb} port.
-Another goal is to forward \gls{ipv6} packets through that same \gls{usb} and this can only be done by the use of an \gls{ipv6} over \gls{usb} tunnel.
-
-\section{Gateway Packet Transformation Process}
-The gateway packet transformation process is one of the main components that hide the heterogeneity from the rest of the system and the outside world.
-This component mainly deals with packet transformation which means that it extracts the data from the packet and create a completely new one that holds the data and forward it.
+The gateway server machine is equipped with two interface cards: the Wi-Fi and the Zigbee network interface cards.
+The Zigbee network interface card is of interest in this section.
+To have a device communicating using Zigbee, one sensor node within the \gls{wsn} was used for this purpose, and the software component implemented in this sensor node is the one described in section 4.6.
+Indeed, the packet is received through \gls{usb} but not read and recognized as a packet belonging to the TCP/IP protocol stack.
+To do so, a tunnel was used for the purpose of extracting packets from the \gls{usb} port.
+In addition, the tunnel was used to forward \gls{ipv6} packets through \gls{usb}.
+
+\section{\gls{ipv4}/\gls{ipv6} Gatewaying Process}
+The \gls{ipv4}/\gls{ipv6} Gatewaying Process is one of the components that hides the heterogeneity that exist in the system.
+This component mainly deals with transforming packets since it extracts the datagram from the packet and create a completely new one that holds the datagram originating from the first one and communicates it.
 This component is composed of three subcomponents:
@@
 \paragraph{}
-Figure \ref{fig:gateway_transformation} depicts the architecture of this component by showing the the different subcomponents that reside in this component and their interaction.
-It also shows the data flow within this component and what are the inputs and outputs of the system.
+Figure \ref{fig:gateway_transformation} depicts the architecture of this subsystem by showing the different subcomponents and their interactions.
+Moreover, it illustrates the data flow within this subsystem by showing the inputs and outputs of the subsystem.
 
@@
 \paragraph{}
-The subsections 4.8.1 - 4.8.3 will go more into details in explaining the three subcomponents of the system.
+The subsections 4.8.1 to 4.8.3 will detail and explaining the three subcomponents of the system.
 
@@
 \paragraph{}
-The packet sniffer subcomponent is the one responsible for sniffing and forwarding the packets to the component.
-It does not forward all the packets that are not going to the gateway machine.
-In other words, all the packets that are not meant to be received by the gateway server are sniffed and forwarded to the packet transformer that will continue the work.
+The packet sniffer subcomponent is the one responsible for sniffing and forwarding the packets to the packet transformer.
+It does not forward the packets whose recipient is the gateway machine.
+Specifically, all the packets to be routed by the gateway server are sniffed and forwarded to the packet transformer.
 \subsection{Packet Transformer}
 \paragraph{}
-Once the packet is received, the first thing this subcomponent does is check from where did it come, was it from the \gls{wsn} or the outside world.
-If it came from the outside world, the packet will only be examined if it is meant to go the \gls{wsn}.
-But an important and crucial problem presents itself:
-\paragraph{Problem}
-\underline{The outside world mostly uses \gls{ipv4} whereas the \gls{wsn} uses exclusively \gls{ipv6}} \\
-Two networks using different network layer technology cannot communicate even though \gls{ipv6} is nothing more but a new version of the very known \gls{ipv4} and this is because \gls{ipv6} has no backward compatibility with \gls{ipv4}.
+Once the packet is delegated to the packet transformer, it checks the source of the packet to differentiate between packets incoming from the \gls{wsn} and the packet going to the mote.
+Concerning the packet not originating from the \gls{wsn}, only those going to \gls{wsn} are processed.
+But a problem presents itself:
+\underline{The outside world mostly uses \gls{ipv4} whereas the \gls{wsn} uses exclusively \gls{ipv6}}
+\paragraph{}
+
+In general, two networks using different network layer technologies cannot communicate despite the fact that \gls{ipv6} is just a new version of the very known \gls{ipv4}, and this is due to the fact that \gls{ipv6} has no backward compatibility with \gls{ipv4}.
 \paragraph{Solution}
-To solve such a problem, we decide to provide all the nodes in the this system with both an \gls{ipv4} and \gls{ipv6} address.
-Some of the nodes have either \gls{ipv4} or \gls{ipv6} addresses, but for the technology missing we assign the missing one virtually.
-For example, a sensor node in the \gls{wsn} with a real \gls{ipv6} address equal to FEC0::1 will be assigned a virtual \gls{ipv4} address equal to 192.168.0.1 but he will not be aware of it.
-Only the Gateway packet transformation component will be aware of it and this is the mechanism used to transform the packets and hide the heterogeneity from the rest of the components of the system.
-\paragraph{}
-Now that a packet is received, it checks whether this component came from the \gls{wsn} or the outside world, depending on that, it extracts the data from the packet which can be a \gls{tcp} or \gls{udp} datagram.
-The source and destination addresses are extracted from the packet and translated to the other network technology.
-Afterwards, a new packet referring to the destination technology used is created by having as a source and destination addresses the translated addresses and as payload the the \gls{tcp} \gls{udp} datagram coming from the original packet.
+To solve this problem, we provided all the nodes with both an \gls{ipv4} and \gls{ipv6} address.
+The nodes have either \gls{ipv4} or \gls{ipv6} addresses, so the one missing is assigned to them virtually.
+For instance, a sensor node in the \gls{wsn} with a real \gls{ipv6} address equal to FEC0::1 will be assigned a virtual \gls{ipv4} address equal to 192.168.0.1 but he will not be aware of it.
+Only the Gateway packet transformation component will be aware of it and based on this, we transform the packets and hide the heterogeneity from the components of the system.
+\paragraph{}
+After the reception of a packet, the Packet Transformer checks whether this component originated from the \gls{wsn} or an external node, and depending on that, it extracts the data from the packet which can be a \gls{tcp} or \gls{udp} datagram.
+The source and destination addresses are extracted from the packet and translated.
+Afterwards, a new packet referring to the destination technology used is created by having as a source and destination addresses the translated IP addresses and as payload the \gls{tcp} \gls{udp} datagram extracted from the original packet.
 Finally, the packet is forwarded to the network.
 \subsection{Address Translator}
-The address translator is a very specialized subcomponent whose sole goal is to translate addresses between \gls{ipv4} and \gls{ipv6}.
+The address translator is a very specialized subcomponent whose sole goal is to translate addresses from \gls{ipv4} and \gls{ipv6} and vice versa.
 That is said, this component receives the address to translate and returns the translated one.
-Its concern is that the translation algorithm yields a bijective function which means that for any \gls{ipv4} address there is one and only one \gls{ipv6} address resulting from that source and vice versa.
+In fact, the translation function yields a bijective function.
+that is, for any \gls{ipv4} address there is one and only one \gls{ipv6} address resulting translation and vice versa.
 This statement compresses a set of statements which are the following:
@@
 \end{itemize}
-\paragraph{}
-The gateway packet transformation process is an important component responsible for dealing with the heterogeneity problem.
-It hides most of the heterogeneity in the system and make communication go independently of the data link layer or network layer technologies used.
+
+\paragraph{}
+The gateway packet transformation process is an important component responsible for hiding the heterogeneous nature of the system.
+It hides most of the heterogeneity in the system and makes communication happen independently of the data link layer or network layer technologies used.
 
@@
 \paragraph{}
-The sensor data storage process is a component small component within the system whose sole purpose is to store such data.
-It seen by the \gls{wsn} as a TCP server, where sensor nodes get connected to it as soon as they are turned ON.
-The sensor node starts sending packet periodically (depending on the sampling rate specified) and whenever the sensor data is received, it is stored into a \gls{rdbms} where the schema is created as shown in figure \ref{fig:erd}.
+The sensor data storage process is a small component within the system whose goal is to store sensory data.
+It is perceived by the \gls{wsn} as a TCP server, where sensor nodes get connected to as soon as they boot.
+The sensor node starts sending packet periodically and whenever the sensor data is received, it is stored into a \gls{rdbms} adopting the schema depicted in figure \ref{fig:erd}.
 
@@
 \section{HTTP Sensor Data Sender} 
-The HTTP sensor data sender is TCP server to which sensor nodes get connected to and start sending their data periodically.
-Once the data is received they are able to send data to the outside world periodically depending on the needs of the clients outside the system.
-This system counts on the deployment of an HTTP server at the level of outside the system client and send an HTTP request containing the sensor data received from the \gls{wsn}.
-\paragraph{}
-The sensor nodes are able to send this data by themselves directly to the outside as this is part of the two-way communication but for convenience, it was decided that the \gls{wsn} sends its data to the gateway and it is up to the gateway to send such data or store it in the \gls{rdbms}.
+The HTTP sensor data sender is TCP server to which sensor nodes get connected to, and start sending their data periodically.
+Once the data is received, this component communicates this data to the external node interested in receiving this data.
+This system relies on the deployment of an HTTP server at the level of external client and sends an HTTP request containing the sensor data received from the \gls{wsn}.
+\paragraph{}
+The sensor nodes communicate this data by themselves directly to the external nodes, as this is part of the two-way communication, but for convenience, it was decided that the \gls{wsn} sends its data to the gateway, and it is up to the gateway to communicate it or store it in the \gls{rdbms}.
 \section{Mobile Client}
-The mobile client is a component of the system but can be seen as an outside world client as it is here to receive sensor data and also send ON/OFF commands to the system.
+The mobile client is a component of the system but can be perceived as an external world client as it is here to receive sensor data and also send ON/OFF commands to the system.
 
@@
 In this chapter, the test-bed will be discussed from the implementation point of view.
-That is said, each component within the system will be given as much details as possible in order for the reader to open the code and understand what is going.
-Chapter 4 discussed the components from the architectural aspect but no implementation and technology decisions were discussed and the goal of this chapter is to explain these details.
+In fact, each component within the system will be given as much details as possible in order for the reader to open the code and understand how it is implemented.
+Chapter 4 discussed the components from the architectural aspect whereas in this chapter, implementation and technology decisions will be clarified.
 \section{Technology Enablers}
-The system's implementation was built using a set of software and hardware technologies and standards that were used throughout the course of this project to make the test-bed achieve its objectives.
-%\glsreset{wsn}
-\subsection{Technologies used in \gls{wsn}}
+The system was built using a set of software and hardware technologies and standards throughout the course of this project to make the test-bed achieve its objectives.
+\glsreset{wsn}
+\subsection{\gls{wsn}}
 
@@
 
-The \gls{wsn} was built using Crossbow professional kit as shown in figure \ref{fig:crossbow} which counts 7 Micaz sensor nodes of type MPR2600 as shows in figure \ref{fig:micaz}.
+The \gls{wsn} was built using Crossbow professional kit as depicted in figure \ref{fig:crossbow} which counts 7 Micaz sensor nodes of type MPR2600 as shows in figure \ref{fig:micaz}.
 \begin{figure}[htbp]
@@
 
-In addition, each mote is supplied with a sensor board of type Crossbow MTS400CA which contains temperature, humidity, barometric pressure, ambient light and infrared light sensors that are used to send different types of sensor data to the gateway.
+In addition, the motes are supplied with sensor boards of type Crossbow MTS400CA that contain temperature, humidity, barometric pressure, ambient light and infrared light sensors that are used to communicate different types of sensor data to the gateway.
 This board is shown in figure \ref{fig:mts400}.
@@
 Programs are implemented using \gls{nesc} which is the sole programming language used in the \gls{tinyos} environment.
-This programming language is event-oriented and its syntax is very influenced by C.
+This programming language is event-oriented and its syntax is similar to C.
 In addition, C libraries can be used in \gls{nesc} programs.
 
-\subsection{Technologies used in the Gateway}
+\subsection{Gateway}
 \paragraph{}
 The gateway's hardware machine is Dell Optiplex 9010 running a Linux environment more specifically the Ubuntu 13.04 distribution.
-The gateway has a set of technologies used to provide numerous services the system counts on.
-The programming languages used in the components hosted in this machine are C and \gls{j2se}.
-Netfilter is also used to control incoming and outgoing packets.
-Concerning the storage of sensor data MySQl is the \gls{rdbms} used.
-
-\subsection{Technologies used in the Mobile Client}
-\paragraph{}
-The mobile used is an Android smart phone with Android 2.3. The application developed is using \gls{j2se}, PHP for database connectivity.
+The gateway has a set of technologies used to provide numerous services the system relies on.
+The programming languages used in the components deployed in this machine are C and \gls{j2se}.
+Netfilter is also used to drop incoming packets.
+Concerning the storage of sensor data, MySQl is the \gls{rdbms} used.
+
+\subsection{Mobile Client}
+\paragraph{}
+The mobile used is a smart phone runnning Android 2.3. The application developed is using \gls{j2se} and PHP for database connectivity.
 
@@
 Section 4.4 and 4.5 were dedicated to explain the TCP Client and TCP server separately.
-In this section, one section will explain both of them because from the implementation point of view,  both of them are implemented in the \gls{tinyos} program.
+In this section, one section will explain both of them because from the implementation point of view,  both are implemented in the \gls{tinyos} program.
 \paragraph{}
 A typical \gls{tinyos} program has two parts: the components definition part and the module part.
-In the components definition part, the main components that will be used.
-These components refer to already implemented modules whose functionalities will be used in the in-development module.
-In the module part, we provide define the interfaces we will be using and the interfaces we will be providing if any \cite{ref14}.
-Back to definition part, after defining components we will be using, we then link the interfaces needed in the module to the components providing that interface.
-This is known as wiring.
+In the components definition part, the main components that will be used are declared.
+These components refer to already-implemented modules whose functionalities will be called in the in-development module.
+In the module part, we define the interfaces we will be using and those we will be providing.
+Back to definitions part, after defining components that will be used, we then link the interfaces needed in the module to the components providing those interfaces.
+This step is called wiring \cite{ref14}.
 \paragraph{}
@@
 \end{itemize}
-Some of the interface do not have all the functions provided and the not implemented function are implemented in the module. The following part will talk about each function that is to be implemented
+The interfaces one uses have some functions that are not implemented and should be written in the module. The following part discusses the main functions:
 \paragraph{After boot of the mote}
-The first thing done after the mote is booted is to initialize the TCP client and server.
-The Client is given a random port by using the rand function.
+The first thing that is done after the mote is booted is to initialize the TCP client and server.
+The Client is binded to a random port.
 The TCP server is binded to port 7.
-The client is set to connect to the Data store in the gateway by providing the \gls{ipv6} address of the gateway and port number of the server.
+The client gets connected to the Data store in the gateway by providing the \gls{ipv6} address of the gateway and port number of the server.
 Afterwards, the timer is set to be triggered every 10 seconds which means that the sampling rate will be 360 sensor readings per hour.
@@
 \paragraph{Timer fired}
-Once the timer is fired, this means it is time to read sensor data from the sensor.
-The command to read from the sensor is called which goes and reads data and once the data is read the event "read done" is triggered.
+Once the timer is fired, it is time to read sensor data from the sensor. The command to read from the sensor is called reads data and once the data is read the event "read done" is triggered.
 \paragraph{Read done}
 Once the data is read and ready, the next thing is to transform it and then send it.
-Concerning the transformation, a series of equations are applied as seen in section 4.4 which build the data value from the original value gotten and it is ready to be sent.
+Concerning the transformation, a series of equations are applied as seen in section 4.4 which builds the data value from the original value.
 A message is created containing two values: the mote NODE\_ID which is the identifier of the mote predefined in the installation of the program in the mote and the sensor data value.
 These two values are separated by a space, encapsulated within a \gls{tcp} datagram and sent.
-\paragraph{}
-During the installation of the software the \glsreset{blip}\gls{blip} package should specified as it is a dependency of the software system.
+The NODE\_ID is not necessary to identify the mote because the mote can be identified using its \gls{ipv6} address.
+\paragraph{}
+During the installation of the software, the \glsreset{blip}\gls{blip} package should specified as the program depends on it.
 \gls{blip} provides the TCP/IP protocol stack that implements \gls{6lowpan} standard by \gls{ietf}.
-In  the installation the NODE\_ID is specified.
+In  the installation, the NODE\_ID is specified.
 If NODE\_ID is 1, this will make the mote have an \gls{ipv6} of FEC0::1, as the NODE\_ID is appended to the predefined prefix of \gls{wsn} during the installation.
@@
 This component is provided as part of the \gls{blip} software package.
-It is \gls{tinyos} program whose goal is to forward packet back and forth between the \gls{wsn} and the outside.
-This mote is connected to the \gls{wsn} using Zigbee and to the outside world using \gls{usb}.
-This special mote has its \gls{ipv6} being FEC0::64.
+It is a \gls{tinyos} program whose goal is to forward packets back and forth between the \gls{wsn} and the external networks.
+This mote is connected to the \gls{wsn} using Zigbee, and to the outside world using \gls{usb}.
+This special mote's \gls{ipv6} address is FEC0::64.
 It receives \gls{ipv6} packets encapsulated in \gls{usb} frames, extracts them and encapsulates them in a Zigbee frame and vice versa.
 \paragraph{}
-During the installation of this \gls{tinyos} program into the mote sink, one should make sure that it is being installed in the mote sink as \gls{usb} is a necessity.
-The \gls{ipv6} address does not need to be provided as it is already predefined in the program.
+During the installation of this \gls{tinyos} program into the mote sink, one should make sure that it is installed in the mote sink as \gls{usb} is a necessity.
+The \gls{ipv6} address does not need to be provided because it is already predefined in the program configuration file.
 
@@
 This is a program that is installed in the gateway server machine.
-As the mote sink is connected to the gateway using \gls{usb}, the gateway will recognize it as a network interface card using \gls{ipv6} as a network technology.
-That is said, this component is part of the \gls{blip} software package and creates the network interface card module and starts extracting incoming \gls{ipv6} packets from the \gls{usb} frames and encapsulates outgoing \gls{ipv6} into \gls{usb} frames.
-\paragraph{}
-The installation and initialization of this component will be given more details in appendix A.
-\section{Gateway Packet Transformation Process}
-\paragraph{}
-This component is in charge of hiding the heterogeneity of the system from the rest of the the components of the system and also the outside world.
-It does that by intercepting all the packet moving and transforms them according to the standards used by the recipient.
-\paragraph{}
-The program works with Netfilter in parallel. This means that Netfilter should drop all incoming and this program will take care of dealing with incoming packets separately. 
-\paragraph{}
-To make use of this process, there is an initial configuration that should be done to all the components of the system in order for the system to work.
+sink gets connected to the gateway using \gls{usb}, the gateway will recognize it as a network interface card that uses \gls{ipv6} as a network technology.
+Moreover, this component is part of the \gls{blip} software package which the network interface card module and starts extracting incoming \gls{ipv6} packets from the \gls{usb} frames and encapsulates outgoing \gls{ipv6} into \gls{usb} frames.
+\paragraph{}
+The installation and initialization of this component will be detailed in appendix A.
+\section{\gls{ipv4}/\gls{ipv6} Gatewaying Process}
+\paragraph{}
+This component is in charge of hiding the heterogeneity of the system from the rest of the components of the system and also the outside world.
+It does that by intercepting all the packets and transforms the ones of interest to meet the standards used by the recipient.
+\paragraph{}
+To make use of this process, there is an initial configuration that should be done to all the components of the system in order for the system to work (explained in Appendix 1).
 The gateway has two network interface cards: the Wi-Fi interface and the Zigbee interface provided through the tunnel.
@@
 It is done by first doing and experiment which will help us gather data and then based on this data, an interpretation of this data will be done to see whether the system is shows an acceptable level of performance.
-\section{Experiment}
+\section{Experimentations}
 \paragraph{}
@@
 We believe that there will be an additional delay to the communication but we are hoping that it will not affect much the whole delay of the communication between different components of the system. 
-\section{Data Gathering}
+\section{Data Gathering and Analysis}
 \paragraph{}
@@
 The result is displayed in the table \ref{table:exp1}.
+\paragraph{}
+The data collected shows that the performance of the system is very good. Even though the delay in the \gls{wsn} is high, 
+ nothing can be done about that as even though the traffic in the \gls{wsn} is low, the delay is still high and this has to do with the computational time taken by the TCP/IP protocol protocol stack provided implemented by \glsreset{blip}\gls{blip} using \glsreset{6lowpan}\gls{6lowpan}.
+To reduce the delay, one should probably rethink the design of the \gls{blip} or find a way to omit it, but that is not a good idea since we want the \gls{wsn} to be part of \gls{iot}.
+ \begin{figure}[htbp]
+ \centering
+ \includegraphics[scale=1]{img/delay_jitter.JPG}
+ \caption{Delay and jitter evolution by Bandwidth}
+ \label{fig:delay_jitter}
+ \end{figure}
+
+ \paragraph{}
+ The good news in experiment 1 is that the priority of packets brings benefit to the network.
+Even though the traffic grows exponentially, the delay does not grow much and the even if the jitter follows to some extent the traffic's trend but still the average delay shows good result.
 
@@
 Depending on the machine's load.
-\section{Interpretation of Results}
-\subsection{Experiment 1}
-\paragraph{}
-The data collected in section 6.2 shows that the performance of the system is very good. Even though the delay in the \gls{wsn} is high, 
- nothing can be done about that as even though the traffic in the \gls{wsn} is low, the delay is still high and this has to do with the computational time taken by the TCP/IP protocol protocol stack provided implemented by \glsreset{blip}\gls{blip} using \glsreset{6lowpan}\gls{6lowpan}.
-To reduce the delay, one should probably rethink the design of the \gls{blip} or find a way to omit it, but that is not a good idea since we want the \gls{wsn} to be part of \gls{iot}.
- \paragraph{}
- The good news in experiment 1 is that the privatization of packets brings benefit to the network.
-Even though the traffic grows exponentially, the delay does not grow much and the even if the jitter follows to some extent the traffic's trend but still the average delay shows good result.
- \subsection{Experiment 2}
 \paragraph{}
@@
 Before conducting experiment 2, we were persuaded that there will be additional delay originating from this process and it could be responsible for the high delay seen in experiment 2 but the results shows otherwise and that the component is not to be put responsible for such important delay seen in experiment 1.
+
 \chapter{Conclusion}
@@
 
-\backmatter
-
-\addcontentsline{toc}{chapter}{Bibliography}
-\renewcommand{\bibname}{References}
+%\addcontentsline{toc}{chapter}{Index}
+\renewcommand{\bibname}{Index}
 \bibliographystyle{IEEEtran} 
@@
 
-\appendix
-\chapter{How to install the system}
+%\backmatter
+\begin{appendices}
+\chapter{System Installation}
 \section{Hardware needed}
@@
 You should be able to get sensor data periodically. The mobile client access for turning On and Off will be explained in the Appendix C.
+\end{appendices}
 

### 2013-07-07 doc/thesis/report.tex
Corrected entirely. Figures will be changes because they are not appropriate

--- 2013-07-02 doc/thesis/report.tex
@@
 \paragraph{}
-The information technology is advancing exponentially and is having bigger impact on our lives as time advances.
-The next revolution in computing is called the \gls{iot}.
-It is basically the internet we know nowadays except that physical objects (fridge, television, air conditioning...) will be the major players in this complex system.
-In this context, the project builds on this idea a test bed for smart home energy optimization.
-The project allows users to monitor their energy consumption via their smart phone and also the users can control their appliances by turning on and off appliances remotely.
-\paragraph{}
-The system is divided into three main components.
-Each of them is using different technologies such as \gls{tinyos}, \gls{nesc}, \gls{j2se}, \gls{http} and other technologies that will be discussed in the thesis report.
-\paragraph{}
-At this stage, the system allows monitoring, but turning on and off appliances is not yet there mainly because of the absence of the equipment but the communication for that purpose is there.
+Information Technology is advancing exponentially and largely impacting our daily lives.
+The current evolution in computing is the \gls{iot}.
+It is smoothly evolving from an Internet of people to and Internet of Things.
+By 2020, it is expected to have 50 billion "Things" connected to the Internet.
+\paragraph{}
+However such a migration induces a strong level of complexity when handling interoperability between heterogeneous Things, i.g. Wireless Sensors, RFID, and mobile devices.
+\paragraph{}
+In this thesis, we focus on the  integration of wireless sensors into \gls{iot}.
+This is q prologue for the future integration of \gls{sg} into \gls{iot}.
+In this work, we deployed a smart building testbed for energy optimization, and for integration into \gls{sg}\glsreset{sg}.
+\paragraph{}
+The testbed built provides IPv6 gatewaying, remote control of wireless sensor nodes which is built on top of \gls{6lowpan}\glsreset{6lowpan}, and mobile access to the system.
+\paragraph{}
+The system is using different technologies such as \gls{tinyos}, \gls{nesc}, \gls{j2se} and  \gls{http}.
 
@@
 \paragraph{}
-The global electrical grid one the bases that gives us this quality of life that most people around the world enjoy.
-The global electrical grid is sometimes called the largest machine ever made by man.
-\cite{ref1}.
-It a widespread machine, distributed but interconnected system that provides power to households, companies and other organizations and therefore the world's economy is counting on it.
-The electrical grid knows numerous problems such as blackouts, current instabilities that costs billions of dollars yearly to both the utilities, utilities and insurances.
-The grid suffers from the problem of peak hours where energy consumption where in many case the one single utility cannot provide such energy.
-\paragraph{}
-The \gls{sg} comes to provide solutions to the problems that suffers from the traditional electrical grid.
-\gls{sg} is an intelligent, auto-balancing, and self-monitoring electrical grid \cite{ref2} and is and whose energy sources are distributed between utilities and consumers.
-The nature of the \gls{sg} as having distributed energy sources reduces numerous losses that arises in the traditional grid such as the energy loss in the energy transmission lines and also in the overload of the electrical lines.
-\paragraph{}
-Another revolution is about to begin and it is taking place in the information technology and is called \glsreset{iot}\gls{iot}.
-\gls{iot} is the era of autonomous machine able to produce and consume information and being able to act autonomously without the guidance of humans to performs their tasks.
-Cisco predicts that the number of devices to be connected in 2020 will break the 50 billions devices \cite{ref3} whereas the world population will be around 7.6 billion.
-That make 6.58 devices connected for each person in the world whereas the number of devices connected in 2010 12.5 billion.
-This clearly shows that the shift to the \gls{iot} has already started.
-\paragraph{}
-\glsreset{sg}\gls{sg} can and will benefit from the \gls{iot}.
-The integration of the \gls{sg} and and \gls{iot} makes this study and thesis of high importance in order to solve as much as possible the problems of the traditional grid but at the same bring a new dimension to the power grid by benefiting from the numerous possibilities offered by the \gls{iot}.
+The global electrical grid is the basis that gives us this quality of life that most people around the world enjoy.
+The global electrical grid is sometimes called the largest machine ever made by man \cite{ref1}.
+It is a widespread and distributed machine that provides power to households, companies and organizations.
+Therefore the whole world's economy depends on it.
+The electrical grid knows numerous problems such as blackouts, and current instabilities that cost billions of dollars yearly to both the utilities and insurances.
+The grid suffers from the problem of peak hours where energy consumption is at its highest level and that one single utility cannot provide.
+\paragraph{}
+The \glsreset{sg}\gls{sg} comes as a solution to the problems of the traditional electrical grid.
+\gls{sg} is an intelligent, auto-balancing, self-monitoring, and self-healing electrical grid \cite{ref2} and whose energy sources are distributed between utilities and consumers.
+The nature of the \gls{sg} as having distributed energy sources, reduces numerous losses that arise from the long distance transmission lines, and also from the stress and the overload of the electrical lines.
+\paragraph{}
+Another revolution began, and it is taking place in the Information Technology.
+It is called the \glsreset{iot}\gls{iot}.
+\gls{iot} is the era of autonomous machine being able to produce and consume information, and being able to act autonomously without the guidance of humans, to perform their tasks.
+Cisco predicts that the number of devices to be connected to the Internet in 2020 will break the 50 billions devices from 500 million in 2003.
+That would make 6.58 devices connected for each person in the world whereas that same ratio was 0.07 17 years earlier \cite{ref3}.
+This clearly shows that the shift to the \gls{iot} will be exponential.
+\paragraph{}
+\glsreset{sg}\gls{sg} should benefit from the \gls{iot}.
+The integration of the \gls{sg} into the \gls{iot} constitutes the main topic of this study and goes along with increasing interest recently raised in both academia and industry, on the inherent correspondence between \gls{sg} and \gls{iot}.
+This  correspondence stems mainly from the need to disseminate information between the \gls{sg} components in order to make optimal decisions about energy use.
+This information dissemination could only be optimal when deploying ubiquitous sensoring and metering, and using the available network infrastructure, i.e.
+the Internet.
 
@@
 This project has been done in order to investigate ways the \gls{iot} can benefit the \gls{sg}.
-The smart grid as a set of philosophical ideas counts on a set of novel functionalities.
-Once of these ideas is the two-way communication that will make possible the utility to communicate with consumers and vice versa.
-This concept is possible with the Internet Of Things as devices in \gls{iot} communicate between them by producing and consuming information.
-Numerous ideas are based on this two-way communication such as the \gls{dr} which offers a set of services for both the power provider and the consumer.
-\paragraph{}
-The \glsreset{iot}\gls{iot} counts on different technologies such as \gls{wsn}, \gls{rfid}, \gls{ipv6} and other technologies which make the Internet of things a very heterogeneous and complex system \cite{ref4}.
-One of the challenges in \gls{iot} is to deals with such heterogeneity and take the most out of it in order to make the communication as smooth and robust as possible.
-This heterogeneity should be made transparent to any two parties trying to communicate in \gls{iot}.
+The Smart Grid needs a set of novel functionalities.
+One of these ideas is the two-way communication that will make it possible for the utility to communicate with consumers and vice versa.
+This is possible in the Internet Of Things as devices in \gls{iot} communicate by producing and consuming information.
+Numerous ideas assume the existence of the two-way communication such as the \gls{dr} which offers a set of services for both the power provider and the consumer.
+\paragraph{}
+The \glsreset{iot}\gls{iot} uses different technologies such as \gls{wsn}, \gls{rfid}, \gls{ipv6} and \gls{6lowpan}\glsreset{6lowpan} which makes the Internet of Things a very heterogeneous and complex system \cite{ref4}.
+One of the challenges in \gls{iot} is to deal with this heterogeneity in order to make the communication smooth and transparent to the parties involved.
+This heterogeneity should become invisible to any two devices willing to communicate in the \gls{iot}.
 
@@
 \paragraph{}
-The \glsreset{sg} \gls{sg} counts on the two-way communication to build to a whole communication and protocol stack namely the \glsreset{dr}\gls{dr} and other services built on top of the \gls{dr}.
-From the consumer's side or more precisely the household's side, there are smart home energy management software in the market nowadays but none of these products offers integration with \gls{sg}.
-To have such an integration with \gls{sg}, there is a need to first build this two-way communication and make it possible for any appliance to report its real-time energy consumption and also to be able to be controlled remotely with this two-way communication.
-Such infrastructure is missing for the time being and therefore the integration of the smart home within the smart grid is not possible.
-Also, in order for \gls{sg} to be integrated into the \gls{iot}, all devices should be able to communicate in the two directions.
-\paragraph{}
-As \gls{sg} counts on numerous technologies to achieve its main goals.
-This usage of different technologies imposes another problem which is the heterogeneity of the system.
+The \glsreset{sg} \gls{sg} relies on the two-way communication to build a whole communication framework and a protocol stack namely the \glsreset{dr}\gls{dr} and other services built on top of this last.
+From the consumer's/household's perspective or more precisely the household's side,there are smart home energy management systems in the market nowadays, but none of them offers an integration with \gls{sg}.
+To have such an integration with \gls{sg}, there is a need for the two-way communication, in order to make it possible for any appliance to report its real-time energy consumption, and to be able to be controlled remotely.
+Such pillar is missing for the time being, and therefore the integration of the smart home within the smart grid is not possible.
+Also, in order for \gls{sg} to be integrated into the \gls{iot}, all devices should be able to communicate in the two ways.
+\paragraph{}
+\gls{sg} counts on various technologies to achieve its main goals.
+This use of different technologies comes with a problem which is the heterogeneous nature of the system.
 As numerous components count on different technologies, there is a need to interface them in order to make this heterogeneity transparent to the all components of the system.
-This means that there is a need to build a software within the whole system that would constitute an important part of it and will take care of making the communication and processes technology-transparent and give the possibility to all components of the system to work hand-in-hand in order their common goals independently of the data link layer technology (Ethernet, Wifi or Zigbee) they are counting on, the network layer technology (\gls{ipv4} or \gls{ipv6}).
+As a matter of fact, a software should be built within the system whose main goal is to take care of making the communication technology-transparent.
+This component will give the possibility to all components of the system to work hand-in-hand in order to achieve their goals independently of the data link layer (Ethernet, Wifi or Zigbee) or the network layer technology (\gls{ipv4} or \gls{ipv6}) they are using.
 
@@
 The goal of the study is to solve the problems introduced in the problem statement section.
-This means creating the two-way communication, and also solving the heterogeneity that arises from integrating smart homes with the smart grid and making each components in the smart home as part of the Internet of things by making it communicate with any device in the Internet and having an \gls{ipv6} address that is built by making use of \gls{6lowpan} whose goal is to provide the TCP/IP to the smallest devices and with the least computational and energy rich.
-\paragraph{}
-The first purpose of this study is to create a testbed having the two-way communication and also being able to make use of it in order to build richer applications based on this possibility.
-To do it, there was a need to first introduce \gls{6lowpan} within every device in the \gls{wsn}.
-This would make every sensor node also known as mote identified and ready to communicate and be part of the \gls{iot}.
-This communication will go in the two directions and based on that the two-way communication will have to be built.
-\paragraph{}
-The second purpose of this study is to solve the heterogeneity problem that arises with the integration of smart homes within the smart grid and also when making the whole system as part of the \glsreset{iot}\gls{iot}.
-Solving such a problem means having an additional component that will track all communication packets and change them to fit the needs of the destination component.
-This will make of this additional component a middleman and therefore make the heterogeneity problem disappear.
+This means creating the two-way communication, and solving the heterogeneity that arises from integrating smart homes with the smart grid.
+In fact, each components in the smart home will become part of the Internet of Things by making it communicate with any device in the Internet by having an \gls{ipv6} address using \gls{6lowpan} whose goal is to provide the TCP/IP to the smallest devices and also those with the least computational resources.
+\paragraph{}
+The first purpose of this study is to create a testbed having the two-way communication and also being able to make use of it in order to build richer applications based on it.
+To do so, \gls{6lowpan} would had to be integrated within every device in the \gls{wsn}.
+This would make every sensor node (mote) identifiable in the network, ready to communicate and be part of the \gls{iot}.
+This communication will go in both directions and based on \gls{6lowpan}; the two-way communication will be built.
+\paragraph{}
+The second purpose of this study is to solve the heterogeneity problem that arises from the integration of smart homes into the grid and also when making the whole system part of the \glsreset{iot}\gls{iot}.
+Solving such a problem means having an extra component that will track all  packets and change them to fit the standards of the targeted component.
+This will make of this additional component a middleman and that would make the heterogeneity problem disappear.
 
@@
 \paragraph{}
-The thesis aims to solves the two problems presented in section 1.3.
-To do so, a variety of tools and technologies were used for this purpose.The two problems were not solved one independently of the other but even though the problems seem separate one cannot solve without having to solve the other.
-In other words, after the creation of the two-way communication in the \glsreset{wsn}\gls{wsn} which supported \gls{ipv6} over \gls{6lowpan}, other network only supported \gls{ipv4} and therefore could not communicate, and this was due to the heterogeneity problem.
-This is why the two-way communication was created and then the heterogeneity problem had to be solved in order to see everything working.
-The system is divided into four components:
+This study aims at solving the two problems presented in section 1.3.
+To do so, a variety of tools and technologies were used for this purpose.The two problems could not be solved one independently of the other.
+Even though the problems seems to be separate, one cannot solve  the first problem without having to solve the second.
+In other words, after the creation of the two-way communication in the \glsreset{wsn}\gls{wsn} which adopts \gls{ipv6} over \gls{6lowpan}, other networks can only support \gls{ipv4} and therefore the communication would not be possible, and this is due to the heterogeneous nature of the system.
+This is why the two-way communication was created, but only after solving the heterogeneity problem, everything worked.
+The system is divided into four subsystems:
 \begin{itemize}
 \item The fully \gls{ipv6} supported \gls{wsn}
-\item The \gls{tinyos} program for sensor transmission and appliance request handler
 \item The heterogeneity handler and middleman gateway server
+\item The middle-ware
 \item The mobile client
@@
 \paragraph{}
-Each of these four component exist to solve part of the problem.
-Each of these components will be discussed extensively in the Chapter 4 from the architectural aspect of the component and in Chapter 5 from the implementation level of the system.
-\paragraph{}
-It is to be noted that some of the concepts will be discussed from the architectural point in Chapter 4 but not from an implementation point in Chapter 5.
-The sole reason behind is the fact that some of the equipment required for this thesis did not arrive on time and therefore the parts in the implementation requiring such equipment was omitted to meet the deadlines.
+Each of these four components is there to solve part of the problem.
+Each of these subsystems will be discussed extensively in Chapter 4 from the architectural aspect of the component, and in Chapter 5 from a deployment point of view.
+\paragraph{}
+It is important to note that some of the components will be discussed from the architectural point in Chapter 4 but not from an implementation aspect in Chapter 5.
+The sole reason behind this is the fact that some of the equipment required for this thesis were not on time and therefore the parts in the implementation requiring such equipment were omitted in order to meet the deadlines.
 \section{Outcome of the Study}
@@
 The outcome of the thesis is to design and implement a test-bed that is able to show the two-way communication in action.
-It will also have to show concepts of the \glsreset{iot}\gls{iot} which are that every node in the system can communicate autonomously and with any other node in the Internet.
-Another \gls{iot} concept is the ability of any node in the \gls{iot} to consume and produce information that is communicated through the network.
-\paragraph{}
-The autonomous production of information is shown by the fact that every sensor node in the \gls{wsn} reads sensory data from its sensors that are attached to appliance in order to sense their consumption and this information is produced and transmitted periodically.
-The consumption of data by is shown by external nodes in the \gls{iot} sending request to the sensor node in the \gls{wsn} in order to control and appliance's status i.e.
-turn on the television.
-This will prove the consumption of information by the nodes in the \gls{iot}.
-\paragraph{}
-The test-bed should also show that communication between any two nodes in the \gls{iot} is successful, robust and reliable and that the heterogeneity of the system is made transparent to the nodes in the system by the creation of the software tool residing in the gateway which is between the \gls{iot} and the outside world.
-This software tool is seen as a heterogeneity handler and also a communication middleman that translates the communication to the appropriate technology needed by the recipient.
-\paragraph{}
-The testbed will able to show the two-way communication in action and will also show that the heterogeneity does not form a barrier by its successful handling.
+It will also have to show its integration into the \glsreset{iot}\gls{iot} by showing that every node in the system can communicate autonomously and with any other node in the Internet.
+Another \gls{iot} concept is the ability of any node in the \gls{wsn} to consume and produce information that is communicated through the network.
+\paragraph{}
+The autonomous production of information is shown by the fact that every sensor node in the \gls{wsn} reads sensory data from its sensors, that are attached to appliance, in order to sense their consumption.
+This information about consumption is produced and communicated periodically.
+The consumption of data is proved by showing the execution of the commands communicated through the network by external nodes in the \gls{iot}.
+These commands request the sensor node to turn On or Off an appliance.
+\paragraph{}
+The test-bed should also show that communication between any two nodes in the \gls{iot} is successful and reliable, and that the heterogeneity of the system is transparent to the nodes in the system.
+This was achieved by creating a software tool deployed in the gateway which is between the \gls{wsn} and the outside world.
+This software tool is perceived as a heterogeneity handler, and also a communication middleman that translates the communication to the appropriate technology needed by the recipient.
+\paragraph{}
+The testbed will be able to show the two-way communication in action and will also show that the heterogeneity is not a barrier.
 
@@
 \paragraph{}
-The thesis is is divided into eight chapters. Each of the chapter approaches thesis from a different aspect. 
+The thesis is divided into eight chapters. Each of the chapter approaches thesis from a different aspect. 
 \paragraph{Chapter 1}
-It is an introductory chapter that gives a flavor of the project that is to be studied in this thesis and also introduces the problem and the solution proposed in this topic.
-This chapter also puts the reader into perspective by introducing some of the main concepts related to the \glsreset{sg}\gls{sg} and the \glsreset{iot}\gls{iot}.
+It is an introductory chapter that gives a flavor of the project that is  studied in this thesis and also introduces the reader to the problem and the solution proposed in this context.
+This chapter also puts into perspective by introducing some of the main concepts related to the \glsreset{sg}\gls{sg} and the \glsreset{iot}\gls{iot}.
 \paragraph{Chapter 2}
 It presents the background of the work by revising the literature related to the topics of \gls{iot} and \gls{sg}.
-It also discussed related work and how it was tackled in different research papers, how does it relate to the work done in this thesis and also how does it differ.
+It also discusses related work and how it was tackled in different research papers, how does it relate to the work done in this thesis and also how does it differ.
+Finally, its presents the commonalities ad differences between this thesis and the work previously done.
 \paragraph{Chapter 3}
-It is dedicated to the \glsreset{iot}\gls{iot}.
-It presents the topic, talks about the importance of \gls{iot}, its impact on the world, the actual Internet, the future of Internet and also how will \gls{sg} benefit from it.
+This part is dedicated to \gls{iot}, its importance and its impact on the world. Is also presents the actual state of the Internet and makes predictions about its future.
 \paragraph{Chapter 4}
-It will present the architecture of the system and the test-bed. It will discuss the system from different aspects, and how were the two problems solved from the architectural aspect.
+It presents the architecture of the test-bed. It also discusses the system from different perspectives, and how the problems stated in section 1.3 were solved from the architectural aspect.
 \paragraph{Chapter 5}
-This chapter will mostly discuss the implementation system.
-It will start by presenting the hardware used in this project, then move to the technology enablers.
-Each component of the system will be discussed extensively and move to the integration of the components to form the test-bed.
+This chapter mostly discusses the system deployment.
+It starts by presenting the hardware used in this project, then moves to the technology enablers.
+Afterwards, each component of the system is discussed extensively and moves along with the integration to form the test-bed.
 \paragraph{Chapter 6}
-It will start first by presenting the findings then move to explain the experiment that was done to measure the performance of the system.
-The data gathering section will show how and what data was gathered and then move to analyzing and interpreting the data.
+It starts first by presenting the findings then move to explain the experiments that were conducted. Afterwards, the data gathered is shown, analyzed and interpreted.
 \paragraph{Chapter 7, 8}
-These are the last two chapters of the thesis that will conclude the thesis, provide directions about future work and recommendations to the reader what parts of the system could be ameliorated and how it can be achieved.
+These are concluding chapters of this study.
+They provide directions about future work, recommendations to the reader about what parts of the system could be improved and how these improvements could be achieved.
+
 \chapter{Literature Review}
@@
 \paragraph{}
-The actual electrical grid is based on a technology that is at least one century old \cite{ref5}.
-As a result, there is a need to change as to meet to 21st century requirement such as having more green energy in order to reduce the pollution caused mainly by the use of fossil energy.
-As a result, research has introduced a new technology called \glsreset{sg}\gls{sg}.
-Although it is still a research project and numerous issues related to \gls{sg} are left unanswered, nations across the globe started to deploy \gls{sg} project.
-It is also noted that in order for the grid to become more reliable, secure and efficient, a bi-directional information and communication infrastructure should be embedded within the grid \cite{ref5}.
-\paragraph{}
-The smart grid brings around new concepts that will achieve efficiency, reliable and security to the power grid.
-The first one is the introduction of the \gls{dg} by exploiting \gls{res} \cite{ref6} and this reduce the losses of energy in the transmission.
-This distribution by itself enables the power grid this to move from the one way energy distribution to a two-way energy distribution where the utility will not be the sole energy producer but households can now produce and even sell the extra power.
-This is known nowadays as the feed-in-tariff which is a policy that allows the consumers to become producer and sell the excess power to the grid \cite{ref7}.
-This will allow the power grid to provide power to consumer in a "free market" fashion which means that it will obey to the laws of supply and demand \cite{ref7}.\glsreset{iot} \glsreset{sg}
+The actual electrical grid is based on a technology that is at least one century old and its weaknesses are becoming more and more costly \cite{ref5}.
+Therefore, there is a need to change it to meet to 21st century requirements such as having more green energy in order to reduce pollution caused mainly by the use of fossil energy.
+As a matter of fact, research introduced us to a new technology: the \glsreset{sg}\gls{sg}.
+Although it is still a research project that has not reached the stage of maturity yet, nations across the globe started to deploy \gls{sg} projects.
+The \gls{sg} is relying on Information Technology to solve many of the problems existing in the traditional grid.
+Consequently, in order for the grid to become more reliable, secure and efficient, a bi-directional information and communication infrastructure should be embedded within the grid \cite{ref5}.
+\paragraph{}
+The smart grid brings new concepts that will make it achieve greater efficiency, reliability and security compared to the traditional grid.
+The first one is the introduction of the \gls{dg} by exploiting \gls{res} \cite{ref6} and this reduces the losses of energy in the transmission.
+This distribution by itself enables the power grid to move from the one way energy distribution to a two-way energy distribution where the utility will not be the sole energy producer.
+Households can now produce and sell the surplus of electricity.
+This is known nowadays as the feed-in-tariff which is a policy that allows the consumers to become producers and sell the excess power to the grid \cite{ref7}.
+This will allow the power grid to provide electricity to consumers in a "free market" fashion which means that it will obey the laws of supply and demand \cite{ref7}.\glsreset{iot} \glsreset{sg}
 \section{\gls{iot} and \gls{sg}}
 \paragraph{}
-The \glsreset{iot}\gls{iot} has known high acceptance by the players in the Internet.
-Nowadays, it is estimated that 12.5 billion devices are connected to the Internet and that this trend will continue to read 50 billion connected devices by the end of the decade \cite{ref3}.
-\gls{iot} brings several benefits to the actual Internet and these benefits will pull \gls{sg} to be part of it.
-Consequently, research has taken this direction which is to integrate the \gls{sg} within the \gls{iot} as most aspects of \gls{sg} will be enhanced by \gls{iot} \cite{ref8}.
-\paragraph{}
-\gls{iot} counts on a set of technologies which are \glsreset{rfid}\gls{rfid}, \gls{m2m} \cite{ref9}, public communication networks such as 2G/3G mobile networks, power lines, WI-FI, Zigbee and other technologies \cite{ref8}.
-The smart grid's architecture based on \gls{iot} will be divided into three layers:
+The \glsreset{iot}\gls{iot} has been highly accepted by the players in the Internet as its adoption is starting smoothly.
+Nowadays, it is estimated that 12.5 billion devices are connected to the Internet and that this trend will continue to reach 50 billion devices by the end of the decade \cite{ref3}.
+\gls{iot} brings several benefits to the actual Internet and will pull the \gls{sg} to make it part of it.
+Consequently, research has taken the direction of integrating the \gls{sg} within the \gls{iot} as most of the aspects of this last favor  the \gls{sg} \cite{ref8}.
+\paragraph{}
+\gls{iot} counts on a set of technologies which are \glsreset{rfid}\gls{rfid}, \gls{m2m} \cite{ref9}, public communication networks such as 2G/3G mobile networks, power line communication, WI-FI and Zigbee\cite{ref8}.
+The smart grid's architecture based on \gls{iot} is divided into three layers:
 \begin{itemize}
-\item Perception layer: Devices will autonomously perform their tasks such as sensing and executing orders.
-\item Communication layer: The communication infrastructure is defined here and issues related to how to communicate information is handled.
-\item Application layer: Applications can be built to make use of the underlying data.
+\item Perception layer: Devices autonomously perform their tasks such as sensing and executing orders.
+\item Communication layer: The communication details are clarified here. It hides the communication issues and provide an easy for upper layers to send and receive information.
+\item Application layer: Applications are built to exploit the data received from the underlying layers.
 \end{itemize}
 \paragraph{}
-Some of the research is discussing the need to replace Zigbee by Wi-Fi as the Wi-Fi offers higher bandwidth, non-line transmission ability, large-scale data collection and is highly cost-effective \cite{ref10} but still Zigbee has the sole advantage of very low energy consumption which is something required in \gls{sg} and \gls{iot}, because in systems that are installed far away from any maintenance team such as smart grid monitoring and self-healing devices, one cannot manage to move for the sole purpose of providing energy often.
+Research discusses the benefits to replace Zigbee by Wi-Fi as this last offers higher bandwidth, non-line transmission ability, large-scale data collection and is highly cost-effective \cite{ref10}.
+Still, Zigbee has the exclusive advantage of consuming the least energy of the communication technologies that exist nowadays which is something required in \gls{sg} and \gls{iot}, especially for systems that are installed far away from any maintenance team such as smart grid monitoring and self-healing devices where one cannot manage to move for the sole purpose of charging these devices.
 \paragraph{}
 From the architectural point of view, integrating \gls{sg} within \gls{iot} means having to address heterogeneity issues.
-An \gls{iot} gateway system that is based on Zigbee and GPRS protocols \cite{ref11}.
-This enables then to deal partly with the heterogeneity problem and therefore enable the \gls{wsn} to communicate with the mobile telecommunication network.
-Another solution to the heterogeneity problem is proposed where the solution is to present a new, light-weight web service transport protocol called Lean Transport Protocol (LTP) that will allow transparent exchange of web service messages between all kinds of devices \cite{ref12}.
-This protocol is platform-independent, low energy consumer.
-Some researchers claim that the heterogeneity is mainly coming from the fact that there are different \gls{wsn} devices that do not necessarily use the same standards and protocols.
-They propose to move all the \glspl{wsn} to all-IP networks as this would remove most of the heterogeneity but there are already deployed legacy \glspl{wsn} \cite{ref13}.
-They sketch an architecture capable of converting all the \glspl{wsn}, new and legacy, to support \gls{ipv6}.\glsreset{sg}
-\section{The \gls{sg} Two-Way Communication}
-\paragraph{}
+An \gls{iot} gateway system based on Zigbee and GPRS protocols helps to deal partly with the heterogeneity problem and therefore enable the \gls{wsn} to communicate with the mobile telecommunication network \cite{ref11}.
+Another solution to the heterogeneity problem is proposed with a new, light-weight web service transport protocol called Lean Transport Protocol (LTP) that will allow transparent exchange of web service messages between all kinds of devices.
+This protocol is platform-independent, low energy communication \cite{ref12}.
+Some researchers claim that the major source of heterogeneity the fact that there are different types of \gls{wsn} devices (Micaz, Mica2, Telosb...) that do not use the same standards and protocols.
+They propose to move the communication in \glspl{wsn} to an "all-IP" as this would remove most of the heterogeneity but as there are already deployed legacy \glspl{wsn}, the task is complicated to conduct.
+Fortunately, the researchers sketch an architecture capable of converting all the \glspl{wsn}, new and legacy, to support \gls{ipv6} \cite{ref13}.\glsreset{sg}
 
 \chapter{Internet Of Things}
-
-\section{The actual Internet}
+\section{The Actual Internet}
+\paragraph{}
+The \glsreset{iot}\gls{iot} refers to the evolution of the actual Internet.The term \gls{iot} was used for the first time by Kevin Ashton in 1999 \cite{ref17}.
+The human beings will represent in the future a very small minority in the Internet and as it will be rules by autonomous machines.
+\gls{iot} will become the biggest machine ever created by human beings as the smart grid will be part of it.
 \section{Introduction to \gls{iot}}
+\paragraph{}
+The \glsreset{iot}\gls{iot} is a network mostly ruled my machines, from the most powerful supercomputer and mainframes to the smallest devices such as sensor nodes.
+All will have one common point, being all connected to the Internet and able to communicate, produce and consume information.
+The \gls{iot} is changing the world of computing.
+New paradigms, architectures, frameworks and prototypes are built and refined everyday to support it \cite{ref15}.
+New technologies such as \glsreset{rfid}\gls{rfid}, \glsreset{m2m}\gls{m2m}, \glsreset{ipv6}\gls{ipv6}, \glsreset{6lowpan}\gls{6lowpan} and \gls{ims} have come to place to establish the infrastructure for \gls{iot}.
+The whole world is now being connected including humans and objects \cite{ref18}.
+\paragraph{}
+The \glsreset{iot}\gls{iot} brings numerous advantages but many issues will have to solved before this network explodes to connecting trillions.
+The first issue is the interconnection of all the devices into the \gls{iot}.
+What hardware will support such an ultra high traffic network with an ultra high performance and availability requirements.
+Will ubiquitous computing or pervasive computing have architectures that deal with such infrastructure?
+\paragraph{}
+To better understand \gls{iot},The following are the three steps in its flow of information:
+\begin{enumerate}
+\item The object is identified by making use of \gls{rfid}. In addition, the objects should be recognized by sensing the environment and this information should be converted to electronic format
+\item  The information is transported to the information consumer.
+\item The information is processed intelligently \cite{ref17}.
+\end{enumerate}
+
 \section{Importance of \gls{iot}}
-\section{The Impact of \gls{iot}}
+\paragraph{}
+The actual Internet is evolving into the \glsreset{iot}\gls{iot} because it opens possibilities that are not in the actual Internet.
+The ability of Things to produce and consume information is a requirement.
+New high-level communication languages such as \gls{owl} were invented and studied extensively to enable objects to communicate freely without the need to interact with humans.
+\paragraph{}
+The world's requirements from the Internet are growing exponentially.
+Such new requirement are the accurate identification of Things, and the ability to get real-time information from devices throughout the globe.
+\paragraph{}
+
 \section{The Benefits of \gls{iot}}
+\paragraph{}
+\gls{iot} is armed with solutions to the new requirements.
+In the Internet of Things, objects will work and communicate freely.
+Any device will be able to send its sensory information, and consume data.
+From this simple idea arises a whole new set of applications that will benefit the world.
+Smart grids exploit this idea and make it part of its requirement in order to solve the problem in the electrical grid\cite{ref8}.
+More and more applications based on \gls{iot} are now being developed.
+Ideas such as smart farming, smart transportation, smart homes, smart health, smart postal and smart refrigerators are only a small subset of a myriad of applications that will come to life with \gls{iot} \cite{ref19}.
+\paragraph{}
+The Internet of Things will be able to help the world's community solve and minimize the impact of hazards it faces.
+With the deployment of applications which will be part of \gls{iot}, we can predict natural disasters by sending real-time sensory data about geological, volcanic, seismic, meteorological and spacial variables and help communities avoid losing their lives.
+We can also monitor water scarcity on a real-time basis.Finally, each one's health can be monitored in real-time by incorporating microscopic sensors that are able to communicate this Information through Internet.
+
 \section{The Future of Internet}
+\paragraph{}
+The Internet counts billions of connected devices nowadays.
+This number is growing exponentially and this is not going to stop in the coming decades.
+It is estimated that there will be trillions of objects connected by the end of the century \cite{ref3}.
+In the next decade, we will watch a major change in the Internet.
+The network architectures we know nowadays will changes radically.
+The hardware architectures used now will evolve to meet the expectations of \gls{iot}.
+Applications based on \gls{iot} will explode and new possibilities will be open to the world community.
+\paragraph{}
+Privacy of information is going to be a serious issue in \gls{iot}.
+As every device will be monitored and controlled, privacy becomes an issue.
+The question is how will the Internet of Things keep its evolution without having to be affected by privacy and security arguments \cite{ref20}.
 
@@
 \item Mote appliance control TCP server
-\item Gateway packet transformation process
+\item \gls{ipv4}/\gls{ipv6} Gatewaying Process
 \item \gls{ipv6} over USB tunnel
-\item mote \gls{ipv6} backstation
+\item mote \gls{ipv6} back station
 \item Sensor data storage process
@@
 The general architecture is depicted in figure \ref{fig:gen_architecture}.
-This figure shows the different components of the system and their interactions.
-Each set of components is appearing within a box and that box is labeled.
-This label refers to the hardware deployment of the component within the whole system.
+This figure shows the existing components and their interactions.
+Each set of components appear within a box as this last refers to what hardware are the components deployed in as it is labeled in the box.
 \begin{figure}[htbp]
 \centering
-\includegraphics[scale=0.40]{img/general_architecture.jpg}
+\includegraphics[width=17cm,height=17cm,keepaspectratio]{img/general_architecture.jpg}
 \caption{General architecture depicting the main components of the system}
@@
 \end{figure}
-\paragraph{}
-Now that the general architecture has been defined, the question that should be asked is whether all these components reside in the same layer in the \gls{osi} model and if not, in which layer does each of these components reside in the \gls{osi} model.
+%\clearpage
+\paragraph{}
+Now that the general architecture has been defined, the question that should be asked is whether all these components operate in the same layer in the \gls{osi} model and if not, in which layer does each of these components reside.
 As an answer to such a question, figure \ref{fig:layered_view} is here to show on what layer does every component act on.
@@
 \end{figure}
-\paragraph{}
-Now that the network diagram has been established in \ref{fig:network_diagram} where the hardware component appears, the next thing is to show where each software component will be deployed.
+
+\paragraph{}
+Now that the network diagram has been established in \ref{fig:network_diagram} where the hardware component are depicted, we need to show where each software component is deployed.
 In other words, what are the software components that exist on which hardware machine.
-This is the goal of table \ref{table:soft_hard_deployment}.
-It clearly shows what software components exist in what hardware.
+This is the goal of the matrix appearing in table \ref{table:soft_hard_deployment}.
+It shows where very software component is deployed.
 From the table \ref{table:soft_hard_deployment}, it is clear that there are hardware machines that have more than one component deployed and this is also the case for small devices such as sensor nodes within the \gls{wsn}.
@@
     Mote appliance control TCP server       & X    & X         & ~       & ~      \\
-    Gateway packet transformation process   & ~    & ~         & X       & ~      \\
+    \gls{ipv4}/\gls{ipv6} Gatewaying Process   & ~    & ~         & X       & ~      \\
     IPv6 over USB tunnel                    & ~    & ~         & X       & ~      \\
-    Mote IPv6 backstation                   & ~    & X         & ~       & ~      \\
+    Mote IPv6 back station                   & ~    & X         & ~       & ~      \\
     Sensor data storage process             & ~    & ~         & X       & ~      \\
@@
 \end{table}
-The transformation function is then: \\
+To transform the initial voltage into the current intensity, we use constant variables provided by the \gls{ct} manufacturer.
+But these value cannot be taken for granted.
+To measure them, we conduct an experiment to using the Voltmeter and Ampere meter to measure the slope, $s$, and the function constant, $i_0$, of the intensity transformation function.
+This function resulted in the following values:
+\begin{itemize}
+\item $s = 66.70$
+\item $u_0 = 0$
+\end{itemize}
+Therefore, the first transformation function is: \\
 \boldmath{$I = 66.7U_0$} \\
@@
 \includegraphics[scale=0.35]{img/appliance_control.jpg}
-\caption{TCP Server controling appliance}
+\caption{TCP Server controlling appliance}
 \label{fig:appliance_control}
@@
 \paragraph{}
-The mote by itself does not have the necessary capabilities needed to control a mote.
-In order for a mote to have such capabilities, it needs to be equipped with an additional board, a Data Acquisition board also called a extension board that enable the mote to have more capabilities and therefore the ability to control the power flowing into the mote.
-The extension board used in the course of this project is the Crossbow MDA320CA as depicted in figure \ref{fig:mda320ca}.
-This board has pins which are electrical inputs and outputs that make it possible to have external analog sensors, digital sensors and also be able to power the appliance.
-But the powering of the appliances is not as straightforward as it seems especially that sensor nodes are very low-power devices that cannot provide enough energy to appliances such as a television or a fridge.
-\paragraph{}
-To make a mote control the appliance, the mote will control the power going in a very small circuit and this circuit is responsible for letting the power flow through the appliance from another source of energy which is naturally far more important than the one provided in the mote.
-In summary, the mote has a the extension board of type MDA320CA that has a small electric circuit attached to one of its digital pins that then controls the flow of electricity going to the appliance.
-By turning off a pin, the electricity stops flowing through the circuit and becomes open which stops the electricity to go to the appliance.
+The mote on itself does not have the necessary hardware capabilities to control an appliance.
+It needs to be equipped with a Data Acquisition board also known as Extension Board that enable the mote to have more capabilities and therefore the ability to control the power flowing into the appliance.
+The extension board used in this project is the Crossbow MDA320CA as depicted in figure \ref{fig:mda320ca}.
+This board has Pins which are electrical inputs and outputs that permit possible to have external analog sensors, digital sensors and also the possibility feeding electricity to the appliance.
+However, the power supply of the appliances from the mote is not straightforward.
+\paragraph{}
+To make a mote control the appliance, the mote will govern the supply of electricity by controlling a small circuit that is responsible for pipelining the power flowing to the appliance.
+In short, the mote is equipped with an extension board to which an electric circuit is attached and whose responsible for the flow of power going to the appliance.
+Thus, the turn Off command will open the circuit and the appliance will stop getting fed electricity.
 
@@
 
-\section{Mote IPv6 Backstation}
-\paragraph{} 
-This component that exist in one special mote in the \glsreset{wsn}\gls{wsn}.
-This mote is able to communicate with the rest of the \gls{wsn} using Zigbee and also with the external world using \gls{usb}.
-This make this mote a sink and a router because it has two interface cards.
+\section{Mote IPv6 Back station}
+\paragraph{} 
+This component is deployed only the sink of the \glsreset{wsn}\gls{wsn}.
+This mote is able to communicate with the sensor nodes using Zigbee and with the gateway using \gls{usb}.
 The software component deployed in this sink is a packet forwarder.
-This means that once it receives an \gls{ipv6} packet that is not in the \gls{wsn}, it is forwarded through \gls{usb}.
-\paragraph{}
-The mote's sole goal is forwarding packets in and outside of the \gls{wsn} while other sensor node do not need to worry about how the packet will arrive to its destination.
-This mote is part of the components aiming to solve the heterogeneity problem stated in the problem statement.
-By making the sensor nodes talk to any device as if it is part of the \gls{wsn}, they do not really have to worry about how it will arrive or even how did the packet get to them in the first place.
+In fact, once it receives an \gls{ipv6} packet that is going outside the \gls{wsn}, it is forwarded through \gls{usb}.
+\paragraph{}
+The component's goal is to forward packets in and out of the \gls{wsn} while other sensor node do not need to worry about how the packet will be communicated.
+This mote as part of the components aiming to solve the heterogeneity problem, provides communication transparency.
 
@@
 \paragraph{}
-The WSN is run using \gls{tinyos} which is one of the most used opensource \gls{os} in \glspl{wsn}.
+The WSN is run using \gls{tinyos} which is one of the most used open source \gls{os} in \glspl{wsn}.
 Programs are implemented using \gls{nesc} which is the sole programming language used in the \gls{tinyos} environment.
-This programming language is event-oriented and its syntax is very influenced by C.
+This programming language is event-oriented and its syntax is similar to C.
 In addition, C libraries can be used in \gls{nesc} programs.
 
-\subsection{Technologies used in the Gateway}
+\subsection{Gateway}
 \paragraph{}
 The gateway's hardware machine is Dell Optiplex 9010 running a Linux environment more specifically the Ubuntu 13.04 distribution.
-The gateway has a set of technologies used to provide numerous services the system counts on.
-The programming languages used in the components hosted in this machine are C and \gls{j2se}.
-Netfilter is also used to control incoming and outgoing packets.
-Concerning the storage of sensor data MySQl is the \gls{rdbms} used.
-
-\subsection{Technologies used in the Mobile Client}
-\paragraph{}
-The mobile used is an Android smart phone with Android 2.3. The application developed is using \gls{j2se}, PHP for database connectivity.
+The gateway has a set of technologies used to provide numerous services the system relies on.
+The programming languages used in the components deployed in this machine are C and \gls{j2se}.
+Netfilter is also used to drop incoming packets.
+Concerning the storage of sensor data, MySQl is the \gls{rdbms} used.
+
+\subsection{Mobile Client}
+\paragraph{}
+The mobile used is a smart phone running Android 2.3. The application developed is using \gls{j2se} and PHP for database connectivity.
 
@@
 \paragraph{Timer fired}
-Once the timer is fired, this means it is time to read sensor data from the sensor.
-The command to read from the sensor is called which goes and reads data and once the data is read the event "read done" is triggered.
+Once the timer is fired, it is time to read sensor data from the sensor. The command to read from the sensor is called reads data and once the data is read the event "read done" is triggered.
 \paragraph{Read done}
 Once the data is read and ready, the next thing is to transform it and then send it.
-Concerning the transformation, a series of equations are applied as seen in section 4.4 which build the data value from the original value gotten and it is ready to be sent.
+Concerning the transformation, a series of equations are applied as seen in section 4.4 which builds the data value from the original value.
 A message is created containing two values: the mote NODE\_ID which is the identifier of the mote predefined in the installation of the program in the mote and the sensor data value.
 These two values are separated by a space, encapsulated within a \gls{tcp} datagram and sent.
-\paragraph{}
-During the installation of the software the \glsreset{blip}\gls{blip} package should specified as it is a dependency of the software system.
+The NODE\_ID is not necessary to identify the mote because the mote can be identified using its \gls{ipv6} address.
+\paragraph{}
+During the installation of the software, the \glsreset{blip}\gls{blip} package should specified as the program depends on it.
 \gls{blip} provides the TCP/IP protocol stack that implements \gls{6lowpan} standard by \gls{ietf}.
-In  the installation the NODE\_ID is specified.
+In  the installation, the NODE\_ID is specified.
 If NODE\_ID is 1, this will make the mote have an \gls{ipv6} of FEC0::1, as the NODE\_ID is appended to the predefined prefix of \gls{wsn} during the installation.
-\section{Mote IPv6 Backstation}
-\paragraph{}
-This component is provided as part of the \gls{blip} software package.
-It is \gls{tinyos} program whose goal is to forward packet back and forth between the \gls{wsn} and the outside.
-This mote is connected to the \gls{wsn} using Zigbee and to the outside world using \gls{usb}.
-This special mote has its \gls{ipv6} being FEC0::64.
-It receives \gls{ipv6} packets encapsulated in \gls{usb} frames, extracts them and encapsulates them in a Zigbee frame and vice versa.
-\paragraph{}
-During the installation of this \gls{tinyos} program into the mote sink, one should make sure that it is being installed in the mote sink as \gls{usb} is a necessity.
-The \gls{ipv6} address does not need to be provided as it is already predefined in the program.
+\section{Mote IPv6 Back station}
+\paragraph{}
+This component is included in the \gls{blip} software package.
+It is a \gls{tinyos} program whose goal is to forward packets back and forth between the \gls{wsn} and the external networks.
+This mote is connected to the \gls{wsn} using Zigbee, and to the outside world using \gls{usb}.
+This mote's \gls{ipv6} address is FEC0::64.
+It receives \gls{ipv6} packets which are encapsulated in \gls{usb} frames, extracts them and encapsulates them in a Zigbee frame and vice versa.
+\paragraph{}
+During the installation of this \gls{tinyos} program into the sink, one should make sure that it is installed in the mote sink device because \gls{usb} is crucial.
+The \gls{ipv6} address does not need to be provided because it is already predefined in the program configuration file.
 
@@
 \paragraph{}
-This is a program that is installed in the gateway server machine.
-As the mote sink is connected to the gateway using \gls{usb}, the gateway will recognize it as a network interface card using \gls{ipv6} as a network technology.
-That is said, this component is part of the \gls{blip} software package and creates the network interface card module and starts extracting incoming \gls{ipv6} packets from the \gls{usb} frames and encapsulates outgoing \gls{ipv6} into \gls{usb} frames.
-\paragraph{}
-The installation and initialization of this component will be given more details in appendix A.
-\section{Gateway Packet Transformation Process}
-\paragraph{}
-This component is in charge of hiding the heterogeneity of the system from the rest of the the components of the system and also the outside world.
-It does that by intercepting all the packet moving and transforms them according to the standards used by the recipient.
-\paragraph{}
-The program works with Netfilter in parallel. This means that Netfilter should drop all incoming and this program will take care of dealing with incoming packets separately. 
-\paragraph{}
-To make use of this process, there is an initial configuration that should be done to all the components of the system in order for the system to work.
+This program is installed in the gateway server machine.
+The sink gets connected to the gateway using \gls{usb}, the gateway will recognize it as a network interface card that uses \gls{ipv6} as a network technology.
+Moreover, this component is part of the \gls{blip} software package which the network interface card module and starts extracting incoming \gls{ipv6} packets from the \gls{usb} frames and encapsulates outgoing \gls{ipv6} into \gls{usb} frames.
+\paragraph{}
+The installation and initialization of this component will be detailed in appendix A.
+\section{\gls{ipv4}/\gls{ipv6} Gatewaying Process}
+\paragraph{}
+This component is in charge of hiding the heterogeneity of the system from the rest of the components of the system and also the outside world.
+It does so by intercepting all the packets and transforms the ones of interest to meet the standards used by the recipient.
+\paragraph{}
+To make use of this process, there is an initial configuration that should be performed on all the components of the system in order for the system to work (explained in Appendix A).
 The gateway has two network interface cards: the Wi-Fi interface and the Zigbee interface provided through the tunnel.
-The first one is using \gls{ipv4} and the second is using \gls{ipv6}.
+The first one is using \gls{ipv4} whereas the second is using \gls{ipv6}.
 \paragraph{}
@@
 \paragraph{}
-The packet sniffer is the one in charge of sniffing packet from both network interface card, as Raw Frames.
+The packet sniffer is the one in charge of sniffing packets from both network interface cards, as Raw Frames.
 The code snippet shows the functions called to sniff the packets.
-We make use of the raw socket library that enables us to sniff data link layer frames before being dropped by Netfilter.
-The main reason of dropping the packets by Netfilter is that the packets will not addressed to the computer and therefore the computer make send ICMP messages to the sender saying that the route is not the best and therefore the sender will never send packets again which is something we need to avoid to have.
+We make use of the Raw Socket library that enables us to sniff the Data Link Layer frames before dropping them using Netfilter.
+The main reason behind dropping the packets is when packets are received by a station who has an empty routing table, this station will automatically send ICMP REDIRECT TO HOST messages to the sender meaning that the route chosen by the sender is not the appropriate one, therefore the sender will stop sending packets to that station.
 
@@
 \paragraph{}
-The first line of code is the initialization of the raw socket, we define the type, scope and name space of the packets to be sniffed.
+The first line of code initializes the Raw Socket, we define the type, scope and name space of the packets to be sniffed.
 This will return a file descriptor that will allow us to call the socket.
@@
 \paragraph{}
-After sniffing a packet, we must check if it is an \gls{ipv4} or \gls{ipv6} packet.
-The problem is that when reading from the Wi-Fi interface card, we get the whole frame where the network packet is encapsulated, but the Zigbee Tunnel only gives the raw \gls{ipv6} packet.
-Once a packet is read by the frame, from which interface card came is a problem and therefore a partial solution is found.
-We cast the packet to Wi-Fi frame and whenever we check for the network protocol, we get two solutions, 0 referring to \gls{ipv6} and 8 referring to \gls{ipv4}.
-If we get zero, the whole data should be casted to the \gls{ipv6} data structure whereas if it is 8, we should move the pointer forward to drop the Wi-Fi header and cast the following part to the \gls{ipv4} data structure.
-\paragraph{}
-In each of those conditions we only select the TCP datagrams and this is why we check if the next header in the IP structure is equal to 6.
-After that we move the pointer to drop the IP header and only keep the TCP datagram that will have to be sent in a the newly transformed packet.
+After sniffing a packet, we must check if it is an \gls{ipv4} or an \gls{ipv6} packet.
+The problem is that when reading from the Wi-Fi interface card, we get the second layer frame where the network packet is encapsulated, but the Zigbee Tunnel only gives the third layer raw \gls{ipv6} packet.
+Once a packet is read, one cannot identify the interface card providing it, and therefore a partial solution is found.
+We cast the packet to Wi-Fi frame and check next header's protocol number.
+There are two solutions: 0 referring to \gls{ipv6} and 8 referring to \gls{ipv4}.
+If we get zero, the whole data should be casted to the \gls{ipv6} data structure whereas if it is 8, we should move the pointer forward to omit the Wi-Fi header and cast the following part to the \gls{ipv4} data structure.
+\paragraph{}
+In each of those conditions, we only select the TCP datagrams, and this is why we check if the next header in the IP structure is equal to 6.
+After that, we move the pointer to ignore the IP header and only keep the TCP datagram that will have to be sent in a the newly created packet.
 Depending on the network technology of the sniffed packet, we should build a new packet with the technology that is recognized by the recipient.
@@
 \paragraph{}
-The packet tranformer subcomponent is in charge of building two types of packets: \gls{ipv4} and \gls{ipv6} packets.
-As stated in section 4.8, all devices in this system have an \gls{ipv4} and \gls{ipv6} address where one of them is not recognized only at the gateway.
-As the \gls{wsn} excelusively supports \gls{ipv6} and they cannot be addressed without using their \gls{ipv6} address and any other device not using \gls{ipv6} may communicate with them by using their virtual \gls{ipv4} address at once the packets reach the gateway, it transform so that the data carried will be delivered and the packet will have instead of \gls{ipv4} source and destination addresses, \gls{ipv6} addresses that will be translated using the Address translator subcomponent.
+The packet transformer subcomponent is in charge of building two types of packets: \gls{ipv4} and \gls{ipv6} packets.
+As stated in section 4.8, all devices in this system have both an \gls{ipv4} and \gls{ipv6} address, where one of them is only recognized at the gateway.
+The \gls{wsn} exclusively supports \gls{ipv6}.
+Therefore, the sensor nodes cannot be addressed without \gls{ipv6}, but any device not using \gls{ipv6} may communicate with them by using their virtual \gls{ipv4} addresses.
+Once the packet reaches the gateway, it is transformed in order for the data carried to be delivered.
+Therefore, the packet will have instead of \gls{ipv4} source and destination addresses, \gls{ipv6} addresses that will be translated using the Address translator subcomponent.
 \paragraph{}
@@
 \paragraph{}
-The idea is the same for except that each technology has its own details such as filling different fields and computing the checksum.
-Another difference is when creating the \gls{ipv4} one should also create the Wi-Fi frame otherwise it won't be able to be sent through the corresponding network interface card whereas when building \gls{ipv6} packet one should just send the packet and the tunnel will take care of encapsulating it into a \gls{usb} frame.
-The function receives the data to be sent which is in all cases a \gls{tcp} datagram along with the size of the datagram, the source and destination addresses that are already translated.
-Afterwards, the transformator creates the packet appends the datagram, fills the source and destination address and create a socket attached to the corresponding interface card and send the packet.
+The idea is the same, except that each technology has its own details such as filling different packet fields and computing the checksum.
+Another difference is when creating the \gls{ipv4} packet, one should also create the Wi-Fi frame otherwise it won't be able to be sent through the Wi-Fi network interface card.
+But when building \gls{ipv6} packet one should only send the packet and the tunnel will take care of encapsulating it into a \gls{usb} frame.
+The sendIPv6Packet() function receives the data to be sent, which is a \gls{tcp} datagram, along with the size of the datagram, and the translated source and destination addresses.
+Afterwards, the transformer creates the packet appends to it the datagram, fills the source and destination addresses and create a Socket linked to the corresponding interface card and sends the packet.
 \begin{lstlisting}
@@
 The address translator is the subcomponent that assigns \gls{ipv4} and \gls{ipv6} to all devices in this system.
-For instance, sensor nodes in the \gls{wsn} have \gls{ipv6} addresses but this subcomponent assigns to them \gls{ipv4} addresses.
+For instance, sensor nodes in the \gls{wsn} have \gls{ipv6} addresses but this subcomponent assigns \gls{ipv4} addresses to them.
 \paragraph{}
@@
 That is said, any \gls{ipv4} is translated to one and only one \gls{ipv6} and vice versa.
-For a mobile client with \gls{ipv4} address of 10.50.1.50 to talk with a mote whose \gls{ipv6} address is FEC0::1, the mobile client should be aware of its translated to \gls{ipv4} address which is 192.168.0.1.
-Once the packet is received and knowing it is not addressed to it.
-The source and destination \gls{ipv4} addresses are translated to \gls{ipv6} making the source address equal to 0A32::0132 and the destination FEC0::1.
+For a mobile client with \gls{ipv4} address of 10.50.1.50 to communicate with a mote whose \gls{ipv6} address is FEC0::1, the mobile client should be aware of the mote's translated \gls{ipv4} address which is 192.168.0.1.
+Once the packet is received, the source and destination \gls{ipv4} addresses are translated to \gls{ipv6} making the source address equal to 0A32::0132 and the destination FEC0::1.
 The \gls{ipv6} packet is created and reaches the destination which is the sensor node.
-Once the response is going back the scenario is the same.
-\paragraph{}
-The translation algorithm has been tested extensively to make sure it is bijective and that any address can be translated to one and no more and than one address.
-When configuring the system, one should make sure that there is no two devices that have the same address real or virtual.
-Any \gls{ipv6} and \gls{ipv4} should be unique throughout the whole system.
-The \gls{wsn} should be granted an interval of \gls{ipv4} address so that no one will be able to have them besides the \gls{wsn} that uses them virtually.
+Once the response is going back, the scenario is similar.
+\paragraph{}
+The translation algorithm has been tested extensively to make sure it is bijective, and that any address can be translated to one and no more and than one address.
+When configuring the system, one should make sure that there are no two devices that have the same address, real or virtual.
+Any \gls{ipv6} and \gls{ipv4} address should be unique throughout the whole system.
+The \gls{wsn} should be granted an interval of \gls{ipv4} addresses to make sure that no one besides the \gls{wsn} has those addresses.
 
 \chapter{Findings}
-
-\section{Experiment}
-%\subsection{Methodology}
-\section{Data Gathering}
-\section{Interpretation of Results}
+\paragraph{}
+The system was built to monitor and control home appliances, and to also provide an infrastructure for future platforms to be integrated in this system.
+An example of infrastructures that can be the integrated is the \glsreset{dr}\gls{dr}.
+To build overlying system's, one should make sure that the system achieves a performance level that one can count on.
+The purpose of this chapter is to measure the performance of the system.
+It is done by conducting two experiments which will help us gather data, and based on that, an interpretation of this data will be done to verify if the system shows an acceptable level of performance.
+\section{Experimentations}
+\paragraph{}
+To measure the performance of the system, two experiment were conducted where the goal is to quantify the performance of the system.
+There is an architectural decision madein \gls{6lowpan}, and we would like to see if it add more performance to the system.
+The architectural decision is the following: There is more priority given to routed packet that the processed ones.
+As a result, this clearly pushes us to think that the delay of the system will not follow the same growth trend as the network traffic.
+\paragraph{Experiment 1:}
+We start with one mote in the \gls{wsn} that will be interacting with the outside world.
+We then start sending ping command periodically (once every 1 second) and record the delay and the jitter.
+Once every two minutes, we will add another mote to the \gls{wsn} and continue pinging the same mote.
+We keep adding motes every 2 minutes until we reach 6 motes which is the maximum number of motes available.
+This will enable us to collect data about the mote delay and jitter as we add more traffic to the system.
+\paragraph{Experiment 2: }
+The second experience is meant to measure the time it takes to transform the packet.
+In other words, measure the extra delay and jitter when having the Gateway Transformation Process as part of the system.
+The timer starts before the sniffing of the packet and ends right after the packet is transformed and communicated.
+We believe that this transformation adds delay, but we are hoping that this extra delay is not significant enough to affect the whole system.
+\section{Data Gathering and Analysis}
+\paragraph{}
+After both experiments were conducted, we collected a significant amount of data that was used to measure the performance.
+In addition, we analyze and interpret the results to make judgments about the overall performance of the system.
+\subsection{Experiment 1}
+\paragraph{}
+In the first experiment, the same mote was pinged once every second and we added new motes to the system every two minutes.
+The experiment as a whole took 12 minutes or 720 pings.
+Table \ref{table:exp1} shows the result of the experiment.
+
+\begin{table}[htbp]
+    \begin{tabular}{lll}
+    \hline
+    number of motes & Delay (ms) & Jitter (ms) \\\hline
+    1               & 87.5         & 1.2         \\ 
+    2               & 88.1         & 2.0         \\
+    3               & 88.3        & 2.7         \\
+    4               & 88.8         & 3.1         \\
+    5               & 90.0         & 3.3         \\
+    6               & 89.5         & 3.5         \\
+    \end{tabular}
+    \caption{Experiment 1 delay and jitter}
+    \label{table:exp1}
+\end{table}
+\paragraph{}
+For each row in the table, the average delay is computed over the pings performed in the 2 minutes interval or the 120 pings gathered.
+The jitter is also computed similarly as the average variation from the mean delay in that 2 minutes interval.
+The result is displayed in the table \ref{table:exp1}.
+\paragraph{}
+The data collected shows even though the delay in the \gls{wsn} is high, 
+ nothing can be done about that because when the traffic in the \gls{wsn} is low, the delay is still high and this has to do with the computational time taken by the TCP/IP protocol protocol stack provided by \gls{blip} using that implements \gls{6lowpan}.
+To reduce the delay, one should probably rethink the design of the \gls{blip} and investigate ways to reduce processing time.
+
+ \begin{figure}[htbp]
+ \centering
+ \includegraphics[scale=1]{img/delay_jitter.JPG}
+ \caption{Delay and jitter evolution by Bandwidth}
+ \label{fig:delay_jitter}
+ \end{figure}
+
+ \paragraph{}
+ The good news in experiment 1 is that the priority of packets brings benefit to the network.
+Even though the traffic grows exponentially, the average delay stays constant.
+However, the jitter follows the traffic's trend, but still the average delay shows good results.
+
+\subsection{Experiment 2}
+\paragraph{}
+The second experiment's goal is to measure the contribution of the Gateway Packet transformation process to the overall communication delay.
+The results gathered reflect the average delay and jitter computed over the elapsed time starting from the sniffing of the packet to the transformation and sending to the recipient.
+Table \ref{table:exp2} shows the result that was collected from this experiment as it was done on 200 packets that were sniffed and transformed by the system.
+
+\begin{table}[htbp]
+    \begin{tabular}{ll}
+    \hline
+    Delay (ms) & Jitter (ms) \\ \hline
+    0.1        & 0.0         \\ 
+    \end{tabular}
+    \caption{Experiment 2 delay and jitter}
+        \label{table:exp2}
+\end{table}
+\paragraph{}
+The transformation process's elapsed time is measured in milliseconds and from table \ref{table:exp2}, one can see that the average delay is on average 0.1 milliseconds whereas the jitter is around 0 milliseconds meaning that the process's time varies between a few nanoseconds to at most 150 nanoseconds.
+Depending on the machine's load.
+\paragraph{}
+ The Packet transformation process is a component that shows excellent results.
+Before conducting experiment 2, we were persuaded that there will be additional delay originating from this process and it could be responsible for the high delay seen in experiment 1.
+However, the results show otherwise and that the component is not to be put responsible for such an important delay seen in experiment 1.
 
 \chapter{Conclusion}
-
+\paragraph{}
+This project enabled us to create an infrastructure that would be part of the \glsreset{iot}\gls{iot} and move towards having the smart home system's integrated within the Smart Grid.
+At this stage and counting solely on this work, one cannot clearly see the integration of smart home systems within the \gls{iot}, but this is an infrastructure that provides two main services which are the establishment of the two-way communication between the smart home devices and the outside world, and the heterogeneity handler.
+The two-way communication enables the \glsreset{wsn}\gls{wsn} to monitor home appliances and send sensor data to a repository and this last can be used to simulate a real world \gls{ami}.
+The second direction in this communication gives the ability for clients to control appliances by turning them On or Off.
+\paragraph{}
+The second contribution of this work is the smart management of the heterogeneity.
+The communication from the mote to the mobile client uses a set of technologies and the journey of a simple \gls{tcp} datagram goes through many different Data Link Layer standards and different Network technologies.
+Starting from the mote sending sensor data, The TCP data is encapsulated into an \gls{ipv6} packet that is encapsulated into a Zigbee frame.
+Afterward, the packet is extracted from the Zigbee frame and encapsulated into a \gls{usb} frame.
+Once it arrives to the gateway, the \gls{tcp} datagram is extracted from the \gls{ipv6} which is extracted from the \gls{usb} frame and encapsulated in an \gls{ipv4} packet that is encapsulated into a Wi-Fi frame and sent until it reaches the mobile.
+That is said, it clear that the system is heterogeneous and making this heterogeneity transparent was a contribution of this work.
+\paragraph{}
+The \glsreset{wsn}\gls{wsn} was built using the very recent \gls{6lowpan} whose goal is to provide the TCP/IP protocol stack to the smallest devices.
+The \gls{wsn} is autonomous in the sense that it can monitor the home appliances independently and indefinitely.
+\paragraph{}
+All in all, this project as a big picture creates a test-bed providing the main components of a smart home management with an intelligent \gls{wsn}, a robust and powerful middle ware application that is able to use the sensory data and use it to control the home's appliances and a mobile application able to provide the users with remote monitoring and control of their smart homes.
+This was possible by making the system part of the \glsreset{iot}\gls{iot}.
 \chapter{Future Work}
 \paragraph{}
-This thesis project's was to establish the infrastructure for integrating smart home within \glsreset{sg}\gls{sg} and make it part of the \glsreset{iot}\gls{iot}.
+The goal of this thesis was to establish the infrastructure for integrating smart home within \glsreset{sg}\gls{sg} and make it part of the \glsreset{iot}\gls{iot}.
 This project enabled the deployment of the two-way communication as an infrastructure that can be used to build the \glsreset{dr}\gls{dr}.
-\gls{dr} is one of the most important topics in smart grids.
+\gls{dr} is one of the most discussed topics in Smart Grids.
 It enables the utility and households to communicate with a set of services that any of them may request from the other.
-The current state of the art does not incorporate this communication as part of the smart home systems \cite{ref15}.
-But as there is a need for this communication, \gls{dr} will have to be implementation as part of the
-
-\backmatter
-
-\appendix
-\chapter{How to install the system}
+The current state of the art does not incorporate this communication interface in smart home systems \cite{ref15}.
+But as there is a need for this communication, \gls{dr} will have to be implemented as part of the smart home package.
+\paragraph{}
+Demand response integration is a future work for this study.
+\gls{dr} includes a set of services that have to be implemented in the smart home packages.
+In \gls{dr}, the consumer has the possibility to ask the utility about the instantaneous price quote for electricity.
+The consumer when in need of high priority in energy, may ask the utility to guarantee the delivery of a certain quantity of electricity.
+Of course this comes with a cost, but this agreement is part of the \gls{dr} and should be implemented at the level of smart home systems.
+Consumers can also help utilities at difficult times.
+For instance, when the utility cannot provide enough energy at peak times, the utility may ask some consumers to reduce their consumption and as a result be billed a reduced cost for the electricity that is consumed.
+In general, the smart home aims to reduce energy consumption and as the utility needs to reduce its energy production in peak hours, the \gls{dr} will help the utility ask the smart homes to reduce their consumption.
+\paragraph{}
+As this project aims to integrate the smart home devices within the \glsreset{iot}\gls{iot}, \glsreset{rfid}\gls{rfid} must be part of this system as a future work.
+The \gls{rfid} is a building block within the \gls{iot} and will equip the devices with a low-power radio identification mechanism hat will give more flexibility to the system.
+therefore, the external systems may be able to recognize uniquely each device in the \gls{iot} without counting primarily on their \gls{ipv6} address, which may change from time to time.
+\paragraph{}
+This system provides near real-time sensory data that is not exploited intelligently.
+The only use of this data is to enable users to monitor their home consumption.
+As far as the system is concerned, one must more intelligently exploit this data that can be used by smart algorithms to automatically and autonomously reduce the household's energy consumption.
+Such algorithms can be based on machine learning and data mining techniques making it a cross disciplinary project aiming to reduce the consumption of households without reducing the utility of users.
+In other words, building algorithms that will reduce consumption of households while the user is not noticing change.
+\paragraph{}
+Before building new services on top of this system, one should test different aspects of the system.
+In this study, the only thing that was tested, measured and quantified was the performance of the system which gave good results about the system's overall performance.
+As a future work, one should test different aspects of the system such as the reliability and availability of the system.
+More experiments should be conducted to see how well the system acts in high traffic networks and how good could it support new overlying systems such as \gls{dr}.
+\paragraph{}
+The system provides different services, but there is no authentication or authorization involved to access the system.
+In fact, any user knowing the mote's \gls{ipv6} address can control it because there is no existing party or component that controls the flow of information.
+As a future work, the system should have a third party component whose goal is to make the system secure.
+
+\renewcommand{\bibname}{Index}
+\bibliographystyle{IEEEtran} 
+\bibliography{IEEEabrv,myrefs}
+
+%\backmatter
+\begin{appendices}
+\chapter{System Installation}
 \section{Hardware needed}
-In order for the user to install the system, one should make sure he/she is having a computer that will be used for the sole purpose of hosting different services.
+In order for the user to install the system, one should dedicate a computer to host the different services of the system.
 The crossbow professional kit is a necessity and a router must be available to connect the different components through Wi-Fi.
-A smartphone or a separate computer with Android emulator may de used.
+A smart phone or a separate computer with Android emulator may de used.
 
@@
 \subsection{Operating system}
-The machine should run Linux Ubuntu distribution (11.10 or more recent). You way use Windows as long as you Install a Virtual machine using Cygwin or other tools
+The machine should run Linux Ubuntu distribution (11.10 or more recent). You way use Windows as long as you Install a Virtual machine such as Cygwin for instance.
 \subsection{TinyOs Installation}
@@
 You should be able to get sensor data periodically. The mobile client access for turning On and Off will be explained in the Appendix C.
-
-\singlespacing
-\bibliographystyle{IEEEtran} 
-\bibliography{IEEEabrv,myrefs}
+\end{appendices}
 

### 2013-07-08 trunk/doc/thesis/cover_page.tex
ARABIC TITLE GRRRRRR

\begin{titlepage}
\begin{center}
\includegraphics[width=0.4\textwidth]{img/aui_logo.jpg}
% Title
\\
{ \huge \bfseries Internet of Things
\\ a Real-World Deployment Towards Smart Grids Integration}\\[1.0cm]

In partial fulfillment of the requirements for the degree of Master of Science in Software Engineering (MSSE) at the School of Science and Engineering at Al Akhawayn University in Ifrane, Morocco \\[1.0cm]

% Author and supervisor

\emph{Author:}\\
{\large \textbf{Nacer \textsc{Khalil}}}\\[1.0cm]

\emph{Supervisors:} \\
{\large \textbf{Dr.~Mohamed Riduan \textsc{Abid}}} \\
{\large \textbf{Dr.~Driss \textsc{Benhaddou}}} \\
{\large \textbf{Dr.~Michael \textsc{Gerndt}}} \\[1.0cm]

\vfill
\emph{Submitted in:}\\
{\large \textbf{Summer 2013}} \\
\end{center}
\end{titlepage}

### 2013-07-08 trunk/doc/thesis/report.tex
ARABIC TITLE GRRRRRR

@@
 
-%\chapter{ملخص}
+%\chapter{RL{ملخص}}
 
@@
 \paragraph{}
-The global electrical grid is the bases that gives us this quality of life that most people around the world enjoy.
+The global electrical grid is the basis that gives us this quality of life that most people around the world enjoy.
 The global electrical grid is sometimes called the largest machine ever made by man \cite{ref1}.
 It is a widespread and distributed machine that provides power to households, companies and organizations.
-Therefore the whole world's economy counts on it.
+Therefore the whole world's economy depends on it.
 The electrical grid knows numerous problems such as blackouts, and current instabilities that cost billions of dollars yearly to both the utilities and insurances.
-The grid suffers from the problem of peak hours where energy consumption is high and in most cases, one single utility cannot provide.
+The grid suffers from the problem of peak hours where energy consumption is at its highest level and that one single utility cannot provide.
 \paragraph{}
@@
 \gls{sg} is an intelligent, auto-balancing, self-monitoring, and self-healing electrical grid \cite{ref2} and whose energy sources are distributed between utilities and consumers.
-The nature of the \gls{sg} as having distributed energy sources reduces numerous losses that arise from the long distance transmission lines, and also from the stress and the overload of the electrical lines.
+The nature of the \gls{sg} as having distributed energy sources, reduces numerous losses that arise from the long distance transmission lines, and also from the stress and the overload of the electrical lines.
 \paragraph{}
@@
 It is called the \glsreset{iot}\gls{iot}.
-\gls{iot} is the era of autonomous machine being able to produce and consume information, and being able to act autonomously, without the guidance of humans, to performs their tasks.
+\gls{iot} is the era of autonomous machine being able to produce and consume information, and being able to act autonomously without the guidance of humans, to perform their tasks.
 Cisco predicts that the number of devices to be connected to the Internet in 2020 will break the 50 billions devices from 500 million in 2003.
-That would make 6.58 devices connected for each person in the world whereas that same ration was 0.07 17 years earlier \cite{ref3}.
+That would make 6.58 devices connected for each person in the world whereas that same ratio was 0.07 17 years earlier \cite{ref3}.
 This clearly shows that the shift to the \gls{iot} will be exponential.
@@
 The integration of the \gls{sg} into the \gls{iot} constitutes the main topic of this study and goes along with increasing interest recently raised in both academia and industry, on the inherent correspondence between \gls{sg} and \gls{iot}.
-This  correspondence stems mainly from the need to disseminate information between the \gls{sg} components in order to make optimal decisions about energy usage.
+This  correspondence stems mainly from the need to disseminate information between the \gls{sg} components in order to make optimal decisions about energy use.
 This information dissemination could only be optimal when deploying ubiquitous sensoring and metering, and using the available network infrastructure, i.e.
@@
 The \glsreset{sg} \gls{sg} relies on the two-way communication to build a whole communication framework and a protocol stack namely the \glsreset{dr}\gls{dr} and other services built on top of this last.
-From the consumer's/household's side or more precisely the household's side,there are smart home energy management systems in the market nowadays, but none of them offers an integration with \gls{sg}.
-To have such an integration with \gls{sg}, there is a demand for the two-way communication, which is to make it possible for any appliance to report its real-time energy consumption, and to be able to be controlled remotely.
+From the consumer's/household's perspective or more precisely the household's side,there are smart home energy management systems in the market nowadays, but none of them offers an integration with \gls{sg}.
+To have such an integration with \gls{sg}, there is a need for the two-way communication, in order to make it possible for any appliance to report its real-time energy consumption, and to be able to be controlled remotely.
 Such pillar is missing for the time being, and therefore the integration of the smart home within the smart grid is not possible.
-Also, in order for \gls{sg} to be integrated into the \gls{iot}, all devices should be able to communicate in the two directions.
+Also, in order for \gls{sg} to be integrated into the \gls{iot}, all devices should be able to communicate in the two ways.
 \paragraph{}
@@
 As numerous components count on different technologies, there is a need to interface them in order to make this heterogeneity transparent to the all components of the system.
-As a matter of fact, a software should be built within the system, and would constitute an important part of it, whose main goal is to take care of making the communication technology-transparent and give the possibility to all components of the system to work hand-in-hand in order to achieve their goals independently of the data link layer (Ethernet, Wifi or Zigbee) or the network layer technology (\gls{ipv4} or \gls{ipv6}) they are using.
+As a matter of fact, a software should be built within the system whose main goal is to take care of making the communication technology-transparent.
+This component will give the possibility to all components of the system to work hand-in-hand in order to achieve their goals independently of the data link layer (Ethernet, Wifi or Zigbee) or the network layer technology (\gls{ipv4} or \gls{ipv6}) they are using.
 
@@
 The goal of the study is to solve the problems introduced in the problem statement section.
-This means creating the two-way communication, and solving the heterogeneity that arises from integrating smart homes with the smart grid and making each components in the smart home as part of the Internet of Things by making it communicate with any device in the Internet by having an \gls{ipv6} address that is built by making use of \gls{6lowpan} whose goal is to provide the TCP/IP to the smallest devices and with the least computational resources.
+This means creating the two-way communication, and solving the heterogeneity that arises from integrating smart homes with the smart grid.
+In fact, each components in the smart home will become part of the Internet of Things by making it communicate with any device in the Internet by having an \gls{ipv6} address using \gls{6lowpan} whose goal is to provide the TCP/IP to the smallest devices and also those with the least computational resources.
 \paragraph{}
 The first purpose of this study is to create a testbed having the two-way communication and also being able to make use of it in order to build richer applications based on it.
-To do so, \gls{6lowpan} had to be integrated within every device in the \gls{wsn}.
-This would make every sensor node (mote) identifiable in the network and ready to communicate and be part of the \gls{iot}.
-This communication will go in both directions and based on \gls{6lowpan}, the two-way communication will be built.
-\paragraph{}
-The second purpose of this study is to solve the heterogeneity problem that arises with the integration of smart homes into the grid and also when making the whole system part of the \glsreset{iot}\gls{iot}.
-Solving such a problem means having an extra component that will track all  packets and change them to fit the standards of the destination component.
-This will make of this additional component a middleman and therefore make the heterogeneity problem disappear.
+To do so, \gls{6lowpan} would had to be integrated within every device in the \gls{wsn}.
+This would make every sensor node (mote) identifiable in the network, ready to communicate and be part of the \gls{iot}.
+This communication will go in both directions and based on \gls{6lowpan}; the two-way communication will be built.
+\paragraph{}
+The second purpose of this study is to solve the heterogeneity problem that arises from the integration of smart homes into the grid and also when making the whole system part of the \glsreset{iot}\gls{iot}.
+Solving such a problem means having an extra component that will track all  packets and change them to fit the standards of the targeted component.
+This will make of this additional component a middleman and that would make the heterogeneity problem disappear.
 
@@
 \paragraph{}
-This study aims to solves the two problems presented in section 1.3.
-To do so, a variety of tools and technologies were used for this purpose.The two problems could not be solved one independently of the other even though the problems seem separate, one cannot solve  the a problem without having to solve the other.
-In other words, after the creation of the two-way communication in the \glsreset{wsn}\gls{wsn} which adopts \gls{ipv6} over \gls{6lowpan}, other networks only support \gls{ipv4} and therefore the communication is not possible, and this was due to the heterogeneous nature of the system.
+This study aims at solving the two problems presented in section 1.3.
+To do so, a variety of tools and technologies were used for this purpose.The two problems could not be solved one independently of the other.
+Even though the problems seems to be separate, one cannot solve  the first problem without having to solve the second.
+In other words, after the creation of the two-way communication in the \glsreset{wsn}\gls{wsn} which adopts \gls{ipv6} over \gls{6lowpan}, other networks can only support \gls{ipv4} and therefore the communication would not be possible, and this is due to the heterogeneous nature of the system.
 This is why the two-way communication was created, but only after solving the heterogeneity problem, everything worked.
@@
 Each of these four components is there to solve part of the problem.
-Each of these subsystems will be discussed extensively in the Chapter 4 from its architectural aspect of the component, and in Chapter 5 from its deployment point of view.
-\paragraph{}
-It is to be noted that some of the components will be discussed from the architectural point in Chapter 4 but not from an implementation aspect in Chapter 5.
-The sole reason behind is the fact that some of the equipment required for this thesis were not delivered on time and therefore the parts in the implementation requiring such equipment were omitted to meet the deadlines.
+Each of these subsystems will be discussed extensively in Chapter 4 from the architectural aspect of the component, and in Chapter 5 from a deployment point of view.
+\paragraph{}
+It is important to note that some of the components will be discussed from the architectural point in Chapter 4 but not from an implementation aspect in Chapter 5.
+The sole reason behind this is the fact that some of the equipment required for this thesis were not on time and therefore the parts in the implementation requiring such equipment were omitted in order to meet the deadlines.
 \section{Outcome of the Study}
@@
 \paragraph{}
-The autonomous production of information is shown by the fact that every sensor node in the \gls{wsn} reads sensory data from its sensors, that are attached to appliance, in order to sense their consumption and this information is produced and communicated periodically.
+The autonomous production of information is shown by the fact that every sensor node in the \gls{wsn} reads sensory data from its sensors, that are attached to appliance, in order to sense their consumption.
+This information about consumption is produced and communicated periodically.
 The consumption of data is proved by showing the execution of the commands communicated through the network by external nodes in the \gls{iot}.
@@
 \paragraph{}
-The test-bed should also show that communication between any two nodes in the \gls{iot} is successful and reliable, and that the heterogeneity of the system is transparent to the nodes in the system and this is done by creating a software tool deployed in the gateway which is between the \gls{wsn} and the outside world.
-This software tool is perceived as a heterogeneity handler and also a communication middleman that translates the communication to the appropriate technology needed by the recipient.
+The test-bed should also show that communication between any two nodes in the \gls{iot} is successful and reliable, and that the heterogeneity of the system is transparent to the nodes in the system.
+This was achieved by creating a software tool deployed in the gateway which is between the \gls{wsn} and the outside world.
+This software tool is perceived as a heterogeneity handler, and also a communication middleman that translates the communication to the appropriate technology needed by the recipient.
 \paragraph{}
@@
 It presents the background of the work by revising the literature related to the topics of \gls{iot} and \gls{sg}.
-It also discussed related work and how it was tackled in different research papers, how does it relate to the work done in this thesis and also how does it differ.
+It also discusses related work and how it was tackled in different research papers, how does it relate to the work done in this thesis and also how does it differ.
+Finally, its presents the commonalities ad differences between this thesis and the work previously done.
 \paragraph{Chapter 3}
-It is dedicated to the \glsreset{iot}\gls{iot}. It presents the topic, talks about the importance of \gls{iot}, its impact on the world, the actual Internet and the future of Internet.
+This part is dedicated to \gls{iot}, its importance and its impact on the world. Is also presents the actual state of the Internet and makes predictions about its future.
 \paragraph{Chapter 4}
-It will present the architecture of the test-bed. It will discuss the system from different aspects, and how were the problems stated in section 1.3 solved from the architectural aspect.
+It presents the architecture of the test-bed. It also discusses the system from different perspectives, and how the problems stated in section 1.3 were solved from the architectural aspect.
 \paragraph{Chapter 5}
-This chapter will mostly discuss the system deployment.
-It will start by presenting the hardware used in this project, then move to the technology enablers.
-Afterward, Each component of the system will be discussed extensively and move to the integration of the components to form the test-bed.
+This chapter mostly discusses the system deployment.
+It starts by presenting the hardware used in this project, then moves to the technology enablers.
+Afterwards, each component of the system is discussed extensively and moves along with the integration to form the test-bed.
 \paragraph{Chapter 6}
-It will start first by presenting the findings then move to explain the experiments that were conducted. Afterwards, the data gathered will be shown, analyzed and interpreted.
+It starts first by presenting the findings then move to explain the experiments that were conducted. Afterwards, the data gathered is shown, analyzed and interpreted.
 \paragraph{Chapter 7, 8}
-These are concluding chapters of this study. They provide directions about future work, recommendations to the reader about what parts of the system could be ameliorated and how it can be achieved.
+These are concluding chapters of this study.
+They provide directions about future work, recommendations to the reader about what parts of the system could be improved and how these improvements could be achieved.
+
 \chapter{Literature Review}
@@
 \paragraph{}
-The actual electrical grid is based on a technology that is at least one century old and its problems are becoming more and more costly \cite{ref5}.
-Therefore, there is a need to change it as to meet to 21st century requirements such as having more green energy in order to reduce pollution caused mainly by the use of fossil energy.
-As a matter of fact, research introduced us to new technology: the \glsreset{sg}\gls{sg}.
-Although it is still a research project and has not reached maturity yet, nations across the globe started to deploy \gls{sg} projects.
-The \gls{sg} is relying on Information Technology to solve many of the problem existing in the traditional grid.
+The actual electrical grid is based on a technology that is at least one century old and its weaknesses are becoming more and more costly \cite{ref5}.
+Therefore, there is a need to change it to meet to 21st century requirements such as having more green energy in order to reduce pollution caused mainly by the use of fossil energy.
+As a matter of fact, research introduced us to a new technology: the \glsreset{sg}\gls{sg}.
+Although it is still a research project that has not reached the stage of maturity yet, nations across the globe started to deploy \gls{sg} projects.
+The \gls{sg} is relying on Information Technology to solve many of the problems existing in the traditional grid.
 Consequently, in order for the grid to become more reliable, secure and efficient, a bi-directional information and communication infrastructure should be embedded within the grid \cite{ref5}.
@@
 The smart grid brings new concepts that will make it achieve greater efficiency, reliability and security compared to the traditional grid.
-The first one is the introduction of the \gls{dg} by exploiting \gls{res} \cite{ref6} and this reduce the losses of energy in the transmission.
-This distribution by itself enables the power grid to move from the one way energy distribution to a two-way energy distribution where the utility will not be the sole energy producer, but households can now produce and even sell the extra electricity.
+The first one is the introduction of the \gls{dg} by exploiting \gls{res} \cite{ref6} and this reduces the losses of energy in the transmission.
+This distribution by itself enables the power grid to move from the one way energy distribution to a two-way energy distribution where the utility will not be the sole energy producer.
+Households can now produce and sell the surplus of electricity.
 This is known nowadays as the feed-in-tariff which is a policy that allows the consumers to become producers and sell the excess power to the grid \cite{ref7}.
-This will allow the power grid to provide electricity to consumer in a "free market" fashion which means that it will obey to the laws of supply and demand \cite{ref7}.\glsreset{iot} \glsreset{sg}
+This will allow the power grid to provide electricity to consumers in a "free market" fashion which means that it will obey the laws of supply and demand \cite{ref7}.\glsreset{iot} \glsreset{sg}
 \section{\gls{iot} and \gls{sg}}
@@
 Nowadays, it is estimated that 12.5 billion devices are connected to the Internet and that this trend will continue to reach 50 billion devices by the end of the decade \cite{ref3}.
-\gls{iot} brings several benefits to the actual Internet and will pull the \gls{sg} to make this part of it.
-Consequently, research has taken the direction to integrate the \gls{sg} within the \gls{iot} as most of its aspects favor  the \gls{sg} \cite{ref8}.
+\gls{iot} brings several benefits to the actual Internet and will pull the \gls{sg} to make it part of it.
+Consequently, research has taken the direction of integrating the \gls{sg} within the \gls{iot} as most of the aspects of this last favor  the \gls{sg} \cite{ref8}.
 \paragraph{}
 \gls{iot} counts on a set of technologies which are \glsreset{rfid}\gls{rfid}, \gls{m2m} \cite{ref9}, public communication networks such as 2G/3G mobile networks, power line communication, WI-FI and Zigbee\cite{ref8}.
-The smart grid's architecture based on \gls{iot} will be divided into three layers:
+The smart grid's architecture based on \gls{iot} is divided into three layers:
 \begin{itemize}
-\item Perception layer: Devices will autonomously perform their tasks such as sensing and executing orders.
-\item Communication layer: The communication infrastructure is defined here and issues related to how to communicate information is handled.
-\item Application layer: Applications can be built to make use of the underlying data.
+\item Perception layer: Devices autonomously perform their tasks such as sensing and executing orders.
+\item Communication layer: The communication details are clarified here. It hides the communication issues and provide an easy for upper layers to send and receive information.
+\item Application layer: Applications are built to exploit the data received from the underlying layers.
 \end{itemize}
@@
 Research discusses the benefits to replace Zigbee by Wi-Fi as this last offers higher bandwidth, non-line transmission ability, large-scale data collection and is highly cost-effective \cite{ref10}.
-Still, Zigbee has the exclusive advantage of comsuming the least energy of the communication technologies that exist nowadays which is something required in \gls{sg} and \gls{iot}, especially for systems that are installed far away from any maintenance team such as smart grid monitoring and self-healing devices where one cannot manage to move for the sole purpose of charging these devices.
+Still, Zigbee has the exclusive advantage of consuming the least energy of the communication technologies that exist nowadays which is something required in \gls{sg} and \gls{iot}, especially for systems that are installed far away from any maintenance team such as smart grid monitoring and self-healing devices where one cannot manage to move for the sole purpose of charging these devices.
 \paragraph{}
@@
 \item \gls{ipv6} over USB tunnel
-\item mote \gls{ipv6} backstation
+\item mote \gls{ipv6} back station
 \item Sensor data storage process
@@
 \centering
-\includegraphics[width=17cm,height=17cm,keepaspectratio]{img/general_architecture.jpg}
+\includegraphics[width=17cm,height=20cm,keepaspectratio]{img/general_architecture.jpg}
 \caption{General architecture depicting the main components of the system}
@@
 \end{figure}
+%\clearpage
 \paragraph{}
 Now that the general architecture has been defined, the question that should be asked is whether all these components operate in the same layer in the \gls{osi} model and if not, in which layer does each of these components reside.
-As an answer to such a question, figure \ref{fig:layered_view} is here to show on what layer does every component act on.
+As an answer to such a question, figure \ref{fig:layered_view} is here to show in what layer does every component operates.
 
@@
 \includegraphics[scale=0.40]{img/layered_view.jpg}
-\caption{Layered view showing where every component appears in the \gls{osi} model}
+\caption{Layered view showing where every component operates in the \gls{osi} model}
 \label{fig:layered_view}
@@
 In other words, what are the software components that exist on which hardware machine.
-This is the goal of the matrix apearing in table \ref{table:soft_hard_deployment}.
+This is the goal of the matrix appearing in table \ref{table:soft_hard_deployment}.
 It shows where very software component is deployed.
@@
     IPv6 over USB tunnel                    & ~    & ~         & X       & ~      \\
-    Mote IPv6 backstation                   & ~    & X         & ~       & ~      \\
+    Mote IPv6 back station                   & ~    & X         & ~       & ~      \\
     Sensor data storage process             & ~    & ~         & X       & ~      \\
@@
 \centering
-\includegraphics[scale=0.8]{img/ct.jpg}
+\includegraphics[scale=0.7]{img/ct.jpg}
 \caption{Current transformer}
@@
 But these value cannot be taken for granted.
-To measure them, we conduct an experiment to using the Voltmeter and Amperemeter to measure the slope, $s$, and the function constant, $i_0$, of the intensity transformation function.
+To measure them, we conduct an experiment to using the Voltmeter and Ampere meter to measure the slope, $s$, and the function constant, $i_0$, of the intensity transformation function.
 This function resulted in the following values:
@@
 The extension board used in this project is the Crossbow MDA320CA as depicted in figure \ref{fig:mda320ca}.
-This board has PINs which are electrical inputs and outputs that permit possible to have external analog sensors, digital sensors and also the possibility feeding electricity to the appliance.
+This board has Pins which are electrical inputs and outputs that permit possible to have external analog sensors, digital sensors and also the possibility feeding electricity to the appliance.
 However, the power supply of the appliances from the mote is not straightforward.
@@
 
-\section{Mote IPv6 Backstation}
+\section{Mote IPv6 Back station}
 \paragraph{} 
@@
 
-\section{\gls{ipv4}/\gls{ipv6} Gatewaying Process}
+\section{IPv4/IPv6 Gatewaying Process}
 The \gls{ipv4}/\gls{ipv6} Gatewaying Process is one of the components that hides the heterogeneity that exist in the system.
@@
 \paragraph{}
-The WSN is run using \gls{tinyos} which is one of the most used opensource \gls{os} in \glspl{wsn}.
+The WSN is run using \gls{tinyos} which is one of the most used open source \gls{os} in \glspl{wsn}.
 Programs are implemented using \gls{nesc} which is the sole programming language used in the \gls{tinyos} environment.
@@
 \paragraph{}
-The mobile used is a smart phone runnning Android 2.3. The application developed is using \gls{j2se} and PHP for database connectivity.
+The mobile used is a smart phone running Android 2.3. The application developed is using \gls{j2se} and PHP for database connectivity.
 
@@
 If NODE\_ID is 1, this will make the mote have an \gls{ipv6} of FEC0::1, as the NODE\_ID is appended to the predefined prefix of \gls{wsn} during the installation.
-\section{Mote IPv6 Backstation}
-\paragraph{}
-This component is provided as part of the \gls{blip} software package.
+\section{Mote IPv6 Back station}
+\paragraph{}
+This component is included in the \gls{blip} software package.
 It is a \gls{tinyos} program whose goal is to forward packets back and forth between the \gls{wsn} and the external networks.
 This mote is connected to the \gls{wsn} using Zigbee, and to the outside world using \gls{usb}.
-This special mote's \gls{ipv6} address is FEC0::64.
-It receives \gls{ipv6} packets encapsulated in \gls{usb} frames, extracts them and encapsulates them in a Zigbee frame and vice versa.
-\paragraph{}
-During the installation of this \gls{tinyos} program into the mote sink, one should make sure that it is installed in the mote sink as \gls{usb} is a necessity.
+This mote's \gls{ipv6} address is FEC0::64.
+It receives \gls{ipv6} packets which are encapsulated in \gls{usb} frames, extracts them and encapsulates them in a Zigbee frame and vice versa.
+\paragraph{}
+During the installation of this \gls{tinyos} program into the sink, one should make sure that it is installed in the mote sink device because \gls{usb} is crucial.
 The \gls{ipv6} address does not need to be provided because it is already predefined in the program configuration file.
@@
 \paragraph{}
-This is a program that is installed in the gateway server machine.
-sink gets connected to the gateway using \gls{usb}, the gateway will recognize it as a network interface card that uses \gls{ipv6} as a network technology.
+This program is installed in the gateway server machine.
+The sink gets connected to the gateway using \gls{usb}, the gateway will recognize it as a network interface card that uses \gls{ipv6} as a network technology.
 Moreover, this component is part of the \gls{blip} software package which the network interface card module and starts extracting incoming \gls{ipv6} packets from the \gls{usb} frames and encapsulates outgoing \gls{ipv6} into \gls{usb} frames.
@@
 The installation and initialization of this component will be detailed in appendix A.
-\section{\gls{ipv4}/\gls{ipv6} Gatewaying Process}
+\section{IPv4/IPv6 Gatewaying Process}
 \paragraph{}
 This component is in charge of hiding the heterogeneity of the system from the rest of the components of the system and also the outside world.
-It does that by intercepting all the packets and transforms the ones of interest to meet the standards used by the recipient.
-\paragraph{}
-To make use of this process, there is an initial configuration that should be done to all the components of the system in order for the system to work (explained in Appendix 1).
+It does so by intercepting all the packets and transforms the ones of interest to meet the standards used by the recipient.
+\paragraph{}
+To make use of this process, there is an initial configuration that should be performed on all the components of the system in order for the system to work (explained in Appendix A).
 The gateway has two network interface cards: the Wi-Fi interface and the Zigbee interface provided through the tunnel.
-The first one is using \gls{ipv4} and the second is using \gls{ipv6}.
+The first one is using \gls{ipv4} whereas the second is using \gls{ipv6}.
 \paragraph{}
@@
 That is said, any \gls{ipv4} is translated to one and only one \gls{ipv6} and vice versa.
-For a mobile client with \gls{ipv4} address of 10.50.1.50 to talk with a mote whose \gls{ipv6} address is FEC0::1, the mobile client should be aware of its translated to \gls{ipv4} address which is 192.168.0.1.
-Once the packet is received and knowing it is not addressed to it.
-The source and destination \gls{ipv4} addresses are translated to \gls{ipv6} making the source address equal to 0A32::0132 and the destination FEC0::1.
+For a mobile client with \gls{ipv4} address of 10.50.1.50 to communicate with a mote whose \gls{ipv6} address is FEC0::1, the mobile client should be aware of the mote's translated \gls{ipv4} address which is 192.168.0.1.
+Once the packet is received, the source and destination \gls{ipv4} addresses are translated to \gls{ipv6} making the source address equal to 0A32::0132 and the destination FEC0::1.
 The \gls{ipv6} packet is created and reaches the destination which is the sensor node.
-Once the response is going back the scenario is the same.
-\paragraph{}
-The translation algorithm has been tested extensively to make sure it is bijective and that any address can be translated to one and no more and than one address.
-When configuring the system, one should make sure that there is no two devices that have the same address real or virtual.
-Any \gls{ipv6} and \gls{ipv4} should be unique throughout the whole system.
-The \gls{wsn} should be granted an interval of \gls{ipv4} address so that no one will be able to have them besides the \gls{wsn} that uses them virtually.
+Once the response is going back, the scenario is similar.
+\paragraph{}
+The translation algorithm has been tested extensively to make sure it is bijective, and that any address can be translated to one and no more and than one address.
+When configuring the system, one should make sure that there are no two devices that have the same address, real or virtual.
+Any \gls{ipv6} and \gls{ipv4} address should be unique throughout the whole system.
+The \gls{wsn} should be granted an interval of \gls{ipv4} addresses to make sure that no one besides the \gls{wsn} has those addresses.
 
@@
 \paragraph{}
-The system was built to provide monitor and control home appliances and also provide an infrastructure for future platforms to be integrated in this system.
-Such infrastructures can be the integration of the \glsreset{dr}\gls{dr}.
-To build overlying system's, one should make sure that the system has a certain level of performance that one can count on to build overlying services.
+The system was built to monitor and control home appliances, and to also provide an infrastructure for future platforms to be integrated in this system.
+An example of infrastructures that can be the integrated is the \glsreset{dr}\gls{dr}.
+To build overlying system's, one should make sure that the system achieves a performance level that one can count on.
 The purpose of this chapter is to measure the performance of the system.
-It is done by first doing and experiment which will help us gather data and then based on this data, an interpretation of this data will be done to see whether the system is shows an acceptable level of performance.
+It is done by conducting two experiments which will help us gather data, and based on that, an interpretation of this data will be done to verify if the system shows an acceptable level of performance.
 \section{Experimentations}
 \paragraph{}
-To measure the performance of the system, an experiment was conducted where the goal is to quantify the performance of the system.
-\gls{6lowpan} was built in order for the mote receiving the packet to forward it is not directed to it before dealing with the one directed to it.
-This means that there is more priority given to routed packet that the processed ones.
-As a result, this clearly pushes us to think that the delay of the system will not be much affected as the traffic in the network grows.
+To measure the performance of the system, two experiment were conducted where the goal is to quantify the performance of the system.
+There is an architectural decision madein \gls{6lowpan}, and we would like to see if it add more performance to the system.
+The architectural decision is the following: There is more priority given to routed packet that the processed ones.
+As a result, this clearly pushes us to think that the delay of the system will not follow the same growth trend as the network traffic.
 \paragraph{Experiment 1:}
-We will first have one mote in the \gls{wsn} that will be interacting with the outside world.
-We will start sending ping command periodically (once every 1 second) and record the delay and the jitter.
-Once every two minutes, we will add another mote to the system and continue pinging the same mote.
-We keep adding motes every 2 minutes until we fill the \gls{wsn} with 6 motes which is the maximum number of motes available in the system.
+We start with one mote in the \gls{wsn} that will be interacting with the outside world.
+We then start sending ping command periodically (once every 1 second) and record the delay and the jitter.
+Once every two minutes, we will add another mote to the \gls{wsn} and continue pinging the same mote.
+We keep adding motes every 2 minutes until we reach 6 motes which is the maximum number of motes available.
 This will enable us to collect data about the mote delay and jitter as we add more traffic to the system.
@@
 The second experience is meant to measure the time it takes to transform the packet.
-In other words, measure the average delay and jitter that is added when having the Gateway Transformation Process as part of the system.
-The timer starts before the sniffing of the packet and ends right after the packet is transformed and sent.
-We believe that there will be an additional delay to the communication but we are hoping that it will not affect much the whole delay of the communication between different components of the system. 
+In other words, measure the extra delay and jitter when having the Gateway Transformation Process as part of the system.
+The timer starts before the sniffing of the packet and ends right after the packet is transformed and communicated.
+We believe that this transformation adds delay, but we are hoping that this extra delay is not significant enough to affect the whole system.
 \section{Data Gathering and Analysis}
 \paragraph{}
-After both experiments were conducted, we collected a significant amount of data that was used to measure the performance and an interpretation was done in section 6.3 to have an idea about the performance of the system.
+After both experiments were conducted, we collected a significant amount of data that was used to measure the performance.
+In addition, we analyze and interpret the results to make judgments about the overall performance of the system.
 \subsection{Experiment 1}
@@
 In the first experiment, the same mote was pinged once every second and we added new motes to the system every two minutes.
-The experiment as a whole took 12 minutes and 720 pings we done.
+The experiment as a whole took 12 minutes or 720 pings.
 Table \ref{table:exp1} shows the result of the experiment.
@@
 \paragraph{}
-For each row in the table, the average delay is computed using the pings that were performed in the 2 minutes interval which means 120 pings per row.
-The jitter is also computed in a similar as the average variation from the mean delay in that 2 minutes interval.
+For each row in the table, the average delay is computed over the pings performed in the 2 minutes interval or the 120 pings gathered.
+The jitter is also computed similarly as the average variation from the mean delay in that 2 minutes interval.
 The result is displayed in the table \ref{table:exp1}.
 \paragraph{}
-The data collected shows that the performance of the system is very good. Even though the delay in the \gls{wsn} is high, 
- nothing can be done about that as even though the traffic in the \gls{wsn} is low, the delay is still high and this has to do with the computational time taken by the TCP/IP protocol protocol stack provided implemented by \glsreset{blip}\gls{blip} using \glsreset{6lowpan}\gls{6lowpan}.
-To reduce the delay, one should probably rethink the design of the \gls{blip} or find a way to omit it, but that is not a good idea since we want the \gls{wsn} to be part of \gls{iot}.
+The data collected shows even though the delay in the \gls{wsn} is high, 
+ nothing can be done about that because when the traffic in the \gls{wsn} is low, the delay is still high and this has to do with the computational time taken by the TCP/IP protocol protocol stack provided by \gls{blip} using that implements \gls{6lowpan}.
+To reduce the delay, one should probably rethink the design of the \gls{blip} and investigate ways to reduce processing time.
+
  \begin{figure}[htbp]
@@
  The good news in experiment 1 is that the priority of packets brings benefit to the network.
-Even though the traffic grows exponentially, the delay does not grow much and the even if the jitter follows to some extent the traffic's trend but still the average delay shows good result.
+Even though the traffic grows exponentially, the average delay stays constant.
+However, the jitter follows the traffic's trend, but still the average delay shows good results.
 
@@
 \paragraph{}
-The second experiment's goal is to measure the additional delay coming from the Gateway Packet transformation process whose goal to handle and hide the heterogeneity of the system.
-The result gathered reflects the average delay and jitter computed over elapsed time from the sniffing of the packet to the transformation and sending to the recipient.
+The second experiment's goal is to measure the contribution of the Gateway Packet transformation process to the overall communication delay.
+The results gathered reflect the average delay and jitter computed over the elapsed time starting from the sniffing of the packet to the transformation and sending to the recipient.
 Table \ref{table:exp2} shows the result that was collected from this experiment as it was done on 200 packets that were sniffed and transformed by the system.
@@
 \paragraph{}
-The transformation process's elapsed time is counted in nanoseconds and from \ref{table:exp2}, we can see that the average delay is on average 0.1 milliseconds whereas the jitter is 0 milliseconds meaning that the process's time varies between a few nanoseconds to at most 150 nanoseconds.
+The transformation process's elapsed time is measured in milliseconds and from table \ref{table:exp2}, one can see that the average delay is on average 0.1 milliseconds whereas the jitter is around 0 milliseconds meaning that the process's time varies between a few nanoseconds to at most 150 nanoseconds.
 Depending on the machine's load.
+The distribution of the delay frequency is shown the histogram depicted in figure \ref{fig:histogram}.
+From this figure one can conclude that the delay is normally distributed
+\begin{figure}[htbp]
+ \centering
+ \includegraphics[scale=0.9]{img/histogram.JPG}
+ \caption{Delay frequency histogram}
+ \label{fig:histogram}
+ \end{figure}
 \paragraph{}
  The Packet transformation process is a component that shows excellent results.
-Before conducting experiment 2, we were persuaded that there will be additional delay originating from this process and it could be responsible for the high delay seen in experiment 2 but the results shows otherwise and that the component is not to be put responsible for such important delay seen in experiment 1.
+Before conducting experiment 2, we were persuaded that there will be additional delay originating from this process and it could be responsible for the high delay seen in experiment 1.
+However, the results show otherwise and that the component is not to be put responsible for such an important delay seen in experiment 1.
 
@@
 \paragraph{}
-This project enabled us to establish an infrastructure as part of the \glsreset{iot}\gls{iot} and move towards having the smart home system's integrated within the smart.
-At this stage and counting solely on this work, one cannot clearly see the integration of smart home systems within the \gls{iot}, but this is an infrastructure that provides two main services which are the establishment of the two-way communication between the smart home devices and the outside world which may be the utility for instance.
-This two-way communication enables the \glsreset{wsn}\gls{wsn} to monitor home appliances and send sensor data to a repository which monitors the home's consumption and this device can be used for simulating a real world \gls{ami}.
-The other part of the communication is the ability of clients to control appliances status by turning them On and Off.
-\paragraph{}
-The second contribution of this work the smart management of the heterogeneity.
-The communication from the mote to the mobile client is built on a set of technologies and the journey of a simple \gls{tcp} datagram move through many different data link layer standards and different network technologies.
+This project enabled us to create an infrastructure that would be part of the \glsreset{iot}\gls{iot} and move towards having the smart home system's integrated within the Smart Grid.
+At this stage and counting solely on this work, one cannot clearly see the integration of smart home systems within the \gls{iot}, but this is an infrastructure that provides two main services which are the establishment of the two-way communication between the smart home devices and the outside world, and the heterogeneity handler.
+The two-way communication enables the \glsreset{wsn}\gls{wsn} to monitor home appliances and send sensor data to a repository and this last can be used to simulate a real world \gls{ami}.
+The second direction in this communication gives the ability for clients to control appliances by turning them On or Off.
+\paragraph{}
+The second contribution of this work is the smart management of the heterogeneity.
+The communication from the mote to the mobile client uses a set of technologies and the journey of a simple \gls{tcp} datagram goes through many different Data Link Layer standards and different Network technologies.
 Starting from the mote sending sensor data, The TCP data is encapsulated into an \gls{ipv6} packet that is encapsulated into a Zigbee frame.
 Afterward, the packet is extracted from the Zigbee frame and encapsulated into a \gls{usb} frame.
-Once it arrives to the gateway, the \gls{tcp} datagram is extracted from the \gls{ipv6} which is extracted from the \gls{usb} frame and encapsulated in an \gls{ipv4} packet that encapsulated into a Wi-Fi frame and sent until it reaches the mobile.
+Once it arrives to the gateway, the \gls{tcp} datagram is extracted from the \gls{ipv6} which is extracted from the \gls{usb} frame and encapsulated in an \gls{ipv4} packet that is encapsulated into a Wi-Fi frame and sent until it reaches the mobile.
 That is said, it clear that the system is heterogeneous and making this heterogeneity transparent was a contribution of this work.
 \paragraph{}
-The \glsreset{wsn}\gls{wsn} was built making use of the very recent \glsreset{6lowpan}\gls{6lowpan} whose goal is to provide the TCP/IP protocol stack to the smallest devices.
-The \gls{wsn} is autonomous in the sense that it can work independently and indefinitely by monitoring the home appliances.
-\paragraph{}
-All in all, this project as a big picture is a testbed providing the main components of a smart home management with an intelligent \gls{wsn}, a robust and powerful middleware application that is able to use the sensory data and use it to control the home's appliances and a mobile application able to provide the users remote monitoring of their smart homes and control.
-This was possible by making the system as part of the \glsreset{iot}\gls{iot}.
+The \glsreset{wsn}\gls{wsn} was built using the very recent \gls{6lowpan} whose goal is to provide the TCP/IP protocol stack to the smallest devices.
+The \gls{wsn} is autonomous in the sense that it can monitor the home appliances independently and indefinitely.
+\paragraph{}
+All in all, this project as a big picture creates a test-bed providing the main components of a smart home management with an intelligent \gls{wsn}, a robust and powerful middle ware application that is able to use the sensory data and use it to control the home's appliances and a mobile application able to provide the users with remote monitoring and control of their smart homes.
+This was possible by making the system part of the \glsreset{iot}\gls{iot}.
 \chapter{Future Work}
 \paragraph{}
-This thesis project's was to establish the infrastructure for integrating smart home within \glsreset{sg}\gls{sg} and make it part of the \glsreset{iot}\gls{iot}.
+The goal of this thesis was to establish the infrastructure for integrating smart home within \glsreset{sg}\gls{sg} and make it part of the \glsreset{iot}\gls{iot}.
 This project enabled the deployment of the two-way communication as an infrastructure that can be used to build the \glsreset{dr}\gls{dr}.
-\gls{dr} is one of the most important topics in smart grids.
+\gls{dr} is one of the most discussed topics in Smart Grids.
 It enables the utility and households to communicate with a set of services that any of them may request from the other.
-The current state of the art does not incorporate this communication as part of the smart home systems \cite{ref15}.
+The current state of the art does not incorporate this communication interface in smart home systems \cite{ref15}.
 But as there is a need for this communication, \gls{dr} will have to be implemented as part of the smart home package.
 \paragraph{}
-Demand reponse integration is a future work for this study.
-\gls{dr} has a set of service that have to be implemented in the smart home packages.
+Demand response integration is a future work for this study.
+\gls{dr} includes a set of services that have to be implemented in the smart home packages.
 In \gls{dr}, the consumer has the possibility to ask the utility about the instantaneous price quote for electricity.
-The consumer when in need of high priority in energy may ask the utility to provide such to provide a minimum of energy that will be guaranteed.
-Of course this means at a cost, but this agreement is part of the \gls{dr} and should be implemented at the level of smart home systems.
+The consumer when in need of high priority in energy, may ask the utility to guarantee the delivery of a certain quantity of electricity.
+Of course this comes with a cost, but this agreement is part of the \gls{dr} and should be implemented at the level of smart home systems.
 Consumers can also help utilities at difficult times.
-For instance, when the utility cannot provide enough energy at peak times, the utility may ask some consumers to significantly cut their consumption and as a result be billed a reduced cost for the electricity that is consumed.
-In general, the smart home aims to reduce energy consumption and as the utility at times may need to reduce its energy production especially in peak hours, the \gls{dr} will help the utility ask the smart home to reduce the home consumption.
-\paragraph{}
-As this project aims to integrate the smart home devices within the \glsreset{iot}\gls{iot}, there is a need to integrate \glsreset{rfid}\gls{rfid} within this system as a future work.
-The \gls{rfid} is a building block within the \gls{iot} and will equip the devices with a low-power radio identification that will allow more flexibility to the system, as the external systems may be able to recognize uniquely each device in the \gls{iot} without counting primarily on the \gls{ipv6} which may change from time to time and therefore cannot keep track of the each device's information and monitor it reliably.
-\paragraph{}
-This system provides near real-time sensory data that is not much exploited intelligently.
+For instance, when the utility cannot provide enough energy at peak times, the utility may ask some consumers to reduce their consumption and as a result be billed a reduced cost for the electricity that is consumed.
+In general, the smart home aims to reduce energy consumption and as the utility needs to reduce its energy production in peak hours, the \gls{dr} will help the utility ask the smart homes to reduce their consumption.
+\paragraph{}
+As this project aims to integrate the smart home devices within the \glsreset{iot}\gls{iot}, \glsreset{rfid}\gls{rfid} must be part of this system as a future work.
+The \gls{rfid} is a building block within the \gls{iot} and will equip the devices with a low-power radio identification mechanism hat will give more flexibility to the system.
+therefore, the external systems may be able to recognize uniquely each device in the \gls{iot} without counting primarily on their \gls{ipv6} address, which may change from time to time.
+\paragraph{}
+This system provides near real-time sensory data that is not exploited intelligently.
 The only use of this data is to enable users to monitor their home consumption.
-As far as the system is concerned, there is a need for a more intelligent exploitation of this very rich data that can be used by intelligent algorithm to automatically and autonomously reduce the household's energy consumption.
-Such algorithms can be based on machine learning and data mining techniques making the direction of this a cross disciplinary project aiming to reduce the consumption of household without reducing the utility of users.
-In other words, build algorithms that will reduce consumption of households while the user is not noticing change.
+As far as the system is concerned, one must more intelligently exploit this data that can be used by smart algorithms to automatically and autonomously reduce the household's energy consumption.
+Such algorithms can be based on machine learning and data mining techniques making it a cross disciplinary project aiming to reduce the consumption of households without reducing the utility of users.
+In other words, building algorithms that will reduce consumption of households while the user is not noticing change.
 \paragraph{}
@@
 \paragraph{}
-The system provides different services, but there is not authentication or authorization involved.
-This means that any users knowing the mote connected to the appliance he/she is interested in, he/she can control it and there is no existing party or component that will control such a communication.
-As a future work, the control of the system's access should be monitored in order for the system to be more secure and robust to external attacks.
-
-%\addcontentsline{toc}{chapter}{Index}
+The system provides different services, but there is no authentication or authorization involved to access the system.
+In fact, any user knowing the mote's \gls{ipv6} address can control it because there is no existing party or component that controls the flow of information.
+As a future work, the system should have a third party component whose goal is to make the system secure.
+
 \renewcommand{\bibname}{Index}
@@
 \section{Hardware needed}
-In order for the user to install the system, one should make sure he/she is having a computer that will be used for the sole purpose of hosting different services.
+In order for the user to install the system, one should dedicate a computer to host the different services of the system.
 The crossbow professional kit is a necessity and a router must be available to connect the different components through Wi-Fi.
-A smartphone or a separate computer with Android emulator may de used.
+A smart phone or a separate computer with Android emulator may de used.
 

### 2013-07-08 doc/thesis/report.tex
Things is in italic and typos fixed

@@
 \paragraph{}
-Information Technology is advancing exponentially and largely impacting our daily lives.
+Information Technology is growing exponentially and largely impacting our daily lives.
 The current evolution in computing is the \gls{iot}.
-It is smoothly evolving from an Internet of people to and Internet of Things.
-By 2020, it is expected to have 50 billion "Things" connected to the Internet.
-\paragraph{}
-However such a migration induces a strong level of complexity when handling interoperability between heterogeneous Things, i.g. Wireless Sensors, RFID, and mobile devices.
+It is smoothly evolving from an Internet of people to and Internet of \textit{Things}.
+By 2020, it is expected to have 50 billion \textit{Things} connected to the Internet.
+\paragraph{}
+However, such a migration induces a strong level of complexity when handling interoperability between heterogeneous \textit{Things}, e.g. Wireless Sensors, RFID, and mobile devices.
 \paragraph{}
 In this thesis, we focus on the integration of wireless sensors into \gls{iot}.
-This is q prologue for the future integration of \gls{sg} into \gls{iot}.
-In this work, we deployed a smart building testbed for energy optimization, and for integration into \gls{sg}\glsreset{sg}.
-\paragraph{}
-The testbed built provides IPv6 gatewaying, remote control of wireless sensor nodes which is built on top of \gls{6lowpan}\glsreset{6lowpan}, and mobile access to the system.
+This is a prologue for the future integration of \gls{sg} into \gls{iot}.
+In this work, we deployed a smart building test-bed for energy optimization, and for integration into \gls{sg}\glsreset{sg}.
+\paragraph{}
+The test-bed built provides IPv6 gatewaying, remote control of wireless sensor nodes which is built on top of \gls{6lowpan}\glsreset{6lowpan}, and mobile access to the system.
 \paragraph{}
@@
 Now that the general architecture has been defined, the question that should be asked is whether all these components operate in the same layer in the \gls{osi} model and if not, in which layer does each of these components reside.
-As an answer to such a question, figure \ref{fig:layered_view} is here to show on what layer does every component act on.
+As an answer to such a question, figure \ref{fig:layered_view} is here to show in what layer does every component operates.
 
@@
 The installation and initialization of this component will be detailed in appendix A.
-\section{\gls{ipv4}/\gls{ipv6} Gatewaying Process}
+\section{IPv4/IPv6 Gatewaying Process}
 \paragraph{}
@@
 Depending on the machine's load.
+The distribution of the delay frequency is shown the histogram depicted in figure \ref{fig:histogram}.
+From this figure one can conclude that the delay is normally distributed
+\begin{figure}[htbp]
+ \centering
+ \includegraphics[scale=0.9]{img/histogram.JPG}
+ \caption{Delay frequency histogram}
+ \label{fig:histogram}
+ \end{figure}
 \paragraph{}

### 2013-07-08 doc/thesis/report.tex

@@
 The current evolution in computing is the \gls{iot}.
-It is smoothly evolving from an Internet of people to and Internet of \textit{Things}.
+The Internet is smoothly evolving from an Internet of people to and Internet of \textit{Things}.
 By 2020, it is expected to have 50 billion \textit{Things} connected to the Internet.
@@
 \paragraph{}
-In this thesis, we focus on the integration of wireless sensors into \gls{iot}.
-This is a prologue for the future integration of \gls{sg} into \gls{iot}.
-In this work, we deployed a smart building test-bed for energy optimization, and for integration into \gls{sg}\glsreset{sg}.
+This thesis focuses on the development of an \gls{iot} test-bed, using wireless sensors, that will be used for smart grid applications.
 \paragraph{}

### 2013-07-09 doc/thesis/report.tex
Abstract in French and Arabic fixed

@@
 Your advices whether they are computer science related, or even every day life related wise words things I will never forget.
-\paragraph*{}
+\paragraph{}
 I am also thankful for having Dr.
@@
 The current evolution in computing is the \gls{iot}.
-The Internet is smoothly evolving from an Internet of people to and Internet of \textit{Things}.
+The Internet is smoothly evolving from an Internet of People to and Internet of \textit{Things}.
 By 2020, it is expected to have 50 billion \textit{Things} connected to the Internet.
@@
 \paragraph{}
+Another migration is taking place in the power grid that is evolving towards the \gls{sg}.
+In fact, the Information Technology will play a key role in implementing needed functions such as \gls{dr}\glsreset{dr} and integrating Distributed and Renewable Energy Sources into \gls{sg}.
+The \gls{iot} is expected to enable a smooth integration of Distributed Renewable Energies within the electric grid.
+\paragraph{}
 This thesis focuses on the development of an \gls{iot} test-bed, using wireless sensors, that will be used for smart grid applications.
-\paragraph{}
 The test-bed built provides IPv6 gatewaying, remote control of wireless sensor nodes which is built on top of \gls{6lowpan}\glsreset{6lowpan}, and mobile access to the system.
 \paragraph{}
-The system is using different technologies such as \gls{tinyos}, \gls{nesc}, \gls{j2se} and  \gls{http}.
-
-\chapter{Résumé}
-\paragraph{}
-La technologie de l'information progresse de façon exponentielle et prend largement plus impact sur nos vies quotidiennes.
-L'évolution actuelle de l'informatique est appelé \glsreset{iot}\gls{iot}.
-il est prévu d'avoir en 2020 50 milliards "choses" connecté à l'Internet.
-\paragraph{}
-Cependant, une telle migration induit un fort niveau de complexité lorsque l'interopérabilité entre en jeu entre differentes choses hétérogènes, à savoir capteurs sans fil, RFID et l'informatique mobile.
-\paragraph{}
-Dans cette thèse, nous nous concentrons sur l'intégration de capteurs sans fil dans \gls{iot}.
-Cela conduit vers une future intégration de Smart Grid (SG) dans \gls{iot}.
-Au cours de ce projet, nous avons déployé un banc d'essai intelligente du bâtiment pour l'optimisation de l'énergie.
-\paragraph{}
-Ce système d'information utilise différentes technologies  telles que TinyOS (TinyOS), \glsreset{nesc}\gls{nesc}, Java Standard Edition (J2SE), HTTP.
-
+This study results in the creation of a test-bed that provides an infrastructure for integrating \gls{sg} into \gls{iot} and build the overlying services such as \gls{dr}.
+Therefore, the system's performance was put to test by conducting two experiments where in each the delay and jitter were measured to make sure the system can handle overlying services in future projects.
+
+\chapter*{Résumé}
+\paragraph{}
+Les nouvelles technologies de l'information connaissent une croissance exponentielle et prennent de plus en plus d'impact sur nos vies.
+L'évolution en cours de l'informatique est \glsreset{iot}\gls{iot}.
+L'Internet est discrètement en train de transitioner d'un réseau pour les personnes vers un reseau pour les \textit{Choses}.
+En 2020, il est prévu d'avoir 50 milliards \textit{Choses} connectées à Internet.
+\paragraph{}
+Cependant, une telle migration induit un niveau élevé de complexité lors de la manipulation de l'interopérabilité entre \textit{Choses} hétérogènes, par exemple, Capteurs sans fil, RFID, et appareils mobiles.
+\paragraph{}
+Une autre migration commence dans le réseau électrique qui est entrain de se transformer en un nouveau réseau appelé \glsreset{sg}\gls{sg}.
+Ce nouveau réseau compte principalement sur les avantages des technologies de l'information qui lui permettra d'intégrer de nouvelles fonctionnalités telle que \glsreset{dr}\gls{dr} et la l'exploitation décentralisée des énergies renouvelables.
+Pour cela, \gls{iot} va permettre une douce intégration des énergies renouvelables dans le réseau électrique.
+\paragraph{} 
+Cette thèse porte à développer un \gls{iot} prototype qui utilise des capteurs sans fil pouvant être utilisés dans le Smart Grid.
+Ce prototype fournit des services tels qu'une passerelle IPv6, le control à distance des capteurs sans fil, et un accès mobile au système.
+\paragraph{}
+Cette étude conduit à la création d'un prototype qui propose une infrastructure for intégrer \gls{sg} au sein de \gls{iot} et donc permettre de développer des services tel que \gls{dr}.
+Pour cela, les performances du système doivent être mises en test pour s'assurer de la possibilités de compter sur ce prototype.
+Dans ce contexte, deux expériences ont été menées mesurant le retard et le gigue pour savoir si le système peut supporter d'autres applications.
 \setcode{utf8}
 
-%\chapter{RL{ملخص}}
-
-\chapter{Arabic}
+\chapter*{\RL{ملخص}}
+%\chapter{Arabic}
 \paragraph{}
 \begin{arabtex}
-تكنولوجيا المعلومات تتقدم باطراد وجود أكبر تأثير على حياتنا كما تقدم الوقت.
-الثورة القادمة في مجال الحوسبة يسمى إنترنت الأشياء (قام المحفل).
-انها اساسا شبكة الانترنت ونحن نعلم في الوقت الحاضر إلا أن الأشياء المادية (ثلاجة، تلفزيون، تكييف هواء ...) وسوف تكون القوى الفاعلة الرئيسية في هذا النظام المعقد.
-وفي هذا السياق، فإن المشروع يبني على هذه الفكرة على سرير اختبار لترشيد استهلاك الطاقة الذكية الرئيسية.
-ال يسمح المشروع للمستخدمين لرصد استهلاك الطاقة لديها عبر الذكية الخاصة بهم الهاتف وأيضا يمكن للمستخدمين التحكم في الأجهزة عن طريق تشغيل وإيقاف الأجهزة عن بعد.
+تكنولوجيا المعلومات تنمو بشكل كبير وتؤثر إلى حد كبير على حياتنا اليومية.
+التطور الحالي في مجال الحاسوب هو  .ان الإنترنت تتطور من إنترنت لناس إلى الإنترنت  للأشياء.
+بحلول عام 2020, نتوقع أن يكون لديها 50 مليار شيء يتصل في شبكة الإنترنت.
 \end{arabtex}
@@
 \begin{arabtex}
-وينقسم النظام إلى ثلاثة عناصر رئيسية. كل واحد منهم يستخدم
-التكنولوجيات المختلفة، مث لTinyOS، NESC، ومعيار جافا
-طبعة J2SE، بروتوكول نقل النص التشعبي HTTP وغيرها من التكنولوجيات
-التي سيتم مناقشتها في التقرير أطروحة.
+ومع ذلك، فإن هذه الهجرة تحث على مستوى قوي من التعقيد في التعامل مع وجود التوافق بين أشياء غير متجانسة،و على سبيل المثال أجهزة الاستشعار اللاسلكية, والأجهزة النقالة.
 \end{arabtex}
@@
 \begin{arabtex}
-
-في هذه المرحلة، ونظام مراقبة يسمح، ولكن تشغيل وإيقاف تطبيق صحيفة-
- ليس هناك حتى الآن وذلك أساسا بسبب غياب المعدات ولكن البلاغ لهذا الغرض هناك.
+هجرة أخرى تقع في شبكة الكهرباء التي تتطور نحو الشبكة الكهربائية الذكية.
+في الواقع، فإن تقنيات الحاسوب تلعب دورا رئيسيا في تنفيذ الوظائف المطلوبة مثل دمج مصادر الطاقة المتجددة والموزعة في الشبكة الكهربائية الذكية.
+ومن المتوقع للإنترنت أن تمكن الإندماج الضريف للطاقات المتجددة الموزعة في الشبكة الكهربائية الحالية.
 \end{arabtex}
-
+\paragraph{}
+\begin{arabtex}
+ويركز هذi العمل على إنجاز سريراً للاختبار، وذلك بإستخدام أجهزة الاستشعار اللاسلكية، التي سيتم استخدامها لتكبصها في الشبكة الذكية.
+في هذا الاختبار، تم بناء عبارة شبكية، جهاز التحكم عن بعد في عقد الاستشعار اللاسلكية، والدخول المتحرك إلى النظام.
+\end{arabtex}
+\paragraph{}
+\begin{arabtex}
+وصلت هذه الدراسة إلى سنع سرير الاختبار الذي يوفر بنية تحتية لدمج الشبكة الذكية إلى شبكة الانترنت وبناء العديد من الخدمات الفوقية.
+ولذلك، وضع هذا النظام إلى الاختبار عن طريق تقويم تجربتين حيث في كل تجربة، قيس التأخير والغضب للتأكد من إمكانية النظام للتعامل مع الخدمات في بنيات فوقية التي تغمر في المشاريع المستقبلية.
+\end{arabtex}
 \singlespacing

### 2013-07-09 trunk/doc/thesis/report.tex
Final modifications done

@@
 \paragraph{}
-Information Technology is advancing exponentially and largely impacting our daily lives.
+Information Technology is growing exponentially and largely impacting our daily lives.
 The current evolution in computing is the \gls{iot}.
-It is smoothly evolving from an Internet of people to and Internet of Things.
-By 2020, it is expected to have 50 billion "Things" connected to the Internet.
-\paragraph{}
-However such a migration induces a strong level of complexity when handling interoperability between heterogeneous Things, i.g. Wireless Sensors, RFID, and mobile devices.
-\paragraph{}
-In this thesis, we focus on the  integration of wireless sensors into \gls{iot}.
-This is q prologue for the future integration of \gls{sg} into \gls{iot}.
-In this work, we deployed a smart building testbed for energy optimization, and for integration into \gls{sg}\glsreset{sg}.
-\paragraph{}
-The testbed built provides IPv6 gatewaying, remote control of wireless sensor nodes which is built on top of \gls{6lowpan}\glsreset{6lowpan}, and mobile access to the system.
-\paragraph{}
-The system is using different technologies such as \gls{tinyos}, \gls{nesc}, \gls{j2se} and  \gls{http}.
-
-\chapter{Résumé}
-\paragraph{}
-La technologie de l'information progresse de façon exponentielle et prend largement plus impact sur nos vies quotidiennes.
-L'évolution actuelle de l'informatique est appelé \glsreset{iot}\gls{iot}.
-il est prévu d'avoir en 2020 50 milliards "choses" connecté à l'Internet.
-\paragraph{}
-Cependant, une telle migration induit un fort niveau de complexité lorsque l'interopérabilité entre en jeu entre differentes choses hétérogènes, à savoir capteurs sans fil, RFID et l'informatique mobile.
-\paragraph{}
-Dans cette thèse, nous nous concentrons sur l'intégration de capteurs sans fil dans \gls{iot}.
-Cela conduit vers une future intégration de Smart Grid (SG) dans \gls{iot}.
-Au cours de ce projet, nous avons déployé un banc d'essai intelligente du bâtiment pour l'optimisation de l'énergie.
-\paragraph{}
-Ce système d'information utilise différentes technologies  telles que TinyOS (TinyOS), \glsreset{nesc}\gls{nesc}, Java Standard Edition (J2SE), HTTP.
-
+The Internet is smoothly evolving from an Internet of People to and Internet of \textit{Things}.
+By 2020, it is expected to have 50 billion \textit{Things} connected to the Internet.
+\paragraph{}
+However, such a migration induces a strong level of complexity when handling interoperability between heterogeneous \textit{Things}, e.g. Wireless Sensors, RFID, and mobile devices.
+\paragraph{}
+Another migration is taking place in the power grid that is evolving towards the \gls{sg}.
+In fact, the Information Technology will play a key role in implementing needed functions such as \gls{dr}\glsreset{dr} and integrating Distributed and Renewable Energy Sources into \gls{sg}.
+The \gls{iot} is expected to enable a smooth integration of Distributed Renewable Energies within the electric grid.
+\paragraph{}
+This thesis focuses on the development of an \gls{iot} test-bed, using wireless sensors, that will be used for smart grid applications.
+The test-bed built provides IPv6 gatewaying, remote control of wireless sensor nodes which is built on top of \gls{6lowpan}\glsreset{6lowpan}, and mobile access to the system.
+\paragraph{}
+This study results in the creation of a test-bed that provides an infrastructure for integrating \gls{sg} into \gls{iot} and build the overlying services such as \gls{dr}.
+Therefore, the system's performance was put to test by conducting two experiments where in each the delay and jitter were measured to make sure the system can handle overlying services in future projects.
+
+\chapter*{Résumé}
+\paragraph{}
+Les nouvelles technologies de l'information connaissent une croissance exponentielle et prennent de plus en plus d'impact sur nos vies.
+L'évolution en cours de l'informatique est \glsreset{iot}\gls{iot}.
+L'Internet est discrètement en train de transitioner d'un réseau pour les personnes vers un reseau pour les \textit{Choses}.
+En 2020, il est prévu d'avoir 50 milliards \textit{Choses} connectées à Internet.
+\paragraph{}
+Cependant, une telle migration induit un niveau élevé de complexité lors de la manipulation de l'interopérabilité entre \textit{Choses} hétérogènes, par exemple, Capteurs sans fil, RFID, et appareils mobiles.
+\paragraph{}
+Une autre migration commence dans le réseau électrique qui est entrain de se transformer en un nouveau réseau appelé \glsreset{sg}\gls{sg}.
+Ce nouveau réseau compte principalement sur les avantages des technologies de l'information qui lui permettra d'intégrer de nouvelles fonctionnalités telle que \glsreset{dr}\gls{dr} et la l'exploitation décentralisée des énergies renouvelables.
+Pour cela, \gls{iot} va permettre une douce intégration des énergies renouvelables dans le réseau électrique.
+\paragraph{} 
+Cette thèse porte à développer un \gls{iot} prototype qui utilise des capteurs sans fil pouvant être utilisés dans le Smart Grid.
+Ce prototype fournit des services tels qu'une passerelle IPv6, le control à distance des capteurs sans fil, et un accès mobile au système.
+\paragraph{}
+Cette étude conduit à la création d'un prototype qui propose une infrastructure for intégrer \gls{sg} au sein de \gls{iot} et donc permettre de développer des services tel que \gls{dr}.
+Pour cela, les performances du système doivent être mises en test pour s'assurer de la possibilités de compter sur ce prototype.
+Dans ce contexte, deux expériences ont été menées mesurant le retard et le gigue pour savoir si le système peut supporter d'autres applications.
 \setcode{utf8}
 
-%\chapter{RL{ملخص}}
-
-\chapter{Arabic}
+\chapter*{\RL{ملخص}}
+%\chapter{Arabic}
 \paragraph{}
 \begin{arabtex}
-تكنولوجيا المعلومات تتقدم باطراد وجود أكبر تأثير على حياتنا كما تقدم الوقت.
-الثورة القادمة في مجال الحوسبة يسمى إنترنت الأشياء (قام المحفل).
-انها اساسا شبكة الانترنت ونحن نعلم في الوقت الحاضر إلا أن الأشياء المادية (ثلاجة، تلفزيون، تكييف هواء ...) وسوف تكون القوى الفاعلة الرئيسية في هذا النظام المعقد.
-وفي هذا السياق، فإن المشروع يبني على هذه الفكرة على سرير اختبار لترشيد استهلاك الطاقة الذكية الرئيسية.
-ال يسمح المشروع للمستخدمين لرصد استهلاك الطاقة لديها عبر الذكية الخاصة بهم الهاتف وأيضا يمكن للمستخدمين التحكم في الأجهزة عن طريق تشغيل وإيقاف الأجهزة عن بعد.
+تكنولوجيا المعلومات تنمو بشكل كبير وتؤثر إلى حد كبير على حياتنا اليومية.
+التطور الحالي في مجال الحاسوب هو  .ان الإنترنت تتطور من إنترنت لناس إلى الإنترنت  للأشياء.
+بحلول عام 2020, نتوقع أن يكون لديها 50 مليار شيء يتصل في شبكة الإنترنت.
 \end{arabtex}
@@
 The electrical grid knows numerous problems such as blackouts, and current instabilities that cost billions of dollars yearly to both the utilities and insurances.
-The grid suffers from the problem of peak hours where energy consumption is at its highest level and that one single utility cannot provide.
-\paragraph{}
-The \glsreset{sg}\gls{sg} comes as a solution to the problems of the traditional electrical grid.
+\paragraph{}
+The global energy demand is significantly growing and the sources of energy such as fossil energies is becoming more scarce.
+Therefore, there is a need to move from depending on fossil energy sources to renewable energies such as Solar energy, wind energy and thermal energy.
+In addition, these renewable energies should be distributed throughout the power grid and permit to have less energy transmission costs.
+\paragraph{}
+The \glsreset{sg}\gls{sg} comes as a solution to the weaknesses of the traditional electrical grid.
 \gls{sg} is an intelligent, auto-balancing, self-monitoring, and self-healing electrical grid \cite{ref2} and whose energy sources are distributed between utilities and consumers.
@@
 \glsreset{sg}\gls{sg} should benefit from the \gls{iot}.
-The integration of the \gls{sg} into the \gls{iot} constitutes the main topic of this study and goes along with increasing interest recently raised in both academia and industry, on the inherent correspondence between \gls{sg} and \gls{iot}.
+This integration should constitute the main topic of this study and goes along with increasing interest recently raised in both academia and industry, on the inherent correspondence between \gls{sg} and \gls{iot}.
 This  correspondence stems mainly from the need to disseminate information between the \gls{sg} components in order to make optimal decisions about energy use.
@@
 The Smart Grid needs a set of novel functionalities.
-One of these ideas is the two-way communication that will make it possible for the utility to communicate with consumers and vice versa.
+One of these is the two-way communication that will make it possible for the utility to communicate with consumers and vice versa.
 This is possible in the Internet Of Things as devices in \gls{iot} communicate by producing and consuming information.
@@
 The \glsreset{iot}\gls{iot} uses different technologies such as \gls{wsn}, \gls{rfid}, \gls{ipv6} and \gls{6lowpan}\glsreset{6lowpan} which makes the Internet of Things a very heterogeneous and complex system \cite{ref4}.
-One of the challenges in \gls{iot} is to deal with this heterogeneity in order to make the communication smooth and transparent to the parties involved.
+One of the challenges is to deal with this heterogeneity in order to make the communication smooth and transparent to the parties involved.
 This heterogeneity should become invisible to any two devices willing to communicate in the \gls{iot}.
@@
 The \glsreset{sg} \gls{sg} relies on the two-way communication to build a whole communication framework and a protocol stack namely the \glsreset{dr}\gls{dr} and other services built on top of this last.
-From the consumer's/household's perspective or more precisely the household's side,there are smart home energy management systems in the market nowadays, but none of them offers an integration with \gls{sg}.
-To have such an integration with \gls{sg}, there is a need for the two-way communication, in order to make it possible for any appliance to report its real-time energy consumption, and to be able to be controlled remotely.
-Such pillar is missing for the time being, and therefore the integration of the smart home within the smart grid is not possible.
+From the consumer's/household's perspective or more precisely the household's side,there are smart energy management systems in the market nowadays, but none of them offers an integration with \gls{sg}.
+Such an integration through the two-way communication will to make it possible for any appliance to report its real-time energy consumption, and to be controlled remotely.
+Such pillar is missing for the time being, and therefore the integration of the smart home within the smart grid is not doable.
 Also, in order for \gls{sg} to be integrated into the \gls{iot}, all devices should be able to communicate in the two ways.
@@
 \gls{sg} counts on various technologies to achieve its main goals.
-This use of different technologies comes with a problem which is the heterogeneous nature of the system.
+The use of different technologies comes at a cost which is the heterogeneous nature of the system.
 As numerous components count on different technologies, there is a need to interface them in order to make this heterogeneity transparent to the all components of the system.
@@
  \paragraph{}
-The goal of the study is to solve the problems introduced in the problem statement section.
-This means creating the two-way communication, and solving the heterogeneity that arises from integrating smart homes with the smart grid.
+The goal of the study is create the two-way communication, and solve the heterogeneity that arises from integrating smart homes with the smart grid.
 In fact, each components in the smart home will become part of the Internet of Things by making it communicate with any device in the Internet by having an \gls{ipv6} address using \gls{6lowpan} whose goal is to provide the TCP/IP to the smallest devices and also those with the least computational resources.
 \paragraph{}
-The first purpose of this study is to create a testbed having the two-way communication and also being able to make use of it in order to build richer applications based on it.
+The first purpose of this study is to create a testbed with the two-way communication.
 To do so, \gls{6lowpan} would had to be integrated within every device in the \gls{wsn}.
@@
 \paragraph{}
-The second purpose of this study is to solve the heterogeneity problem that arises from the integration of smart homes into the grid and also when making the whole system part of the \glsreset{iot}\gls{iot}.
-Solving such a problem means having an extra component that will track all  packets and change them to fit the standards of the targeted component.
-This will make of this additional component a middleman and that would make the heterogeneity problem disappear.
+The second purpose of this study is to hide the heterogeneity that arises from the integration of smart homes into the grid and also when making the whole system part of the \glsreset{iot}\gls{iot}.
+To solve such a problem, we add an extra component that will track all  packets and change them to fit the standards of the targeted component.
 
@@
 \paragraph{}
-This study aims at solving the two problems presented in section 1.3.
-To do so, a variety of tools and technologies were used for this purpose.The two problems could not be solved one independently of the other.
+A variety of tools and technologies were used for this study.The two problems could not be solved one independently of the other.
 Even though the problems seems to be separate, one cannot solve  the first problem without having to solve the second.
@@
 \paragraph{}
-The outcome of the thesis is to design and implement a test-bed that is able to show the two-way communication in action.
-It will also have to show its integration into the \glsreset{iot}\gls{iot} by showing that every node in the system can communicate autonomously and with any other node in the Internet.
+The outcome of the thesis is to design and implement a test-bed that will be integrated into the \glsreset{iot}\gls{iot} by showing that every node in the system can communicate autonomously and with any other node in the Internet.
 Another \gls{iot} concept is the ability of any node in the \gls{wsn} to consume and produce information that is communicated through the network.
@@
 Each set of components appear within a box as this last refers to what hardware are the components deployed in as it is labeled in the box.
+There are are five hardware components:
+\begin{itemize}
+\item Sensor Node
+\item Mote Sink
+\item Gateway
+\item Middle-ware
+\item Smart phone
+\end{itemize}
+Each hardware component hosts a set of software subsystems that will be explained more clearly in the sections 4.4 to 4.11.
 \begin{figure}[htbp]
@@
 \includegraphics[width=17cm,height=20cm,keepaspectratio]{img/general_architecture.jpg}
-\caption{General architecture depicting the main components of the system}
+\caption{General architecture depicting the main components of the system and the information flow}
 \label{fig:gen_architecture}
@@
 \paragraph{}
-Now that the general architecture has been defined, the question that should be asked is whether all these components operate in the same layer in the \gls{osi} model and if not, in which layer does each of these components reside.
+Now that the general architecture has been established, the question that should be asked is whether all these components operate in the same layer in the \gls{osi} model and if not, in which layer does each of these components reside.
 As an answer to such a question, figure \ref{fig:layered_view} is here to show in what layer does every component operates.
@@
 \includegraphics[scale=0.35]{img/appliance_control.jpg}
-\caption{TCP Server controlling appliance}
+\caption{TCP Server controlling appliance information flow}
 \label{fig:appliance_control}
@@
 However, the jitter follows the traffic's trend, but still the average delay shows good results.
+ \paragraph{}
+ It is to be noted that this experiment did not get to the point where the network was saturated.
+In fact, Zigbee has a bandwidth of 250 Kbps.
+However, experiment one attained less than 1 Kbps which does not reflect clearly how the network would react to such a high bandwidth utilization.
+The point is that even though the delay seems constant in experiment 1, that does not give any indications about how the delay and jitter varies as a function of the percentage of the bandwidth used.
 
@@
     \hline
-    Delay (ms) & Jitter (ms) \\ \hline
-    0.1        & 0.0         \\ 
+    Delay ($\mu s$) & Jitter ($\mu s$) \\ \hline
+    0.1        & 0.03         \\ 
     \end{tabular}
@@
 \paragraph{}
-The transformation process's elapsed time is measured in milliseconds and from table \ref{table:exp2}, one can see that the average delay is on average 0.1 milliseconds whereas the jitter is around 0 milliseconds meaning that the process's time varies between a few nanoseconds to at most 150 nanoseconds.
-Depending on the machine's load.
-The distribution of the delay frequency is shown the histogram depicted in figure \ref{fig:histogram}.
+The transformation process's elapsed time is measured in microseconds and from table \ref{table:exp2}, one can see that the average delay is on average 0.1 milliseconds whereas the jitter is around 0.03 milliseconds meaning that the process's time varies between a few microseconds to at most 150 microseconds.
+In addition, depending on the machine's load, the distribution of the delay frequency is shown the histogram depicted in figure \ref{fig:histogram}.
 From this figure one can conclude that the delay is normally distributed
@@
 
-\paragraph{}
-You should be able to get sensor data periodically. The mobile client access for turning On and Off will be explained in the Appendix C.
 \end{appendices}

### 2013-07-09 doc/thesis/report.tex
tiny modifications done

@@
   \chapter*{\bibname}\@mkboth{\MakeUppercase\bibname}{\MakeUppercase\bibname}}{%
-  \chapter{Index}}{}{}
+  \chapter{References}}{}{}
 \makeatother
@@
 The current evolution in computing is the \gls{iot}.
-The Internet is smoothly evolving from an Internet of People to and Internet of \textit{Things}.
-By 2020, it is expected to have 50 billion \textit{Things} connected to the Internet.
+It is smoothly evolving from an Internet of People to and Internet of \textit{Things}.
+By 2020, the Internet is expected to have 50 billion \textit{Things} connected to it.
 \paragraph{}
@@
 Another migration is taking place in the power grid that is evolving towards the \gls{sg}.
-In fact, the Information Technology will play a key role in implementing needed functions such as \gls{dr}\glsreset{dr} and integrating Distributed and Renewable Energy Sources into \gls{sg}.
+In fact, Information Technology will play a key role in implementing needed functions such as \gls{dr}\glsreset{dr} and integrating Distributed and Renewable Energy Sources into \gls{sg}.
 The \gls{iot} is expected to enable a smooth integration of Distributed Renewable Energies within the electric grid.
 \paragraph{}
-This thesis focuses on the development of an \gls{iot} test-bed, using wireless sensors, that will be used for smart grid applications.
-The test-bed built provides IPv6 gatewaying, remote control of wireless sensor nodes which is built on top of \gls{6lowpan}\glsreset{6lowpan}, and mobile access to the system.
-\paragraph{}
-This study results in the creation of a test-bed that provides an infrastructure for integrating \gls{sg} into \gls{iot} and build the overlying services such as \gls{dr}.
-Therefore, the system's performance was put to test by conducting two experiments where in each the delay and jitter were measured to make sure the system can handle overlying services in future projects.
+This thesis focuses on the development of an \gls{iot} test-bed, using wireless sensors, that will be used for Smart Grid applications.
+The test-bed built provides an IPv6 gateway, remote control of wireless sensor nodes which is built on top of \gls{6lowpan}\glsreset{6lowpan}, and mobile access to the system.
+\paragraph{}
+This study results in the creation of a test-bed that provides an infrastructure for integrating \gls{sg} into \gls{iot} and building overlying services such as \gls{dr}.
+Therefore, the system's performance was put to test by conducting two experiments where the delay and jitter were measured to make sure the system can handle overlying services in future projects.
 
@@
 L'évolution en cours de l'informatique est \glsreset{iot}\gls{iot}.
-L'Internet est discrètement en train de transitioner d'un réseau pour les personnes vers un reseau pour les \textit{Choses}.
+L'Internet est discrètement en train de transitioner d'un réseau pour les personnes vers un réseau pour les \textit{Choses}.
 En 2020, il est prévu d'avoir 50 milliards \textit{Choses} connectées à Internet.
@@
 Cette thèse porte à développer un \gls{iot} prototype qui utilise des capteurs sans fil pouvant être utilisés dans le Smart Grid.
-Ce prototype fournit des services tels qu'une passerelle IPv6, le control à distance des capteurs sans fil, et un accès mobile au système.
+Ce prototype fournit des services tels qu'une passerelle IPv6, le contrôle à distance des capteurs sans fil, et un accès mobile au système.
 \paragraph{}
@@
 The global electrical grid is sometimes called the largest machine ever made by man \cite{ref1}.
-It is a widespread and distributed machine that provides power to households, companies and organizations.
+It is a widespread and distributed machine that provides electricity to households, companies and organizations.
 Therefore the whole world's economy depends on it.
 The electrical grid knows numerous problems such as blackouts, and current instabilities that cost billions of dollars yearly to both the utilities and insurances.
-The grid suffers from the problem of peak hours where energy consumption is at its highest level and that one single utility cannot provide.
-\paragraph{}
-The \glsreset{sg}\gls{sg} comes as a solution to the problems of the traditional electrical grid.
-\gls{sg} is an intelligent, auto-balancing, self-monitoring, and self-healing electrical grid \cite{ref2} and whose energy sources are distributed between utilities and consumers.
+\paragraph{}
+The global energy demand is significantly growing and the sources of energy such as fossil energies are becoming scarce.
+Therefore, there is a need to move from exploiting fossil energy sources to using renewable energies such as Solar energy, wind energy and thermal energy.
+In addition, these renewable energies should be distributed throughout the power grid and permit to have less energy transportation losses.
+\paragraph{}
+The \glsreset{sg}\gls{sg} comes as a solution to the weaknesses of the traditional electrical grid.
+\gls{sg} is an intelligent, auto-balancing, self-monitoring, and self-healing electrical grid \cite{ref2} whose energy sources are distributed between utilities and consumers.
 The nature of the \gls{sg} as having distributed energy sources, reduces numerous losses that arise from the long distance transmission lines, and also from the stress and the overload of the electrical lines.
 \paragraph{}
-Another revolution began, and it is taking place in the Information Technology.
+Another revolution began, and it is taking place in Information Technology.
 It is called the \glsreset{iot}\gls{iot}.
@@
 \glsreset{sg}\gls{sg} should benefit from the \gls{iot}.
-The integration of the \gls{sg} into the \gls{iot} constitutes the main topic of this study and goes along with increasing interest recently raised in both academia and industry, on the inherent correspondence between \gls{sg} and \gls{iot}.
+This integration constitutes the main topic of this study and goes along with increasing interest recently raised in both academia and industry, on the inherent correspondence between \gls{sg} and \gls{iot}.
 This  correspondence stems mainly from the need to disseminate information between the \gls{sg} components in order to make optimal decisions about energy use.
-This information dissemination could only be optimal when deploying ubiquitous sensoring and metering, and using the available network infrastructure, i.e.
+This information dissemination could only be optimal when deploying ubiquitous sensors and meters, and using the available network infrastructure, i.e.
 the Internet.
@@
 The first item refers to the led number(1,2,3) and the second refers to the command 0 meaning turn Off and 1 means turn On.
-The string is tokenized and the data items are extracted and the command is executed accordingly.
+The string is tokenised and the data items are extracted and the command is executed accordingly.
 \paragraph{Timer fired}
@@
 To measure the performance of the system, two experiment were conducted where the goal is to quantify the performance of the system.
-There is an architectural decision madein \gls{6lowpan}, and we would like to see if it add more performance to the system.
+There is an architectural decision made in \gls{6lowpan}, and we would like to see if it add more performance to the system.
 The architectural decision is the following: There is more priority given to routed packet that the processed ones.
@@
 \paragraph{}
-This project enabled us to create an infrastructure that would be part of the \glsreset{iot}\gls{iot} and move towards having the smart home system's integrated within the Smart Grid.
+This project enabled us to create an infrastructure that would integrated into \glsreset{iot}\gls{iot} and move towards having the Smart Home systems' integrated within the Smart Grid.
 At this stage and counting solely on this work, one cannot clearly see the integration of smart home systems within the \gls{iot}, but this is an infrastructure that provides two main services which are the establishment of the two-way communication between the smart home devices and the outside world, and the heterogeneity handler.
@@
 \subsection{Install Software in Motes}
-The first thing to do is to install the backstation within the Mote sink (the mote with the \gls{usb} IO) To install it we do the following:
+The first thing to do is to install the back station within the Mote sink (the mote with the \gls{usb} IO) To install it we do the following:
 \begin{enumerate}

### 2013-07-09 trunk/doc/thesis/TUM/report_TUM.tex
AUI AND TUM REPARATED IN TWO FOLDERS

--- 2013-07-09 doc/thesis/report.tex
@@
 Mohamed Riduan Abid for his very close supervision, guidance and for the advice he kept giving me during the lifetime of the project.
-Your advices whether they are computer science related, or even every day life related wise words things I will never forget.
+Your advices whether they are computer science related, or even every day life related, are wise words that I will never forget.
 \paragraph{}
@@
 Une autre migration commence dans le réseau électrique qui est entrain de se transformer en un nouveau réseau appelé \glsreset{sg}\gls{sg}.
-Ce nouveau réseau compte principalement sur les avantages des technologies de l'information qui lui permettra d'intégrer de nouvelles fonctionnalités telle que \glsreset{dr}\gls{dr} et la l'exploitation décentralisée des énergies renouvelables.
+Ce nouveau réseau compte principalement sur les avantages des technologies de l'information qui lui permettra d'intégrer de nouvelles fonctionnalités telle que \glsreset{dr}\gls{dr} et l'exploitation décentralisée des énergies renouvelables.
 Pour cela, \gls{iot} va permettre une douce intégration des énergies renouvelables dans le réseau électrique.

### 2013-07-14 trunk/doc/paper- towards the internet of things/paper v1.tex

@@
 
-\title{Towards the Internet of Things: A Real-World Testbed}
+\title{Towards the Internet of Things: A Real-World Test-bed}
 \author{\IEEEauthorblockN{Nacer Khalil}
@@
 \begin{abstract}
-The internet has seen its importance grow exponentially throughout the previous two decades.
-It has started as a set of computers sharing static information written by human beings and accessed by other human beings.
-It followed mainly the publisher subscriber model.
-Later on, with the arrival of internet 2.0, the subscriber had now the ability to affect the content at the publisher's level and this basic idea gave rise to a whole new world of applications such as wikis, blogs, social networks and other applications.
-The next revolution in the internet is the introduction of machines as players within the internet.
-This will allow the machines to intercommunicate between them or what is known as machine-to-machine communication.
-More and more devices will be able to communicate and thus be part of the internet.
-This will give rise to a whole new dimension of possibilities as the focus of the internet will not only be human interactions but also machine interactions.
-As an example of autonomous machines communicating through, the components of smart grids will be part of the internet of things.
-Devices such as AMI, smart homes, wireless sensor networks will all be part of the internet of things.
-Any device within the smart grid will be uniquely identified through the network and therefore will be able to communicate.
-In this paper, we present a real-world testbed for integrating wireless sensors into the internet.
-The solution uses 6lowPAN, and addresses relevant issues at the level of the gateway.
-we used legacy hardware and proved communication.
+The Internet is smoothly migrating from an Internet of people to an Internet of Things. By 2050, it is expected to have 50 billions “Things” connected to the Internet. 
+However, such a migration induces a strong level of complexity when handling interoperability between the heterogeneous things, e.g., Wireless sensor networks, RFID, and mobile computing.
+In this context, a couple of standards have been already set, and they are currently reached the mature stage, e.g., IPv6, 6loWPAN, M2M, and IMS.
+In this paper, we focus on the integration of wireless sensor networks to the Internet.
+We shed further on the subtleties of such integration, and we deployed a real-world testbed where wireless sensors are used to control electrical appliances in a smart building.
 \end{abstract}
@@
 \begin{IEEEkeywords}
-WSN, 6lowPAN, TinyOS, IOT, cyber physical systems
+WSN, 6LoWPAN, TinyOS, IoT, Cyber Physical System
 \end{IEEEkeywords}
@@
 \section{Introduction}
-The direction that is taken by the global society is a direction leading to what is known by "always connected" model.
-People are eager to receive the information as fast as possible and whenever available.
-They also want to send information as fast as possible and whenever wanted.
-Also, these requests and responses that are dictated by the human are not all the time directed towards humans , but can be sent also to devices and machines but that is not always possible at this time being.
-The future  of internet goes into this direction: being able to connect human beings as well as machine in order to provide robust and flexible communication for all.
-The machines on the other hand need an interface to be able to communicate on the internet.There will be millions of devices in the internet both producing and consuming information \cite{ref8}.
+The Internet of Things is the future of internet.
+It is an Internet where physical devices communicate by producing and consuming information.
+This makes these devices more intelligent as they will be able to provide services but at the same time use others.
+These physical devices will be uniquely identified in the internet by making use of technologies such as IPv6 and RFID.
+There will be millions of devices in the Internet both producing and consuming information \cite{ref8}.
+To build the internet of things, there are numerous points that should be added to the actual internet to make all this possible.
+The first thing is to make the TCP/IP protocol stack based on 6LoWPAN adopted by all physical devices and wireless sensor networks.
+The identification of each physical object is possible by using technologies such as RFID which is a low power identification chip that allows one to identify each object uniquely.
+Concerning communication between physical objects, M2M provides standards to make the communication possible and robust.
+\\
+To make these physical devices cyber physical systems, they will not only need to communicate but also to sense the environment and this is possible when making use of the WSN that allow physical devices to sense the environment based on their requirements.
+\\
 In this paper, we create a small network where a machine-to-machine communication is possible.
-This can be part of what is known as internet of things \cite{ref6}.
-The testbed created is composed of a wireless sensor network(WSN), a middleware station and a mobile client within the field of smart grids where data is collected from the motes within the WSN and sent to the middleware.
+This can be part of what is known as Internet of things \cite{ref1}.
+The deployed tested  is composed of a wireless sensor network(WSN), a middleware station and a mobile client within the field of smart grids where data is collected from the motes within the WSN and sent to the middleware.
 The mobile client accesses this data and the mobile client can send messages directly to any mote within the WSN network in order to turn on or off any appliance.
-The issues that were handled are the following:
+The contributions of this paper are:
 \begin{itemize}
@@
 \end{itemize}
-To solve these issues, different programs have been used to deal with these issues.
+To solve these issues, different programs have been used to deal with these issues.  \\
+The Rest of the paper is organized as follows: Section II presents related work.
+In Section III, we provide a review of the internet of things concepts and issues.
+Section IV presents 6LoWPAN technology for low power devices such as wireless sensor networks that is used in this paper.
+Section V presents the system architecture of the testbed.
+In section VI the implementation of the system is discussed more into details and section VII presents the the measurements that are done on the testbed.
 
@@
 They also use GSM technology to transmit information whereas in our case, internet is used for this purpose.
-\
+\\
+Research discusses the benefits to replace Zigbee by Wi-Fi as this last offers higher bandwidth, non-line transmission ability, large-scale data collection and is highly cost-effective \cite{ref11}.
+Still, Zigbee has the exclusive advantage of consuming the least energy of the communication technologies that exist nowadays which is something required in SG and IoT, especially for systems that are installed far away from any maintenance team such as smart grid monitoring and self-healing devices where one cannot manage to move for the sole purpose of charging these devices.
+\\
+From the architectural point of view, integrating SG within IoT means having to address heterogeneity issues.
+An IoT gateway system based on Zigbee and GPRS protocols helps to deal partly with the heterogeneity problem and therefore enable the WSN to communicate with the mobile telecommunication network \cite{ref11}.
+Another solution to the heterogeneity problem is proposed with a new, light-weight web service transport protocol called Lean Transport Protocol (LTP) that will allow transparent exchange of web service messages between all kinds of devices.
+This protocol is platform-independent, low energy communication \cite{ref12}.
+Some researchers claim that the major source of heterogeneity the fact that there are different types of WSN devices (Micaz, Mica2, Telosb...) that do not use the same standards and protocols.
+They propose to move the communication in WSNs to an "all-IP" as this would remove most of the heterogeneity but as there are already deployed legacy WSNs, the task is complicated to conduct.
+Fortunately, the researchers sketch an architecture capable of converting all the \glspl{wsn}, new and legacy, to support IPv6 \cite{ref13}.
 
@@
 The middleware by making use of the sensor data will communicate with motes to turn on or off the light bulb.
+\\
+Ubiquitous computing and pervasive computing is an important component of the Internet of things.
+Physical objects will be able to communicate in the Internet by producing information as well as consuming information \cite{ref9}.
+The main technological components needed in the Internet of things are the Radio Frequency identification (RFID) and is a technology that enables physical devices to identify themselves through this low power technology.
+Another important technological component is the wireless sensor network (WSN) and it is a technology that gives the physical objects to sense the environment and communicate such information.
+Mobile computing and cloud computing are important components as most of the storage and processing power will come from the cloud that will provide such services to the different physical devices in the internet of things.
+\\
+The internet of things has a great impact on the every day life of people as it will affect our lives inside our homes where all physical devices will be able to communicate information by providing services and consuming others.
+More tasks will be automated, as most of the physical devices will communicate their needs to each other.
+Smart grid is another important technology made possible with the Internet of Things.
+The US national intelligence listed the Internet of things as being on six disruptive civil technologies with potential impact on the US national power \cite{ref10}.
+\\
 \section{6lowpan}
@@
 This project makes use of BLIP to provide the TCP/IP protocol stack to the devices within the wireless sensor network.
+
 \section{Architecture}
@@
 \centering
-\includegraphics[scale=0.5]{images/network_diagram.jpg}
+\includegraphics[scale=0.48]{images/network_diagram.jpg}
 \caption{Network diagram}
@@
 \centering
-\includegraphics[height=110mm,width=90mm]{images/send_sensor_data_data_flow.jpg}
+\includegraphics[scale=3.0]{images/send_sensor_data_data_flow.jpg}
 \caption{Data flow diagram for the mobile client sending on/off commands}
@@
 \section{Measurements}
+The goal of this part is to measure to what extent is the system able to operate reliable and offering some acceptable level of performance.
+As this system is to be deployed in a small environment such as an apartment or a house where each mote will be connected to one or more appliances, and motes will be closely positioned and this positioning will make it possible to make use of the multi-hop routing and therefore placing no more than one sink in each house.
+In this experiment, we try to simulate the normal work of this system where each sensor is read once every second generating traffic in the Zigbee network.
+Our goal is to measure the delay and see how does it vary in this system with the amount of traffic generated.
+As the multi-hop routing protocol gives more priority to packets to be routed than those addressed to that mote, we would like to see if this prioritization reduces the delay when having more and more motes sending packets through the network.
+TABLE \ref{table:exp1} illustrates how the delay and jitter varies as we add more and more nearby motes to the network.
+\begin{table}[htbp]
+    \begin{tabular}{lll}
+    \hline
+    number of motes & Delay (ms) & Jitter (ms) \\\hline
+    1               & 87.5         & 1.2         \\ 
+    2               & 88.1         & 2.0         \\
+    3               & 88.3        & 2.7         \\
+    4               & 88.8         & 3.1         \\
+    5               & 90.0         & 3.3         \\
+    6               & 89.5         & 3.5         \\
+    \end{tabular}
+    \caption{Experiment 1 delay and jitter}
+    \label{table:exp1}
+\end{table}
+It is clear that the average delay is not much affected by the number of motes in the network, but the jitter on the other hand is clearly affected by the number of motes in the network.
+This is mainly due to the fact that there are timeframes where the network is less busy and other timeframes where the network becomes busier.
+This is why the jitter grows as the number of motes in the network grows with it.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.6]{images/delay_jitter.JPG}
+\caption{Delay and jitter evolution by Traffic}
+\label{fig:delay_jitter}
+\end{figure}
 
 \section{Conclusion}
-The conclusion goes here.
+The Internet of things is the next revolution in computing.
+The testbed presented in this paper illustrates how can smart grid by making use of numerous technologies such as TinyOs based WSN, 6lowPan can integrate within the Internet of things and benefit from it.
+The test bed presents the 2-way communication as needed by smart grids and also discusses the architecture of the system and the issues that were handled to make the communication possible and smooth.
+\\
+The issues that were handled are mainly related to making the communication possible even if it is on a homogeneous system where the data link layer is not the same and the network layer is not the same.
+The packet is transmitted from the an IPV4 network based on the Wifi network and moves to an IPV6 network based on USB then to Zigbee.
+These were one of the numerous issues handled in this paper.
+\\
+Smart grids can and will benefit from the internet of things and making it as one of the most important applications of the internet of things and therefore enable more efficiency in the generation and consumption of data.
 

### 2013-07-15 doc/paper- towards the internet of things/paper v1.tex

--- 2013-06-18 doc/paper- towards the internet of things/paper v1.tex
@@
 \usepackage[noadjust]{cite}
+\usepackage[acronym]{glossaries} 
+\makeglossaries
 
 \begin{document}
-
-\title{Towards the Internet of Things: A Real-World Testbed}
+\input{acronyms}
+\title{Towards the Internet of Things: A Real-World Test-bed}
 \author{\IEEEauthorblockN{Nacer Khalil}
@@
 \begin{abstract}
-The internet has seen its importance grow exponentially throughout the previous two decades.
-It has started as set of computers sharing static information written by human beings and accessed by other human beings.
-It followed mainly the publisher subscriber model.
-Later on, with the arrival of internet 2.0, the subscriber had now the ability to affect the content at the publisher's level and this basic idea gave rise to a whole new world of applications such as wikis, blogs, social networks and other applications.
-The next revolution in the internet is the introduction of machines as players within the internet.
-This will allow the machines to intercommunicate between them or what is known as machine-to-machine communication.
-More and more devices will be able to communicate and thus be part of the internet.
-This will give rise to a whole new dimension of possibilities as the focus of the internet will not only be human interactions but also machine interactions.
-As an example of autonomous machines communicating through, the components of smart grids will be part of the internet of things.
-Devices such as AMI, smart homes, wireless sensor networks will all be part of the internet of things.
-Any device within the smart grid will be uniquely identified through the network and therefore will be able to communicate.
-In this paper, we present a real-world testbed for integrating wireless sensors into the internet.
-The solution uses 6lowPAN, and addresses relevant issues at the level of the gateway.
-we used legacy hardware and proved communication.
+The Internet is smoothly migrating from an Internet of people to an Internet of Things. By 2050, it is expected to have 50 billions “Things” connected to the Internet. 
+However, such a migration induces a strong level of complexity when handling interoperability between the heterogeneous things, e.g., Wireless sensor networks, \gls{rfid}\glsreset{rfid}, and mobile computing.
+In this context, a couple of standards have been already set, and they are currently reached the mature stage, e.g., \gls{ipv6}\glsreset{ipv6}, \gls{6lowpan}\glsreset{6lowpan}, \gls{m2m}\glsreset{m2m}, and \gls{ims}\glsreset{ims}.
+In this paper, we focus on the integration of wireless sensor networks to the Internet.
+We shed further on the subtleties of such integration, and we deployed a real-world testbed where wireless sensors are used to control electrical appliances in a smart building.
 \end{abstract}
@@
 \section{Introduction}
-The direction that is taken by the global society is a direction leading to what is known by "always connected" model.
-People are eager to receive the information as fast as possible and whenever available.
-They also want to send information as fast as possible and whenever wanted.
-Also, these requests and responses that are dictated by the human are not all the time directed towards humans , but can be sent also to devices and machines but that is not always possible at this time being.
-The future  of internet goes into this direction: being able to connect human beings as well as machine in order to provide robust and flexible communication for all.
-The machines on the other hand need an interface to be able to communicate on the internet.
-In this paper, we create a small network where a machine-to-machine communication is possible.
-This can be part of what is known as internet of things.
-The testbed created is composed of a wireless sensor network(WSN), a middleware station and a mobile client within the field of smart grids where data is collected from the motes within the WSN and sent to the middleware.
-The mobile client accesses this data and the mobile client can send messages directly to any mote within the WSN network in order to turn on or off any appliance.
-The issues that were handled are the following:
+The \gls{iot} is the future of Internet.
+It is an Internet where \textit{Things} communicate by producing and consuming information.
+This makes these \textit{Things} more intelligent as they are able to provide services but at the same time use others.
+These physical devices are uniquely identified in the Internet by making use of technologies such as \gls{ipv6} and \gls{rfid}.
+There will be millions of devices in the Internet both producing and consuming information \cite{ref8}.
+To build the \gls{iot}, there are key points that should be added to the actual Internet to make all this possible.
+The first point is to make the TCP/IP protocol stack based on \gls{6lowpan} adopted by all physical devices and \glspl{wsn}.
+The identification of each \textit{Thing} is possible by using \gls{rfid} which is a low power identification chip that allows one to identify objects uniquely.
+The second point is to standardize the communication and \gls{m2m} is adopted for this purpose.
+
+\
+To make of these \textit{Things} \glspl{cps}, they must have three main functionalities: sensing, computation and communication.Sensing is made possible using \glspl{wsn} that provide a basis to sense the environment.
+In addition, each mote is equipped with processing power that provides the computation functionality.
+Finally, communication is made using \gls{ipv6} as a network technology.
+
+\
+In this paper, we create a small network where a \gls{m2m} is possible.
+This network can be part of what is known as Internet of things \cite{ref1}.
+The deployed tested  is composed of a \gls{wsn}, a middle-ware, and a mobile client for smart home energy monitoring and control where data is collected from the motes within the \gls{wsn} and communicated to the middle-ware.
+The mobile client is able to monitor and visualize this data and control appliances remotely.
+The contributions of this paper are:
 \begin{itemize}
-\item The system is developed using heterogeneous networks, different data link layer technologies
-\item The WSN uses IPV6 over 6lowPAN whereas other components of the system use IPV4
+\item The system is developed in a heterogeneous system,and makes it transparent
+\item The \gls{wsn} uses \gls{ipv6} over \gls{6lowpan} and is able to communicate in the Internet
+\item The Two-Way communication is built in the system 
 \end{itemize}
-To solve these issues, different programs have been used to deal with these issues.
+The Rest of the paper is organized as follows: Section II presents related work.
+In Section III, a review of \gls{iot} concepts and issues are presented.
+Section IV introduces \gls{6lowpan} standard.
+Section V describes the system architecture.
+In section VI, the deployment of the system is discussed in details and section VII presents the experiments conducted to measure the system's performance and the findings of the system.
 
 \section{Literature review}
-Smart grid as being one of the most important applications of the internet of things makes use of different technologies such as Zigbee in the WSN \cite{ref1}.
-\cite{ref2} discusses the need to introduce IPV6 within the wireless sensor networks and discusses the the existing approaches as well as the issues related to introducing IPV6 on top of Zigbee.
-Such issues are fragmentation, frame size, addressing, and security issues.
-This paper introduces a real testbed that includes the whole TCP/IP protocol provided by the BLIP project that takes into consideration most of those issues and implement the two-way communication as needed by smart grids and measures the performance of such a system.
+\gls{sg} is perceived as one of the key applications based on \gls{iot}.
+It uses different technologies such as Zigbee, \gls{6lowpan}, \gls{m2m} and \gls{ims} \cite{ref1}.
+The need to introduce \gls{ipv6} within the wireless sensor networks is discussed as well as the existing approaches and issues related to introducing IPV6 on top of Zigbee.
+Such issues are fragmentation, frame size, addressing, and security issues \cite{ref2}.
+This paper introduces a real testbed that includes the whole TCP/IP protocol implemented by \gls{blip} that takes into consideration most of those issues and implements the two-way communication as needed by smart grids and measures the performance of such a system.
+In addition, El Kouche investigates on the widely used architectures and technologies and explains which architecture is the most suitable for a WSN within the \gls{iot} \cite{ref3}.
+\cite{ref4} discusses the requirement of an \gls{iot} gateway and propose an architecture for the system to be deployed in the gateway.
+A rather similar architecture to what is done in this paper is proposed by Yerra, Bharathi, Rajalakshmi and Desai \cite{ref7}.
+They use \gls{gsm} to communicate information whereas in our case, Internet is used for this purpose.
+
+\
+Research discusses the benefits of replacing Zigbee by Wi-Fi as this last offers higher bandwidth, non-line transmission ability, large-scale data collection, and is highly cost-effective.
+Still, Zigbee has the exclusive advantage of consuming the least energy of existing communication technologies which is something required in \gls{sg} and \gls{iot}, especially for systems deployed in difficult-to-access environments \cite{ref11}.
+
+\
+From the architectural point of view, integrating \glspl{sg} within \gls{iot} imposes the need to address heterogeneity.
+An \gls{iot} gateway system based on Zigbee and GPRS protocols helps partly in dealing with the heterogeneity problem and therefore enable the \glspl{wsn} communicate with the mobile telecommunication network \cite{ref11}.
+Another solution to the heterogeneity problem is proposed with a new, light-weight web service transport protocol called Lean Transport Protocol (LTP) that will allow transparent exchange of web service messages between all kinds of devices.
+This protocol is platform-independent, low-energy communication \cite{ref12}.
+Other researchers claim that the major source of heterogeneity arises from the fact that there are different types of WSN devices e.g.
+Micaz, Mica2, and Telosb that do not use the same standards.
+They propose to move the communication in \glspl{wsn} to "all-IP" as this would remove most of the heterogeneity.
+However, there are already deployed legacy \glspl{wsn} which makes the task is complicated.
+Fortunately, an architecture is sketched that is capable of converting all the \glspl{wsn}, new and legacy, to support IPv6 \cite{ref13}.
+
 \section{Internet of things}
-The internet of things is a new concept based on the idea that there will be much more users that are computers, machine than human beings.
-This means that machines will play an important role in the internet and they will be able to communicate autonomously without the interaction of human beings.
-Wireless sensor network will be part of the internet of things and therefore will be actively communicating in the internet by sending sensor data.
-Nowadays, the internet is mainly based on a human-to-human.
-In the future of internet called the internet of things, we will have in addition to human-to-human communication, human-to-machine communication and machine-to-machine communication.
-An example of human-to-machine communication is when the human will be able to control the wireless sensor network's sampling time.
-An example of machine-to-machine communication which will be discussed in this paper is the middleware being able to turn off the light because no one is in the room.
-The middleware by making use of the sensor data will communicate with motes to turn on or off the light bulb.
-\section{6lowpan}
-6loWPAN stands for  "IPv6 over Low power Wireless Personal Area Networks".
-It stands for the name of the working group within IETF.
-It is based on the idea that all the devices independently of their processing power or energy resources should be able to have the TCP/IP protocol stack and be part of the internet of things.
-In order to be able to build the TCP/IP protocol stack in low-power devices, there have been multiple aspects of the protocol stack that has to be optimized.
-As an example, the IP MTU(maximum transmission unit) is fixed at 1280 bytes whereas the Zigbee MTU is only 127 bytes.
-This means that not every IPV6 packet can be inserted within a Zigbee frame.
-Another issue is related to the addressing with the 128 bits address.
-in 6lowpan, the addressing via the IPV6 address is performed hierarchically to first identify is the packet is sent to a device within that network but afterwards, addressing is done by counting mainly on the 16 last bits of the IPV6 address.
-These were just two of a large set of issues that 6lowpan tries to solve in order to enable even the low-power devices to join the internet of things.
-TinyOS as being one of the most used operating system comes with its implementation of 6lowpan through a project called BLIP (Berkeley Low-power IP stack) that provides a lightweight implementation of the 6lowpan and that can be installed and used with a variety of motes as long as they support IPV6.
-This project makes use of BLIP to provide the TCP/IP protocol stack to the devices within the wireless sensor network.
-\section{Architecture}
-The project is composed of three building blocks. 
+\gls{iot} is new concept based on the idea that there will more \textit{Things} than humans connected to the Internet.
+This means that machines will play a major role in the Internet.
+The \textit{Things} will be able to communicate autonomously without the need to interact with human beings.
+\glspl{wsn} will be included in most \textit{Things} and therefore the sensory data will make a significant portion of the information flowing in the Internet.
+Nowadays, humans are the major players in the Internet but this trend is changing as there are over 12.5 billion devices connected nowadays in the Internet \cite{ref14}.
+In \gls{iot}, \gls{m2m} will be the main communication standard between different \textit{Things}.
+\gls{m2m} is used in this project between the middle-ware and the \gls{wsn} when monitoring and controlling an appliance.
+
+\
+Ubiquitous computing and pervasive computing are key paradigms in \gls{iot}.
+Physical objects will be able to communicate in the Internet by producing and consuming information \cite{ref9}.
+\gls{rfid} is a technology that identifies physical devices while consuming very little energy.
+Another important technological component is \gls{wsn} that gives the physical objects the ability to sense the environment and communicate such information.
+Mobile computing and cloud computing are important components as most of the storage and processing power will come from the cloud.
+
+\
+\gls{iot} will have a major impact on the every day life of people as it will affect our lives inside and outside our homes.
+The physical devices will be able to communicate and consume different types of information.
+More and more applications based on \gls{iot} are now being developed.
+Ideas such as smart farming, smart transportation, smart homes, smart health, smart postal and smart refrigerators are only a small subset of a myriad of applications that will come to life with \gls{iot} \cite{ref15}.
+Smart grid is another key application that benefits from \gls{iot}.
+The US national intelligence listed the Internet of things as being on six disruptive civil technologies with potential impact on the US national power \cite{ref10}.
+\section{6LoWPAN}
+\gls{6lowpan} stands for  "IPv6 over Low power Wireless Personal Area Networks".
+It stands for the name of the working group within the \gls{ietf}.
+It is based on the idea that all the \textit{Things} should be able to have the TCP/IP protocol stack and join the \gls{iot}.
+In order to be able to build the TCP/IP protocol stack in all devices, there have been multiple aspects of IP that have been modified.
+the IP maximum transmission unit(MTU) is fixed at 1280 Bytes, but the Zigbee MTU is only 127 Bytes.
+This means that not every \gls{ipv6} packets can be encapsulated within the Zigbee frame, and such issue has been addressed by \gls{6lowpan}.
+Another issue is related to the addressing with the 128 bits address; In \gls{6lowpan}, addressing via the \gls{ipv6} is performed hierarchically.
+The main purpose behind is to  identify the packet's destination network Id before forwarding it to the network.
+These were just two of a large set of issues that \gls{6lowpan} solves in order to enable even the low-power devices to join \gls{iot}.
+\gls{tinyos} as being one of the most used operating system in \glspl{wsn} comes with an implementation of \gls{6lowpan} through \gls{blip} that provides a lightweight implementation of the \gls{6lowpan}.
+This project makes use of \gls{blip} to provide the TCP/IP protocol stack to the \gls{wsn}.
+
+\section{System Deployment}
+The system is composed of four building blocks. 
 \begin{itemize}
-\item Wireless sensor network (WSN)
-\item Gateway server
-\item Middleware
-\item Mobile client
+\item Wireless Sensor Network (WSN)
+\item Gateway Server
+\item Middle-ware
+\item Mobile Client
 \end{itemize}
-The WSN communicates using zigbee as well as with the gateway server using IPV6 technology whereas whereas the communication between the gateway server, the middleware and the mobile client is done via Wifi using IPV4 technology.
-The architecture enables any component within this system to communicate with any other system independently of the communication medium used (Zigbee or Wifi) or the network technology used (IPV4 or IPV6).
-Figure 1 is a network diagram that depicts the different components of the system as well as the routes that exist for any component to communicate with another.
+The WSN uses Zigbee as the communication medium and uses \gls{ipv6} in the network layer.
+However, the communication between the gateway server, the middle-ware and the mobile client is done via Wi-Fi and \gls{ipv4} technology.
+This architecture enables any device within the system to communicate with any other device independently of the communication medium used (Zigbee or Wi-Fi) or the network technology used (\gls{ipv4} or \gls{ipv6}).
+Figure 1 is a network diagram that depicts the different components of the system as well as the interconnections that exist between the different components of the system.
 
@@
 
-\subsection{Wireless sensor network}
-The hardware is composed of motes of type crossbow MPR2600 that are used to build the wireless sensor network as shown in figure 2.From the network aspect, the wireless sensor network is based on the mesh topology.
-It is also designed in order to be an ad-hoc network.
-This means that you can place the mote any where as long as there is one link it can attain to communicate with.
-The routing is multi-hop and links are created dynamically between different motes of the WSN provided that there frames can reach the destination.
+\subsection{Wireless Sensor Network}
+The test-bed's \gls{wsn} is composed of seven motes of type Crossbow MPR2600 as shown in figure 2.From the network topology perspective, the \gls{wsn} is a multi-hop mesh network that uses \gls{aodv} for routing.
+It is an ad-hoc network which means that one can place the mote anywhere as long as there is one link it can attain to communicate with.
+The communication links are created and refreshed dynamically between different motes of the WSN provided that there frames can reach the destination.
+In addition to the seven motes in the test-bed, there is an additional mote that plays the role of a sink because it connects the \gls{wsn} to the gateway server machine.
+The connection between the sink and the gateway is based on \gls{usb}.
 
@@
 
-\subsection{Gateway server}
-The gateway server is a crucial component within the system.
-It is responsible for extracting Wifi frames and forwarding them as Zigbee frames and vice versa.
-It is also responsible for receiving IPV4 packet and transforming them into IPV6 and vice versa.
-The gateway server has other functionalities which are the ability to receive sensor data from all the motes in the wireless sensor network and forwarding them to the middleware.
-In case the link between the gateway server and the middleware is lost, the gateway server stores the received data and sends them when the link is up again.
-This feature serves as a way to avoid losing data.
+\subsection{Gateway Server}
+The gateway server is a key component in the system.
+It is responsible for extracting Wi-Fi frames and forwarding them as Zigbee frames and vice versa.
+It is also responsible for receiving IPV4 packets and transforming them into IPV6 and vice versa.
+The gateway server has other functionalities such as the ability to receive sensor data from \gls{wsn} and forward them to the middle-ware.
+In case the link between the gateway server and the middle-ware is lost, the gateway server stores the received data in a temporary data store and communicate this data once the link is up again.
+This feature serves as a way to avoid losing sensory data.
 
-\subsection{Middleware}
-The middleware is a special program that is used to hide the heterogeneity of the system from external users.
-The middleware also provides automation mechanisms to control and reduce the energy consumption, but the main features are the ability to receive data, filter it, transform it and store it in a coherent fashion in order to make the most of this data and use it smartly to reduce consumption.
-Concerning the client access, the middleware provides a set of web services that enable the clients to access all sorts of information such as the real-time consumption, daily consumption and other information inferred from the sensor data coming from the wireless sensor network.
+\subsection{Middle-ware}
+The middle-ware is a software component that is used to hide the heterogeneity of the system from external users.
+The middle-ware also provides automation mechanisms in order to control and reduce the energy consumption, but the main features are the ability to receive data, filter it, transform it and store it in a coherent fashion in order to use it smartly to reduce consumption.
+In addition, the middle-ware provides a set of web services that enable the clients to access all sorts of information such as the real-time consumption, daily, weekly and monthly consumption information inferred from the sensory data coming from the \gls{wsn}.
 
-\subsection{Mobile client}
-The mobile client application is an interactive application for android phones that enable users to get access to the real-time consumption of their homes and also to remotely control their appliance by turning on and off any appliance linked to some mote in the wireless sensor network.
-The mobile client when wanting to turn on or off some appliance sends a request directly towards the mote responsible for that action and designate the mote using its "virtual" IPV4 address that it does not own in really since it has an IPV6 address but that virtual IPV4 address is reserved for that mote and the translation is dealt with at the level of the gateway.
+\subsection{Mobile Client}
+The mobile client application is an application for Android phones that enable users to access the real-time consumption of their homes, and also to remotely control their appliance by turning On and Off any appliance linked to a mote in the \gls{wsn}.
+The mobile client when wanting to turn on or off some appliance sends a command directly to the mote responsible for controlling the mote and addresses the mote using its "virtual" \gls{ipv4} address that it does not own in really since it has an \gls{ipv6} address.
+This virtual \gls{ipv4} address is reserved for that mote and the translation is made at the gateway level.
 
-\ Now that all components have been explained, the data flow of the information is to be explained.
-As it was stated before, the any component of the system can communicate with any other independently of the data link layer technology or network layer technology.
+\ Now that all components have been introduced, the data flow of the information is to be explained.
+As it was stated, any component in the system can communicate with any other independently of the data link layer technology or network layer technology.
 
@@
 \
-As one of the main goals of this project is to provide a two-way communication between the client and sensor, we will explain both way's flow of information.
-Figure 3 depicts the flow of information going from the client to the appliance.
-As the mobile client in the system is connected to a Wifi network that uses IPV4 and the mote is a network that uses a IPV6, then there should be a process that does some transformation to incoming and outgoing packets.
-So the client sends the packet to the virtual IPV4 address of the mote.
-The gateway receives it translates the virtual IPV4 address into the real IPV6 address of the mote and sets as source address the virtual IPV6 address of the mobile client.
-The new IPV6 packet is created holding the same payload as the original IPV4 packet.
-This new IPV6 packet is to be forwarded to the wireless sensor network.
-To do so, it is passed to the tunnel that encapsulates the packet into a USB frame and sends it to the mote sink that extracts the IPV6 packet from the USB frame and encapsulates it into a Zigbee frame.
-Once the Zigbee frame arrives to the destination mote, the TCP datagram is extracted from the IPV6 packet and passed to the TCP server of the mote that reads the message and executes it by turning on/off the appliance using I2C protocol.
+One of the main goals of this project is to build the two-way communication between the client and sensor nodes.
+To do so, the flow of data in both directions is depicted in figure \ref{fig:control_appliance} and figure \ref{fig:send_data}.
+The mobile client is connected to a Wi-Fi network that uses \gls{ipv4} whereas the \gls{wsn} uses \gls{ipv6}.
+Therefore, there should be a process that control and transforms the incoming and outgoing packets.
+The client starts by sending packet to the virtual \gls{ipv4} address of the mote.
+Afterwards, the gateway receives it, translates the virtual \gls{ipv4} address into the real \gls{ipv6} address of the mote by setting as source address the virtual \gls{ipv6} address of the mobile client.
+The new \gls{ipv6} packet is created carrying the payload coming from the original packet.
+This new \gls{ipv6} packet is forwarded to the wireless sensor network using an \gls{ipv6}-over-\gls{usb} tunnel that encapsulates the packet into a \gls{usb} frame and communicates it to the mote sink that extracts the \gls{ipv6} packet from the frame and encapsulates it into a Zigbee frame.
+Once the Zigbee frame arrives to the destination mote, the TCP datagram is extracted from the packet and passed to the TCP server of the mote that reads the message and executes it by turning on/off the appliance using I2C.
 
 \
-As a second part of the two-way communication, the mote sends periodically sensor data to the middleware but again the communication is goes through several steps as depicted in figure 4.
-\linebreak 
-The mote reads periodically sends sensor data from the sensor, transforms the data and sends it.
-To send it the mote client connects to a TCP server hosted at the gateway.
-The IPV6 packet is encapsulated in a Zigbee frame that is forwarded to the mote  sink that extracts the IPV6 packet and  encapsulates it into a USB frame and forwarded to the gateway that reads the TCP datagram.
-Once the sensor data is at the gateway, the gateway forwards the sensor data to the middleware and if the link is down, the sensor data is temporarily stored in a database hosted in the gateway and once the link is up, all the stored data is sent to the middleware and cleared from the database.
+The mote on the other hand sends periodically sensory data to the middle-ware, but again the communication passes through several steps which are depicted in figure 4.
+
+\
+The mote reads periodically sensory data from the sensor, transforms the data and communicates it.
+To send it, the mote client connects to a TCP server hosted at the gateway server.
+An \gls{ipv6} packet is encapsulated in a Zigbee frame that is forwarded to the mote sink that extracts the IPV6 packet, encapsulates it into a \gls{usb} frame and forwards it to the gateway that reads the TCP datagram.
+Once the sensory data is in the gateway, it communicates it to the middle-ware, and if the link is down, the sensory data is temporarily stored in a database hosted in the gateway.
+Once the link is up, all the stored data is sent to the middle-ware and cleared from the database.
 
@@
 \centering
-\includegraphics[height=110mm,width=90mm]{images/send_sensor_data_data_flow.jpg}
-\caption{Data flow diagram for the mobile client sending on/off commands}
-\label{fig:control_appliance}
+\includegraphics[scale=2.6]{images/send_sensor_data_data_flow.jpg}
+\caption{Data flow diagram depicting the sending of the data}
+\label{fig:send_data}
 \end{figure}
 
-\section{Implementation}
-To have a system that is capable of providing a two-way communication between any host and the any mote within the wireless sensor network, four building blocks had to be provided and which are the following:
+\section{System Deployment}
+To have a system that is capable of providing a two-way communication between any host and any mote in the \gls{wsn}, the following components have been built:
 \begin{itemize}
@@
 
-\subsection{Mote programming}
+\subsection{Mote Programming}
 Each mote within the WSN network is equipped with a current transformer that is attached to data acquisition board through which data is read and transformed to the appropriate format and sent to the network host.
@@
 
-\subsubsection{TCP server}
+\subsubsection{TCP Server}
 The TCP server is an important component in the mote's program.
@@
 
-\subsection{Mote sink packet forwarding}
+\subsection{Mote Sink Packet Forwarding}
 The mote sink packet forwarding module is a special program installed within a mote that is equipped with a USB port that plays the role of a network interface card.
@@
 
-\subsection{Network gateway packet transformation}
+\subsection{IPv4/IPv6 Gateway}
 This is a very important component in the system.
 The issue is the following.
-The WSN network supports only IPV6 while other components such as the middleware and the mobile client do not necessary have an IPV6 address, but we still want all the components to communicate independently of the IP technology to be used.
+The WSN network supports only IPV6 while other components such as the middle-ware and the mobile client do not necessary have an IPV6 address, but we still want all the components to communicate independently of the IP technology to be used.
 To do so, we have created a network packet transformation program.
@@
 So whenever a request is coming, the source and destination address will be converted using this algorithm and this will make avoid the whole request response tracking part.
-\section{Measurements}
+\section{Findings}
+The goal of this part is to measure to what extent is the system able to operate reliably by offering an acceptable level of performance.
+Two experiments were conducted to measure the system's performance.
+In the first experiment, the normal work of the system is simulated, where each mote reads sensory data every second and generates traffic in the \gls{wsn}.
+Our goal is to measure the delay and observe its variance as the traffic varies.
+The routing protocol in principle gives more priority to routing packets rather than processing them.
+Therefore, this priority must affect the network's delay.
+Table \ref{table:exp1} illustrates how the delay and jitter vary as more and more nearby motes are added to the network.
+\begin{table}[htbp]
+    \begin{tabular}{lll}
+    \hline
+    number of motes & Delay (ms) & Jitter (ms) \\\hline
+    1               & 87.5         & 1.2         \\ 
+    2               & 88.1         & 2.0         \\
+    3               & 88.3        & 2.7         \\
+    4               & 88.8         & 3.1         \\
+    5               & 90.0         & 3.3         \\
+    6               & 89.5         & 3.5         \\
+    \end{tabular}
+    \caption{Experiment 1 delay and jitter}
+    \label{table:exp1}
+\end{table}
 
+\
+It is clear that the average delay is not significantly affected by the number of motes in the network, but the jitter on the other hand is clearly affected by the number of motes in the network as it is depicted in figure \ref{fig:delay_jitter}.
+The delay is much more affected because there are time interval where the network is less busy and other intervals where the network becomes busier.
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.6]{images/delay_jitter.JPG}
+\caption{Delay and jitter evolution by Traffic}
+\label{fig:delay_jitter}
+\end{figure}
+
+The second experiment's goal is to measure the contribution of the Gateway Packet transformation process to the overall communication delay.
+The results gathered reflect the average delay and jitter computed over the elapsed time starting from the sniffing of the packet to the transformation and sending to the recipient.
+Table \ref{table:exp2} shows the result that was collected from this experiment as it was done on 200 packets that were sniffed and transformed by the system.
+
+\begin{table}[htbp]
+    \begin{tabular}{ll}
+    \hline
+    Delay ($\mu s$) & Jitter ($\mu s$) \\ \hline
+    100        &  30        \\ 
+    \end{tabular}
+    \caption{Experiment 2 delay and jitter}
+        \label{table:exp2}
+\end{table}
+
+\
+The transformation process's elapsed time is measured in microseconds and from table \ref{table:exp2}, one can see that the average delay is on average 0.1 milliseconds whereas the jitter is around 0.03 milliseconds meaning that the process's time varies between a few microseconds to at most 150 microseconds.
+In addition, depending on the machine's load, the distribution of the delay frequency is shown the histogram depicted in figure \ref{fig:histogram}.
+From this figure one can conclude that the delay is normally distributed
+\begin{figure}[htbp]
+ \centering
+ \includegraphics[scale=0.7]{images/histogram.JPG}
+ \caption{Delay frequency histogram}
+ \label{fig:histogram}
+ \end{figure}
 \section{Conclusion}
-The conclusion goes here.
+\glsreset{iot}\gls{iot} is the next revolution in computing.
+The test-bed presented in this paper shows how can Smart Grids by making use of numerous technologies such as \gls{iot} based \glspl{wsn}, \gls{6lowpan} in order to integrate it into \gls{iot} and benefit from it.
+The test-bed illustrates the two-way communication and also discusses the architecture of this system.
+
+\
+The issues that were handled are mainly related to making the communication possible on a heterogeneous system where the data link layer and network standards used by different components of the system differs.
+The packet is transmitted from the an \gls{ipv4} network over Wi-Fi and moves to an \gls{ipv6} network over USB then to Zigbee.
+
+\
+Smart Grids should benefit from the Internet of Things and will become one of the most important applications of \gls{iot} and therefore enable more efficient generation and consumption of data.
 

### 2013-07-22 trunk/doc/thesis/AUI/report_AUI.tex
MODIFICATIONS DONE! YA RABBI

--- 2013-07-09 trunk/doc/thesis/TUM/report_TUM.tex
@@
-\documentclass[oneside,12pt,a4paper,final]{book}
-%\usepackage[margin=3cm]{geometry}
+\documentclass[12pt,a4paper,final]{report}
+\usepackage[margin=2.5cm]{geometry}
 \usepackage[utf8]{inputenc}
@@
 \makeatother
+\usepackage{chngcntr}
+\counterwithout{figure}{chapter}
+\counterwithout{table}{chapter}
 
@@
 \input{acronym}
-\frontmatter
 
@@
 \input{cover_page}
-\chapter{} % dedications chapter
+\chapter*{} % dedications chapter
 \begin{dedication}
@@
 To my friends \\
-
 \end{dedication}
-\chapter{Acknowledgments}
+
+% display page numbers in the headings. Start with roman numerals %
+\pagestyle{headings}
+\setcounter{page}{1}
+\pagenumbering{roman}
+
+\chapter*{Acknowledgments}
 \paragraph{}
@@
 
-\chapter{Abstract}
+\chapter*{Abstract}
 \paragraph{}
@@
 The current evolution in computing is the \gls{iot}.
-It is smoothly evolving from an Internet of People to and Internet of \textit{Things}.
-By 2020, the Internet is expected to have 50 billion \textit{Things} connected to it.
+The Internet is smoothly evolving from an Internet of People to and Internet of \textit{Things}.
+By 2020, it is expected to have 50 billion \textit{Things} connected to it.
 \paragraph{}
@@
 Another migration is taking place in the power grid that is evolving towards the \gls{sg}.
-In fact, Information Technology will play a key role in implementing needed functions such as \gls{dr}\glsreset{dr} and integrating Distributed and Renewable Energy Sources into \gls{sg}.
+In fact, Information Technology (IT) is playing a key role in implementing needed functions such as \gls{dr}\glsreset{dr} and integrating Distributed and Renewable Energy Sources into \gls{sg}.
 The \gls{iot} is expected to enable a smooth integration of Distributed Renewable Energies within the electric grid.
@@
 This thesis focuses on the development of an \gls{iot} test-bed, using wireless sensors, that will be used for Smart Grid applications.
-The test-bed built provides an IPv6 gateway, remote control of wireless sensor nodes which is built on top of \gls{6lowpan}\glsreset{6lowpan}, and mobile access to the system.
-\paragraph{}
-This study results in the creation of a test-bed that provides an infrastructure for integrating \gls{sg} into \gls{iot} and building overlying services such as \gls{dr}.
-Therefore, the system's performance was put to test by conducting two experiments where the delay and jitter were measured to make sure the system can handle overlying services in future projects.
+The test-bed provides an IPv6 gateway, remote control of wireless sensor nodes which is built on top of \gls{6lowpan}\glsreset{6lowpan}, and mobile access to the system.
+\paragraph{}
+The project contributes to SG by settling the IT infrastructure  (i.e.
+Network and protocol) that is needed to disseminate information the SG.
+This infrastructure allows a remote monitoring and control of appliances connected to such infrastructure and therefore the two-way communication needed by SG.
+\paragraph{}
+This study leads to the creation of a test-bed that provides an infrastructure for integrating \gls{sg} into \gls{iot} and building overlying services such as \gls{dr}.
+The system's performance was put to test by conducting two experiments where the delay and jitter were measured to make sure the system can handle overlying services in future projects.
 
@@
 Ce prototype fournit des services tels qu'une passerelle IPv6, le contrôle à distance des capteurs sans fil, et un accès mobile au système.
+\paragraph{}
+Le projet contribue à SG en construisant l'infrastructure IT (ex.
+réseau et protocole) qui est nécessaire pour diffuser l'information au SG.
+Cette infrastructure permet une surveillance et le contrôle à distance des appareils connectés à cette infrastructure et donc la communication bidirectionnelle nécessaires pour SG.
 \paragraph{}
@@
 
-\mainmatter
 \doublespacing
+\chapter{Preamble - Thesis Research Methodology}
+\setcounter{page}{1}
+\pagenumbering{arabic}
+\paragraph{}
+ In the course of this study, a set of rules and standards dictated the research methodology.
+With the adoption of this methodology and mainly the scientific method, the study resulted in a test-bed that is presented in this report along with the experiments conducted and measurements collected.
+This chapter will briefly explain the phases through which the study passed through and how it influenced the different directions taken throughout the course of this project.
+\section{Introduction}
+\paragraph{}
+Everything started with an idea: to be able to monitor and control appliances along with integrating it in the Smart Grid.
+As an idea, many things are very abstract at this point.
+the what question is not clear: What are we trying to build?
+What are we trying to get at the end?
+What should it do?
+Afterwards comes the how questions: How are building such a system?
+How can we monitor and control appliances?
+How should it be part of smart Grids?
+Such questions lead to the second step in this research methodology which getting to look at the state of the art, how such things are done nowadays and what already exists in the market.
+
+\section{State of the Art}
+\paragraph{}
+In this step, the goal is to get acquainted with the main concepts that are used in this study such as Smart Grid and the Internet of Things.
+To do so, a review of the work done in these two fields has been performed.
+Such review includes reading the various papers from different journals, learning more about the projects that are in development, their approaches, the existing architectures, the standards adopted and the tools used.
+\paragraph{}
+The academic perspective gives a deep understanding of the topic, the options, issues, challenges and outcomes that exist but there is a need to see what industry is actually building.
+As a matter of fact, a research has been conducted to have an idea about the current commercial applications that exist nowadays, how are they built and what do they perform.
+This step gives us a clear idea about what has been achieved both in academia and industry.
+\paragraph{}
+Now that we have a clear understanding of the main concepts and tools that exist, it is time to start bringing our contribution to the field but before that, the objectives should be clearer and more accurate than the initial proposed idea at the beginning of the study
+\section{Development and Analysis}
+\paragraph{}
+The problem statement is restated as well the purpose and scope of the study.
+It is based on the initial objectives but takes into consideration numerous ideas, issues and challenges that arise from looking at the state of the art.
+An initial architecture is sketched, that shows the main blocks of the work.
+After building such architecture comes different directions and one should choose and defend the most appropriate for this study.
+In this study, numerous alternatives were raised (e.g.
+whether the wireless sensor network should be mesh-based or tree-based) and we had to pick the most appropriate one with respect to our objectives and constraints namely time and cost.
+\paragraph{}
+Afterwards, in each component that have to be developed, numerous directions and alternatives appeared, and again the most appropriate that fulfils our needs was chosen as it will be discussed in this thesis.
+The rest of the development followed the same trend which is an iterative development model but at the same time modular.
+Each component, was studied separately where different research decisions were taken and as it was divided into several subcomponents, we studied each of them separately.
+The interconnection between different components made this work iterative because each decision in a component affects the rest of the components and therefore this decision had to be studied to be aware of its effect on the rest of the system.
+For example, when we decided to have a TCP client in the mote, this automatically affected the gateway component, because there had to be a TCP server there too.
+If UDP was chosen in the mote, this decision would have affected the gateway component too.
+\paragraph{}
+This approach was followed until the whole system was developed and analysed.
+Once the system was ready, the next step is to see if is solves the problem stated in the first step and if it fulfils the outcomes of this study
+\section{Measurements of the Results}
+\paragraph{}
+In this step, the goal is to check if the system fulfils its requirements.
+To do so, an initial testing phase was performed where the system was put to test.
+The functional requirements were tested extensively in order to identify the main problems the system could suffer from.
+\paragraph{}
+Afterwards, the system's various non functional requirements were put to test.
+The first one is to check is to check if the system could host overlying services.
+To do so, it should meet some performance qualities and criteria.
+As a matter of fact, these performance criteria were identified based on measurement and past studies.
+Once these criteria have been identified, the system's performance had to tested.
+To do so, experiments were conducted to be able to collect data that can show whether the system has met these performance criteria or not.
+\paragraph{}
+After these experiments has been conducted, the next step is to report the results of the this study which includes all the steps through which the study through.
+
+\section{Report}
+In this step, a report has been written showing all the steps through which the study has passed through.
+This includes the introduction, the review of the state of the art, the development and analysis and the measurements of the result.
+The report is a concise document showing all the information, alternatives, decisions and outcomes dealt with throughout the course of this study.
+
 \chapter{Introduction}
+
+% set the page numbers to be arabic, starting at page 1 %
 
@@
 \paragraph{}
-The \glsreset{iot}\gls{iot} uses different technologies such as \gls{wsn}, \gls{rfid}, \gls{ipv6} and \gls{6lowpan}\glsreset{6lowpan} which makes the Internet of Things a very heterogeneous and complex system \cite{ref4}.
+The \gls{iot} uses different technologies such as \gls{wsn}, \gls{rfid}, \gls{ipv6} and \gls{6lowpan} which makes the Internet of Things a very heterogeneous and complex system \cite{ref4}.
 One of the challenges is to deal with this heterogeneity in order to make the communication smooth and transparent to the parties involved.
@@
 \paragraph{}
-The \glsreset{sg} \gls{sg} relies on the two-way communication to build a whole communication framework and a protocol stack namely the \glsreset{dr}\gls{dr} and other services built on top of this last.
+The \gls{sg} relies on the two-way communication to build a whole communication framework and a protocol stack namely the \gls{dr} and other services built on top of this last.
 From the consumer's/household's perspective or more precisely the household's side,there are smart energy management systems in the market nowadays, but none of them offers an integration with \gls{sg}.
-Such an integration through the two-way communication will to make it possible for any appliance to report its real-time energy consumption, and to be controlled remotely.
-Such pillar is missing for the time being, and therefore the integration of the smart home within the smart grid is not doable.
+Such an integration through the two-way communication will make it possible for any appliance to report its real-time energy consumption, and to be controlled remotely.
+Such pillar is absent for the time being, and therefore the integration of the smart home within the smart grid is not doable.
 Also, in order for \gls{sg} to be integrated into the \gls{iot}, all devices should be able to communicate in the two ways.
@@
  \paragraph{}
-The goal of the study is create the two-way communication, and solve the heterogeneity that arises from integrating smart homes with the smart grid.
+The goal of the study is to create the two-way communication, and solve the heterogeneity that arises from integrating smart homes with the smart grid.
 In fact, each components in the smart home will become part of the Internet of Things by making it communicate with any device in the Internet by having an \gls{ipv6} address using \gls{6lowpan} whose goal is to provide the TCP/IP to the smallest devices and also those with the least computational resources.
@@
 \paragraph{}
-The second purpose of this study is to hide the heterogeneity that arises from the integration of smart homes into the grid and also when making the whole system part of the \glsreset{iot}\gls{iot}.
+The second purpose of this study is to hide the heterogeneity that arises from the integration of smart homes into the grid and also when making the whole system part of the \gls{iot}.
 To solve such a problem, we add an extra component that will track all  packets and change them to fit the standards of the targeted component.
@@
 Even though the problems seems to be separate, one cannot solve  the first problem without having to solve the second.
-In other words, after the creation of the two-way communication in the \glsreset{wsn}\gls{wsn} which adopts \gls{ipv6} over \gls{6lowpan}, other networks can only support \gls{ipv4} and therefore the communication would not be possible, and this is due to the heterogeneous nature of the system.
+In other words, after the creation of the two-way communication in the \gls{wsn} which adopts \gls{ipv6} over \gls{6lowpan}, other networks can only support \gls{ipv4} and therefore the communication would not be possible, and this is due to the heterogeneous nature of the system.
 This is why the two-way communication was created, but only after solving the heterogeneity problem, everything worked.
@@
 \paragraph{}
-It is important to note that some of the components will be discussed from the architectural point in Chapter 4 but not from an implementation aspect in Chapter 5.
-The sole reason behind this is the fact that some of the equipment required for this thesis were not on time and therefore the parts in the implementation requiring such equipment were omitted in order to meet the deadlines.
+It is important to note that some of the components will be discussed from the architectural point in Chapter 4 but not from the deployment perspective in Chapter 5.
+The reason behind this is the fact that some of the equipment required for this thesis were not on time and therefore the parts in the implementation requiring such equipment were omitted in order to meet the deadlines.
 \section{Outcome of the Study}
 \paragraph{}
-The outcome of the thesis is to design and implement a test-bed that will be integrated into the \glsreset{iot}\gls{iot} by showing that every node in the system can communicate autonomously and with any other node in the Internet.
+The outcome of the thesis is to design and implement a test-bed that will be integrated into the \gls{iot} by showing that every node in the system can communicate autonomously and with any other node in the Internet.
 Another \gls{iot} concept is the ability of any node in the \gls{wsn} to consume and produce information that is communicated through the network.
@@
 The thesis is divided into eight chapters. Each of the chapter approaches thesis from a different aspect. 
-\paragraph{Chapter 1}
-It is an introductory chapter that gives a flavor of the project that is  studied in this thesis and also introduces the reader to the problem and the solution proposed in this context.
-This chapter also puts into perspective by introducing some of the main concepts related to the \glsreset{sg}\gls{sg} and the \glsreset{iot}\gls{iot}.
-\paragraph{Chapter 2}
+\paragraph{Chapter 3}
 It presents the background of the work by revising the literature related to the topics of \gls{iot} and \gls{sg}.
@@
 Finally, its presents the commonalities ad differences between this thesis and the work previously done.
-\paragraph{Chapter 3}
+\paragraph{Chapter 4}
 This part is dedicated to \gls{iot}, its importance and its impact on the world. Is also presents the actual state of the Internet and makes predictions about its future.
-\paragraph{Chapter 4}
+\paragraph{Chapter 5}
 It presents the architecture of the test-bed. It also discusses the system from different perspectives, and how the problems stated in section 1.3 were solved from the architectural aspect.
-\paragraph{Chapter 5}
+\paragraph{Chapter 6}
 This chapter mostly discusses the system deployment.
@@
 Afterwards, each component of the system is discussed extensively and moves along with the integration to form the test-bed.
-\paragraph{Chapter 6}
+\paragraph{Chapter 7}
 It starts first by presenting the findings then move to explain the experiments that were conducted. Afterwards, the data gathered is shown, analyzed and interpreted.
-\paragraph{Chapter 7, 8}
+\paragraph{Chapter 8, 9}
 These are concluding chapters of this study.
 They provide directions about future work, recommendations to the reader about what parts of the system could be improved and how these improvements could be achieved.
-
+\section*{}
+\paragraph{}
+Now that the problem statement has been defined in addition to the purpose of the study and the scope, a review of the current work and the state of the art will be discussed in chapter 3 and 4.
 \chapter{Literature Review}
@@
 \paragraph{}
-The actual electrical grid is based on a technology that is at least one century old and its weaknesses are becoming more and more costly \cite{ref5}.
+The current electrical grid is based on a technology that is at least one century old and its weaknesses are becoming more and more costly \cite{ref5}.
 Therefore, there is a need to change it to meet to 21st century requirements such as having more green energy in order to reduce pollution caused mainly by the use of fossil energy.
-As a matter of fact, research introduced us to a new technology: the \glsreset{sg}\gls{sg}.
+As a matter of fact, research introduced us to a new technology: the \gls{sg}.
 Although it is still a research project that has not reached the stage of maturity yet, nations across the globe started to deploy \gls{sg} projects.
@@
 This is known nowadays as the feed-in-tariff which is a policy that allows the consumers to become producers and sell the excess power to the grid \cite{ref7}.
-This will allow the power grid to provide electricity to consumers in a "free market" fashion which means that it will obey the laws of supply and demand \cite{ref7}.\glsreset{iot} \glsreset{sg}
+This will allow the power grid to provide electricity to consumers in a "free market" fashion which means that it will obey the laws of supply and demand \cite{ref7}.
 \section{\gls{iot} and \gls{sg}}
@@
 \gls{iot} brings several benefits to the actual Internet and will pull the \gls{sg} to make it part of it.
-Consequently, research has taken the direction of integrating the \gls{sg} within the \gls{iot} as most of the aspects of this last favor  the \gls{sg} \cite{ref8}.
-\paragraph{}
-\gls{iot} counts on a set of technologies which are \glsreset{rfid}\gls{rfid}, \gls{m2m} \cite{ref9}, public communication networks such as 2G/3G mobile networks, power line communication, WI-FI and Zigbee\cite{ref8}.
+Consequently, research has taken the direction of integrating the \gls{sg} within the \gls{iot} as most of the aspects of this last favour  the \gls{sg} \cite{ref8}.
+\paragraph{}
+\gls{iot} counts on a set of technologies which are \gls{rfid}, \gls{m2m} \cite{ref9}, public communication networks such as 2G/3G mobile networks, power line communication, WI-FI and Zigbee\cite{ref8}.
 The smart grid's architecture based on \gls{iot} is divided into three layers:
@@
 \paragraph{}
-The \glsreset{iot}\gls{iot} refers to the evolution of the actual Internet.The term \gls{iot} was used for the first time by Kevin Ashton in 1999 \cite{ref17}.
+The \gls{iot} refers to the evolution of the actual Internet.The term \gls{iot} was used for the first time by Kevin Ashton in 1999 \cite{ref17}.
 The human beings will represent in the future a very small minority in the Internet and as it will be rules by autonomous machines.
@@
 \paragraph{}
-The \glsreset{iot}\gls{iot} is a network mostly ruled my machines, from the most powerful supercomputer and mainframes to the smallest devices such as sensor nodes.
+\gls{iot} is a network mostly ruled my machines, from the most powerful supercomputer and mainframes to the smallest devices such as sensor nodes.
 All will have one common point, being all connected to the Internet and able to communicate, produce and consume information.
@@
 New paradigms, architectures, frameworks and prototypes are built and refined everyday to support it \cite{ref15}.
-New technologies such as \glsreset{rfid}\gls{rfid}, \glsreset{m2m}\gls{m2m}, \glsreset{ipv6}\gls{ipv6}, \glsreset{6lowpan}\gls{6lowpan} and \gls{ims} have come to place to establish the infrastructure for \gls{iot}.
+New technologies such as \gls{rfid}, \gls{m2m}, \gls{ipv6}, \gls{6lowpan} and \gls{ims} have come to place to establish the infrastructure for \gls{iot}.
 The whole world is now being connected including humans and objects \cite{ref18}.
 \paragraph{}
-The \glsreset{iot}\gls{iot} brings numerous advantages but many issues will have to solved before this network explodes to connecting trillions.
+The \gls{iot} brings numerous advantages but many issues will have to solved before this network explodes to connecting trillions.
 The first issue is the interconnection of all the devices into the \gls{iot}.
@@
 \paragraph{}
-The actual Internet is evolving into the \glsreset{iot}\gls{iot} because it opens possibilities that are not in the actual Internet.
+The actual Internet is evolving into \gls{iot} because it opens possibilities that are not in the actual Internet.
 The ability of Things to produce and consume information is a requirement.
@@
 
+\section*{}
+\paragraph{}
+After reviewing the state of the art and the current work done in academia and industry, the following chapter will present the test-bed built in this study by first presenting the architecture of the system and all of the components that were built as part of the whole test-bed.
+
 \chapter{System Architecture}
 \paragraph{}
-The thesis is based on the study and creation of a test-bed whose goal is to interconnect smart homes with the \glsreset{sg}\gls{sg} and integrate both into the \glsreset{iot}\gls{iot}.
+The thesis is based on the study and creation of a test-bed whose goal is to interconnect smart homes with the \gls{sg} and integrate both into the \gls{iot}.
 To do so, we first define the system's objectives and goals.
@@
 The goal of the system is to present a working test-bed that is able to interact with any appliance in the smart home.
-The system makes this possible by making all appliances part of the \glsreset{wsn}\gls{wsn} where each sensor node is addressable and may be used to control the appliance.
+The system makes this possible by making all appliances part of the \gls{wsn} where each sensor node is addressable and may be used to control the appliance.
 \paragraph{Objective 2}
@@
 \paragraph{}
-The system is divided into nine components, as depicted in figure \ref{fig:gen_architecture}, where every component is in charge of a part of the system's objectives.
+The system is divided into nine components, as depicted in Figure \ref{fig:gen_architecture}, where every component is in charge of a part of the system's objectives.
 When these nine components work hand-in-hand, they achieve all the system's and project's objectives and goals.
@@
 \begin{itemize}
-\item Full \gls{ipv6} \glsreset{wsn}\gls{wsn}
+\item Full \gls{ipv6} \gls{wsn}
 \item Mote sensor data collection TCP client
 \item Mote appliance control TCP server
-\item \gls{ipv4}/\gls{ipv6} Gatewaying Process
+\item \gls{ipv4}/\gls{ipv6} Gateway Process
 \item \gls{ipv6} over USB tunnel
@@
 \paragraph{}
-The general architecture is depicted in figure \ref{fig:gen_architecture}.
-This figure shows the existing components and their interactions.
-Each set of components appear within a box as this last refers to what hardware are the components deployed in as it is labeled in the box.
-There are are five hardware components:
+The general architecture is depicted in Figure \ref{fig:gen_architecture}. This Figure shows main blocks of the system and their interactions. There are are 3 main blocks:
 \begin{itemize}
-\item Sensor Node
-\item Mote Sink
+\item Wireless Sensor Network
 \item Gateway
-\item Middle-ware
-\item Smart phone
+\item External Users
 \end{itemize}
-Each hardware component hosts a set of software subsystems that will be explained more clearly in the sections 4.4 to 4.11.
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=17cm,height=20cm,keepaspectratio]{img/general_architecture.jpg}
+Each block is explained and detailed in a separate figure. These figures will present the hardware and the software components of the the system
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=17cm,height=20cm,keepaspectratio]{img/architecture_general.jpg}
 \caption{General architecture depicting the main components of the system and the information flow}
@@
 \end{figure}
-%\clearpage
+\clearpage
+
+\paragraph{}
+Each block that appear in Figure \ref{fig:gen_architecture} is depicted in a separate figure to show the different software components that exist in it.
+\paragraph{}
+Figure \ref{fig:part1} shows the different software and hardware components that exist in the Wireless Sensor Network.
+The figure also shows the inputs and outputs of the system in terms of information flow.
+There are two types of hardware components in the WSN: the mote and the mote sink.
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=12cm,height=20cm,keepaspectratio]{img/part1.jpg}
+\caption{WSN main software components and their interconnections}
+\label{fig:part1}
+\end{figure}
+\paragraph{}
+The mote is the sensor node that controls and monitors the appliance whereas the mote sink is the sensor node that makes the communication between the WSN nodes and the gateway possible.
+The mote on the one hand has a TCP Client and a TCP Server.
+The TCP client is the software component in charge of monitoring the appliance's consumption and communicating this information,  whereas the TCP server receives commands and executes them.
+The mote sink on the other hand has one component which is in charge of making the communication between the WSN and the gateway possible.
+The mote sink has \gls{ipv6} back station which forwards the \gls{ipv6} packets in and out of the \gls{wsn}.
+
+\paragraph{}
+In Figure \ref{fig:part2},  the main software components in the gateway are depicted.
+There are four components: \gls{ipv6} over USB tunnel, Packet transformation Process or IPv6/IPv4 gateway process, Sensor Date Storage Unit and HTTP sensor data sender.
+In addition the arrows show the inputs and outputs of this component.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=17cm,height=20cm,keepaspectratio]{img/part2.jpg}
+\caption{Gateway main software components and their interconnections}
+\label{fig:part2}
+\end{figure}
+
+\paragraph{}
+Figure \ref{fig:part3} depicts the main components that exist in the external users box in Figure \ref{fig:gen_architecture}.
+There are two hardware components: The middle-ware and the mobile client.
+In the middle-ware component, there are three software components, the data store, the HTTP server and the command request dispatcher.
+In the mobile client, there is one software component which is mobile client App.
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=17cm,height=20cm,keepaspectratio]{img/part3.jpg}
+\caption{External users main software components and their interconnections}
+\label{fig:part3}
+\end{figure}
+
 \paragraph{}
 Now that the general architecture has been established, the question that should be asked is whether all these components operate in the same layer in the \gls{osi} model and if not, in which layer does each of these components reside.
-As an answer to such a question, figure \ref{fig:layered_view} is here to show in what layer does every component operates.
+As an answer to such a question, Figure \ref{fig:layered_view} is here to show in what layer does every component operates.
 
@@
 \paragraph{}
-From figure \ref{fig:layered_view}, one can see that there are three components in performing in the data link layer and network layer.
-The first one is the \gls{ipv4}/\gls{ipv6} gatewaying process whose goal is to hide the communication heterogeneity by sniffing incoming packets, transforming them in order to meet the recipient's communication technology and standards.
-The second component performing in the data link layer and network is the \gls{ipv6} over USB tunnel; This component is in charge of receiving raw \gls{ipv6} packets and encapsulate them in  USB frames to reach the destination.
+From Figure \ref{fig:layered_view}, one can see that there are three components in performing in the data link layer and network layer.
+The first one is the \gls{ipv4}/\gls{ipv6} gateway process whose goal is to hide the communication heterogeneity by sniffing incoming packets, transforming them in order to meet the recipient's communication technology and standards.
+The second component performing in the data link layer and network is the \gls{ipv6} over USB tunnel; this component is in charge of receiving raw \gls{ipv6} packets and encapsulate them in  USB frames to reach the destination.
 The third component is deployed in the sink which is the entry and exit point of the \gls{wsn}.
@@
 The first one is the client mobile App.
-This component is implements a TCP client and uses most of the services provided by the layers it is operating in.
-The second component is the sensor data storage which constitutes of a TCP server receiving data from the \gls{wsn}, transforming and storing it in a \gls{rdbms}.
+This component implements a TCP client and uses most of the services provided by the layers it is operating in.
+The second component is the sensor data storage which consists of a TCP server receiving data from the \gls{wsn}, transforming and storing it in a \gls{rdbms}.
 The third and fourth components are the Mote TCP client and server that send sensor data and through the TCP client and receive commands using the TCP server.
@@
 This component receives data and communicates it to the destination using HTTP.
-The middle-ware application used HTTP to get sensor data from this system.
+The middle-ware application uses HTTP to get sensor data from this system.
 \paragraph{}
 Now that the general structure of the software system has been established, the hardware structure will be presented and explained further.
-figure \ref{fig:network_diagram} is a network diagram that shows the hardware components in this system.
-The \glsreset{wsn}\gls{wsn} appears on the right of the figure.
+Figure \ref{fig:network_diagram} is a network diagram that shows the hardware components in this system.
+The \gls{wsn} appears on the right of the Figure.
 Attached to it the sink that which constitutes and entry and exit point for the \gls{wsn}.
@@
 \end{figure}
-
-\paragraph{}
-Now that the network diagram has been established in \ref{fig:network_diagram} where the hardware component are depicted, we need to show where each software component is deployed.
+\clearpage
+
+\paragraph{}
+Now that the network diagram has been established in Figure \ref{fig:network_diagram} where the hardware component are depicted, we need to show where each software component is deployed.
 In other words, what are the software components that exist on which hardware machine.
-This is the goal of the matrix appearing in table \ref{table:soft_hard_deployment}.
+This is the goal of the matrix appearing in Table \ref{table:soft_hard_deployment}.
 It shows where very software component is deployed.
-From the table \ref{table:soft_hard_deployment}, it is clear that there are hardware machines that have more than one component deployed and this is also the case for small devices such as sensor nodes within the \gls{wsn}.
+From the Table \ref{table:soft_hard_deployment}, it is clear that there are hardware machines that have more than one component deployed and this is also the case for small devices such as sensor nodes within the \gls{wsn}.
 
@@
     Mote appliance control TCP server       & X    & X         & ~       & ~      \\
-    \gls{ipv4}/\gls{ipv6} Gatewaying Process   & ~    & ~         & X       & ~      \\
+    \gls{ipv4}/\gls{ipv6} Gateway Process   & ~    & ~         & X       & ~      \\
     IPv6 over USB tunnel                    & ~    & ~         & X       & ~      \\
@@
 \paragraph{}
-The mote sensor data collection TCP client is a simple \gls{tcp} client deployed in a mote that is equipped with a \gls{ct} attached to an appliance that reads sensory data periodically as shown in figure \ref{fig:current_transformer}.
+The mote sensor data collection TCP client is a simple \gls{tcp} client deployed in a mote that is equipped with a \gls{ct} attached to an appliance that reads sensory data periodically as shown in Figure \ref{fig:current_transformer}.
 This data needs to be transformed using a set of transformation functions to get the consumption in \gls{wh}.
@@
 \boldmath{$I = 66.7U_0$} \\
-I: Intensity in \glsreset{a}\gls{a} \\
+$I$: Intensity in \glsreset{a}\gls{a} \\
 $U_0$: The voltage read from the \gls{ct}
@@
 \boldmath{$ P = 220I $} \\
-P: Power in \gls{w} \\
-I: Intensity in \glsreset{a}\gls{a}
-\paragraph{}
-To get the consumption we use an the sampling rate.
+$P$: Power in \gls{w} \\
+$I$: Intensity in \glsreset{a}\gls{a}
+\paragraph{}
+To get the consumption we use the sampling rate.
 The sampling rate is as the number of sensory readings per hour.
@@
 $ C = \cfrac{P_0 + \cfrac{P_1 - P_0}{2}}{S} $ \\
-C: Current Consumption in \glsreset{wh}\gls{wh} \\
-P: Power sample number 1 in \glsreset{w}\gls{w} \\
-P': Power sample number 2 in \glsreset{w}\gls{w} \\
-$U_1$: Electrical potential difference or Voltage in households in \glsreset{v}\gls{v} \\
-S: Sampling rate \\
+$C$: Current Consumption in \gls{wh} \\
+$P$: Power sample number 1 in \gls{w} \\
+$P'$: Power sample number 2 in \gls{w} \\
+$U_1$: Electrical potential difference or Voltage in households in \gls{v} \\
+$S$: Sampling rate \\
 The assumption made is that if the power varies between two consecutive samples, the change was linear throughout the interval of time between the two sensory readings.
 \paragraph{}
-Once the consumption data in hand after this series of transformation, the next thing left to do is to communicate this data.
+Once the consumption data is in hand after this series of transformation, the next thing left to do is to communicate this data.
 To send it, the mote uses the \gls{tcp} client built on top of the \gls{blip} package that implements the TCP/IP protocol stack.
@@
 \paragraph{}
-The mote appliance control \glsreset{tcp}\gls{tcp} server is deployed in all the sensor nodes of the \gls{wsn} including the sink .
-The goal of this TCP server is be ready to receive commands from external node in the Internet in order to turn On or Off the appliance it oversees.
-To control an appliance.
+The mote appliance control \gls{tcp} server is deployed in all the sensor nodes of the \gls{wsn} including the mote sink .
+The goal of this TCP server is to be ready to receive commands from external node in the Internet in order to turn On or Off the appliance it oversees.
 \paragraph{}
@@
 It needs to be equipped with a Data Acquisition board also known as Extension Board that enable the mote to have more capabilities and therefore the ability to control the power flowing into the appliance.
-The extension board used in this project is the Crossbow MDA320CA as depicted in figure \ref{fig:mda320ca}.
-This board has Pins which are electrical inputs and outputs that permit possible to have external analog sensors, digital sensors and also the possibility feeding electricity to the appliance.
+The extension board used in this project is the Crossbow MDA320CA as depicted in Figure \ref{fig:mda320ca}.
+This board has Pins which are electrical inputs and outputs that permit to have external analog sensors, digital sensors and also the possibility of indirectly feeding electricity to the appliance.
 However, the power supply of the appliances from the mote is not straightforward.
+To indirectly feed electricity to an appliance using a mote, the system should look as follows: The mote is equipped with MDA230CA board, to which a circuit is attached to one of the digital pins of the mote.
+The appliance is also attached to this same circuit.
+when one wants to turn off an appliance, the circuit is opened by controlling the digital pin attached to the circuit and therefore the appliance is turned Off.
+To turn on the appliance, one should close the circuit.
 \paragraph{}
@@
 \paragraph{} 
-This component is deployed only the sink of the \glsreset{wsn}\gls{wsn}.
+This component is deployed only in the mote sink of the \gls{wsn} as shown in Figure \ref{fig:gen_architecture}.
 This mote is able to communicate with the sensor nodes using Zigbee and with the gateway using \gls{usb}.
@@
 The Zigbee network interface card is of interest in this section.
-To have a device communicating using Zigbee, one sensor node within the \gls{wsn} was used for this purpose, and the software component implemented in this sensor node is the one described in section 4.6.
+To have a device communicating using Zigbee, one sensor node within the \gls{wsn} was used for this purpose, and the software component implemented in this sensor node is the one described in section 5.6.
 Indeed, the packet is received through \gls{usb} but not read and recognized as a packet belonging to the TCP/IP protocol stack.
@@
 \paragraph{}
-The subsections 4.8.1 to 4.8.3 will detail and explaining the three subcomponents of the system.
+The subsections 5.8.1 to 5.8.3 will detail and explaining the three subcomponents of the system.
 
@@
 The packet sniffer subcomponent is the one responsible for sniffing and forwarding the packets to the packet transformer.
-It does not forward the packets whose recipient is the gateway machine.
 Specifically, all the packets to be routed by the gateway server are sniffed and forwarded to the packet transformer.
+The sniffer uses Netfilter which is a kernel module in Linux.
+Netfilter has a set of points in which the packets appear.
+These points are known as Hooks.
+In the case of this subcomponent, the PRE-ROUTING hook was targeted by this sniffer and all the packets appearing in this hook were sniffed and dropped except for those targeting the gateway machine.
 \subsection{Packet Transformer}
@@
 Concerning the packet not originating from the \gls{wsn}, only those going to \gls{wsn} are processed.
-But a problem presents itself:
+But a problem presents itself: \\
 \underline{The outside world mostly uses \gls{ipv4} whereas the \gls{wsn} uses exclusively \gls{ipv6}}
@@
 In general, two networks using different network layer technologies cannot communicate despite the fact that \gls{ipv6} is just a new version of the very known \gls{ipv4}, and this is due to the fact that \gls{ipv6} has no backward compatibility with \gls{ipv4}.
-\paragraph{Solution}
+\paragraph{}
 To solve this problem, we provided all the nodes with both an \gls{ipv4} and \gls{ipv6} address.
 The nodes have either \gls{ipv4} or \gls{ipv6} addresses, so the one missing is assigned to them virtually.
-For instance, a sensor node in the \gls{wsn} with a real \gls{ipv6} address equal to FEC0::1 will be assigned a virtual \gls{ipv4} address equal to 192.168.0.1 but he will not be aware of it.
+For instance, a sensor node in the \gls{wsn} with a real \gls{ipv6} address equal to FEC0::1 will be assigned a virtual \gls{ipv4} address equal to 192.168.0.1 but it will not be aware of it.
 Only the Gateway packet transformation component will be aware of it and based on this, we transform the packets and hide the heterogeneity from the components of the system.
@@
 The source and destination addresses are extracted from the packet and translated.
-Afterwards, a new packet referring to the destination technology used is created by having as a source and destination addresses the translated IP addresses and as payload the \gls{tcp} \gls{udp} datagram extracted from the original packet.
+Afterwards, a new packet referring to the destination technology used is created by having as a source and destination addresses the translated IP addresses and as payload the \gls{tcp}/\gls{udp} datagram extracted from the original packet.
 Finally, the packet is forwarded to the network.
@@
 The address translator is a very specialized subcomponent whose sole goal is to translate addresses from \gls{ipv4} and \gls{ipv6} and vice versa.
-That is said, this component receives the address to translate and returns the translated one.
+That is this component receives the address to translate and returns the translated one.
 In fact, the translation function yields a bijective function.
-that is, for any \gls{ipv4} address there is one and only one \gls{ipv6} address resulting translation and vice versa.
+That is, for any \gls{ipv4} address there is one and only one \gls{ipv6} address resulting translation and vice versa.
 This statement compresses a set of statements which are the following:
@@
 It is perceived by the \gls{wsn} as a TCP server, where sensor nodes get connected to as soon as they boot.
-The sensor node starts sending packet periodically and whenever the sensor data is received, it is stored into a \gls{rdbms} adopting the schema depicted in figure \ref{fig:erd}.
+The sensor node starts sending packets periodically and whenever the sensor data is received, it is stored into a \gls{rdbms} adopting the schema depicted in Figure \ref{fig:erd}.
 
@@
 \section{HTTP Sensor Data Sender} 
-The HTTP sensor data sender is TCP server to which sensor nodes get connected to, and start sending their data periodically.
+The HTTP sensor data sender is TCP server to which sensor nodes get connected, and start sending their data periodically.
 Once the data is received, this component communicates this data to the external node interested in receiving this data.
@@
 The mobile client is a component of the system but can be perceived as an external world client as it is here to receive sensor data and also send ON/OFF commands to the system.
+\section*{}
+\paragraph{}
+In this chapter, the architecture of the system was established, showing all the different architectural decisions made in addition to the different tradeoffs that were present when deciding what architectural decision to make.
+Now that the architecture was presented and discussed, the next chapter will talk about the deployment of the system.
 
@@
 The system was built using a set of software and hardware technologies and standards throughout the course of this project to make the test-bed achieve its objectives.
-\glsreset{wsn}
 \subsection{\gls{wsn}}
@@
 
-The \gls{wsn} was built using Crossbow professional kit as depicted in figure \ref{fig:crossbow} which counts 7 Micaz sensor nodes of type MPR2600 as shows in figure \ref{fig:micaz}.
+The \gls{wsn} was built using Crossbow professional kit as depicted in Figure \ref{fig:crossbow} which counts 7 Micaz sensor nodes of type MPR2600 as shows in Figure \ref{fig:micaz}.
 \begin{figure}[htbp]
@@
 In addition, the motes are supplied with sensor boards of type Crossbow MTS400CA that contain temperature, humidity, barometric pressure, ambient light and infrared light sensors that are used to communicate different types of sensor data to the gateway.
-This board is shown in figure \ref{fig:mts400}.
+This board is shown in Figure \ref{fig:mts400}.
 
@@
 \paragraph{}
-Section 4.4 and 4.5 were dedicated to explain the TCP Client and TCP server separately.
-In this section, one section will explain both of them because from the implementation point of view,  both are implemented in the \gls{tinyos} program.
+Section 5.4 and 5.5 were dedicated to explain the TCP Client and TCP server separately.
+In this section, both othe TCP Client and TCP Server will be explained because from the implementation point of view,  both are implemented in the \gls{tinyos} program.
 \paragraph{}
@@
 In the module part, we define the interfaces we will be using and those we will be providing.
-Back to definitions part, after defining components that will be used, we then link the interfaces needed in the module to the components providing those interfaces.
+Back to the definitions part, after defining components that will be used, we then link the interfaces needed in the module to the components providing those interfaces.
 This step is called wiring \cite{ref14}.
@@
 \item LedsC: provides access to the leds existing in the mote
-\item Taos2550C: is a light sensor component that enable to read data
-\item TimerMillic: A Timer that will be used for sampling
-\item TCPSocket as client: Is the TCP client
-\item TCPSocket as Server: Is the TCP server
+\item Taos2550C: a light sensor component that enables reading data
+\item TimerMillic: a Timer that will be used for sampling
+\item TCPSocket as client: the TCP client
+\item TCPSocket as Server: the TCP server
 \end{itemize}
@@
 \begin{itemize}
-\item Boot: This is the interface that provides main functions of booting and closing
+\item Boot: the interface that provides main functions of booting and closing
 \item TCP as server: The main interface for TCP
@@
 The first thing that is done after the mote is booted is to initialize the TCP client and server.
-The Client is binded to a random port.
-The TCP server is binded to port 7.
+The Client is bound to a random port.
+The TCP server is to port 7.
 The client gets connected to the Data store in the gateway by providing the \gls{ipv6} address of the gateway and port number of the server.
@@
 \paragraph{Timer fired}
-Once the timer is fired, it is time to read sensor data from the sensor. The command to read from the sensor is called reads data and once the data is read the event "read done" is triggered.
+Once the timer is fired, it is time to read sensor data from the sensor. The command to read from the sensor is called and once the data is read the event "read done" is triggered.
 \paragraph{Read done}
 Once the data is read and ready, the next thing is to transform it and then send it.
-Concerning the transformation, a series of equations are applied as seen in section 4.4 which builds the data value from the original value.
+Concerning the transformation, a series of equations are applied as seen in section 5.4 which builds the data value from the original value.
 A message is created containing two values: the mote NODE\_ID which is the identifier of the mote predefined in the installation of the program in the mote and the sensor data value.
@@
 \paragraph{}
-During the installation of the software, the \glsreset{blip}\gls{blip} package should specified as the program depends on it.
+During the installation of the software, the \gls{blip} package should specified as the program depends on it.
 \gls{blip} provides the TCP/IP protocol stack that implements \gls{6lowpan} standard by \gls{ietf}.
@@
 \paragraph{}
-The installation and initialization of this component will be detailed in appendix A.
+The installation and initialization of this component will be detailed in Appendix A.
 \section{IPv4/IPv6 Gatewaying Process}
@@
 The gateway has two network interface cards: the Wi-Fi interface and the Zigbee interface provided through the tunnel.
-The first one is using \gls{ipv4} whereas the second is using \gls{ipv6}.
-\paragraph{}
-The program has three subcomponents as seen in section 4.8. The three components are the following:
+The first one uses \gls{ipv4} whereas the second uses \gls{ipv6}.
+\paragraph{}
+The program has three subcomponents as seen in section 5.8. The three components are the following:
 \begin{itemize}
@@
 The main reason behind dropping the packets is when packets are received by a station who has an empty routing table, this station will automatically send ICMP REDIRECT TO HOST messages to the sender meaning that the route chosen by the sender is not the appropriate one, therefore the sender will stop sending packets to that station.
-
+\newpage
+\begin{singlespacing}
 \begin{lstlisting}
@@
 \end{lstlisting}
+\end{singlespacing}
 \paragraph{}
@@
 There are two solutions: 0 referring to \gls{ipv6} and 8 referring to \gls{ipv4}.
-If we get zero, the whole data should be casted to the \gls{ipv6} data structure whereas if it is 8, we should move the pointer forward to omit the Wi-Fi header and cast the following part to the \gls{ipv4} data structure.
+If we get 0, the whole data should be cast to the \gls{ipv6} data structure whereas if it is 8, we should move the pointer forward to omit the Wi-Fi header and cast the following part to the \gls{ipv4} data structure.
 \paragraph{}
@@
 Depending on the network technology of the sniffed packet, we should build a new packet with the technology that is recognized by the recipient.
-
+\newpage
+\begin{singlespacing}
 \begin{lstlisting}
@@
 \end{lstlisting}
+\end{singlespacing}
 
@@
 The packet transformer subcomponent is in charge of building two types of packets: \gls{ipv4} and \gls{ipv6} packets.
-As stated in section 4.8, all devices in this system have both an \gls{ipv4} and \gls{ipv6} address, where one of them is only recognized at the gateway.
+As stated in section 5.8, all devices in this system have both an \gls{ipv4} and \gls{ipv6} address, where one of them is only recognized at the gateway.
 The \gls{wsn} exclusively supports \gls{ipv6}.
@@
 Afterwards, the transformer creates the packet appends to it the datagram, fills the source and destination addresses and create a Socket linked to the corresponding interface card and sends the packet.
+
+\begin{singlespacing}
 \begin{lstlisting}
@@
 To do so, the translation algorithm is purely bijective.
-That is said, any \gls{ipv4} is translated to one and only one \gls{ipv6} and vice versa.
+That is, any \gls{ipv4} is translated to one and only one \gls{ipv6} and vice versa.
 For a mobile client with \gls{ipv4} address of 10.50.1.50 to communicate with a mote whose \gls{ipv6} address is FEC0::1, the mobile client should be aware of the mote's translated \gls{ipv4} address which is 192.168.0.1.
@@
 
+\section*{}
+\paragraph{}
+This chapter discussed how the system was deployed; this includes the hardware used, the software technology used and the implementation decisions made to deploy the test-bed.
+The next chapter will present the experiments that were conducted to test the performance of the system and the reasons behind conducting such experiments.
+The results will be presented and analyzed.
+
 \chapter{Findings}
@@
 The system was built to monitor and control home appliances, and to also provide an infrastructure for future platforms to be integrated in this system.
-An example of infrastructures that can be the integrated is the \glsreset{dr}\gls{dr}.
-To build overlying system's, one should make sure that the system achieves a performance level that one can count on.
-The purpose of this chapter is to measure the performance of the system.
+An example of infrastructures that can be the integrated is the \gls{dr}.
+To build overlying system's, one should make sure that the system achieves a performance level that one can rely on.
+In addition to performance, the availability and security of the system should be put to test to see whether the system can be adopted by overlying services.In this chapter, only performance will be tested whereas the availability and security experiments are left as future work.
+\paragraph{}
+The purpose of this chapter is to assess the performance of the system.
 It is done by conducting two experiments which will help us gather data, and based on that, an interpretation of this data will be done to verify if the system shows an acceptable level of performance.
-\section{Experimentations}
+\section{Experimentation}
 \paragraph{}
@@
 There is an architectural decision made in \gls{6lowpan}, and we would like to see if it add more performance to the system.
-The architectural decision is the following: There is more priority given to routed packet that the processed ones.
+The architectural decision is the following: There is more priority given to routed packet that the local ones.
 As a result, this clearly pushes us to think that the delay of the system will not follow the same growth trend as the network traffic.
@@
 \paragraph{}
-After both experiments were conducted, we collected a significant amount of data that was used to measure the performance.
-In addition, we analyze and interpret the results to make judgments about the overall performance of the system.
+After both experiments were conducted, we collected data that was used to measure the performance.
+In addition, we analyzed and interpreted the results to make judgments about the overall performance of the system.
+Each experiment was conducted only once.
 \subsection{Experiment 1}
@@
  \includegraphics[scale=1]{img/delay_jitter.JPG}
- \caption{Delay and jitter evolution by Bandwidth}
+ \caption{Delay and jitter evolution by Traffic}
  \label{fig:delay_jitter}
@@
 In fact, Zigbee has a bandwidth of 250 Kbps.
-However, experiment one attained less than 1 Kbps which does not reflect clearly how the network would react to such a high bandwidth utilization.
+However, experiment 1 attained less than 1 Kbps which does not reflect clearly how the network would react to such a high bandwidth utilization.
 The point is that even though the delay seems constant in experiment 1, that does not give any indications about how the delay and jitter varies as a function of the percentage of the bandwidth used.
@@
 The results gathered reflect the average delay and jitter computed over the elapsed time starting from the sniffing of the packet to the transformation and sending to the recipient.
-Table \ref{table:exp2} shows the result that was collected from this experiment as it was done on 200 packets that were sniffed and transformed by the system.
-
-\begin{table}[htbp]
-    \begin{tabular}{ll}
-    \hline
-    Delay ($\mu s$) & Jitter ($\mu s$) \\ \hline
-    0.1        & 0.03         \\ 
-    \end{tabular}
-    \caption{Experiment 2 delay and jitter}
-        \label{table:exp2}
-\end{table}
-\paragraph{}
-The transformation process's elapsed time is measured in microseconds and from table \ref{table:exp2}, one can see that the average delay is on average 0.1 milliseconds whereas the jitter is around 0.03 milliseconds meaning that the process's time varies between a few microseconds to at most 150 microseconds.
-In addition, depending on the machine's load, the distribution of the delay frequency is shown the histogram depicted in figure \ref{fig:histogram}.
-From this figure one can conclude that the delay is normally distributed
+The result that was collected from this experiment was done on 200 packets that were sniffed and transformed by the system.
+
+\paragraph{}
+The transformation process's elapsed time is measured in microseconds, one can see that the average delay is on average 100 $\mu s$ whereas the jitter is around 30 $\mu s$ meaning that the process's time varies between a 70 $\mu s$ to 130 $\mu s$.
+In addition, depending on the machine's load, the distribution of the delay frequency is shown the histogram depicted in Figure \ref{fig:histogram}.
+From this Figure one can conclude that the delay is normally distributed.
 \begin{figure}[htbp]
@@
 
+\section*{}
+\paragraph{}
+This chapter presented the two experiments that were conducted to test the performance of the system.
+An analysis of the result was made and conclusions were made about the system's overall performance and its ability to host overlying services.
+The next two chapters are concluding chapters that summarize the study and discuss the future work based on this study.
 \chapter{Conclusion}
 \paragraph{}
-This project enabled us to create an infrastructure that would integrated into \glsreset{iot}\gls{iot} and move towards having the Smart Home systems' integrated within the Smart Grid.
+This project enabled us to create an infrastructure that would integrated into \gls{iot} and move towards having the Smart Home systems' integrated within the Smart Grid.
 At this stage and counting solely on this work, one cannot clearly see the integration of smart home systems within the \gls{iot}, but this is an infrastructure that provides two main services which are the establishment of the two-way communication between the smart home devices and the outside world, and the heterogeneity handler.
-The two-way communication enables the \glsreset{wsn}\gls{wsn} to monitor home appliances and send sensor data to a repository and this last can be used to simulate a real world \gls{ami}.
+The two-way communication enables the \gls{wsn} to monitor home appliances and send sensor data to a repository and this last can be used to simulate a real world \gls{ami}.
 The second direction in this communication gives the ability for clients to control appliances by turning them On or Off.
@@
 Once it arrives to the gateway, the \gls{tcp} datagram is extracted from the \gls{ipv6} which is extracted from the \gls{usb} frame and encapsulated in an \gls{ipv4} packet that is encapsulated into a Wi-Fi frame and sent until it reaches the mobile.
-That is said, it clear that the system is heterogeneous and making this heterogeneity transparent was a contribution of this work.
-\paragraph{}
-The \glsreset{wsn}\gls{wsn} was built using the very recent \gls{6lowpan} whose goal is to provide the TCP/IP protocol stack to the smallest devices.
+That is said, it is clear that the system is heterogeneous and making this heterogeneity transparent was a contribution of this work.
+\paragraph{}
+The \gls{wsn} was built using the very recent \gls{6lowpan} whose goal is to provide the TCP/IP protocol stack to the smallest devices.
 The \gls{wsn} is autonomous in the sense that it can monitor the home appliances independently and indefinitely.
@@
 All in all, this project as a big picture creates a test-bed providing the main components of a smart home management with an intelligent \gls{wsn}, a robust and powerful middle ware application that is able to use the sensory data and use it to control the home's appliances and a mobile application able to provide the users with remote monitoring and control of their smart homes.
-This was possible by making the system part of the \glsreset{iot}\gls{iot}.
+This was possible by making the system part of the \gls{iot}.
 \chapter{Future Work}
 \paragraph{}
-The goal of this thesis was to establish the infrastructure for integrating smart home within \glsreset{sg}\gls{sg} and make it part of the \glsreset{iot}\gls{iot}.
-This project enabled the deployment of the two-way communication as an infrastructure that can be used to build the \glsreset{dr}\gls{dr}.
+The goal of this thesis was to establish the infrastructure for integrating smart home within \gls{sg} and make it part of the \gls{iot}.
+This project enabled the deployment of the two-way communication as an infrastructure that can be used to build the \gls{dr}.
 \gls{dr} is one of the most discussed topics in Smart Grids.
@@
 \paragraph{}
-As this project aims to integrate the smart home devices within the \glsreset{iot}\gls{iot}, \glsreset{rfid}\gls{rfid} must be part of this system as a future work.
+As this project aims to integrate the smart home devices within the \gls{iot}, \gls{rfid} must be part of this system as a future work.
 The \gls{rfid} is a building block within the \gls{iot} and will equip the devices with a low-power radio identification mechanism hat will give more flexibility to the system.
@@
 Before building new services on top of this system, one should test different aspects of the system.
-In this study, the only thing that was tested, measured and quantified was the performance of the system which gave good results about the system's overall performance.
+In this study, the only thing that was tested, measured and quantified was the performance of the system which gave accurate results about the system's overall performance.
 As a future work, one should test different aspects of the system such as the reliability and availability of the system.
 More experiments should be conducted to see how well the system acts in high traffic networks and how good could it support new overlying systems such as \gls{dr}.
+\paragraph{}
+The test-bed uses TCP as a transport layer protocol.
+TCP has numerous advantages but the main advantage of being slow compared to UDP.
+One should investigate the benefits of replacing UDP by TCP in the current system and see if it brings any benefits to the overall system's performance without having to lose reliability.
 \paragraph{}
@@
 
-%\backmatter
 \begin{appendices}
@@
 \subitem sudo driver/ip-driver /dev/ttyUSB0 micaz. \\
- You must check that the mote sink is attached to USB0 by using dmesg, if not change it accordingly. After that, you should have an output that looks similar to the figure \ref{fig:tunnel_output}
+ You must check that the mote sink is attached to USB0 by using dmesg, if not change it accordingly. After that, you should have an output that looks similar to the Figure \ref{fig:tunnel_output}
 
